### 2021-09-04 paper/article.tex
Initial paper outline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\begin{document}

\title[Modeling a Neural Co-Processor]
{Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept}

\author{Matthew J Bryan$^{1}$, Linxing Preston Jiang$^{1}$, Rajesh P N Rao$^{1}$}

\address{$^{1}$ Neural Systems Laboratory, Department of Computer
Science and Engineering, University of Washington, Box 352350,
Seattle, WA 98105, USA}

\ead{matthew.bryan@dell.com}
\vspace{10pt}
\begin{indented}
\item[]September 2021
\end{indented}

\begin{abstract}
Brain co-processors\cite{rao.coproc} are devices which use artificial intelligence (AI) and closed-loop
neurostimulation to shape neural activity and to bridge injured neural circuits for
targeted repair and rehabilitation.
\end{abstract}

\vspace{2pc}
\noindent{\it Keywords}: brain-computer interface, neural co-processor, ai, machine learning, stimulation
%
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
%
% Uncomment if a separate title page is required
\maketitle
% 
% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
%\ioptwocol
%

\section{Introduction}

Closed-loop control:
\begin{itemize}
	\item \cite{walker.inception} Inception Loops: driving brain states
	\item \cite{kahana.biomarker} “Closed loop” here refers to when to apply stimulation, always of the same type and at the same site, based on memory performance.
	\item \cite{tafazoli.acls} ACLS
\end{itemize}

Simulation:
\begin{itemize}
	\item \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime
\end{itemize}

\section{Method}
asdf

\section{Results}
asdf

\section{Discussion}
asdf

\section{Conclusion}
asdf

\section{Acknowledgements}
asdf

\section{Ethical Statement}
asdf

\section{References}
\bibliographystyle{iopart-num}
\bibliography{refs}
\end{document}

### 2021-09-05 class_stdevs.py
Add class separation.

@@ -1,0 +1,80 @@
+import torch
+
+
+def collate_data(collected):
+    # Map class to tensor
+    out = {}
+    for label, recs in collected.items():
+
+        t_out = torch.cat(recs, 0)
+        out[label] = t_out
+
+    return out
+
+
+def rotate_dset(raw):
+    # Map class to records list
+    collected = {}
+    for rec in raw:
+        _, _, _, dout, label = rec
+
+        label = label.item()
+        if label not in collected:
+            collected[label] = []
+        collected[label].append(dout.unsqueeze(0))
+
+    return collate_data(collected)
+
+
+def batch_variance(batch):
+    """
+    For a given batch of data, we find the average of stdevs across
+    feature dims and time. We use this to characterize the typical
+    variance within and between object classes.
+    """
+    assert len(batch.shape) == 3
+    std = torch.std(batch, 0)
+    mean = torch.mean(std)
+    mean = torch.mean(mean)
+    return mean
+
+
+def calc_raw_dset_class_vars(dset):
+    """
+    dset: a MichaelsDataset, loaded with with_labels=True
+    """
+
+    dset = rotate_dset(dset)
+    stds = []
+    for label, batch in dset.items():
+        std = batch_variance(batch)
+        stds.append(std.item())
+
+    var_within = sum(stds) / len(stds)
+
+    whole = torch.cat(list(dset.values()), 0)
+    var_whole = batch_variance(whole).item()
+
+    return var_whole, var_within
+
+
+def calc_class_vars(actuals, labels):
+    collected = {}
+    for idx in range(labels.shape[0]):
+        label = labels[idx].item()
+
+        if label not in collected:
+            collected[label] = []
+
+        collected[label].append(actuals[idx:idx+1, :, :])
+
+    collated = collate_data(collected)
+
+    stdevs_within = []
+    for batch in collated.values():
+        std = batch_variance(batch).item()
+        stdevs_within.append(std)
+    var_within = sum(stdevs_within) / len(stdevs_within)
+
+    var_whole = batch_variance(actuals).item()
+    return var_whole, var_within

### 2021-09-05 loss_funcs.py

@@ -4,11 +4,14 @@ import json
 import attr
 import torch.nn
 
+import class_stdevs
+
+
 _G_MSELOSS = torch.nn.MSELoss()
 
 LOSS_RENDER_FMT = "%0.6f"
 def _render_none_or_float(val, fmt=LOSS_RENDER_FMT):
-    if val is None:
+    if val is None or val != val:
         return "-"
     else:
         return fmt % val
@@ -42,10 +45,11 @@ class LossRec:
     pred_val_loss: float
     task_val_loss: float
     pct_recov: float
+    class_separation: float
 
 
 class LossHistory:
-    def __init__(self, lesioned_loss, healthy_loss):
+    def __init__(self, lesioned_loss, healthy_loss, data_set, var_report_cadence=10):
         self._recs = []
 
         self.lesioned_loss = lesioned_loss
@@ -60,6 +64,12 @@ class LossHistory:
         self._last_pred_val_loss = None
         self._last_task_val_loss = None
 
+        var_whole, var_within = \
+                class_stdevs.calc_raw_dset_class_vars(data_set)
+        self._healthy_class_separation = var_whole / var_within
+
+        self._var_report_cadence = var_report_cadence
+
         self._max_pct_recov = 0
 
     def calc_pct_recov(self, task_loss):
@@ -68,16 +78,21 @@ class LossHistory:
         recov_pct = 1.0 - (dh / dl)
         return recov_pct
 
+    def calc_class_vars(self, actuals, labels):
+        var_whole, var_within = class_stdevs.calc_class_vars(actuals, labels)
+        return var_whole, var_within
+
     def render_rec(self, rec):
         rec_rendered = {
             "eidx": rec.eidx,
             "type": rec.type.name,
             "pred_loss": _render_none_or_float(rec.pred_loss),
             "task_loss": _render_none_or_float(rec.task_loss),
-            "train_loss": LOSS_RENDER_FMT % rec.train_loss,
-            "pred_val_loss": LOSS_RENDER_FMT % rec.pred_val_loss,
-            "task_val_loss": LOSS_RENDER_FMT % rec.task_val_loss,
+            "train_loss": _render_none_or_float(rec.train_loss),
+            "pred_val_loss": _render_none_or_float(rec.pred_val_loss),
+            "task_val_loss": _render_none_or_float(rec.task_val_loss),
             "pct_recov": _render_none_or_float(rec.pct_recov, fmt="%0.3f"),
[…]

### 2021-09-05 mRNN.py

@@ -373,7 +373,12 @@ class MichaelsRNN(nn.Module):
 
 
 class MichaelsDataset(Dataset):
-    def __init__(self, data_file_path, with_label=False, limit=None, cuda=None):
+    def __init__(self, data_file_path, with_label=False, limit=None, class_filter=None,
+            cuda=None):
+        """
+        class_filter: optional iterable of class labels to allowlist
+        """
+
         f = michaels_load.load_from_path(data_file_path)
         inps = f["inp"]
         outs = f["targ"]
@@ -391,11 +396,14 @@ class MichaelsDataset(Dataset):
 
         self.with_label = with_label
 
+        self.class_filter = class_filter
+
         self.data = []
-        self._load_data(inps, outs, trial_info=ti, cuda=cuda)
+        self._load_data(inps, outs, trial_info=ti,
+                class_filter=self.class_filter, cuda=cuda)
 
     def __len__(self):
-        return self.num_samples
+        return len(self.data)
 
     def _load_data_single(self, idx, inps, outs, trial_info=None, cuda=None):
         cur_in = inps[idx]
@@ -435,11 +443,15 @@ class MichaelsDataset(Dataset):
 
         return din, trial_end, trial_len, dout, label
 
-    def _load_data(self, inps, outs, trial_info=None, cuda=None):
+    def _load_data(self, inps, outs, trial_info=None, class_filter=None, cuda=None):
+        if trial_info is None and class_filter is not None:
+            raise ValueError("Must provide a trial_info if providing class_filter")
+
         for i in range(self.num_samples):
-            self.data.append(
-                self._load_data_single(i, inps, outs, trial_info=trial_info, cuda=cuda)
-            )
+            datum = self._load_data_single(i, inps, outs, trial_info=trial_info, cuda=cuda)
+
+            if class_filter is None or (datum[-1].item() in class_filter):
+                self.data.append(datum)
 
     def __getitem__(self, idx):
         return self.data[idx]

### 2021-09-05 train.py

@@ -88,6 +88,7 @@ def train_loop(
     din,
     dout,
     trial_end,
+    labels,
     observer,
     loss_history,
     epoch_type,
@@ -113,13 +114,14 @@ def train_loop(
     if is_val:
         loss_history.report_val_last_result(actuals, preds, dout)
     else:
-        loss_history.report_by_result(epoch_type, actuals, preds, dout)
+        loss_history.report_by_result(epoch_type, actuals, preds, dout, labels)
 
     return actuals, preds, stims
 
 
-def train_new_en(
-    mike, observer, cpn, data_loader, loss_history, en_num_neurons=None, cuda=None
+def train_en(
+    mike, observer, cpn, data_loader, loss_history, en=None, opt_en=None,
+    en_num_neurons=None, cuda=None
 ):
     """
     mike: a Michaels modular RNN (a torch Module)
@@ -137,28 +139,33 @@ def train_new_en(
     # Stim: mike.stimulus.num_stim_channels
     # +1 for trial_end
     en_in_dim = obs_dim + mike.stimulus.num_stim_channels + 1
-    en = stim_model.StimModelLSTM(
-        en_in_dim,
-        mike.output_dim,
-        num_neurons=en_num_neurons or (en_in_dim + 50),
-        activation_func=torch.nn.Tanh,
-        cuda=cuda,
-    )
+
+    if en is None:
+        en = stim_model.StimModelLSTM(
+            en_in_dim,
+            mike.output_dim,
+            num_neurons=en_num_neurons or (en_in_dim + 50),
+            activation_func=torch.nn.Tanh,
+            cuda=cuda,
+        )
+
+        assert opt_en is None
+
+        opt_en = AdamW(en.parameters(), lr=9e-3, weight_decay=0.04)
 
     RECENT_EN = en
     vl = torch.tensor(1.0)
-    opt_en = AdamW(en.parameters(), lr=9e-3, weight_decay=0.04)
 
     checkpoint_eidx = 0
     eidx = -1
     while True:
         for batch in data_loader:
-            din, trial_end, _, dout = batch
+            din, trial_end, _, dout, labels = batch
             eidx += 1
             batch_size = din.shape[0]
             opt_en.zero_grad()
 
-            remaining_loss = loss_history.recent_task_loss
+            remaining_loss = loss_history.recent_train_loss
             if remaining_loss is None:
                 # Just some high-ish number for the first time
                 # through this function.
@@ -193,6 +200,7 @@ def train_new_en(
                 din,
[…]

### 2021-09-05 blah.py
Start experimental harness

@@ -1,0 +1,81 @@
+import torch
+from torch.optim import AdamW
+
+import cpn_model
+import experiment
+
+
+class CPN_EN_CoProc(experiment.CoProc):
+    def __init__(self, cfg, retain_stim_grads=False, cuda=None):
+        self.cfg = cfg
+
+        in_dim, stim_dim, out_dim, cuda = cfg.unpack()
+
+        self.cpn = cpn_model.CPNModelLSTM(
+            in_dim,
+            stim_dim,
+            num_neurons=in_dim,
+            activation_func=torch.nn.Tanh,
+            cuda=cuda,
+        )
+
+        en_in_dim = cfg.observer_instance.out_dim + stim_dim + 1
+        self.en = stim_model.StimModelLSTM(
+            en_in_dim,
+            out_dim,
+            num_neurons=en_in_dim + 50,
+            activation_func=torch.nn.Tanh,
+            cuda=cuda,
+        )
+
+        self.opt_cpn = AdamW(self.cpn.parameters(), lr=1e-3)
+        self.opt_en = AdamW(self.en.parameters(), lr=1e-3)
+
+        self.current_noisey_cpn = None
+        self.retain_stim_grads = retain_stim_grads
+
+        self.stims = None
+        self.preds = None
+        self.reset()
+
+    def reset(self):
+        self.stims = []
+        self.preds = torch.tensor
+
+    def ensure_preds(self, ...):
+        pass
+
+    def forward(self, brain_data):
+        cpn_in = torch.cat(brain_data, axis=1).detach()
+
+        new_stim = self.cpn(cpn_in)
+        if self.retain_stim_grads:
+            new_stim.retain_grad()
+        self.stims.append(new_stim)
+
+        # en receives (obs, stims, trial_end)
+        # (detaching just in case; we don't want to backprop here)
+        new_obs_en = brain_data[0].detach()
+        en_in = torch.cat((new_obs_en, new_stim, obs[-1]), axis=1)
+        cur_pred = en(en_in)
+
+        preds[:, tidx, :] = cur_pred[:, :]
+
+
+        raise NotImplementedError()
+
+    def feedback(self, actuals, targets, loss_history):
+        """
+        Args:
+            loss_history: a loss.LossHistory
+        Returns:
+            True to stop training; False otherwise
+        """
+        raise NotImplementedError()
+
+    def finish(self):
+        """
+        Called between batches. User can add any cleanup logic here, e.g.
+        resetting internal state of RNNs.
+        """
+        pass

### 2021-09-05 config.py

@@ -1,12 +1,7 @@
-import enum
 import logging
-import sys
+import typing
 
-LOG_FORMAT = "%(asctime)s %(message)s"
-LOG_DATEFMT = "%m-%d %H:%M:%S"
-logging.basicConfig(format=LOG_FORMAT, datefmt=LOG_DATEFMT)
-
-
+import attr
 import torch
 import torch.nn
 
@@ -14,32 +9,79 @@ import activation
 import lesion
 import observer
 import stim
-import utils
+
+
+LOG_FORMAT = "%(asctime)s %(message)s"
+LOG_DATEFMT = "%m-%d %H:%M:%S"
+logging.basicConfig(format=LOG_FORMAT, datefmt=LOG_DATEFMT)
+
+
+@attr.s(auto_attribs=True, frozen=True)
+class Config:
+    observer_instance: observer.Observer
+    stim_instance: stim.Stimulus
+    lesion_instance: lesion.Lesion
+    en_activation: torch.nn.Module
+    cpn_activation: torch.nn.Module
+    cfg_str: str
+    cfg_str_short: str
+    out_dim: int
+    cuda: typing.Any
+
+    def unpack(self):
+        # 3x due to 3 modules in the mRNN
+        # +1 for trial_end
+        return (
+            3 * self.observer_instance.out_dim + 1,
+            self.stim_instance.out_dim,
+            self.out_dim,
+            self.cuda,
+        )
+
 
 DEFAULT_OBSERVER_TYPE = observer.ObserverType.gaussian
 DEFAULT_STIMULATION_TYPE = stim.StimulationType.gaussian_exp
 DEFAULT_LESION_TYPE = lesion.LesionType.connection
-DEFAULT_LESION_MODULE_ID = lesion.LesionModule.M1
-DEFAULT_LESION_PCT = 1.0
+DEFAULT_LESION_ARGS = (
+    [
+        # No F5->M1
+        (1, 2, 0, 1),
+        # No M1->F5
+        (0, 1, 1, 2),
+    ],
+)
 DEFAULT_ACTIVATION_TYPE = activation.ActivationType.Tanh
 DEFAULT_NUM_NEURONS_PER_MODULE = 100
-DEFAULT_BATCH_SIZE = 64
+DEFAULT_OBS_OUT_DIM = 20
+DEFAULT_OBS_SIGMA = 1.75
+DEFAULT_NUM_STIM_CHANNELS = 16
[…]

### 2021-09-05 experiment.py

@@ -1,0 +1,224 @@
+import logging
+import os
+
+import torch
+from torch.utils.data import DataLoader
+
+
+# Local imports
+import config
+import michaels_load
+import mRNN
+import utils
+
+
+# TODO: loss_history = run(co_proc, loss_history=None)
+# TODO: score(loss_history)
+
+g_logger = logging.getLogger("experiment")
+
+
+def get_config(cuda=None):
+    """
+    Args:
+        - cuda (str, torch.device, or None): None for CPU, or a string like "0" specifying a GPU
+                              device. This will be passed to the CUDA_VISIBLE_DEVICES env var.
+                              Reference: https://docs.nvidia.com/cuda/cuda-c-programming-guide/index.html#env-vars
+    Returns:
+        - in_dim (int): dimensionality of the data coming from the brain
+        - stim_dim (int): dimensionality of stimulation vector, which we get
+                          from the co-proc
+        - out_dim (int): dimensionality of the task: muscle velocities in our
+                         case
+        - cuda (torch.device, or None): will be a torch.device if we are using
+                                        GPU. Note your model/Module will need to
+                                        use this to push any parameters to GPU
+                                        if the experiment is running on GPU.
+    """
+    if isinstance(cuda, str):
+        os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID"
+        os.environ["CUDA_VISIBLE_DEVICES"] = cuda
+        cuda_out = torch.device(0)
+    elif cuda is None:
+        cuda_out = None
+    else:
+        cuda_out = cuda
+
+    cfg = config.get_default(cuda=cuda_out)
+
+    return cfg
+
+
+def get_raw_data(cuda=None, **kwargs):
+    dataset = mRNN.MichaelsDataset(
+        michaels_load.get_default_path(), cuda=cuda, with_label=True, **kwargs
+    )
+    return dataset
+
+
+class CoProc:
+    def forward(self, brain_data):
+        """
+        Args:
+            brain_data: 4-tuple (M1 data, F5 data, AIP data, trial_end signal)
+                        The first three are Tensors (batch_size, cfg.observer_instance.out_dim)
+                        The last is (batch_size, 1)
+                        The sum of second dimensions is in_dim.
+        Returns:
+            a Tensor(batch_size, stim_dim)
+        """
+        raise NotImplementedError()
+
[…]

### 2021-09-05 stim.py

@@ -17,7 +17,7 @@ class Stimulus(object):
         self._batch_size = batch_size
 
     @property
-    def num_stim_channels(self):
+    def out_dim(self):
         return self._num_stim_channels
 
     @property
@@ -169,7 +169,7 @@ class StimulusGaussian(Stimulus1to1):
         return torch.tensor(wout)
 
     def __str__(self):
-        return f"gaussian{self.num_stim_channels}.{self._sigma}"
+        return f"gaussian{self.out_dim}.{self._sigma}"
 
 
 class StimulusGaussianExp(Stimulus):
@@ -277,7 +277,7 @@ class StimulusGaussianExp(Stimulus):
         return stim_out
 
     def __str__(self):
-        return f"gaussianExp{self.num_stim_channels}.{self._sigma}"
+        return f"gaussianExp{self.out_dim}.{self._sigma}"
 
 
 class StimulationType(enum.Enum):

### 2021-09-25 paper/article.tex
Paper with complete abstract

@@
 \begin{abstract}
-Brain co-processors\cite{rao.coproc} are devices which use artificial intelligence (AI) and closed-loop
-neurostimulation to shape neural activity and to bridge injured neural circuits for
-targeted repair and rehabilitation.
+\textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
+intelligence (AI) and closed-loop neurostimulation to shape neural activity and to bridge
+injured neural circuits for targeted repair and rehabilitation. The co-processor framework
+offers a flexible approach to learning stimulation policies that optimize for
+(a) specific regimes of neural activities, or (b) external task performance.
+For example, it may seek to learn a complex stimulation policy, to aid a
+stroke victim in reaching to grasp an object.
+Through the use of artificial neural networks (ANNs) and deep learning, the
+co-processor co-adapts with the neural circuit, allowing it to seek optimal
+stimulation policies, and adapt them as the neural circuit changes. The results
+presented here demonstrate a neural co-processor for the first time through
+the use of a simulation, and explore some of the core algorithms that may allow them
+to succeed. \textit{Approach} We provide the first proof-of-concept of a neural
+co-processor, through the use of a simulated neural circuit, based on (ANNs).
+The simulated circuit performs a reach-to-grasp task, based on visual
+input which is designed to closely resemble a similar circuit in a primate brain
+\cite{michaels.mrnn}. We simulate a variety of lesions by altering the model, and
+then demonstrate a co-processor's ability to restore lost function through
+``stimulation'' of that model. We further test the ability of a
+co-processor to adapt its stimulation as the simulated brain
+undergoes changes. \textit{Main results} Our simulated co-processor
+successfully co-adapts with the neural circuit to accomplish the external reaching
+task. The co-processor framework demonstrated here adapts to a variety of lesion
+types, and to ongoing changes in the simulated brain. \textit{Significance} The proof-of-concept here
+outlines a co-processor model, as well as our approach to training it, leading to
+insights on how such a model may be developed for \textit{in vivo} use. We believe this
+co-processor design will allow for learning complex stimulation policies that
+help restore function to a stroke victim. This initial proof-of-concept lays
+the foundation for a future demonstration \textit{in vivo}.
 \end{abstract}
@@
 
-Closed-loop control:
+* Introduce the notion of a co-processor
+
+* BCIs: a single direction
+* Increasingly sophisticated ML techniques applied to decoding
+** See 2019Raj
+* Encoding models, stimulation models (cite)
+
+* Bi-directional BCIs (BBCIs)
+** Closed loop control
 \begin{itemize}
@@
 \end{itemize}
+
+Need for more complex stimulation policies, and need for optimizing task performance.
+
+Demonstration by way of simulation.
 
@@
 \section{Method}
-asdf
+
+* Overview of test bed
+** Michaels model
+** Lesion designs
+** Stimulation design
+
+* EN architecture
+* CPN architecture
 
 \section{Results}
-asdf
+
+* Task performance improves drastically
+* Fine tuning takes a long time
+* Object classes separate
+* Can co-adapt with the brain, as it changes
+
+* White noise and noised params
+** Learning instability
+** Predictive power isn't enough
 
 \section{Discussion}
-asdf
+* Training efficiency
+* Spectrum from simple low dimensional stimulation vectors today to higher dimensional future
+* Toward in vivo application
 
@@
 \section{Acknowledgements}
-asdf
+Ganguly, Priya, Anca, Justin, Luciano
 

### 2021-09-27 paper/article.tex
Paper progress

@@
 stimulation policies, and adapt them as the neural circuit changes. The results
-presented here demonstrate a neural co-processor for the first time through
+presented here demonstrate a neural co-processor for the first time, through
 the use of a simulation, and explore some of the core algorithms that may allow them
 to succeed. \textit{Approach} We provide the first proof-of-concept of a neural
-co-processor, through the use of a simulated neural circuit, based on (ANNs).
-The simulated circuit performs a reach-to-grasp task, based on visual
-input which is designed to closely resemble a similar circuit in a primate brain
-\cite{michaels.mrnn}. We simulate a variety of lesions by altering the model, and
+co-processor, through the use of a simulated neural circuit, which  performs a
+reach-to-grasp task, based on visual input. It is designed to closely resemble
+a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate a
+variety of lesions by altering the model, and
 then demonstrate a co-processor's ability to restore lost function through
@@
 \section{Introduction}
+* Aided in part by application of advanced AI techniques, BCIs have made
+advancements over the last several decads, allowing for decoded
+brain signals to be used for control of a wide variety of virtual and physical
+prostheses.
+** Separately: advances in stimulation techniques and modeling have allowed us to probe neural circuit dynamics and learn to better drive neural circuits towards target dynamics (citations)
+** Increasing interesting in combining decoding and encoding, for closed-loop control. BBCIs
+** They have the potential to apply real-time, fine-grained control of neural circuits.
+** Allows stimulation to be conditioned on a user's intention, sensory feedback, error signals, etc., to drive e.g. improvement in motor control after impairment due to TBI.
 
 * Introduce the notion of a co-processor
+** Use of AI and deep learning to determine a stimulation policy.
+** Co-adapts policy to solve a task, together with the neural circuit
+** Flexible model, allows for optimization of either brain activities or external task
+** "These ``neural co-processors'' can be used to jointly optimize cost functions with the
+   nervous system to achieve desired behaviors ranging from targeted neuro-rehabilitation
+   to augmentation of brain function."
+** Proposed approach here combines a learned stimulation model with a stimulation agent
+** Learning allows for active adaptation to the neural circuit as it changes.
 
@@
 * Increasingly sophisticated ML techniques applied to decoding
-** See 2019Raj
+** See 2019Raj citations
 * Encoding models, stimulation models (cite)
+
+* Open loop control examples.  \cite{khanna.openloop}
+
+* Learning stimulation models \cite{shanechi.stimmodel}
 
@@
 
-Need for more complex stimulation policies, and need for optimizing task performance.
+* Need for more complex stimulation policies, and need for optimizing task performance.
+** i.e. what if we don't have a specific pre-recorded activation we want to drive towards?
+** And how do we find activations that drive an intended task solution?
 
-Demonstration by way of simulation.
-
-Simulation:
-\begin{itemize}
-	\item \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime
-\end{itemize}
+Begin with a proof-of-concept in simulation.
 
@@
 
+* Proof-of-concept by way of simulation.
+** Simulation allows for rapid, cheap exploration of learning algorithms.
+** Potentially helpful to look for a simulated neural circuit which has internal dynamics, i.e. information propagation, which is demonstrably similar to natural circuits.
+** \cite{kao.sim} Likeness of RNNs to natural circuits re: dynamics
+** Cite Michaels, and those upstream from him (Susillo?) re: RNNs
+** \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime. Use of a biomimetic
+  simulation to develop and evaluate a neural controller. Use of lesion model.
+
 * Overview of test bed
-** Michaels model
+** Michaels model \cite{michaels.mrnn}
 ** Lesion designs
+*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
+*** Disconnect modules vs lesion M1, and the effects
 ** Stimulation design
+*** Spatio-temporal smoothing
 

### 2021-10-03 paper/article.tex
Paper progress

@@
 \textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
-intelligence (AI) and closed-loop neurostimulation to shape neural activity and to bridge
+intelligence (AI) for closed-loop neurostimulation to shape neural activity and to bridge
 injured neural circuits for targeted repair and rehabilitation. The co-processor framework
 offers a flexible approach to learning stimulation policies that optimize for
-(a) specific regimes of neural activities, or (b) external task performance.
+(a) specific regimes of neural activity, or (b) external task performance.
 For example, it may seek to learn a complex stimulation policy, to aid a
-stroke victim in reaching to grasp an object.
-Through the use of artificial neural networks (ANNs) and deep learning, the
-co-processor co-adapts with the neural circuit, allowing it to seek optimal
-stimulation policies, and adapt them as the neural circuit changes. The results
-presented here demonstrate a neural co-processor for the first time, through
-the use of a simulation, and explore some of the core algorithms that may allow them
-to succeed. \textit{Approach} We provide the first proof-of-concept of a neural
-co-processor, through the use of a simulated neural circuit, which  performs a
-reach-to-grasp task, based on visual input. It is designed to closely resemble
-a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate a
-variety of lesions by altering the model, and
-then demonstrate a co-processor's ability to restore lost function through
-``stimulation'' of that model. We further test the ability of a
-co-processor to adapt its stimulation as the simulated brain
-undergoes changes. \textit{Main results} Our simulated co-processor
-successfully co-adapts with the neural circuit to accomplish the external reaching
-task. The co-processor framework demonstrated here adapts to a variety of lesion
-types, and to ongoing changes in the simulated brain. \textit{Significance} The proof-of-concept here
-outlines a co-processor model, as well as our approach to training it, leading to
-insights on how such a model may be developed for \textit{in vivo} use. We believe this
-co-processor design will allow for learning complex stimulation policies that
-help restore function to a stroke victim. This initial proof-of-concept lays
-the foundation for a future demonstration \textit{in vivo}.
+stroke victim in reaching to grasp an object. Through the use of artificial neural
+networks (ANNs) and deep learning, the co-processor co-adapts with the
+neural circuit, allowing it to seek optimal stimulation policies, and adapt them
+as the neural circuit changes. The results presented here demonstrate a
+neural co-processor for the first time, through the use of a simulation, and
+explore some of the core algorithms that may allow them to succeed.
+\textit{Approach} We provide the first proof-of-concept of a neural co-processor,
+through the use of a simulated neural circuit, which  performs a reach-to-grasp task,
+based on visual input. It is designed to closely resemble a similar circuit in
+a primate brain \cite{michaels.mrnn}. We simulate a variety of lesions by altering
+the model, and then demonstrate a co-processor's ability to restore lost function
+through ``stimulation'' of that model. We further test the ability of a co-processor to
+adapt its stimulation as the simulated brain undergoes changes.
+\textit{Main results} Our simulated co-processor successfully co-adapts with the neural
+circuit to accomplish the external reaching task. The co-processor framework
+demonstrated here adapts to a variety of lesion types, and to ongoing changes in the
+simulated brain.
+\textit{Significance} The proof-of-concept here outlines a co-processor model, as well
+as our approach to training it, leading to insights on how such a model may be
+developed for \textit{in vivo} use. We believe this co-processor design will allow for
+learning complex stimulation policies that help restore function to a stroke victim.
 \end{abstract}
@@
 \section{Introduction}
-* Aided in part by application of advanced AI techniques, BCIs have made
-advancements over the last several decads, allowing for decoded
-brain signals to be used for control of a wide variety of virtual and physical
-prostheses.
-** Separately: advances in stimulation techniques and modeling have allowed us to probe neural circuit dynamics and learn to better drive neural circuits towards target dynamics (citations)
-** Increasing interesting in combining decoding and encoding, for closed-loop control. BBCIs
-** They have the potential to apply real-time, fine-grained control of neural circuits.
-** Allows stimulation to be conditioned on a user's intention, sensory feedback, error signals, etc., to drive e.g. improvement in motor control after impairment due to TBI.
+Aided in part by application of advanced AI techniques, brain-computer interfaces (BCIs) have made
+advancements over the last several decads, allowing for decoded brain signals to be used for
+control of a wide variety of virtual and physical prostheses \cite{rao.bcibook} \cite{wolpaw.bcibook} \cite{moritz.neuro}
+\cite{lebedev.bmi}. Separately: advances in stimulation techniques and modeling have allowed us to probe
+neural circuit dynamics \textbf{TODO: citations} and learn to better drive neural circuits towards target
+dynamics, by encoding and delivering information through stimulation \textbf{TODO: citations}.
+Recently, there has been increasing interest in building on these advances to combine decoding
+and encoding in a single system, for closed-loop stimulation of a neural circuit. Bi-directional
+BCIs (BBCIs) allow stimulation to be conditioned by decoded brain activity as well as external
+sensor data (e.g. camera), which can allow for the application of real-time, fine-grained control of
+neural circuits and prosthetic devices \textbf{TODO: citations}. These may lead, for example, to neuro-
+prostheses that are capable of restoring movement which was lost due to traumatic brain injury (TBI),
+to a degree not previously possible.
 
-* Introduce the notion of a co-processor
-** Use of AI and deep learning to determine a stimulation policy.
-** Co-adapts policy to solve a task, together with the neural circuit
-** Flexible model, allows for optimization of either brain activities or external task
-** "These ``neural co-processors'' can be used to jointly optimize cost functions with the
-   nervous system to achieve desired behaviors ranging from targeted neuro-rehabilitation
-   to augmentation of brain function."
-** Proposed approach here combines a learned stimulation model with a stimulation agent
-** Learning allows for active adaptation to the neural circuit as it changes.
+Motivated by that progress, we demonstrate here a flexible and unifying framework for combining encoding
+and decoding, which we term ``neural co-processors'' \cite{rao.coproc}. Neural co-processors leverage
+AI and deep learning to identify optimal, closed-loop stimulation patterns. The approach is flexible
+enough to optimize not only for particular neural activities, but also for tasks external to the
+subject. For example, they may be able to aid a stroke victim by finding a stimulation pattern of
+the motor cortex which helps restore lost limb function.
 
-* BCIs: a single direction
-* Increasingly sophisticated ML techniques applied to decoding
-** See 2019Raj citations
-* Encoding models, stimulation models (cite)
+Additionally, the co-processor framework allows a neuro-prosthesis to actively adapt to
+a neural circuit as it changes with time. This framework is capable of co-adapting with
+the circuit, i.e. brain, by updating its stimulation regime, while at the same time the
+brain is updating its response to the stimulation, and changing due to natural plasticity,
+aging, etc. This allows the co-processor to continually optimize for the intended cost
+function, despite the signficant non-stationarity of the target circuit.
 
-* Open loop control examples.  \cite{khanna.openloop}
+Here we provide a proof-of-concept in simulation for a co-processor that restores
+movement to a limb, after a subject has suffered a stroke affecting its ability to
+use that limb. It combines:
+\begin{itemize}
+	\item A stimulation model, which models the relationship between decoded brain activity,
+	      stimulation, and task performance.
+	\item An AI agent which determines the stimulation to apply, in real time.
+\end{itemize}
+Proposed approach here combines a learned stimulation model with a stimulation agent
 
-* Learning stimulation models \cite{shanechi.stimmodel}
+* Encoding models \cite{shanechi.stimmodel} (others?)
+** Decades of advancement in encoding models, used for:
+*** Better understanding the brain
+*** Learning to understand the effects fo stimulation regimes
+** Optogenetic vs TMS vs eCog
+** In our case, designed to aid in the training of the stimulation agent.
+
+* Open loop control
+** Example: \cite{khanna.openloop}
 
@@
 	\item \cite{tafazoli.acls} ACLS
+	\item cite Bolus
 \end{itemize}
@@
 *** Spatio-temporal smoothing
+** Observation model
 
@@
 * CPN architecture
+
+* Training
+** Alternate training EN and CPN
+*** Train an EN
+**** Training the EN until predictive power reaches a threshold, based on current task performance
+**** Training regime based on most recent CPN, similar CPNS (noise added to params), and white noise stimulation
+**** Single batch of data from applying current CPN set and white noise to the brain, EN fit to ``offline'' it over maining epochs
+**** Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
+*** Train the CPN
+**** Backprop through the EN, effectively using the EN's prediction as ground truth
 
@@
 
+* Split by lesion design
+
+* Show example where AIP/F5 lesioned and how that affects performance
+
+* Any point in showing sensitivity to observation or stim dim?
+
+* Try regularization variation?
+
 * White noise and noised params
@@
 ** Predictive power isn't enough
+** Illustrate this
+
+* Training efficiency analysis
 

### 2021-10-04 cpn.py
EN closed loop / offline training

@@ -1,3 +1,4 @@
+import logging
 import os
 import json
 import uuid
@@ -12,6 +13,8 @@ import cpn_epoch_en
 from cpn_utils import EpochType
 from experiment import experiment
 import stim_model
+
+g_logger = logging.getLogger("cpn")
 
 
 class CPN_EN_CoProc(experiment.CoProc):
@@ -35,6 +38,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         )
 
         self.stims = None
+        self.brain_data = None
 
         self.epoch_type = EpochType.EN
         for param in self.cpn.parameters():
@@ -51,15 +55,23 @@ class CPN_EN_CoProc(experiment.CoProc):
 
         self.uuid = uuid.uuid4()
         if log_dir is not None:
-            self.log_dir = os.path.join(log_dir, str(self.uuid))
+            self.log_dir = os.path.join(log_dir, cfg.cfg_str, str(self.uuid))
             os.makedirs(self.log_dir)
         else:
             self.log_dir = None
+
+        self.round_thresh = 3
+        # A list of up to round_thresh length
+        #  Each item is a tuple (actuals, targets, trial_end, stim list, obs list)
+        #   The list is trial_len length.
+        #     Each item is a stim or obs
+        self.saved_data = []
 
         self.reset_soft()
 
     def reset_soft(self):
         self.stims = []
+        self.brain_data = []
         self.en_epoch.reset()
         self.cpn_epoch.reset()
 
@@ -79,16 +91,19 @@ class CPN_EN_CoProc(experiment.CoProc):
 
     def forward(self, brain_data, loss_history):
         if self.epoch_type in (EpochType.EN, EpochType.VAL_EN):
-            new_stim = self.en_epoch.forward(brain_data, loss_history,
-                    self.epoch_type == EpochType.VAL_EN)
+            new_stim = self.en_epoch.forward(
+                brain_data, loss_history, self.epoch_type == EpochType.VAL_EN
+            )
         elif self.epoch_type in (EpochType.CPN, EpochType.VAL_CPN):
-            new_stim = self.cpn_epoch.forward(brain_data, loss_history,
-                    self.epoch_type == EpochType.VAL_CPN)
+            new_stim = self.cpn_epoch.forward(
+                brain_data, loss_history, self.epoch_type == EpochType.VAL_CPN
+            )
         else:
             # For now: no other thing
             raise ValueError(self.epoch_type)
 
[…]

### 2021-10-04 cpn_epoch_en.py

@@ -46,6 +46,7 @@ class CPNEpochEN:
 
     def reset_en(self):
         self.en, self.opt_en = self.new_en(self.en)
+
         return self.en, self.opt_en
 
     def new_en(self, old_en):
@@ -73,17 +74,20 @@ class CPNEpochEN:
         if self.preds is None:
             self.preds = []
 
-    def forward(self, brain_data, loss_history, is_validation):
+    def forward(self, brain_data, loss_history, is_validation, stim=None):
         batch_size = brain_data[0].shape[0]
         self.ensure_preds(batch_size)
 
-        cpn_in = torch.cat(brain_data, axis=1).detach()
+        if stim is None:
+            cpn_in = torch.cat(brain_data, axis=1).detach()
 
-        if is_validation:
-            new_stim = self.cpn(cpn_in)
+            if is_validation:
+                new_stim = self.cpn(cpn_in)
+            else:
+                self.ensure_noisey_cpn(batch_size)
+                new_stim = self.cpn_noise(cpn_in)
         else:
-            self.ensure_noisey_cpn(batch_size)
-            new_stim = self.cpn_noise(cpn_in)
+            new_stim = stim
 
         # en receives (obs, stims, trial_end)
         # (detaching just in case; we don't want to backprop here)
@@ -110,7 +114,7 @@ class CPNEpochEN:
             pred_loss.backward(inputs=list(self.en.parameters()))
             self.opt_en.step()
 
-    def finish(self, loss_history, is_validation):
+    def finish(self, loss_history, is_validation, reused_data=False):
 
         vl = self.recent_pred_val_loss
         for p in self.opt_en.param_groups:
@@ -147,7 +151,7 @@ class CPNEpochEN:
         elif (
             vl < max(0.02 * self.recent_train_loss, 0.0003)
             and self.checkpoint_eidx > 100
-        ) or self.checkpoint_eidx == 2000:
+        ) or self.checkpoint_eidx >= 2000:
             en_is_ready = True
             self.checkpoint_eidx = 0
         else:
@@ -155,7 +159,7 @@ class CPNEpochEN:
             en_is_ready = False
 
         user_data = CPNENStats(
-            "en",
+            "en_offline" if reused_data else "en",
             EpochType.EN,
             train_loss_out,
             train_val_loss_out,

### 2021-10-06 cpn.py
Cleanup offline loop and logging.

@@ -60,7 +60,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         else:
             self.log_dir = None
 
-        self.round_thresh = 3
+        self.round_thresh = 1
         # A list of up to round_thresh length
         #  Each item is a tuple (actuals, targets, trial_end, stim list, obs list)
         #   The list is trial_len length.
@@ -144,7 +144,9 @@ class CPN_EN_CoProc(experiment.CoProc):
         is_validation = next_is_validation
 
         done = False
-        while not done:
+        checkpoint_eidx = 0
+        en_is_ready = False
+        while not en_is_ready and checkpoint_eidx < 1000:
             for bidx in range(self.round_thresh):
                 actuals, targets, trial_end, stims, brain_data = self.saved_data[bidx]
 
@@ -185,13 +187,16 @@ class CPN_EN_CoProc(experiment.CoProc):
                 loss_history.report_user_data(user_data)
                 loss_history.log(g_logger, msg=user_data.msg)
 
-                done = en_is_ready
                 is_validation = next_is_validation
 
-                if done:
+                if en_is_ready:
                     break
 
+                checkpoint_eidx += 1
+
         self.saved_data = []
+        self.en_epoch.reset_period()
+        return en_is_ready
 
     def finish(self, loss_history):
         if self.epoch_type in (EpochType.EN, EpochType.VAL_EN):
@@ -220,16 +225,19 @@ class CPN_EN_CoProc(experiment.CoProc):
                 self.epoch_type = EpochType.EN
 
             if len(self.saved_data) == self.round_thresh and not en_is_ready:
-                self.train_en_closed_loop(loss_history, next_is_validation)
-                en_is_ready = True
+                en_is_ready = self.train_en_closed_loop(loss_history,
+                    next_is_validation)
 
-                self.epoch_type = EpochType.CPN
-                self.cpn_epoch.set_en(self.en, self.opt_en)
+                if en_is_ready:
+                    self.epoch_type = EpochType.CPN
+                    self.cpn_epoch.set_en(self.en, self.opt_en)
 
-                for param in self.cpn.parameters():
-                    param.requires_grad = True
-                for param in self.en.parameters():
-                    param.requires_grad = False
+                    for param in self.cpn.parameters():
+                        param.requires_grad = True
+                    for param in self.en.parameters():
+                        param.requires_grad = False
+                else:
[…]

### 2021-10-06 cpn_epoch_en.py

@@ -37,6 +37,9 @@ class CPNEpochEN:
         self.tidx = 0
         self.reset_models()
 
+    def reset_period(self):
+        self.checkpoint_eidx = 0
+
     def reset_models(self):
         self.en.reset()
         self.cpn.reset()
@@ -144,23 +147,23 @@ class CPNEpochEN:
 
         if (
             vl != vl or vl == float("inf") or vl > 1.5
-        ):  # or self.checkpoint_eidx > 5000:
+        ):
             self.en, self.opt_en = self.new_en(self.en)
             en_is_ready = False
-            self.checkpoint_eidx = 0
+            self.reset_period()
         elif (
             vl < max(0.02 * self.recent_train_loss, 0.0003)
             and self.checkpoint_eidx > 100
         ) or self.checkpoint_eidx >= 2000:
             en_is_ready = True
-            self.checkpoint_eidx = 0
+            self.reset_period()
         else:
             self.checkpoint_eidx += 1
             en_is_ready = False
 
         user_data = CPNENStats(
             "en_offline" if reused_data else "en",
-            EpochType.EN,
+            EpochType.EN_OFFLINE if reused_data else EpochType.EN,
             train_loss_out,
             train_val_loss_out,
             self.recent_pred_loss,

### 2021-10-06 cpn_utils.py

@@ -9,10 +9,11 @@ from experiment.utils import render_none_or_float
 
 class EpochType(enum.Enum):
     EN = 0
-    CPN = 1
-    CPN_AND_EN = 2
-    VAL_EN = 3
-    VAL_CPN = 4
+    EN_OFFLINE = 1
+    CPN = 2
+    CPN_AND_EN = 3
+    VAL_EN = 4
+    VAL_CPN = 5
 
 
 @attr.s(auto_attribs=True)

### 2021-10-06 experiment/config.py

@@ -223,9 +223,9 @@ def get(
         f"cpnAct{cpn_activation_type.name}",
     ]
     if recover_after_lesion:
-        cfg_toks += "recovPre"
+        cfg_toks.append("recovPre")
     if coadapt:
-        cfg_toks += "coadapt"
+        cfg_toks.append("coadapt")
 
     cfg_str = "_".join(cfg_toks)
 
@@ -237,9 +237,9 @@ def get(
         f"cpnAct{cpn_activation_type.name}",
     ]
     if recover_after_lesion:
-        cfg_toks_short += "recovPre"
+        cfg_toks_short.append("recovPre")
     if coadapt:
-        cfg_toks_short += "coadapt"
+        cfg_toks_short.append("coadapt")
     cfg_str_short = "_".join(cfg_toks_short)
 
     dataset, loader_train, loader_test = _get_dataset(

### 2021-10-06 experiment/experiment.py

@@ -45,8 +45,12 @@ def get_m1_lesion_config(cuda=None, **kwargs):
     lesion_type = lesion.LesionType.outputs
     lesion_args = (0, 50)
     return get_config(
-        cuda=cuda, coadapt=True, drop_m1=True, lesion_type=lesion_type,
-        lesion_args=lesion_args, **kwargs
+        cuda=cuda,
+        coadapt=True,
+        drop_m1=True,
+        lesion_type=lesion_type,
+        lesion_args=lesion_args,
+        **kwargs,
     )
 
 
@@ -138,7 +142,7 @@ class CoProc:
         """
         pass
 
-    def report(self, loss_history):
+    def report(self, loss_history, mike):
         """
         Final chance to report/calc stats, records, logs, etc., after finish.
         This step is separate from finish(), since the most recent loss history
@@ -276,8 +280,8 @@ class Experiment:
     def _coproc_finish(self, loss_history):
         return self.coproc.finish(loss_history)
 
-    def _coproc_report(self, loss_history):
-        return self.coproc.report(loss_history)
+    def _coproc_report(self, loss_history, mike):
+        return self.coproc.report(loss_history, mike)
 
     def run(self):
         is_validation = False
@@ -365,7 +369,7 @@ class Experiment:
 
             self.loss_history.log(g_logger, msg=msg)
 
-            self._coproc_report(self.loss_history)
+            self._coproc_report(self.loss_history, self.mike)
 
             if should_stop:
                 break

### 2021-10-30 paper/article.tex
Paper progress

@@
 \textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
-intelligence (AI) for closed-loop neurostimulation to shape neural activity and to bridge
+intelligence (AI) for closed-loop neurostimulation, to shape neural activity and to bridge
 injured neural circuits for targeted repair and rehabilitation. The co-processor framework
-offers a flexible approach to learning stimulation policies that optimize for
+offers a flexible approach to learning closed-loop stimulation policies that optimize for
 (a) specific regimes of neural activity, or (b) external task performance.
-For example, it may seek to learn a complex stimulation policy, to aid a
-stroke victim in reaching to grasp an object. Through the use of artificial neural
+For example, it may seek to learn to stimulate the motor cortex of a stroke patient,
+conditioning the stimulation on upstream visual information, aiding the patient's
+attempt to grasp an object. Through the use of artificial neural
 networks (ANNs) and deep learning, the co-processor co-adapts with the
@@
 as the neural circuit changes. The results presented here demonstrate a
-neural co-processor for the first time, through the use of a simulation, and
-explore some of the core algorithms that may allow them to succeed.
-\textit{Approach} We provide the first proof-of-concept of a neural co-processor,
-through the use of a simulated neural circuit, which  performs a reach-to-grasp task,
-based on visual input. It is designed to closely resemble a similar circuit in
-a primate brain \cite{michaels.mrnn}. We simulate a variety of lesions by altering
-the model, and then demonstrate a co-processor's ability to restore lost function
-through ``stimulation'' of that model. We further test the ability of a co-processor to
-adapt its stimulation as the simulated brain undergoes changes.
+neural co-processor for the first time, through the use of a simulation. We
+explore some of the core algorithms that may allow co-processors to successfully
+learn and to adapt to non-stationarity in both the brain and sensors.
+\textit{Approach} We provide the first proof-of-concept of a neural co-processor that
+leverages deep learning, through the use of a simulated neural circuit
+That circuit performs reach-to-grasp task, based on visual input, and is designed to
+closely resemble a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate
+a variety of lesions by altering the model, and then demonstrate a co-processor's ability
+to restore lost function through ``stimulation'' of that model. We further test the
+ability of a co-processor to adapt its stimulation as the simulated brain undergoes changes.
 \textit{Main results} Our simulated co-processor successfully co-adapts with the neural
@@
 advancements over the last several decads, allowing for decoded brain signals to be used for
-control of a wide variety of virtual and physical prostheses \cite{rao.bcibook} \cite{wolpaw.bcibook} \cite{moritz.neuro}
-\cite{lebedev.bmi}. Separately: advances in stimulation techniques and modeling have allowed us to probe
-neural circuit dynamics \textbf{TODO: citations} and learn to better drive neural circuits towards target
-dynamics, by encoding and delivering information through stimulation \textbf{TODO: citations}.
+control of a wide variety of virtual and physical prostheses \cite{rao.bcibook, wolpaw.bcibook,
+moritz.neuro, lebedev.bmi}. Separately: advances in stimulation techniques and modeling have allowed
+us to probe neural circuit dynamics (e.g. \cite{walker.inception}) and learn to better drive neural
+circuits towards target dynamics, by encoding and delivering information through
+stimulation \cite{niparko.cochlear, weiland.retinal, tomlinson.propr, tabot.tact, tyler.tact,
+dadarlat.tact, sharlene.tact, cronin.tact}.
 Recently, there has been increasing interest in building on these advances to combine decoding
@@
 sensor data (e.g. camera), which can allow for the application of real-time, fine-grained control of
-neural circuits and prosthetic devices \textbf{TODO: citations}. These may lead, for example, to neuro-
-prostheses that are capable of restoring movement which was lost due to traumatic brain injury (TBI),
-to a degree not previously possible.
+neural circuits and prosthetic devices, e.g. Nicolelis et al. \cite{nicolelis.bmbi}. These may lead,
+for example, to neuro-prostheses that are capable of restoring movement which was lost due to
+traumatic brain injury (TBI), to a degree not previously possible.
 
-Motivated by that progress, we demonstrate here a flexible and unifying framework for combining encoding
+Motivated by that progress, we demonstrate here a flexible framework for combining encoding
 and decoding, which we term ``neural co-processors'' \cite{rao.coproc}. Neural co-processors leverage
@@
 subject. For example, they may be able to aid a stroke victim by finding a stimulation pattern of
-the motor cortex which helps restore lost limb function.
+the motor cortex which helps restore lost limb function. Likewise, the framework generalizes enough
+to condition stimulation on both brain activity, and external sensors, e.g. cameras or LIDAR, in
+order to incorporate feedback for realtime control.
 
@@
 	      stimulation, and task performance.
-	\item An AI agent which determines the stimulation to apply, in real time.
-\end{itemize}
-Proposed approach here combines a learned stimulation model with a stimulation agent
-
-* Encoding models \cite{shanechi.stimmodel} (others?)
-** Decades of advancement in encoding models, used for:
-*** Better understanding the brain
-*** Learning to understand the effects fo stimulation regimes
-** Optogenetic vs TMS vs eCog
-** In our case, designed to aid in the training of the stimulation agent.
-
-* Open loop control
-** Example: \cite{khanna.openloop}
-
-* Bi-directional BCIs (BBCIs)
-** Closed loop control
-\begin{itemize}
-	\item \cite{walker.inception} Inception Loops: driving brain states
-	\item \cite{kahana.biomarker} “Closed loop” here refers to when to apply stimulation, always of the same type and at the same site, based on memory performance.
-	\item \cite{tafazoli.acls} ACLS
-	\item cite Bolus
+	\item An AI agent which determines the stimulation to apply in a closed-loop fashion, in real time.
 \end{itemize}
 
-* Need for more complex stimulation policies, and need for optimizing task performance.
-** i.e. what if we don't have a specific pre-recorded activation we want to drive towards?
-** And how do we find activations that drive an intended task solution?
+Significant advances have been made in modeling the effects of electrical stimulation
+of the brain, some of which can be leveraged for our co-processor design, as we outline
+below. Researchers have explored how information can be biomimetically or
+artificially encoded and delivered via stimulation to neuronal networks in the brain and
+other regions of the nervous system for auditory \cite{niparko.cochlear}, visual \cite{weiland.retinal},
+proprioceptive \cite{tomlinson.propr}, and tactile
+\cite{tabot.tact, tyler.tact, dadarlat.tact, sharlene.tact, cronin.tact} perception.
+Advancements have also been made in modeling the effects of stimulation over large scale, multi-region
+networks, and across time \cite{shanechi.stimmodel}. Some have additionally designed models which
+can adapt to ongoing changes in the brain, including changes due to the stimulation itself
+\cite{tafazoli.acls}. In our proof-of-concept outlined below, we will use a stimulation
+model, not unlike those cited here, which seeks to account for both network dynamics
+and non-stationarity. In addition to training the model to have a strong ability to predict
+the effect of stimulation, we additionally train it to be useful for then learning an
+optimal stimulation policy, which is a property somewhat distinct from predictive
+power alone.
 
-Begin with a proof-of-concept in simulation.
+Advances have also been made in both open- and closed-loop stimulation for
+treating a variety of disorders. Open loop stimulation has been effective in
+treating Parkinson's Disease \cite{benabid.parkinsons}, as well as various
+psychiatric disorders \cite{holtzheimer.psy, kisely.psy, fraint.psy}.
+More directly related to this paper, we see in Khanna et al. \cite{khanna.openloop},
+the use of open loop stimulation in restoring dexterity after a lesion
+occurs affecting a primate's motor cortex. The authors demonstrate that
+the use of low-frequency alternating current, applied epidurally and set to
+certain phases, can improve grasp performance.
+
+While open loop stimulation techniques have yielded clinically useful results,
+their results in many domains have been mixed, such as use in visual
+prostheses \cite{bosking.visual}, and use in invoking somatosensory feedback
+\cite{cronin.tact}. Likely this is due to the stimulation not being conditioned
+on the ongoing dynamics of the circuit being stimulated. Moment-to-moment and
+throughout the day, the circuit will respond differently to the same stimulus,
+as a result of differing inputs and ongoing activity. Stimulation
+therefore needs to be proactively adapted in response. This need is even
+greater over longer time scales as the effects of plasticity and ageing change
+the connectivity of the brain.
+
+Closed-loop stimulation conditions stimulation on observations of brain activity,
+possibly allowing it to shape the neural activity more precisely, and to adapt to changes
+in the circuit over time. This opens the door to real-time, targeted control of the neural
+circuit. It has been used to aid in learning new memories after some impairment
+\cite{berger.closedloop, kahana.biomarker}, to replay visually-invoked
+activations \cite{tafazoli.acls}, and for optogenetic control of
+a thalamocortical circuit \cite{bolus.opto}, among others.
+
+Something that remains unclear is how to leverage closed-loop control for real-time
+co-adaption with the brain to accomplish an external task. ``Co-adaption'' here refers
+to the ability of a neuro-prosthesis to adapt its stimulation regime to the ongoing
+changes in the circuit it is stimulating, and to adapt with that circuit to accomplish
+the external task, such as grasping. The neural co-processor we present here provides one
+potential model for accomplishing that. Through the use of deep learning, the
+co-processor model we present co-adapts an AI agent, which governs the
+stimulation, with both a stimulation model, and the neural circuit being stimulated.
+
+For a neurologically complex task such as grasping, we cannot identify
+\textit{a priori} a real time controller of the neural circuit. That is
+due in large part to the variability of circuits from subject-to-subject,
+as well as variations in the placement of sensors and stimulators in the
+brain. The only plausible path to such a real time controller is to parameterize
+it in a subject- and time-specific way. Our model seeks to accomplish that using
+deep learning, together with a data-efficient approach to training.
+
+Before attempting \textit{in vivo} experiments using such a model, we first demonstrate
+here a number of crucial design elements of it, through the use of a simulated
+grasping circuit, presented previously by Michaels et al. \cite{michaels.mrnn}. We
+explore:
+\begin{itemize}
+	\item The properties of the artificial neural networks (ANNs) that are needed to
+	      successfully adapt to the long-running dynamics of a stimulated neural circuit,
+	      as well as to adapt to that circuit's ongoing changes.
+	\item Data-efficient methods for training these models, to better ensure we
+	      can train them with a biologically-realistic amount of data.
+	\item How we must train our stimulation model to make it effective in later
+	      training our stimulation agent.
+\end{itemize}
 
@@
 
+TODO: arch figure here (from Slide 5)
+
+\subsection{Architecture Overview}
+First, we present the architecture of our co-processor design. The co-processor is
+composed of two artifical neural networks:
+\begin{itemize}
+	\item A stimulation and neural dynamics model, known as an ``Emulator Network''.  It models the relationship
+\end{itemize}
+* EN architecture
+** Single layer LSTM with tanh activations, following by linear readout
+** LSTMs proved to be necessary for tracking long-running neural dynamics in a way that gave us a useful training signal
+*** Explore why? Revisit
+** Inputs: brain observations, stimulation vector, trial end indicator
+** Outputs: predicted muscle velocity
+* CPN architecture
+** Single layer LSTM with tanh activations, following by linear readout
+** Inputs: brain observations, trial end indicator
+** Outputs: stimulation vector
+
+\subsection{Simulation Overview}
 * Proof-of-concept by way of simulation.
@@
 ** \cite{kao.sim} Likeness of RNNs to natural circuits re: dynamics
-** Cite Michaels, and those upstream from him (Susillo?) re: RNNs
+** Cite Michaels, and those upstream from him (Susillo?) re: RNNs. Networks have internal dynamics that result from both inputs and internal connectivity. Perturbations at t=0 affect outputs at t=N.
 ** \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime. Use of a biomimetic
@@
 
+The stimulation model in our case adapts itself to the task of training our closed-
+loop AI stimulation agent. For a neurologically complex task such as grasping, we
+don't know \textit{a priori} what stimulation regimes to to explore while searching
+for an optimal solution. Our stimulation model plays the role of generating learning
+signals, via error backpropagation (<TODO? Citation?>), for our stimulation agent.
+It continuously updates itself based on observations of brain activity, stimulation,
+and external task queues, such as hand position.
+
+Example: Use of actual stimulation of an actual brain, together with offline deep learning,
+to search for an optimal visual input: \cite{walker.inception}
+
 * Overview of test bed
-** Michaels model \cite{michaels.mrnn}
+** Michaels model \cite{michaels.mrnn} \cite{susillo.mrnn}.
 ** Lesion designs
@@
 
-* EN architecture
-* CPN architecture
-
-* Training
+\subsection{Training the Model}
+Training
 ** Alternate training EN and CPN

### 2021-10-31 paper/article.tex
Paper progress

@@
 \documentclass[12pt]{iopart}
-\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
+\usepackage{graphicx}
+\graphicspath{ {./figs/} }
 %Uncomment next line if AMS fonts required
@@
 
-TODO: arch figure here (from Slide 5)
+\begin{figure}
+\includegraphics[width=\textwidth]{weill_arch.png}
+\caption{Using a co-processor to drive external task performance after a traumatic brain injury}
+\centering
+\label{fig:weill}
+\end{figure}
 
 \subsection{Architecture Overview}
-First, we present the architecture of our co-processor design. The co-processor is
-composed of two artifical neural networks:
+First, we present the architecture of our co-processor design. This design aims to solve two
+fundamental challenges in using neural stimulation to improve external task performance.
+First, neural networks exhibit long-running and highly nonlinear dependencies, necessitating
+the need for a stimulation agent to account for-far distant effects of the stimulation it
+applies. Second, with neurologically complex tasks, such as grasping, the mapping between
+neural activity as-measured and the external task cannot be determined \textit{a priori}.
+
+Our co-processor attempts to solve this with a pair of artificial neural networks:
 \begin{itemize}
-	\item A stimulation and neural dynamics model, known as an ``Emulator Network''.  It models the relationship
+	\item A stimulation and neural dynamics model, known as an ``Emulator Network'' (EN). It models the relationship
+	      between the stimulation, neural dynamics, and external task. It's purpose is for training the second network.
+	\item A stimulation agent, known as the ``co-processor network'' (CPN), which maps neural activity, and possibly
+	      data from external sensors, to stimulation parameters.
 \end{itemize}
-* EN architecture
-** Single layer LSTM with tanh activations, following by linear readout
-** LSTMs proved to be necessary for tracking long-running neural dynamics in a way that gave us a useful training signal
-*** Explore why? Revisit
-** Inputs: brain observations, stimulation vector, trial end indicator
-** Outputs: predicted muscle velocity
-* CPN architecture
-** Single layer LSTM with tanh activations, following by linear readout
-** Inputs: brain observations, trial end indicator
-** Outputs: stimulation vector
+
+We co-train the EN and CPN, with the goal of training a CPN whose output stimulation parameters
+improve task performance. By continually training both, they adapt to the brain as it changes,
+effectively causing brain-stimulator co-adaptation. The EN is a tool for training the CPN,
+giving us a way to back-propagate task error to the CPN. It outputs task-relevant
+metrics - a prediction of muscle velocities in our case - given measurements of neural
+activity, and the stimulation parameters output from the CPN. If the EN is trained in a
+particular way, and to a sufficient level of precision, it can be used to overcome the
+challenge of not knowing a mapping between stimulation and the task. When training the
+CPN, we in-effect treat the EN's output as the true task performance, or a related
+metric, and then backpropagate the loss defined in terms of that metric. See Figs X, Y.
+We will illustrate the details of the training algorithm below.
+
+In our present demonstration, the EN is constituted as a single layer, fully connected,
+long short-term memory (LSTM) recurrent neural network (RNN), with hyperbolic tangent ($tanh$)
+activations, and a linear readout. The general notion of a co-processor does
+not require this precise architecture, but for our example, we found that the LSTM
+approach allows the network to continuously adapt to long-running dependencies in
+the simulated neural dynamics, far better than a vanilla RNN. The CPN is constituted
+as an almost identical network, though with a different dimensionality, as we explain
+below. There is no strict requirement for the EN and CPN to be so similar, but we
+found this simple architecture to work well in our example.
+
+Note that the EN effectively constitutes a stimulation model. Its design is
+somewhat motivated by the common linear time-invariant state space model of
+stimulation, as in e.g. Yang et al. \cite{shanechi.stimmodel}, which is a helpfully
+simple (linear) approach. That approach seeks to model neural network dynamics in
+terms of a stationary linear model, which, due in part to its simplicity and the
+number of techniques developed for it, lends itself to useful interpretations and
+analyses of those neural dynamics.
+
+There are some key differences between that approach and the EN architecture we
+present here. Our EN's purpose is simply to train the CPN, so we pick an
+architecture we find to be best suited to that job. First, our EN architecture
+is not linear, through the use of a $tanh$ activation function. That allows us
+to escape some of the limits in representational power of linear models.
+Second, LSTM cells are designed to better capture long-running network
+dynamics, and in our case we found that they were crucial for that reason.
+Compare that to the simpler cells commonly used in vanilla RNNs, which
+yield functionally the same approach as the linear state space model, less
+the use of a nonlinear activation function.
+
+// TODO? Try having no nonlinearity?
 
@@
 ** Observation model
+
+* Overview of the four experiments:
+** Vanilla
+** Co-adaptive with connectivity loss
+** Co-adaptive with partial M1 loss
+** Co-adaptive with data reuse
 

### 2021-11-02 cpn.py
Clean up reuse and renable it. Models were generated while it was disabled.

@@ -60,7 +60,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         else:
             self.log_dir = None
 
-        self.round_thresh = 1
+        self.round_thresh = 3
         # A list of up to round_thresh length
         #  Each item is a tuple (actuals, targets, trial_end, stim list, obs list)
         #   The list is trial_len length.
@@ -146,7 +146,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         done = False
         checkpoint_eidx = 0
         en_is_ready = False
-        while not en_is_ready and checkpoint_eidx < 1000:
+        while not en_is_ready and checkpoint_eidx < 3000:
             for bidx in range(self.round_thresh):
                 actuals, targets, trial_end, stims, brain_data = self.saved_data[bidx]
 

### 2021-11-02 cpn_epoch_en.py

@@ -48,6 +48,7 @@ class CPNEpochEN:
         self.opt_cpn.zero_grad()
 
     def reset_en(self):
+        self.recent_pred_val_loss = 0.05
         self.en, self.opt_en = self.new_en(self.en)
 
         return self.en, self.opt_en
@@ -128,12 +129,6 @@ class CPNEpochEN:
             else:
                 p["lr"] = 4e-3
 
-        # Every 10 epochs let's validate/test
-        if not is_validation and (loss_history.eidx % 10) == 0:
-            next_is_validation = True
-        else:
-            next_is_validation = False
-
         last_rec = loss_history.get_recent_record(-2)
         train_loss_out = float("nan")
         train_val_loss_out = float("nan")
@@ -154,12 +149,18 @@ class CPNEpochEN:
         elif (
             vl < max(0.02 * self.recent_train_loss, 0.0003)
             and self.checkpoint_eidx > 100
-        ) or self.checkpoint_eidx >= 2000:
+        ) or self.checkpoint_eidx >= 3000:
             en_is_ready = True
             self.reset_period()
         else:
             self.checkpoint_eidx += 1
             en_is_ready = False
+
+        # Every 10 epochs let's validate/test
+        if not en_is_ready and not is_validation and (self.checkpoint_eidx % 10) == 0:
+            next_is_validation = True
+        else:
+            next_is_validation = False
 
         user_data = CPNENStats(
             "en_offline" if reused_data else "en",

### 2021-11-02 cpn_utils.py

@@ -11,9 +11,10 @@ class EpochType(enum.Enum):
     EN = 0
     EN_OFFLINE = 1
     CPN = 2
-    CPN_AND_EN = 3
-    VAL_EN = 4
-    VAL_CPN = 5
+    CPN_OFFLINE = 3
+    CPN_AND_EN = 4
+    VAL_EN = 5
+    VAL_CPN = 6
 
 
 @attr.s(auto_attribs=True)

### 2021-11-06 paper/article.tex
Paper progress

@@
 First, neural networks exhibit long-running and highly nonlinear dependencies, necessitating
-the need for a stimulation agent to account for-far distant effects of the stimulation it
+the need for a stimulation agent to account for far-distant effects of the stimulation it
 applies. Second, with neurologically complex tasks, such as grasping, the mapping between
 neural activity as-measured and the external task cannot be determined \textit{a priori}.
+As a result, the co-processor must somehow learn what stimulation is appropriate for
+aiding the user in the external task they are attempting to perform.
 
-Our co-processor attempts to solve this with a pair of artificial neural networks:
+Our co-processor attempts to solve these with a pair of artificial neural networks:
 \begin{itemize}
@@
 improve task performance. By continually training both, they adapt to the brain as it changes,
-effectively causing brain-stimulator co-adaptation. The EN is a tool for training the CPN,
+effectively allowing for brain-stimulator co-adaptation. The EN is a tool for training the CPN,
 giving us a way to back-propagate task error to the CPN. It outputs task-relevant
@@
 activity, and the stimulation parameters output from the CPN. If the EN is trained in a
-particular way, and to a sufficient level of precision, it can be used to overcome the
-challenge of not knowing a mapping between stimulation and the task. When training the
-CPN, we in-effect treat the EN's output as the true task performance, or a related
-metric, and then backpropagate the loss defined in terms of that metric. See Figs X, Y.
-We will illustrate the details of the training algorithm below.
+particular way, and to a sufficient level of precision, it can be used as a function
+approximator relating stimulation and a task, and do so in a way that allows us to train
+the CPN with it. When training the CPN, we in-effect treat the EN's output as the true
+task performance, or a related metric, and then backpropagate the loss defined in terms
+of that metric in order to train the CPN. See Fig. \ref{fig:weill}. We will illustrate the details
+of the training algorithm below.
 
@@
 is not linear, through the use of a $tanh$ activation function. That allows us
-to escape some of the limits in representational power of linear models.
+to escape some of the limits in representational power of linear models, though
+we didn't deeply test a strictly linear approach.
 Second, LSTM cells are designed to better capture long-running network
@@
 
-// TODO? Try having no nonlinearity?
+\subsection{Simulation Overview}
+We demonstrate our approach here using a simulated grasping circuit. A detailed
+simulation such as this allows us to explore some of the critical details in
+how to train a co-processor of this design. By first doing such exploration in
+simulation, we are able to rapidly and cheaply iterate on our design, prior to
+any \textit{in vivo} experiments. In order for the simulation to be
+admissible, we need to ensure it has some properties that allow it to strongly
+indicate if our design is improving in a direction that will later allow for
+real deployments. Otherwise, our design may be adapting to the pecularities of
+the simulation, witheout becoming more useful.
 
-\subsection{Simulation Overview}
-* Proof-of-concept by way of simulation.
-** Simulation allows for rapid, cheap exploration of learning algorithms.
 ** Potentially helpful to look for a simulated neural circuit which has internal dynamics, i.e. information propagation, which is demonstrably similar to natural circuits.

### 2021-11-07 paper/article.tex
Paper progress

@@
 fundamental challenges in using neural stimulation to improve external task performance.
-First, neural networks exhibit long-running and highly nonlinear dependencies, necessitating
+First, neural networks exhibit long-running and nonlinear dependencies, necessitating
 the need for a stimulation agent to account for far-distant effects of the stimulation it
@@
 Second, LSTM cells are designed to better capture long-running network
-dynamics, and in our case we found that they were crucial for that reason.
+dynamics, and in our case we found that they were crucial for that reason:
+when using vanilla RNN cells, our co-processor wasn't able to train well.
 Compare that to the simpler cells commonly used in vanilla RNNs, which
@@
 \subsection{Simulation Overview}
-We demonstrate our approach here using a simulated grasping circuit. A detailed
-simulation such as this allows us to explore some of the critical details in
-how to train a co-processor of this design. By first doing such exploration in
+We demonstrate our approach here using a simulated grasping circuit.
+A detailed simulation such as this allows us to explore some of the critical details in
+training a co-processor of this design. By first doing such exploration in
 simulation, we are able to rapidly and cheaply iterate on our design, prior to
@@
 real deployments. Otherwise, our design may be adapting to the pecularities of
-the simulation, witheout becoming more useful.
+the simulation, without becoming more useful for the real world.
+
+The simulated circuit, from Michaels et al. \cite{michaels.mrnn, susillo.mrnn}, was trained
+to resemble the grasping circuits of monkey subjects engaged in a reach-to-grasp task. It
+consists of a ``modular'' vanilla RNN, and a linear readout layer. The inputs are visual
+features intended to capture the view the monkeys had during the task, i.e. 3D renderings of
+the same objects which the monkeys grasped. The outputs are muscle length velocities for the
+should, arm, and hand of the monkey during the trial. The natural velocities were captured
+with a motion capturing glove, and the artificial neural network is trained to recapitulate
+those grasping motions. See Fig. \ref{fig:michaels}.
+
+The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
+representing the visual processing needed to reach the hand to the appropriate position for
+the grasp, and to form the hand properly for grasping the particular shape of object. The
+emergent dynamics of the artificial network's ``modules'', once train, correspond roughly
+to the AIP, F5, and M1 portions of the monkey subjects' brains. Notably, the activity of
+the module receiving the visual inputs resembles the AIP activity of the monkey subject
+from the same trials. Likewise, the second and third modules resemble the natural activity
+from the monkey's F5 and M1 regions, respectively. The network dynamics show a number of
+other correspondences to the monkeys' natural brain activity as well, as detailed in the paper.
+
+Notably, the simulated circuit's activity shows a relatively clear separation of the
+object shapes.  That is - the visual information input to the network differentiates one trial
+from another, and the circuit is able to leverage that information to properly condition
+the hand shape trajectory for grasping the object of that trial's particular shape, size, and location.
+The visual information forward-propagates through the network, through successive processing
+steps, where it is eventually leveraged to recapitulate the grasp appropriate for the object
+represented by the input visual features. As we will note below: in the absense of this
+visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
+stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
+grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
+waiting period, but performance suffers drastically.
+
+As a result, simulating a brain lesion in terms of this artificial network results in error modes
+which resemble a natural lesion of certain regions of a primate brain.
+
+...
+* Lesioning properties
+
+* Long-running dynamics
 
@@
 
-The stimulation model in our case adapts itself to the task of training our closed-
-loop AI stimulation agent. For a neurologically complex task such as grasping, we
-don't know \textit{a priori} what stimulation regimes to to explore while searching
-for an optimal solution. Our stimulation model plays the role of generating learning
-signals, via error backpropagation (<TODO? Citation?>), for our stimulation agent.
-It continuously updates itself based on observations of brain activity, stimulation,
-and external task queues, such as hand position.
-
 Example: Use of actual stimulation of an actual brain, together with offline deep learning,
@@
 * Overview of test bed
-** Michaels model \cite{michaels.mrnn} \cite{susillo.mrnn}.
+** Michaels model .
 ** Lesion designs

### 2021-11-07 experiment/mRNN.py

@@ -474,58 +474,3 @@ def load_from_file(data_path, pretrained=False, **kwargs):
     return mrnn
 
 
-# TODO: refactor
-#def recover(mike, data_loader, cuda=None):
-#    """
-#    This function can be used to refine a Michaels model, simulating
-#    some amount of recovery.
-#    """
-#    loss_func = torch.nn.MSELoss()
-#
-#    dset = self.cfg.dataset
-#    dset_size = len(dset)
-#    dset_samp_len = dset.sample_len
-#    loader = DataLoader(dset, batch_size=dset_size, shuffle=True)
-#
-#    for p in self.mike.parameters():
-#        p.requires_grad = True
-#
-#    opt_mike = AdamW(
-#        [p for p in self.mike.parameters() if p is not self.mike.I], lr=1e-4
-#    )
-#
-#    loss = 1
-#    global mm
-#    mm = self.mike
-#    while loss > 0.0007:
-#        for batch in loader:
-#            if loss >= 0.00385:
-#                for p in opt_mike.param_groups:
-#                    p["lr"] = 1e-4
-#            else:
-#                for p in opt_mike.param_groups:
-#                    p["lr"] = 1e-5
-#
-#            self.mike.reset()
-#            self.opt_mike.zero_grad()
-#
-#            din, _, _, dout, _ = batch
-#
-#            preds = torch.zeros((dset_size, dset_samp_len, self.cfg.out_dim))
-#            if cuda is not None:
-#                preds = preds.cuda(cuda)
-#
-#            for tidx in range(dout.shape[1]):
-#                cur_din = din[:, tidx, :].T
-#                p = self.mike(cur_din)
-#                preds[:, tidx, :] = p[:, :]
-#
-#            loss = loss_func(preds, dout)
-#            loss.backward()
-#            self.mike.set_coadap_grads()
-#            opt_mike.step()
-#
-#            g_logger.info("Brain recovery loss: %0.7f", loss.item())
-#
-#    for p in self.mike.parameters():
-#        p.requires_grad = False

### 2021-11-07 experiment/stim.py

@@ -230,6 +230,10 @@ class StimulusGaussianExp(Stimulus):
         if self._retain_grad:
             self.W.retain_grad()
 
+    def get_neuron_weights(self):
+        self._calc_neuron_weights()
+        return self.W
+
     def reset(self, batch_size=None):
         super(StimulusGaussianExp, self).reset(batch_size=batch_size)
         self._vals = torch.zeros(

### 2021-11-08 paper/article.tex
Paper progress

@@
 \documentclass[12pt]{iopart}
-\usepackage{graphicx}
+\usepackage{graphicx,epstopdf}
+\usepackage{caption}
+\usepackage{subcaption}
 \graphicspath{ {./figs/} }
@@
 to resemble the grasping circuits of monkey subjects engaged in a reach-to-grasp task. It
-consists of a ``modular'' vanilla RNN, and a linear readout layer. The inputs are visual
+consists of a ``modular'' vanilla RNN (mRNN), and a linear readout layer. The inputs are visual
 features intended to capture the view the monkeys had during the task, i.e. 3D renderings of
@@
 with a motion capturing glove, and the artificial neural network is trained to recapitulate
-those grasping motions. See Fig. \ref{fig:michaels}.
+those grasping motions. Data and trained models from this work were supplied to us by the lead
+author Jonathan Michaels, for the purpose of our present simulation. See Fig. \ref{fig:michaels}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{michaels.eps}
+\caption{Michaels Modular RNN (mRNN), a simulated grasping circuit. The emergent dynamics of the
+three modules correspond well to natural neural activity measured from primate AIP, F5, M1 regions,
+respectively, from the same task. Visual (VGGNet) features forward propagate through the
+network, conditioning the grasp for the object of the particular size, shape, and location.}
+\centering
+\label{fig:michaels}
+\end{figure}
 
@@
 the grasp, and to form the hand properly for grasping the particular shape of object. The
-emergent dynamics of the artificial network's ``modules'', once train, correspond roughly
-to the AIP, F5, and M1 portions of the monkey subjects' brains. Notably, the activity of
-the module receiving the visual inputs resembles the AIP activity of the monkey subject
-from the same trials. Likewise, the second and third modules resemble the natural activity
-from the monkey's F5 and M1 regions, respectively. The network dynamics show a number of
-other correspondences to the monkeys' natural brain activity as well, as detailed in the paper.
+emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
+to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
+refers to the fact that the activity of the module receiving the visual inputs resembles
+the AIP activity of the monkey subject from the same trials. Likewise, the second and
+third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
+The emergent network dynamics show a number of other correspondences to the monkeys' natural
+brain activity as well, as detailed in the paper.
 
-Notably, the simulated circuit's activity shows a relatively clear separation of the
+Importantly, the simulated circuit's activity shows a relatively clear separation of the
 object shapes. That is - the visual information input to the network differentiates one trial
 from another, and the circuit is able to leverage that information to properly condition
-the hand shape trajectory for grasping the object of that trial's particular shape, size, and location.
-The visual information forward-propagates through the network, through successive processing
-steps, where it is eventually leveraged to recapitulate the grasp appropriate for the object
-represented by the input visual features. As we will note below: in the absense of this
+the hand shape trajectory for grasping the object of that trial's particular shape, size, and
+location. The visual information forward-propagates through the network, through successive
+processing steps, where it is eventually leveraged to recapitulate the grasp appropriate
+for the object represented by the input visual features. It follows, then, that the classes
+are separable in some way by observing the distinct ways that the mRNN is activated by the
+classes' corresponding visual inputs. As we will note below: in the absense of this
 visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
@@
 
-As a result, simulating a brain lesion in terms of this artificial network results in error modes
-which resemble a natural lesion of certain regions of a primate brain.
+As a result, simulating a brain lesion in terms of this artificial network results in error
+modes which resemble a natural lesion of certain regions of a primate brain. For example, if we
+alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
+the reaching motion generally succeeds, but the finger muscle velocities show a high degree
+of error - effectively implying that the subject can reach to grasp, but cannot form a
+grasp appropriate for the object. That suggests that losing a portion of the cortical
+machinery needed to translate or forward-propagate visual object shape information
+to movement-related cortex can result in a reduced ability to form the hand properly for
+the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
+the hand is also a common symptom of certain strokes in primates, and indeed many who
+suffer from it are still able to position their hand, even while being unable to form
+it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
+simulated lesion closely resembles natural stroke symptoms. Likewise, if we simulate a
+``stroke'' of a portion of the output module of the Michaels mRNN, roughly
+corresponding to M1, we see a more wholesale loss of movement, affecting even the
+ability to reach for the grasp. See Fig. \ref{fig:lesion} for examples.
 
-...
-* Lesioning properties
+\begin{figure}
+	\centering
+	\begin{subfigure}[b]{0.46\textwidth}
+	    \centering
+	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
+	    \caption{inner1}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.46\textwidth}
+	    \centering
+	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
+	    \caption{inner1}
+	\end{subfigure}
+	\hfill
+	\caption{Two simple graphs}
+	\label{fig:lesion}
+\end{figure}
 
@@
 
-* Show example where AIP/F5 lesioned and how that affects performance
-
 * Any point in showing sensitivity to observation or stim dim?

### 2021-11-08 cpn.py

@@ -27,14 +27,15 @@ class CPN_EN_CoProc(experiment.CoProc):
             in_dim,
             stim_dim,
             num_neurons=in_dim,
-            activation_func=torch.nn.Tanh,
+            activation_func=cfg.cpn_activation,
             cuda=cuda,
         )
         self.opt_cpn = AdamW(self.cpn.parameters(), lr=1e-3)
 
         en_in_dim = cfg.observer_instance.out_dim + stim_dim + 1
         self.en, self.opt_en = stim_model.get_stim_model(
-            en_in_dim, out_dim, cuda=cfg.cuda
+            en_in_dim, out_dim, activation=cfg.en_activation,
+            cuda=cfg.cuda
         )
 
         self.stims = None

### 2021-11-08 experiment/activation.py

@@ -9,3 +9,4 @@ class ActivationType(enum.Enum):
     ReLU = torch.nn.ReLU
     ReTanh = utils.ReTanh
     Tanh = torch.nn.Tanh
+    Linear = utils.NonAct

### 2021-11-08 experiment/lesion.py

@@ -31,7 +31,7 @@ class Lesion(object):
 
 
 class LesionOutputs(Lesion):
-    def __init__(self, num_neurons_per_module, module_id, pct):
+    def __init__(self, num_neurons_per_module, module_id, pct, cuda=None):
         self.lesion_mask = torch.ones((num_neurons_per_module * 3,))
         start_idx, end_idx = module_id_to_idxs(num_neurons_per_module, module_id)
         kill_idxs = random.sample(
@@ -40,6 +40,9 @@ class LesionOutputs(Lesion):
         self.lesion_mask[kill_idxs] = 0.0
         self.module_id = module_id
         self.pct = pct
+
+        if cuda is not None:
+            self.lesion_mask = self.lesion_mask.cuda(cuda)
 
     def lesion(self, network, pre_response):
         batch_size = pre_response.shape[0]

### 2021-11-08 paper/article.tex
Paper progress

@@
 
-The simulated circuit, from Michaels et al. \cite{michaels.mrnn, susillo.mrnn}, was trained
-to resemble the grasping circuits of monkey subjects engaged in a reach-to-grasp task. It
-consists of a ``modular'' vanilla RNN (mRNN), and a linear readout layer. The inputs are visual
-features intended to capture the view the monkeys had during the task, i.e. 3D renderings of
-the same objects which the monkeys grasped. The outputs are muscle length velocities for the
-should, arm, and hand of the monkey during the trial. The natural velocities were captured
-with a motion capturing glove, and the artificial neural network is trained to recapitulate
-those grasping motions. Data and trained models from this work were supplied to us by the lead
+An example of such an approach is Bernal et al. \cite{bernal.sim}. In this work, the
+authors use a simulated spiking neural network to train a stimulation agent, much
+like the work we present here. Their stimulation agent sought to restore the network's
+control of a simulated arm, to reach a target. As they point out, we have a limited
+ability to probe a neural circuit \textit{in vivo} in order to perform learning. As
+a result, we first need to design our approach through the use of an admissible
+simulation. The authors in this case simulated lesions by effectively removing parts
+of their simulated network, or by cutting connections between parts of the network.
+Likewise here, we use an established design for an artificial neural circuit which performs
+a grasping task, whose architecture and training methods were designed specifically
+to result in naturalistic dynamics.
+
+The simulated circuit, from Michaels et al. \cite{michaels.mrnn}, was trained
+to resemble the grasping circuits of monkey subjects engaged in a delayed reach-to-grasp task.
+Its design draws on a body of literature focused on architectures and training methods for RNNs which
+seek to create artificial neural circuits that have activation dynamics similar to natural circuits,
+including delayed grasping tasks \cite{susillo.mrnn}. The Michaels circuit consists of a
+``modular'' vanilla RNN (mRNN), and a linear readout layer. The inputs are visual features
+intended to capture the view the monkeys had during the task, i.e. 3D renderings of the same
+objects which the monkeys grasped. The outputs are muscle length velocities for the should,
+arm, and hand of the monkey during the trial. The natural velocities were captured with a
+motion capturing glove, and the artificial neural network was trained to recapitulate those
+grasping motions. Data and trained models from this work were supplied to us by the lead
 author Jonathan Michaels, for the purpose of our present simulation. See Fig. \ref{fig:michaels}.
@@
 The emergent network dynamics show a number of other correspondences to the monkeys' natural
-brain activity as well, as detailed in the paper.
+brain activity as well, as detailed in that paper.
 
 Importantly, the simulated circuit's activity shows a relatively clear separation of the
-object shapes. That is - the visual information input to the network differentiates one trial
-from another, and the circuit is able to leverage that information to properly condition
+object shapes. That is - the visual information input to the network that differentiates
+one trial from another is leveraged by the circuit to properly condition
 the hand shape trajectory for grasping the object of that trial's particular shape, size, and
@@
 
-As a result, simulating a brain lesion in terms of this artificial network results in error
-modes which resemble a natural lesion of certain regions of a primate brain. For example, if we
+\subsubsection{Simulated lesions cause real world failure modes}
+Simulating a brain lesion in terms of this artificial network results in error
+modes that resemble a natural lesion of certain regions of a primate brain. For example, if we
 alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
 the reaching motion generally succeeds, but the finger muscle velocities show a high degree
-of error - effectively implying that the subject can reach to grasp, but cannot form a
+of error - effectively implying that the subject can somewhat reach to grasp, but cannot form a
 grasp appropriate for the object. That suggests that losing a portion of the cortical
-machinery needed to translate or forward-propagate visual object shape information
+machinery needed to translate or forward-propagate object shape information
 to movement-related cortex can result in a reduced ability to form the hand properly for
@@
 it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
-simulated lesion closely resembles natural stroke symptoms. Likewise, if we simulate a
+simulated lesion closely resembles natural stroke symptoms. Conversely, if we simulate a
 ``stroke'' of a portion of the output module of the Michaels mRNN, roughly
@@
 	\centering
-	\begin{subfigure}[b]{0.46\textwidth}
+	\begin{subfigure}[c]{0.65\textwidth}
 	    \centering
 	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
-	    \caption{inner1}
+	    \caption{Example muscle trajectories during a single trial, for one hand and shoulder muscle}
 	\end{subfigure}
 	\hfill
-	\begin{subfigure}[b]{0.46\textwidth}
+	\begin{subfigure}[c]{0.32\textwidth}
 	    \centering
 	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
-	    \caption{inner1}
+	    \caption{Proportion of loss attributable to hand muscles}
 	\end{subfigure}
 	\hfill
-	\caption{Two simple graphs}
+	\caption{Lesion designs which prevent forward propagation of object shape information
+	         differentially impact hand pose. Loss here is a mean squared error (MSE) loss,
+		 measured relative to the circuit's trajectories prior to the lesion.}
 	\label{fig:lesion}
@@
 
-* Long-running dynamics
+The co-processor's task in our simulation will be to identify the appropriate information,
+read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
+to effectively bridge across the lesion, forward propatating the object shape information,
+indirectly, through the use of stimulation.
 
-** Potentially helpful to look for a simulated neural circuit which has internal dynamics, i.e. information propagation, which is demonstrably similar to natural circuits.
-** \cite{kao.sim} Likeness of RNNs to natural circuits re: dynamics
-** Cite Michaels, and those upstream from him (Susillo?) re: RNNs. Networks have internal dynamics that result from both inputs and internal connectivity. Perturbations at t=0 affect outputs at t=N.
-** \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime. Use of a biomimetic
-  simulation to develop and evaluate a neural controller. Use of lesion model.
+In our experiments below, we demonstrate our co-processor design on two types of
+simulated lesions:
+\begin{itemize}
+	\item \textbf{Connection-based:} we prevent the forward and backward propagation of
+	      information between the ``F5'' and ``M1'' modules, effectively representing a
+	      severing of the connections between the two.
+	\item \textbf{Output-based:} we force the output of some proportion of ``M1'' neurons
+	      to zero, effectively removing them from the network.
+\end{itemize}
 
-Example: Use of actual stimulation of an actual brain, together with offline deep learning,
-to search for an optimal visual input: \cite{walker.inception}
+TODO: show example trajectories for both
+
+\subsubsection{Simulated network exhibits long running dynamics}
+Just as the co-processor must adapt to
+
+\subsubsection{Sensor model}
+asdf
+
+\subsubsection{Stimulation model}
+asdf
 

### 2021-11-09 cpn.py
Tweak reuse params

@@ -61,7 +61,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         else:
             self.log_dir = None
 
-        self.round_thresh = 3
+        self.round_thresh = 5
         # A list of up to round_thresh length
         #  Each item is a tuple (actuals, targets, trial_end, stim list, obs list)
         #   The list is trial_len length.
@@ -147,7 +147,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         done = False
         checkpoint_eidx = 0
         en_is_ready = False
-        while not en_is_ready and checkpoint_eidx < 3000:
+        while not en_is_ready and checkpoint_eidx < 1000:
             for bidx in range(self.round_thresh):
                 actuals, targets, trial_end, stims, brain_data = self.saved_data[bidx]
 

### 2021-11-09 paper/article.tex
Paper progress

@@
 corresponding to M1, we see a more wholesale loss of movement, affecting even the
-ability to reach for the grasp. See Fig. \ref{fig:lesion} for examples.
+ability to reach for the grasp. Finally, if we ``disconnect'' communications between the
+F5 and M1 modules, we see a failure similar to an AIP lesion: movement is generally
+achieved, but we see a disproportionate impact on hand pose.
+See Fig. \ref{fig:lesion} for examples.
 
@@
 	\centering
-	\begin{subfigure}[c]{0.65\textwidth}
+	\begin{subfigure}[c]{0.62\textwidth}
 	    \centering
@@
 	\caption{Lesion designs which prevent forward propagation of object shape information
-	         differentially impact hand pose. Loss here is a mean squared error (MSE) loss,
+	         differentially impact hand pose. Loss here is an L2 distance
 		 measured relative to the circuit's trajectories prior to the lesion.}
@@
 read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
-to effectively bridge across the lesion, forward propatating the object shape information,
+to effectively bridge across the lesion, forward propagating the object shape information,
 indirectly, through the use of stimulation.
 
-In our experiments below, we demonstrate our co-processor design on two types of
+In our experiments below, we demonstrate our co-processor design on three types of
 simulated lesions:
@@
 	      severing of the connections between the two.
-	\item \textbf{Output-based:} we force the output of some proportion of ``M1'' neurons
-	      to zero, effectively removing them from the network.
+	\item \textbf{Output-based (AIP):} we force the output of some proportion of ``AIP''
+	      neurons to zero, effectively removing them from the network. This results in a
+	      loss of object shape information and, as we will see below, causes the co-processor
+	      to be unable to differentiate objects.
+	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
+	      neurons to zero, causing a more holisitic loss of movement.
 \end{itemize}
 
-TODO: show example trajectories for both
+\subsubsection{Simulated network exhibits long running dynamics}
+Just as the co-processor must adapt to the lesion, it must also adapt to the network
+dynamics.
 
-\subsubsection{Simulated network exhibits long running dynamics}
-Just as the co-processor must adapt to
+\begin{figure}
+	\centering
+	\begin{subfigure}[c]{0.48\textwidth}
+	    \centering
+	    \includegraphics[width=\textwidth]{perturbe_single.pdf}
+	    \caption{}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[c]{0.48\textwidth}
+	    \centering
+	    \includegraphics[width=\textwidth]{perturbe_dist.pdf}
+	    \caption{}
+	\end{subfigure}
+	\hfill
+	\caption{Instantaneous perturbations of a random sample (n=10) of M1 neurons at time t=200 results in long-running effects on output.}
+	\label{fig:dynamics}
+\end{figure}
 
-\subsubsection{Sensor model}
-asdf
+\subsubsection{Observation model}
+** Observation model
+*** Spatio-temporal smoothing
+
+\begin{figure}
+\includegraphics[width=\textwidth]{stim_and_obs.png}
+\caption{Example of stimulation and observation for a single trial}
+\centering
+\label{fig:stim_and_obs}
+\end{figure}
 
 \subsubsection{Stimulation model}
-asdf
+** Stimulation design
 
-* Overview of test bed
-** Michaels model .
-** Lesion designs
-*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
-*** Disconnect modules vs lesion M1, and the effects
-** Stimulation design
-*** Spatio-temporal smoothing
-** Observation model
+\subsection{Experiments}
+Summary of tunables:
+\begin{itemize}
+	\item Learning rate schedules
+	\item Activation functions
+	\item Obs function dimensionality
+	\item Stim dimensionality
+	\item Dimensionality of EN, CPN
+\end{itemize}
 
-* Overview of the four experiments:
-** Vanilla
-** Co-adaptive with connectivity loss
-** Co-adaptive with partial M1 loss
-** Co-adaptive with data reuse
+Experiment types:
+\begin{itemize}
+	\item Connectivity lesion
+	\item Co-adaptive with connectivity lesion
+	\item Co-adaptive with partial M1 loss
+	\item Co-adaptive with partial AIP loss
+	\item Co-adaptive with data reuse
+\end{itemize}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
+\caption{Experiment architecture overview}
+\centering
+\label{fig:exp_overview}
+\end{figure}
 
 \subsection{Training the Model}
-Training
-** Alternate training EN and CPN
-*** Train an EN
-**** Training the EN until predictive power reaches a threshold, based on current task performance
-**** Training regime based on most recent CPN, similar CPNS (noise added to params), and white noise stimulation
-**** Single batch of data from applying current CPN set and white noise to the brain, EN fit to ``offline'' it over maining epochs
-**** Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
-*** Train the CPN
-**** Backprop through the EN, effectively using the EN's prediction as ground truth
+Alternate training EN and CPN.
+\begin{itemize}
+	\item Train an EN
+		\begin{itemize}
+			\item Training the EN until predictive power reaches a threshold, based on current task performance
+			\item Training regime based on most recent CPN, similar CPNS (noise added to params), and white noise stimulation
+			\item Single batch of data from applying current CPN set and white noise to the brain, EN fit to ``offline'' it over maining epochs
+			\item Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
+		\end{itemize}
+	\item Train the CPN
+		\begin{itemize}
+			\item Backprop through the EN, effectively using the EN's prediction as ground truth
+		\end{itemize}
+\end{itemize}
 
@@
 * Split by lesion design
-
-* Any point in showing sensitivity to observation or stim dim?
-
-* Try regularization variation?
 

### 2021-11-09 experiment/mRNN.py

@@ -52,6 +52,7 @@ class MichaelsRNN(nn.Module):
         self.activation_func = activation_func()
         self.output_dim = output_dim
         self._cuda = cuda
+        self.last_stimulus = None
 
         npm = self.num_neurons_per_module
         numn = self.num_neurons
@@ -371,7 +372,8 @@ class MichaelsRNN(nn.Module):
 
     def get_next_stimulus(self):
         # (batch_size, num_neurons)
-        return self.stimulus.get_next()
+        self.last_stimulus = self.stimulus.get_next()
+        return self.last_stimulus
 
 
 class MichaelsDataset(Dataset):

### 2021-11-09 paper/article.tex
Paper progress

@@
 \textit{Approach} We provide the first proof-of-concept of a neural co-processor that
-leverages deep learning, through the use of a simulated neural circuit
+leverages deep learning, through the use of a simulated neural circuit.
 That circuit performs reach-to-grasp task, based on visual input, and is designed to
 closely resemble a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate
-a variety of lesions by altering the model, and then demonstrate a co-processor's ability
+a variety of lesions by altering the model, and then demonstrate our co-processor's ability
 to restore lost function through ``stimulation'' of that model. We further test the
-ability of a co-processor to adapt its stimulation as the simulated brain undergoes changes.
+ability of our co-processor to adapt its stimulation as the simulated brain undergoes changes.
 \textit{Main results} Our simulated co-processor successfully co-adapts with the neural
@@
 We demonstrate our approach here using a simulated grasping circuit.
-A detailed simulation such as this allows us to explore some of the critical details in
-training a co-processor of this design. By first doing such exploration in
-simulation, we are able to rapidly and cheaply iterate on our design, prior to
-any \textit{in vivo} experiments. In order for the simulation to be
-admissible, we need to ensure it has some properties that allow it to strongly
-indicate if our design is improving in a direction that will later allow for
-real deployments. Otherwise, our design may be adapting to the pecularities of
-the simulation, without becoming more useful for the real world.
+A detailed simulation such as this allows us to explore some of the critical
+architectural details and training algorithms for a co-processor of this type.
+By first doing such exploration in simulation, we are able to rapidly and
+cheaply iterate on our design, prior to any \textit{in vivo} experiments.
+In order for the simulation to be admissible, we need to ensure it has some
+properties that allow it to strongly indicate if our design is improving in
+a direction that will later allow for real deployments. Otherwise, our
+design may be adapting to the pecularities of the simulation, without
+becoming more useful for the real world.
 
-An example of such an approach is Bernal et al. \cite{bernal.sim}. In this work, the
+An example of such a simulation approach is Bernal et al. \cite{bernal.sim}. In this work, the
 authors use a simulated spiking neural network to train a stimulation agent, much
 like the work we present here. Their stimulation agent sought to restore the network's
-control of a simulated arm, to reach a target. As they point out, we have a limited
-ability to probe a neural circuit \textit{in vivo} in order to perform learning. As
-a result, we first need to design our approach through the use of an admissible
-simulation. The authors in this case simulated lesions by effectively removing parts
-of their simulated network, or by cutting connections between parts of the network.
-Likewise here, we use an established design for an artificial neural circuit which performs
-a grasping task, whose architecture and training methods were designed specifically
-to result in naturalistic dynamics.
+control of a simulated arm, to reach a target, after a simulated lesion was applied.
+As they point out, we have a limited ability to probe a neural circuit \textit{in vivo}
+in order to perform learning. As a result, we first need to design our approach through
+the use of an admissible simulation. The authors in this case simulated lesions by
+effectively removing parts of their simulated network, or by cutting connections between
+parts of the network. Likewise here, we use an established design for an artificial
+neural circuit which performs a grasping task, whose architecture and training methods
+were designed specifically to result in naturalistic dynamics.
 
@@
 including delayed grasping tasks \cite{susillo.mrnn}. The Michaels circuit consists of a
-``modular'' vanilla RNN (mRNN), and a linear readout layer. The inputs are visual features
-intended to capture the view the monkeys had during the task, i.e. 3D renderings of the same
-objects which the monkeys grasped. The outputs are muscle length velocities for the should,
-arm, and hand of the monkey during the trial. The natural velocities were captured with a
-motion capturing glove, and the artificial neural network was trained to recapitulate those
-grasping motions. Data and trained models from this work were supplied to us by the lead
-author Jonathan Michaels, for the purpose of our present simulation. See Fig. \ref{fig:michaels}.
+``modular'' vanilla RNN (mRNN), and a linear readout layer. Each ``module'' consists of 100 vanilla
+RNN neurons, with a non-linearity applied on the outputs. The modules are internally fully connected,
+and are connected to each other sparsely (10\% connectivity). The inputs are visual features
+intended to capture the view the monkeys had during the task, specifically VGGNet features from 3D
+renderings of the same objects which the monkeys grasped. The outputs are muscle length
+velocities for the should, arm, and hand of the monkey during the trial. The natural
+velocities were captured with a motion capturing glove, and the artificial neural network
+was trained to recapitulate those grasping motions. Data and trained models from this work
+were supplied to us by the lead author Jonathan Michaels, for the purpose of our present
+simulation. See Fig. \ref{fig:michaels}.
 
@@
 The emergent network dynamics show a number of other correspondences to the monkeys' natural
-brain activity as well, as detailed in that paper.
+brain activity as well, as detailed in the paper.
 
@@
 it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
-simulated lesion closely resembles natural stroke symptoms. Conversely, if we simulate a
-``stroke'' of a portion of the output module of the Michaels mRNN, roughly
+simulated lesion closely resembles natural stroke symptoms, though the simulation is not
+intended to constitute a physical model of a lesion, i.e. to directly explain the
+connection between hand spasticity and the lesion. Conversely, if we
+lesion of a portion of the output module of the Michaels mRNN, roughly
 corresponding to M1, we see a more wholesale loss of movement, affecting even the
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 \begin{itemize}
-	\item \textbf{Connection-based:} we prevent the forward and backward propagation of
-	      information between the ``F5'' and ``M1'' modules, effectively representing a
-	      severing of the connections between the two.
 	\item \textbf{Output-based (AIP):} we force the output of some proportion of ``AIP''
@@
 	      to be unable to differentiate objects.
+	\item \textbf{Connection-based:} we prevent the forward and backward propagation of
+	      information between the ``F5'' and ``M1'' modules, effectively representing a
+	      severing of the connections between the two.
 	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
@@
 \subsubsection{Simulated network exhibits long running dynamics}
-Just as the co-processor must adapt to the lesion, it must also adapt to the network
-dynamics.
+Just as the co-processor must adapt to the lesion, it must also adapt to the dynamics of
+the network it is stimulating. Natural neural networks as well as our stimulated network
+exhibit long running dynamics, which our CPN and EN must account for in their learning.
+A perturbation of a network (i.e. due to stimulation) will cause changes in neuron
+activations long after the stimulation is applied, sometimes far from the site of
+stimulation. Our simulated grasping circuit exhibits the same behavior.
 
-\begin{figure}
+To illustrate this, suppose we applied a small, one-time, instantaneous perturbation
+of the hidden state of 10 randomly chosen neurons in the output module at some point
+in time during a trial. If we repeat that experiment many times, we can see what the
+distribution of long- running effects tends to look like.
+
+In Fig. \ref{fig:dynamics} we can see that even a single, one-time perturbation in
+the network has effects dozens of time steps later. Our co-processor will need to
+account for these, since stimulation is intended to cause perturbations in a network.
+To successfully aid in grasping, the co-processor will need to account for these
+dynamics. As we will show in the next subsection, the problem the co-processor faces
+is in-fact even harder than this, due to a stimulation model that includes both
+spatial and temporal smoothing.
+
+\begin{figure}[h]
 	\centering
@@
 	    \includegraphics[width=\textwidth]{perturbe_single.pdf}
-	    \caption{}
+	    \caption{Example trial}
 	\end{subfigure}
@@
 	    \includegraphics[width=\textwidth]{perturbe_dist.pdf}
-	    \caption{}
+	    \caption{Distribution of effects across n=1000 samples}
 	\end{subfigure}
 	\hfill
-	\caption{Instantaneous perturbations of a random sample (n=10) of M1 neurons at time t=200 results in long-running effects on output.}
+	\caption{Instantaneous perturbations of a random sample (n=10) of M1
+	         neurons at time t=200 results in long-running effects on output.}
 	\label{fig:dynamics}
@@
 
+\subsubsection{Stimulation model}
+TODO: overview of spatio-temporal smoothing, connect to figure.
+
 \subsubsection{Observation model}
-** Observation model
-*** Spatio-temporal smoothing
+TODO: overview of spatial / gaussian convolution concept, connect to $fig:stim_and_obs$.
 
@@
 \end{figure}
-
-\subsubsection{Stimulation model}
-** Stimulation design
 
@@
 
-\subsection{Training the Model}
+\subsection{Training the Co-Processor}
 Alternate training EN and CPN.
@@
 			\item Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
+			\item White noise and noised params
+			\item ** Learning instability
+			\item ** Predictive power isn't enough
+			\item ** Illustrate this
 		\end{itemize}
@@
 		\end{itemize}
+
+	\item Data recycling/reuse
 \end{itemize}
 
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{pseudocode.png}
+\caption{Pseudocode of the Training Method}
+\centering
+\label{fig:pseudocode}
+\end{figure}
+
 \section{Results}
+\begin{itemize}
+	\item Task performance improves drastically.
+		\begin{itemize}
+			\item Fig: normalized task loss vs CPN training epoch
+			\item Fig: normalized task loss vs training epoch
+			\item Fig: normalized task loss vs training epoch
+			\item Split by experiment
+		\end{itemize}
+	\item Fine tuning takes a long time
+	\item Object classes separate. Fig: sep vs experiment vs epoch
+	\item Can co-adapt with the brain, as it changes
+	\item Training efficiency analysis. Show num epochs with and without recycling.
+\end{itemize}
 
-* Task performance improves drastically
-* Fine tuning takes a long time
-* Object classes separate
-* Can co-adapt with the brain, as it changes
-
-* Split by lesion design
-
-* White noise and noised params
-** Learning instability
-** Predictive power isn't enough
-** Illustrate this
-
-* Training efficiency analysis
-
-\section{Discussion}
-* Training efficiency
-* Spectrum from simple low dimensional stimulation vectors today to higher dimensional future
-* Toward in vivo application
-
-\section{Conclusion}
-asdf
+\section{Discussion and Conclusion}
+\begin{itemize}
+	\item Training efficiency
+	\item Spectrum from simple low dimensional stimulation vectors today to higher dimensional future
+	\item Toward in vivo application
+\end{itemize}
 

### 2021-11-11 cpn.py
Update M1 paradigm

@@ -18,7 +18,7 @@ g_logger = logging.getLogger("cpn")
 
 
 class CPN_EN_CoProc(experiment.CoProc):
-    def __init__(self, cfg, log_dir=None):
+    def __init__(self, cfg, log_dir=None, recycle_thresh=None):
         self.cfg = cfg
 
         in_dim, stim_dim, out_dim, cuda = cfg.unpack()
@@ -61,8 +61,8 @@ class CPN_EN_CoProc(experiment.CoProc):
         else:
             self.log_dir = None
 
-        self.round_thresh = 5
-        # A list of up to round_thresh length
+        self.recycle_thresh = recycle_thresh
+        # A list of up to recycle_thresh length
         #  Each item is a tuple (actuals, targets, trial_end, stim list, obs list)
         #   The list is trial_len length.
         #     Each item is a stim or obs
@@ -119,7 +119,7 @@ class CPN_EN_CoProc(experiment.CoProc):
             )
             update_task_loss = False
 
-            if self.epoch_type == EpochType.EN:
+            if self.epoch_type == EpochType.EN and self.recycle_thresh:
                 new_save = (actuals, targets, trial_end, self.stims, self.brain_data)
                 self.saved_data.append(new_save)
 
@@ -148,7 +148,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         checkpoint_eidx = 0
         en_is_ready = False
         while not en_is_ready and checkpoint_eidx < 1000:
-            for bidx in range(self.round_thresh):
+            for bidx in range(self.recycle_thresh):
                 actuals, targets, trial_end, stims, brain_data = self.saved_data[bidx]
 
                 for tidx in range(trial_len):
@@ -225,7 +225,7 @@ class CPN_EN_CoProc(experiment.CoProc):
             else:
                 self.epoch_type = EpochType.EN
 
-            if len(self.saved_data) == self.round_thresh and not en_is_ready:
+            if len(self.saved_data) == self.recycle_thresh and not en_is_ready:
                 en_is_ready = self.train_en_closed_loop(loss_history,
                     next_is_validation)
 

### 2021-11-11 experiment/config.py

@@ -85,6 +85,7 @@ DEFAULT_OBS_OUT_DIM = 20
 DEFAULT_OBS_SIGMA = 1.75
 DEFAULT_NUM_STIM_CHANNELS = 16
 DEFAULT_STIM_SIGMA = 2.175
+DEFAULT_STIM_PAD_RIGHT_NEURONS = 200
 DEFAULT_OUT_DIM = 50
 DEFAULT_HOLDOUT_PCT = 0.2
 
@@ -146,7 +147,9 @@ def get(
     cpn_activation_type=DEFAULT_ACTIVATION_TYPE,
     num_neurons_per_module=DEFAULT_NUM_NEURONS_PER_MODULE,
     num_stim_channels=DEFAULT_NUM_STIM_CHANNELS,
+    num_stim_neurons=None,
     stim_sigma=DEFAULT_STIM_SIGMA,
+    stim_pad_right_neurons=DEFAULT_STIM_PAD_RIGHT_NEURONS,
     stim_retain_grad=False,
     obs_out_dim=DEFAULT_OBS_OUT_DIM,
     obs_sigma=DEFAULT_OBS_SIGMA,
@@ -167,10 +170,14 @@ def get(
     else:
         raise ValueError(f"Unrecognized observer type: {observer_type}")
 
+    if num_stim_neurons is None:
+        num_stim_neurons = num_neurons_per_module
+
     if stimulation_type is stim.StimulationType.one_to_one:
         stimulus = stimulation_type.value(
-            num_neurons_per_module,
-            num_neurons_per_module,
+            num_stim_neurons,
+            num_stim_neurons,
+            pad_right_neurons=stim_pad_right_neurons,
         )
     elif stimulation_type is stim.StimulationType.gaussian_alpha:
         if cuda:
@@ -179,16 +186,18 @@ def get(
         # NOTE: can add the num_stim_channels and sigma arg above
         stimulus = stimulation_type.value(
             num_stim_channels,
-            num_neurons_per_module,
+            num_stim_neurons,
+            pad_right_neurons=stim_pad_right_neurons,
             sigma=stim_sigma,
         )
     elif stimulation_type is stim.StimulationType.gaussian_exp:
         stimulus = stimulation_type.value(
             num_stim_channels,
-            num_neurons_per_module,
+            num_stim_neurons,
             batch_size=1,  # Will be reset before use
             sigma=stim_sigma,
             retain_grad=stim_retain_grad,
[…]

### 2021-11-11 experiment/experiment.py

@@ -43,33 +43,30 @@ def get_coadapt_config(cuda=None, **kwargs):
 
 def get_m1_lesion_config(cuda=None, **kwargs):
     lesion_type = lesion.LesionType.outputs
-    lesion_args = (0, 50)
+    # Lesion the neurons with idxs 50-100, which are in M1
+    lesion_args = (50, 100)
     return get_config(
         cuda=cuda,
         coadapt=True,
         drop_m1=True,
         lesion_type=lesion_type,
         lesion_args=lesion_args,
+        num_stim_neurons=50,
+        stim_pad_right_neurons=250,
         **kwargs,
     )
 
 
-def get_config(recover_after_lesion=False, coadapt=False, cuda=None, **kwargs):
+def get_config(recover_after_lesion=False, coadapt=False, num_stim_neurons=None,
+        stim_pad_right_neurons=config.DEFAULT_STIM_PAD_RIGHT_NEURONS,
+        cuda=None, **kwargs):
     """
     Args:
         - cuda (str, torch.device, or None): None for CPU, or a string like "0" specifying a GPU
                               device. This will be passed to the CUDA_VISIBLE_DEVICES env var.
                               Reference: https://docs.nvidia.com/cuda/cuda-c-programming-guide/index.html#env-vars
     Returns:
-        - in_dim (int): dimensionality of the data coming from the brain
-        - stim_dim (int): dimensionality of stimulation vector, which we get
-                          from the co-proc
-        - out_dim (int): dimensionality of the task: muscle velocities in our
-                         case
-        - cuda (torch.device, or None): will be a torch.device if we are using
-                                        GPU. Note your model/Module will need to
-                                        use this to push any parameters to GPU
-                                        if the experiment is running on GPU.
+        A config.Config, which contains the parameterization of this experiment.
     """
     if isinstance(cuda, str):
         os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID"
@@ -82,6 +79,8 @@ def get_config(recover_after_lesion=False, coadapt=False, cuda=None, **kwargs):
 
     cfg = config.get(
         recover_after_lesion=recover_after_lesion,
+        num_stim_neurons=num_stim_neurons,
+        stim_pad_right_neurons=stim_pad_right_neurons,
         coadapt=coadapt,
         cuda=cuda_out,
         **kwargs,

### 2021-11-11 paper/article.tex
Paper progress

@@
 \begin{figure}
+	\centering
+	\begin{subfigure}[c]{0.69\textwidth}
+		\centering
 		\includegraphics[width=\textwidth]{michaels.eps}
@@
 		network, conditioning the grasp for the object of the particular size, shape, and location.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[c]{0.30\textwidth}
 		\centering
+		\includegraphics[width=\textwidth]{mRNN_J.png}
+		\caption{Connectivity matrix $J$. Note the within-module connections
+		(on the diagonal) are fully connected, and connections to adjacent
+		modules are sparse $10\%$.}
+	\end{subfigure}
+	\hfill
+\caption{Architecture of the Michaels mRNN}
 \label{fig:michaels}
@@
 Just as the co-processor must adapt to the lesion, it must also adapt to the dynamics of
-the network it is stimulating. Natural neural networks as well as our stimulated network
+the network it is stimulating. Natural neural networks as well as our simulated network
 exhibit long running dynamics, which our CPN and EN must account for in their learning.
@@
 \subsection{Training the Co-Processor}
-Alternate training EN and CPN.
+In all experiments, training the co-processor requires a careful interleaving of
+EN and CPN training epochs. EN training epochs concentrate on updating the EN,
+based on observations of the effect of stimulation on the simulated network.
+The loss function when training the EN is the MSE loss between the actual and
+predicted output of the mRNN (i.e. muscle velocities). We then use the EN
+to train the CPN during the CPN training epochs. During that time, we backpropagate
+the MSE loss between the EN output, and target output, to the CPN. See Fig.
+\ref{fig:training}.
+
+TODO: explain EN training
+
+\begin{figure}
+	\centering
+	\begin{subfigure}[c]{0.48\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{backprop_en.eps}
+		\caption{EN training phase: backpropagate actual vs predicted}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[c]{0.48\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{backprop_cpn.eps}
+		\caption{CPN training phase: backpropagate actual vs target MSE loss, via EN, to CPN. This effectively treats the EN output as the actual, and thus the EN must be trained first.}
+	\end{subfigure}
+	\hfill
+\caption{Two Phase Training Regime}
+\label{fig:training}
+\end{figure}
+
 \begin{itemize}

### 2021-11-11 experiment/lesion.py

@@ -71,6 +71,19 @@ class LesionOutputsByIdxs(Lesion):
     def __str__(self):
         return f"outputsIdxs{self.start_idx}.{self.end_idx}"
 
+class LesionOutputsByMask(Lesion):
+    def __init__(self, num_neurons_per_module, lesion_mask, cuda=None):
+        assert lesion_mask.shape == (num_neurons_per_module * 3,)
+        self.lesion_mask = lesion_mask
+
+    def lesion(self, network, pre_response):
+        batch_size = pre_response.shape[0]
+        out = pre_response * torch.tile(self.lesion_mask, (batch_size, 1))
+        return out
+
+    def __str__(self):
+        return f"outputsIdxs{self.start_idx}.{self.end_idx}"
+
 
 class LesionConnectionsByIdxs(Lesion):
     application = "connection"

### 2021-11-11 paper/article.tex
Paper progress

@@
 were supplied to us by the lead author Jonathan Michaels, for the purpose of our present
-simulation. See Fig. \ref{fig:michaels}.
+simulation. We re-implemented his models' logic in PyTorch, and loaded his trained parameters
+into it, for one of his subjects, arbitrarily chosen. See Fig. \ref{fig:michaels}.
 
@@
 \subsubsection{Stimulation model}
-TODO: overview of spatio-temporal smoothing, connect to figure.
 
-\subsubsection{Observation model}
-TODO: overview of spatial / gaussian convolution concept, connect to $fig:stim_and_obs$.
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{stim_single.png}
+	\caption{Spatial and temporal smoothing of a single 16 dimensional,
+	         randomized $\theta$, onto 100 simulated neurons.
+		 Color values indicate the magnitude of the value summed into
+		 each neuron's hidden state in that time step. After $t=0$,
+		 $\theta$ is the zero vector.}
+	\centering
+	\label{fig:stim_single}
+\end{figure}
 
-\begin{figure}
+In order to inject another aspect of realism into our simulation, we subject
+the co-processor to a stimulation model, rather than allowing it to directly
+influence the hidden state of the simulated network's neurons. Our stimulation
+model includes aspects of both spatial and temporal smoothing.
+
+In our experiments, we stimulate only the output module, since this co-processor's
+purpose is to improve external task performance, which it is able to do with
+only stimulation of the output module of the network. It is conceivable that
+a co-processor could stimulate other areas of the brain to improve task performance
+downstream, or to probe the brain to better reveal the user's intent (i.e. object
+shape), but we did not explore those possibilities in this paper.
+
+The stimulation function $S$ receives as inputs the stimulation parameters $\theta$
+provided by the CPN, which it accumulates in an internal memory. Each time step,
+it outputs a modifier to the internal states of all neurons in the output module,
+based on the current state of memory in the function object. That memory allows
+the stimulation function to perform temporal smoothing. Specifically, we used a
+simple exponential decay model where the function reads its current state of memory,
+sums in the new parameters, and then decays each memory element towards $0.0$ at some
+rate. Roughly speaking, this design intends to approximate the notion of dissipation: the
+effect of stimulation is not instantaneous, but rather decays with time, as charge
+dissipates into the surrounding area.
+
+Likewise, the stimulation function maps the relatively low dimensional stimulation
+parameter vector $\theta$ onto the M1 cells which it stimulates. In our experiments,
+our stimulation parameters $\theta$ had 16 dimensions, which we found was
+sufficient to allow significant improvement in task performance. The stimulation
+function uses a simple Gaussian smoothing to map those parameters onto the simulated
+M1 neurons. Each parameter represents, in a sense, an electrode located along a single
+spatial dimension, whose stimulation affects the neurons in its vicinity more than
+it affects others, according to each neuron's Mahalanobis distance from it. The neurons
+are aligned along that dimension arbitrarily. We fix the width of the Gaussians
+arbitrarily to $\sigma=1.75$. Note that we do not attempt to model the physical
+layout of our simulated neurons, but simply want to disallow our co-processor
+from having an unrealistically high degree of spatial resolution.
+
+Thus, the governing equations of the stimulation become:
+
+\begin{equation}
+\alpha_{t} = \tau\alpha_{t-1} + \theta_{t}
+\end{equation}
+\begin{equation}
+s_{t} = C(\alpha_{t})
+\end{equation}
+
+\begin{itemize}
+	\item $\alpha$: the 16 dimensional internal activation, or memory
+        of our stimulation
+	\item $\tau$: our decay rate, which we set arbitrarily to $0.7$
+	\item $C$: a precalculated $100 x 16$ matrix which provides our
+	Gaussian smoothing
+	\item $s_{t}$: the stimulation we apply to each neuron at the
+	given time step.
+\end{itemize}
+
+The governing equations of the simulated network then become:
+
+\begin{equation}
+x_{t+1} = Jx_{t} + Iv_{t} + s_{t}
+\end{equation}
+\begin{equation}
+a_{t} = tanh(x_{t})
+\end{equation}
+\begin{equation}
+y_{t} = La_{t} + b
+\end{equation}
+
+\begin{itemize}
+	\item $x$: the hidden state of each neuron
+	\item $J$: the recurrence weight matrix
+	\item $I$: the input response matrix
+	\item $L$, $b$: the parameters of the linear readout
+	\item $y$: the output of the network
+\end{itemize}
+
+See Fig. \ref{fig:stim_single} for a visual depiction of an example where
+$\theta$ is non-zero at $t=0$, and zero for all other times, in order to see how
+stimulation is applied across space and time. In Fig. \ref{fig:stim_and_obs},
+we see stimulation applied during the simulation, in a real trial. In this case,
+the trial is one where $50\%$ of M1 neurons are lesioned to have zero output.
+Observations of the network's hidden state (explained in the next section) are taken
+from the first two modules, and stimulation is applied to the last module. This fact
+is intended to capture the difficulties with observing the same cortex that you are
+stimulating, due to stimulation artifacts. The trial we show depicts stimulation
+supplied by a highly trained co-processor, where task performance has been
+largely restored. We explain this experiment in more detail below.
+
+\begin{figure}[h]
+	\centering
 	\includegraphics[width=\textwidth]{stim_and_obs.png}
-\caption{Example of stimulation and observation for a single trial}
-\centering
+	\caption{Example of stimulation and observation for a single trial. Here, M1
+	has been lesioned 50\%, as seen in the zero (light blue) hidden states. We
+	gather observations only from the AIP and F5 modules, as we explain in
+	Section \ref{sec:experiments}. Stimulation is applied to M1, to drive the
+	network output.}
 	\label{fig:stim_and_obs}
@@
 
-\subsection{Experiments}
-Summary of tunables:
-\begin{itemize}
-	\item Learning rate schedules
-	\item Activation functions
-	\item Obs function dimensionality
-	\item Stim dimensionality
-	\item Dimensionality of EN, CPN
-\end{itemize}
+\subsubsection{Observation model}
+Our observation model likewise relies on a notion of electrodes spread along
+a single spatial dimension. As with stimulation, this is not intended to
+capture true physical relationships between neurons and electrodes, but
+rather to act as a dimensionality reduction that isn't designed to specifically
+to favor task success, such as PCA. We treat the neurons of each module as
+occupying their own space, with 20 electrodes arrayed along each.
 
-Experiment types:
-\begin{itemize}
-	\item Connectivity lesion
-	\item Co-adaptive with connectivity lesion
-	\item Co-adaptive with partial M1 loss
-	\item Co-adaptive with partial AIP loss
-	\item Co-adaptive with data reuse
-\end{itemize}
+We use a similar Gaussian-based approach as with stimulation, but in this
+case our approach looks much like a Gaussian convolution, where the Gaussian
+kernel is centered at each electrode position. Each `electrode' is read out
+as a weighted average of all neurons in the given module, with weights being
+provided by the Gaussian kernel. As above, refer to Fig. \ref{fig:stim_and_obs}
+for a visual example.
 
-\begin{figure}
-\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
-\caption{Experiment architecture overview}
-\centering
-\label{fig:exp_overview}
-\end{figure}
+\subsubsection{Brain co-adaptation}
+To demonstrate the co-processor's ability to adapt to the non-stationarity
+of the brain, we also change it throughout the co-processor's training.
+Specifically, we simulate the brain's co-adaptation with the stimulation
+to cooperatively solve the problem, as the stimulation is also changing
+through the CPN's training.
 
-\subsection{Training the Co-Processor}
+We achieve this through a simple error backpropagation, using PyTorch's
+implementation of the $Adam$ optimizer. With each trial, task loss is
+calculated and backpropagated into the mRNN, allowing it to adapt to
+its inputs, one of which is the stimulation being applied. The learning
+rate for the optimizer is set arbitrarily to a relatively low rate of
+$1e-7$.
+
+\subsection{Training Algorithm}
 In all experiments, training the co-processor requires a careful interleaving of
@@
 
+Alternate training EN and CPN.
 \begin{itemize}
@@
 
+\subsection{Experiments}
+\label{sec:experiments}
+
+Using the simulation described above, we perform three experiments to
+demonstrate the co-processor's ability to learn:
+\begin{itemize}
+	\item 50\% AIP loss
+	\item 100\% connectivity lesion between the F5 and M1 modules
+	\item 50\% M1 loss
+\end{itemize}
+
+Additionally, we perform each of the ...
+
+Summary of tunables:
+\begin{itemize}
+	\item Learning rate schedules
+	\item Activation functions
+	\item Obs function dimensionality
+	\item Stim dimensionality
+	\item Dimensionality of EN, CPN
+\end{itemize}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
+\caption{Experiment architecture overview}
+\centering
+\label{fig:exp_overview}
+\end{figure}
+
 \section{Results}
+
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{results_standin.png}
+\caption{Loss and class separation vs training epoch}
+\centering
+\label{fig:results}
+\end{figure}
+
 \begin{itemize}

### 2021-11-11 cpn.py
Training update

@@ -147,7 +147,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         done = False
         checkpoint_eidx = 0
         en_is_ready = False
-        while not en_is_ready and checkpoint_eidx < 1000:
+        while not en_is_ready and checkpoint_eidx < 100:
             for bidx in range(self.recycle_thresh):
                 actuals, targets, trial_end, stims, brain_data = self.saved_data[bidx]
 

### 2021-11-12 experiment/experiment.py
Add AIP lesion

@@ -56,6 +56,19 @@ def get_m1_lesion_config(cuda=None, **kwargs):
         **kwargs,
     )
 
+def get_aip_lesion_config(cuda=None, **kwargs):
+    lesion_type = lesion.LesionType.outputs
+    # Lesion the neurons with idxs 200-250, which are in AIP
+    lesion_args = (200, 250)
+    return get_config(
+        cuda=cuda,
+        coadapt=True,
+        drop_m1=True,
+        lesion_type=lesion_type,
+        lesion_args=lesion_args,
+        **kwargs,
+    )
+
 
 def get_config(recover_after_lesion=False, coadapt=False, num_stim_neurons=None,
         stim_pad_right_neurons=config.DEFAULT_STIM_PAD_RIGHT_NEURONS,

### 2021-11-12 paper/article.tex
Training update
Paper progress

@@
 	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
-	    \caption{Proportion of loss attributable to hand muscles}
+	    \caption{Ratio of hand muscle MSE losses vs. overall MSE losses, for different lesion types}
 	\end{subfigure}
@@
 \begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{stim_single.png}
+	\begin{subfigure}[c]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{stim_single.png}
+		\caption{$\sigma=1.875$}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[c]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{stim_single2.png}
+		\caption{$\sigma=2.875$}
+	\end{subfigure}
+
 	\caption{Spatial and temporal smoothing of a single 16 dimensional,
@@
 		 $\theta$ is the zero vector.}
-	\centering
 	\label{fig:stim_single}
@@
 rate for the optimizer is set arbitrarily to a relatively low rate of
-$1e-7$.
+$1e-7$. The learning rate was not chosen in a principled way, since
+we are not aware of a principled way of choosing it. However, we reason
+that it should be lower than the co-processor by orders of magnitude due
+to the rapidity by which an ANN can be retrained.
+
+\subsubsection{Simulating recovery prior to co-processor use}
+We sought to additonally pre-train the mRNN after applying the lesion in
+order to simulate stroke recovery. For simulated lesions which zero the outputs
+of neurons, the mRNN is, unfortunately, not an appropriate model.
+We know that strokes can cause damage from which the subject, in many cases,
+cannot recover. The mRNN model has sufficient redundancy built into it that
+lesioning the outputs of neurons leaves enough remaining degrees of freedom
+that a nearly full recovery can occur, unless so many neurons are lesioned
+that no stimulation could be effective. We go into this further in the
+Discussion section below.
+
+We can, however, simulate pre-recovery in the case of a lesion that excludes
+communication between the F5 and M1 modules. In that case, object shape information
+cannot propagate forward in the network, in order to condition the hand for
+grasping. The mRNN can, in pre-recovery, learn a stereotyped grasp, at which
+point the co-processor's job will be to forward-propagate that information.
+We explore this in one of our experiments, explained below.
 
@@
 EN and CPN training epochs. EN training epochs concentrate on updating the EN,
-based on observations of the effect of stimulation on the simulated network.
+based on observations of the effect of stimulation on the simulated network's output.
 The loss function when training the EN is the MSE loss between the actual and
@@
 to train the CPN during the CPN training epochs. During that time, we backpropagate
-the MSE loss between the EN output, and target output, to the CPN. See Fig.
-\ref{fig:training}.
+the MSE loss between the EN's predicted output, and target output, to the CPN.
+See Fig. \ref{fig:training}.
 
-TODO: explain EN training
+\begin{figure}
+\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
+\caption{Experiment architecture overview}
+\centering
+\label{fig:exp_overview}
+\end{figure}
 
@@
 
-Alternate training EN and CPN.
-\begin{itemize}
-	\item Train an EN
-		\begin{itemize}
-			\item Training the EN until predictive power reaches a threshold, based on current task performance
-			\item Training regime based on most recent CPN, similar CPNS (noise added to params), and white noise stimulation
-			\item Single batch of data from applying current CPN set and white noise to the brain, EN fit to ``offline'' it over maining epochs
-			\item Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
-			\item White noise and noised params
-			\item ** Learning instability
-			\item ** Predictive power isn't enough
-			\item ** Illustrate this
-		\end{itemize}
-	\item Train the CPN
-		\begin{itemize}
-			\item Backprop through the EN, effectively using the EN's prediction as ground truth
-		\end{itemize}
+\subsubsection{EN training}
+For the EN to be useful, it must accurately predict the effect of stimulation
+produced by the CPN. In addition to that, backpropagating through it must yield
+gradients which train the CPN to output more useful stimulus. We discovered that
+this latter property does not necessarily occur simply by virtue of the former.
+We have found that an EN can be trained to extreme levels of predictive power, even
+on white noise stimulation - to orders of magnitude lower loss than the task loss -
+while at the same time backpropagation through it yields gradients which are not
+useful for training the CPN. As a result, the CPN exhibits unstable training.
+We have thus far been unable to discover a way to measure and optimize for this second
+property, though pursuing it may be a fruitful subject of future ML research. Instead
+of attempting to train an EN directly to exhibit this property, we discovered a
+training regime that reliably makes it appear.
 
-	\item Data recycling/reuse
-\end{itemize}
+The central concept of our training regime is to provide a carefully chosen variety of
+training examples, and to use heavy weight regularization, in order to force
+the EN to generalize well to stimulation which it hasn't yet seen. We hypothesize that
+the crux of our EN training problem is one of over-fitting: that the EN may be trained
+to great predictive power on a set of stimulation examples, but that the effective
+dimensionality of the problem it is learning is comparatively quite high, due to both
+the dimensionality of our stimulation parameters $\theta$, and the dynamics of the
+network being stimulated. As a result, the hypersurface which it has fit to the training
+examples may exhibit gradients not reflective of reality.
 
-\begin{figure}[h]
-\includegraphics[width=0.5\textwidth]{pseudocode.png}
-\caption{Pseudocode of the Training Method}
-\centering
-\label{fig:pseudocode}
-\end{figure}
+First, and most significantly, we structure the training data set for each EN training epoch
+in a very specific way. Each epoch, we include examples from the current CPN, the current CPN
+with Gaussian-distributed mean-0 noise added to its parameters, and examples of white noise
+stimulation. We developed this structure under the believe that we needed to cover a sufficient
+variety of examples to prevent overfit, and to do so in a way that emphasizes the neighborhood
+in CPN parameter space of the current CPN. We initially attempted to pad our dataset with white
+noise examples alone, but that was not sufficient to stabilize CPN training, despite the EN's
+ability to reach a high predictive power. Adding in the addition of samples from the parameter
+space near the current CPN appears to help ensure that the EN is localized enough to give
+useful gradients for that area. As we will explain below, this produces an additional problem
+that the EN is to specialized to the local area of CPN parameter space, but we then solve
+that problem using EN retraining, interleaved with the CPN, as we will explain further below.
+
+Thus, we composed each batch of training data in the following way:
+
+\begin{table}[]
+\begin{tabular}{|l|l|}
+\hline
+\textbf{Source}                                                                                                                            & \textbf{Proportion of dataset} \\ \hline
+Examples produced by the current CPN                                                                                                       & 10\%                           \\ \hline
+\begin{tabular}[c]{@{}l@{}}Examples produced by a copy of the current CPN,\\ with mean-0 random noise added to its parameters\end{tabular} & 60\%                           \\ \hline
+Examples of white noise stimulation                                                                                                        & 30\%                           \\ \hline
+\end{tabular}
+\end{table}
+
+Second, we use of PyTorch's $AdamW$ optimizer, which includes an option for weight regularization,
+and a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
+one of the standard mitigations for overfit. To converge, the learning rate schedule also needs
+to be chosen carefully: a relatively high learning rate at first, dropping to a very low rate soon
+thereafter. Using too-low or too-high of a learning rate in any phase causes the EN learning
+to not converge. EN training proceeds until it reaches a threshold of prediction error, defined as
+a fraction of the current CPN's task loss.
+
+\subsubsection{CPN training}
+Once we discovered an effective way to train an EN, CPN training was straightforward. For this phase,
+We generate training examples using the current CPN alone. The EN generates predictions of the
+effect of those stimulations, and we backpropagate through the EN to generate training gradients
+for the CPN. As with the EN, we constitute the dataset with a large random sample of trials from
+the original Michaels task. See Fig. \ref{fig:training}.
+
+As with the EN, the choice of learning rate schedule for training the CPN is somewhat important.
+The CPN appears to train in two phases. In the first phase, it is largely learning the structure
+of the reach-to-grasp task, e.g. that the muscles need to stay still until the reach begins. During
+that phase, the learning rate can be quite high. Later, the CPN begins to learn the mapping between
+the object shape information observed from the simulated brain, and how that should map onto
+stimulation. That phase of the training takes much longer, and requires a learning rate 2-3
+orders of magnitude lower than in the first phase. We will cover this in some more detail in
+Section \ref{sec:results}.
+
+\subsubsection{Interleaved CPN/EN training, and adapting to non-stationarity}
+Having defined the training epochs for the EN and CPN, we can define a training
+algorithm on the basis of those two. We train the two in alteration, creating
+a new EN each time we enter an EN training period. We explored the possibility
+of reusing an existing EN by retraining it, for the purpose of training efficiency.
+However, we found that retraining an EN was no more fast than training a new one,
+and in-fact usually took longer, despite significant experimentation with learning
+rates.
+
+The key challenge we face is determining when an EN is no longer suitable for training the
+current CPN. Expiring an EN at the right time accomplishes two things. First, it ensures
+that our current EN is trained properly for the current CPN; without that, the CPN training
+will become unstable. Second, expiring the EN is how we adapt our learning to the brain's
+non-stationarity. At some point, the brain will have changed sufficiently for our EN to be
+out-of-date, and therefore requiring replacement.
+
+We solve this challenge through a simple set of metrics, which, when satisfied, indicate
+we need to transition to an EN training epoch. Our first metric simply captures the EN's
+declining predictive power. When the EN's prediction error reaches a point which is
+sufficiently above some ratio to the CPN's task loss, we expire the EN immediately.
+Second, we expire the EN if CPN loss does not improve (or gets worse) across a sufficient
+number of recent training steps. This resembles common stopping conditions in
+iterative learning: one stops when learning is no longer improving the model. Together,
+these two metrics appear effective in ensuring we expire the EN at the right time.
+
+Unfortunately, because we expire the EN rather than retraining it, and because we need
+to expire it so often, training is less efficient than it otherwise could be. We explore
+this some in Section \ref{sec:results}. In the next subsection, we illustrate one way we
+mitigate training inefficiency.
+
+TODO: pseudocode
+
+\subsubsection{Optimization: data reuse}
+One obvious
+
+\section{Results}
+\label{sec:results}
 
@@
 	\item 100\% connectivity lesion between the F5 and M1 modules
+	\item 100\% connectivity lesion between the F5 and M1 modules without brain co-adpatation
 	\item 50\% M1 loss
@@
 \end{itemize}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
-\caption{Experiment architecture overview}
-\centering
-\label{fig:exp_overview}
-\end{figure}
-
-\section{Results}
 

### 2021-11-13 paper/article.tex
Paper progress

@@
 \begin{abstract}
-\textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
+\textit{Objective} Brain co-processors \cite{rao.coproc} are devices which use artificial
 intelligence (AI) for closed-loop neurostimulation, to shape neural activity and to bridge
@@
 leverages deep learning, through the use of a simulated neural circuit.
-That circuit performs reach-to-grasp task, based on visual input, and is designed to
+That circuit performs a reach-to-grasp task, based on visual input, and is designed to
 closely resemble a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate
@@
 	      as well as to adapt to that circuit's ongoing changes.
-	\item Data-efficient methods for training these models, to better ensure we
+	\item Training methods for our stimulation model, to make it effective in later
+	      training our stimulation agent.
+	\item Data reuse methods, to improve learning efficiency, to better ensure we
 	      can train them with a biologically-realistic amount of data.
-	\item How we must train our stimulation model to make it effective in later
-	      training our stimulation agent.
 \end{itemize}
@@
 effectively allowing for brain-stimulator co-adaptation. The EN is a tool for training the CPN,
-giving us a way to back-propagate task error to the CPN. It outputs task-relevant
+giving us a way to backpropagate task error to the CPN. It outputs task-relevant
 metrics - a prediction of muscle velocities in our case - given measurements of neural
@@
 we didn't deeply test a strictly linear approach.
-Second, LSTM cells are designed to better capture long-running network
-dynamics, and in our case we found that they were crucial for that reason:
-when using vanilla RNN cells, our co-processor wasn't able to train well.
-Compare that to the simpler cells commonly used in vanilla RNNs, which
-yield functionally the same approach as the linear state space model, less
-the use of a nonlinear activation function.
+Second, LSTM cells are designed to better capture the relevant asepcts of
+long-running network dynamics, and in our case we found that they were
+crucial for that reason: when using vanilla RNN cells, our co-processor
+wasn't able to train well. Compare that to the simpler cells commonly
+used in vanilla RNNs, which yield functionally the same approach as the
+linear state space model, less the use of a nonlinear activation function.
 
@@
 ``modular'' vanilla RNN (mRNN), and a linear readout layer. Each ``module'' consists of 100 vanilla
-RNN neurons, with a non-linearity applied on the outputs. The modules are internally fully connected,
+RNN neurons, with a nonlinearity applied on the outputs. The modules are internally fully connected,
 and are connected to each other sparsely (10\% connectivity). The inputs are visual features
-intended to capture the view the monkeys had during the task, specifically VGGNet features from 3D
-renderings of the same objects which the monkeys grasped. The outputs are muscle length
-velocities for the should, arm, and hand of the monkey during the trial. The natural
+intended to capture the view the monkeys had during the task, specifically VGGNet features \cite{simonyan.vgg},
+from 3D renderings of the same objects which the monkeys grasped. The outputs are muscle length
+velocities for the shoulder, arm, and hand of the monkey during the trial. The natural
 velocities were captured with a motion capturing glove, and the artificial neural network
@@
 	      neurons to zero, effectively removing them from the network. This results in a
-	      loss of object shape information and, as we will see below, causes the co-processor
-	      to be unable to differentiate objects.
-	\item \textbf{Connection-based:} we prevent the forward and backward propagation of
+	      loss of object shape information until the brain co-adapts to the lesion.
+	\item \textbf{Connection-based:} we prevent the propagation of
 	      information between the ``F5'' and ``M1'' modules, effectively representing a
-	      severing of the connections between the two.
+	      severing of the connections between the two. Note that the sparse connections
+	      between them run both directions, and we are lesioning both.
 	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
-	      neurons to zero, causing a more holisitic loss of movement.
+	      neurons to zero, causing a more holisitic loss of movement. Here, the lesion
+	      may make it impossible for the co-processor to find a solution, unless the
+	      brain co-adapts with it to re-map the M1 neurons onto the muscle space.
 \end{itemize}
@@
 To illustrate this, suppose we applied a small, one-time, instantaneous perturbation
-of the hidden state of 10 randomly chosen neurons in the output module at some point
+of the hidden state of 10 randomly chosen neurons in the output (M1) module at some point
 in time during a trial. If we repeat that experiment many times, we can see what the
@@
 the network has effects dozens of time steps later. Our co-processor will need to
-account for these, since stimulation is intended to cause perturbations in a network.
-To successfully aid in grasping, the co-processor will need to account for these
-dynamics. As we will show in the next subsection, the problem the co-processor faces
+account for these dynamics, since stimulation is intended to cause perturbations
+in a network. As we will show in the next subsection, the problem the co-processor faces
 is in-fact even harder than this, due to a stimulation model that includes both
@@
 
+The intent with this approach is not to create a realistic biophysical model. The
+mRNN model does not contain sufficient information to arrange the neurons in a
+simulated space, and to simulate how that may change the effects of stimulation.
+Instead, our stimulation model is merely attempting to make the co-processors task
+more realistic: it cannot precisely set the hidden state of every neuron,
+instantaneously and in each time step. Any stimulation it applies will diffuse
+across ``space'' and time.
+
 In our experiments, we stimulate only the output module, since this co-processor's
@@
 are aligned along that dimension arbitrarily. We fix the width of the Gaussians
-arbitrarily to $\sigma=1.75$. Note that we do not attempt to model the physical
-layout of our simulated neurons, but simply want to disallow our co-processor
-from having an unrealistically high degree of spatial resolution.
+arbitrarily to $\sigma=1.75$. We are not aware of a principled way of picking this
+value, and clearly if we set it to extreme values, it makes our simulation less useful.
+For example, if it is set to an extreme high value, the stimulation is effectively
+lower dimensional, resulting in a model corresponding to a stimulation technology that
+is inappropriately low resolution for our task, for example epidural stimulation.
+We experimented with this value, and found we could, unsurprisingly, set it to a high
+enough or low enough value that the co-processor cannot succeed, though the co-processor
+could succeed over a broad range of values. However, we could not conclude much from
+those experiments, because it isn't clear what those experiments ultimately indicate.
+As a result, we don't have a principled way of choosing an appropriate $\sigma$ value;
+we set it to a moderate, balanced value, and don't vary it.
 
@@
 \begin{equation}
-x_{t+1} = Jx_{t} + Iv_{t} + s_{t}
+x_{t+1} = Jx_{t} + Iv_{t} + s_{t} + b
 \end{equation}
@@
 \begin{equation}
-y_{t} = La_{t} + b
+y_{t} = La_{t} + l
 \end{equation}
@@
 	\item $I$: the input response matrix
-	\item $L$, $b$: the parameters of the linear readout
+	\item $b$: the neuron activation bias
+	\item $L$, $l$: the parameters of the linear readout
 	\item $y$: the output of the network
@@
 from the first two modules, and stimulation is applied to the last module. This fact
-is intended to capture the difficulties with observing the same cortex that you are
+is intended to capture the difficulties with observing the same region that you are
 stimulating, due to stimulation artifacts. The trial we show depicts stimulation
@@
 rather to act as a dimensionality reduction that isn't designed to specifically
-to favor task success, such as PCA. We treat the neurons of each module as
+favor task success, such as PCA. We treat the neurons of each module as
 occupying their own space, with 20 electrodes arrayed along each.
@@
 cannot recover. The mRNN model has sufficient redundancy built into it that
-lesioning the outputs of neurons leaves enough remaining degrees of freedom
-that a nearly full recovery can occur, unless so many neurons are lesioned
-that no stimulation could be effective. We go into this further in the
-Discussion section below.
+lesioning the outputs of large numbers of neurons leaves enough remaining
+degrees of freedom that a nearly full recovery can occur, unless so many
+neurons are lesioned that no stimulation could be effective. We go into
+this further in the Discussion section below.
 
@@
 to train the CPN during the CPN training epochs. During that time, we backpropagate
-the MSE loss between the EN's predicted output, and target output, to the CPN.
+the MSE loss between the EN's predicted output, and target output, to the CPN. In
+other words: we treat the EN output as a proxy to the true brain, allowing us
+to train the CPN, since we have no way to ``backpropagate through the brain'' itself.
 See Fig. \ref{fig:training}.
@@
 We have found that an EN can be trained to extreme levels of predictive power, even
-on white noise stimulation - to orders of magnitude lower loss than the task loss -
+on random stimulation, and to orders of magnitude lower loss than the task loss,
 while at the same time backpropagation through it yields gradients which are not
-useful for training the CPN. As a result, the CPN exhibits unstable training.
-We have thus far been unable to discover a way to measure and optimize for this second
-property, though pursuing it may be a fruitful subject of future ML research. Instead
-of attempting to train an EN directly to exhibit this property, we discovered a
-training regime that reliably makes it appear.
+useful for training the CPN. As a result, if the EN is not trained in specific ways,
+the CPN exhibits unstable training. We have thus far been unable to discover a way
+to measure and optimize for this second property, though pursuing it may be a
+fruitful subject of future applied ML research. Instead of attempting to train an EN
+directly to exhibit this property, we discovered a training regime that reliably makes
+it occur.
 
@@
 First, and most significantly, we structure the training data set for each EN training epoch
-in a very specific way. Each epoch, we include examples from the current CPN, the current CPN
-with Gaussian-distributed mean-0 noise added to its parameters, and examples of white noise
-stimulation. We developed this structure under the believe that we needed to cover a sufficient
+in a very specific way. Each epoch, we include:
+\begin{itemize}
+	\item examples using stimulation from the current CPN
+	\item a small (100s) collection of CPNs which are copies of the current CPN, with
+	      Gaussian-distributed mean-0 noise added to their parameters
+	\item examples of white noise stimulation
+\end{itemize}
+We developed this structure under the belief that we needed to cover a sufficient
 variety of examples to prevent overfit, and to do so in a way that emphasizes the neighborhood
@@
 useful gradients for that area. As we will explain below, this produces an additional problem
-that the EN is to specialized to the local area of CPN parameter space, but we then solve
-that problem using EN retraining, interleaved with the CPN, as we will explain further below.
+that the EN is too specialized to the local area of CPN parameter space, but we then solve
+that problem using EN retraining, interleaved with the CPN training.
 
-Thus, we composed each batch of training data in the following way:
+After some experimentation, we composed each batch of EN training data in the following way:
 
-\begin{table}[]
+\begin{table}[h]
 \begin{tabular}{|l|l|}
@@
 \textbf{Source}                                                                                                                            & \textbf{Proportion of dataset} \\ \hline
-Examples produced by the current CPN                                                                                                       & 10\%                           \\ \hline
-\begin{tabular}[c]{@{}l@{}}Examples produced by a copy of the current CPN,\\ with mean-0 random noise added to its parameters\end{tabular} & 60\%                           \\ \hline
-Examples of white noise stimulation                                                                                                        & 30\%                           \\ \hline
+Current CPN                                                                                                       & 10\%                           \\ \hline
+\begin{tabular}[c]{@{}l@{}}Current CPN with parameter noise\end{tabular} & 60\%                           \\ \hline
+White noise stimulation                                                                                                        & 30\%                           \\ \hline
 \end{tabular}
@@
 
-Second, we use of PyTorch's $AdamW$ optimizer, which includes an option for weight regularization,
-and a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
+Second, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
+Additionally, we use a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
 one of the standard mitigations for overfit. To converge, the learning rate schedule also needs
@@
 Once we discovered an effective way to train an EN, CPN training was straightforward. For this phase,
-We generate training examples using the current CPN alone. The EN generates predictions of the
+we generate training examples using the current CPN alone. The EN generates predictions of the
 effect of those stimulations, and we backpropagate through the EN to generate training gradients
@@
 
+TODO: train vs test
+
 Summary of tunables:

### 2021-11-13 cpn.py
Exp progress. Added AIP and M1 coadapt tunables

@@ -147,7 +147,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         done = False
         checkpoint_eidx = 0
         en_is_ready = False
-        while not en_is_ready and checkpoint_eidx < 100:
+        while not en_is_ready and checkpoint_eidx < 30:
             for bidx in range(self.recycle_thresh):
                 actuals, targets, trial_end, stims, brain_data = self.saved_data[bidx]
 
@@ -195,8 +195,10 @@ class CPN_EN_CoProc(experiment.CoProc):
 
                 checkpoint_eidx += 1
 
+            if en_is_ready:
+                break
+
         self.saved_data = []
-        self.en_epoch.reset_period()
         return en_is_ready
 
     def finish(self, loss_history):

### 2021-11-13 experiment/experiment.py

@@ -18,8 +18,6 @@ from . import mRNN
 from . import utils
 
 
-# TODO: score(loss_history)
-
 g_logger = logging.getLogger("experiment")
 
 
@@ -41,13 +39,13 @@ def get_coadapt_config(cuda=None, **kwargs):
     return get_config(cuda=cuda, coadapt=True, **kwargs)
 
 
-def get_m1_lesion_config(cuda=None, **kwargs):
+def get_m1_lesion_config(cuda=None, coadapt=True, **kwargs):
     lesion_type = lesion.LesionType.outputs
     # Lesion the neurons with idxs 50-100, which are in M1
     lesion_args = (50, 100)
     return get_config(
         cuda=cuda,
-        coadapt=True,
+        coadapt=coadapt,
         drop_m1=True,
         lesion_type=lesion_type,
         lesion_args=lesion_args,
@@ -56,13 +54,14 @@ def get_m1_lesion_config(cuda=None, **kwargs):
         **kwargs,
     )
 
-def get_aip_lesion_config(cuda=None, **kwargs):
+def get_aip_lesion_config(cuda=None, coadapt=True, **kwargs):
     lesion_type = lesion.LesionType.outputs
     # Lesion the neurons with idxs 200-250, which are in AIP
     lesion_args = (200, 250)
+
     return get_config(
         cuda=cuda,
-        coadapt=True,
+        coadapt=coadapt,
         drop_m1=True,
         lesion_type=lesion_type,
         lesion_args=lesion_args,

### 2021-11-13 paper/article.tex
Paper progress

@@
 waiting period, but performance suffers drastically.
+
+Clearly, training and deploying a neural co-processor in a true brain involves challenges we
+don't capture with this simulation. We don't intend this simulation to constitute evidence
+that our method will, for example, immediately translate to true restoration of fine-grained
+control for individual fingers of a stroke patient. One clear challenge with doing so would be the
+sheer dimensionality of natural brain dynamics, and how that compares to the resolution
+of stimulation the co-processor can learn to apply. There exists a clear mismatch between
+the dimensionality of the underlying problem, electrode technologies, and the amount of training
+data which can reasonably be collected to train a closed-loop neural controller. We argue
+that while this fact clearly creates a challenge for learning-based closed-loop stimulation,
+that nevertheless the insights we are generating through this simulation are likely to be
+applicable. As we show next, the simulation exhibits a number of properties that reflect
+a real world application, which our co-processor design must contend with. In the Discussion
+section below, we will explore how those insights are likely to lead to initial demonstrations
+of the co-processor approach on lower dimensional problems.
 
@@
 Having defined the training epochs for the EN and CPN, we can define a training
-algorithm on the basis of those two. We train the two in alteration, creating
+algorithm on the basis of those two. We train the two in alternation, creating
 a new EN each time we enter an EN training period. We explored the possibility
@@
 out-of-date, and therefore requiring replacement.
+
+TODO: talk about retraining more
 
@@
 	\item Spectrum from simple low dimensional stimulation vectors today to higher dimensional future
-	\item Toward in vivo application
+	\item Toward in vivo application: example of lower dimensional problems this may help with
 \end{itemize}

### 2021-11-14 paper/article.tex
Paper progress

@@
 greater over longer time scales as the effects of plasticity and ageing change
-the connectivity of the brain.
+the connectivity of the brain. Closed-loop stimulation may also provide means
+to better regulate the energy use of an implanted stimulator, allowing it to
+intelligently regulate when to apply stimulation, and to balance implant
+battery life and stimulation effect. Closed-loop stimulation also offers an
+opportunity minimize side-effects of stimulation, through real time regulation
+of the stimulation parameters, such as in the use of deep brain stimulation in
+Parkinson's patients.
 
@@
 
-Note that the EN effectively constitutes a stimulation model. Its design is
-somewhat motivated by the common linear time-invariant state space model of
+Note that the EN effectively constitutes a stimulation model. We initially attempted
+an EN architecture resembling the common linear time-invariant state space model of
 stimulation, as in e.g. Yang et al. \cite{shanechi.stimmodel}, which is a helpfully
@@
 number of techniques developed for it, lends itself to useful interpretations and
-analyses of those neural dynamics.
+analyses of those neural dynamics. We found, however, that the linear approach,
+as well as the related vanilla RNN approach with nonlinearities added, was not
+sufficiently expressive of long-term dependencies, such that the co-processor
+was capable of learning well.
 
-There are some key differences between that approach and the EN architecture we
-present here. Our EN's purpose is simply to train the CPN, so we pick an
-architecture we find to be best suited to that job. First, our EN architecture
-is not linear, through the use of a $tanh$ activation function. That allows us
-to escape some of the limits in representational power of linear models, though
-we didn't deeply test a strictly linear approach.
-Second, LSTM cells are designed to better capture the relevant asepcts of
-long-running network dynamics, and in our case we found that they were
-crucial for that reason: when using vanilla RNN cells, our co-processor
-wasn't able to train well. Compare that to the simpler cells commonly
-used in vanilla RNNs, which yield functionally the same approach as the
-linear state space model, less the use of a nonlinear activation function.
+Separately, nonlinear LSTM models have also been applied to modeling natural
+neural networks, such as use in predicting local field potentials
+\cite{kim.lstm}, and for stimulation modeling \cite{guclu.lstm}.
+LSTM cells are designed to better capture the relevant aspects of
+long-running network dynamics, and in our case we found that they were crucial
+for that reason. As a result, both our EN and CPN models leverage LSTM cells.
 
@@
 based on observations of the effect of stimulation on the simulated network's output.
-The loss function when training the EN is the MSE loss between the actual and
-predicted output of the mRNN (i.e. muscle velocities). We then use the EN
+The loss function when training the EN is the mean squared error (MSE) loss between the
+actual and predicted output of the mRNN (i.e. muscle velocities). We then use the EN
 to train the CPN during the CPN training epochs. During that time, we backpropagate
@@
 produced by the CPN. In addition to that, backpropagating through it must yield
-gradients which train the CPN to output more useful stimulus. We discovered that
+gradients which train the CPN to output more useful stimulation. We discovered that
 this latter property does not necessarily occur simply by virtue of the former.
@@
 the CPN exhibits unstable training. We have thus far been unable to discover a way
-to measure and optimize for this second property, though pursuing it may be a
+to measure and optimize for this second property directly, though pursuing it may be a
 fruitful subject of future applied ML research. Instead of attempting to train an EN
-directly to exhibit this property, we discovered a training regime that reliably makes
-it occur.
+directly to exhibit this property, we discovered a training regime that indirectly,
+but reliably, makes it occur.
 
@@
 the crux of our EN training problem is one of over-fitting: that the EN may be trained
-to great predictive power on a set of stimulation examples, but that the effective
+to high predictive power on a set of stimulation examples, but that the effective
 dimensionality of the problem it is learning is comparatively quite high, due to both
 the dimensionality of our stimulation parameters $\theta$, and the dynamics of the
-network being stimulated. As a result, the hypersurface which it has fit to the training
+network being stimulated. As a result, the hypersurface which it fits to the training
 examples may exhibit gradients not reflective of reality.
 
-First, and most significantly, we structure the training data set for each EN training epoch
+First, and most significantly, we structure the training dataset for each EN training epoch
 in a very specific way. Each epoch, we include:
@@
 	\item examples using stimulation from the current CPN
-	\item a small (100s) collection of CPNs which are copies of the current CPN, with
-	      Gaussian-distributed mean-0 noise added to their parameters
+	\item examples from a small (100s) collection of CPNs which are copies of the
+	      current CPN, with Gaussian-distributed mean-0 noise added to their parameters
 	\item examples of white noise stimulation
@@
 
-Second, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
-Additionally, we use a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
-one of the standard mitigations for overfit. To converge, the learning rate schedule also needs
+Addiontally, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
+With that, we use a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
+one of the standard mitigations for overfit. To converge, the learning rate schedule needs
 to be chosen carefully: a relatively high learning rate at first, dropping to a very low rate soon
@@
 \subsubsection{CPN training}
-Once we discovered an effective way to train an EN, CPN training was straightforward. For this phase,
+Once an EN is properly trainedx, CPN training is straightforward. For this phase,
 we generate training examples using the current CPN alone. The EN generates predictions of the
@@
 and in-fact usually took longer, despite significant experimentation with learning
-rates.
+rates. It is unclear if that fact is an artifact of our simulation, or of the
+problem more generally, and so it needs more examination in the future, for the
+purpose of training efficiency.
 
-The key challenge we face is determining when an EN is no longer suitable for training the
+The key remaining challenge is determining when an EN is no longer suitable for training the
 current CPN. Expiring an EN at the right time accomplishes two things. First, it ensures
@@
 out-of-date, and therefore requiring replacement.
-
-TODO: talk about retraining more
 
@@
 
-Unfortunately, because we expire the EN rather than retraining it, and because we need
-to expire it so often, training is less efficient than it otherwise could be. We explore
-this some in Section \ref{sec:results}. In the next subsection, we illustrate one way we
-mitigate training inefficiency.
-
-TODO: pseudocode
-
-\subsubsection{Optimization: data reuse}
-One obvious
-
-\section{Results}
-\label{sec:results}
+\subsubsection{Improving learning efficiency with data reuse}
+TODO. May drop this section if I can't get learning stabilized with this approach.
 
@@
 
-Using the simulation described above, we perform three experiments to
-demonstrate the co-processor's ability to learn:
-\begin{itemize}
-	\item 50\% AIP loss
-	\item 100\% connectivity lesion between the F5 and M1 modules
-	\item 100\% connectivity lesion between the F5 and M1 modules without brain co-adpatation
-	\item 50\% M1 loss
-\end{itemize}
+Using the simulation described above, we perform several experiments to
+demonstrate the co-processor's ability to learn. In each experiment, we apply one
+of the three lesion types described above. We additionally vary whether we simulate
+brain co-adaptation, in order to better understand what the co-processor is learning.
+Finally, we perform one experiment where we attempt to simulate natural recovery from
+the lesion, prior to training the co-processor. In sum, our experiments are:
 
-Additionally, we perform each of the ...
+\begin{table}[]
+\begin{tabular}{|l|l|c|c|}
+\hline
+\textbf{Index} & \textbf{Lesion}                                           & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ co-adaptation?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ pre-recovery?\end{tabular}}} \\ \hline
+1              & 50\% AIP loss                                             &                                                                                                 &                                                                                                \\ \hline
+2              & 50\% AIP loss                                             & X                                                                                               &                                                                                                \\ \hline
+3              & 50\% M1 loss                                              &                                                                                                 &                                                                                                \\ \hline
+4              & 50\% M1 loss                                              & X                                                                                               &                                                                                                \\ \hline
+5              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 &                                                                                                 &                                                                                                \\ \hline
+6              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 & X                                                                                               &                                                                                                \\ \hline
+7              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 & X                                                                                               & X                                                                                              \\ \hline
+\end{tabular}
+\end{table}
 
-TODO: train vs test
+In each experiment, we use a single mRNN model, sourced directly from the results
+cited in Michaels \cite{michaels.mrnn}. We drive the experiment using the same set of
+input and output data that the mRNN was trained on. The dataset contains a total of
+502 trials, spread uniformally among 42 object classes. In each experiment, we separate
+a random sample of 20\% of the dataset for validation.
 
-Summary of tunables:
-\begin{itemize}
-	\item Learning rate schedules
-	\item Activation functions
-	\item Obs function dimensionality
-	\item Stim dimensionality
-	\item Dimensionality of EN, CPN
-\end{itemize}
+\section{Results}
+\label{sec:results}
 
@@
 
+TODO: summary of conclusions that will go here:
 \begin{itemize}
-	\item Task performance improves drastically.
+	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
 		\begin{itemize}
-			\item Fig: normalized task loss vs CPN training epoch
-			\item Fig: normalized task loss vs training epoch
-			\item Fig: normalized task loss vs training epoch
-			\item Split by experiment
+			\item 90\% on connection lesion, both with and without co-adaptation
+			\item M1 50\%: 81\% recovery on first attempt with co-adaptation
+			\item AIP 50\%: 35\% recovery w/co-adaptation. Presumably, this is because we've lost encoding functionality, and co-adaptation is very slow. If we ran the simulation for 1 month, who knows?
+			\item Waiting on all other results.
 		\end{itemize}
-	\item Fine tuning takes a long time
-	\item Object classes separate. Fig: sep vs experiment vs epoch
+	\item Initially, performance improves orders of magnitude more quickly than later.
+50-75\% recovery during this phase.
+This appears to be the co-processor learning the task structure (e.g.
+not to move the hand until after the hold signal is lifted).
+	\item Fine tuning takes a far longer time
+	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
 	\item Can co-adapt with the brain, as it changes
-	\item Training efficiency analysis. Show num epochs with and without recycling.
+	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
+	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
+	\item Show pre-recovery results; co-processor will improve beyond what recovery can achieve. (will run results this week)
 \end{itemize}
@@
 \begin{itemize}
-	\item Training efficiency
-	\item Spectrum from simple low dimensional stimulation vectors today to higher dimensional future
+	\item Co-processors can be trained, and they can adapt to a difficult problem
+	\item One central issue that still needs work is data/training efficiency
+	\item Pitch a couple ideas to help solve it:
+		\begin{itemize}
+			\item Continuous retraining of EN, rather than throwing it away
+			\item Retain recent data and reuse it during training
+		\end{itemize}
 	\item Toward in vivo application: example of lower dimensional problems this may help with
@@
 
+\subsection{Learning efficiency}
+Learning efficiency is a key concern for our design. Due to limits on patient fatigue,
+limited time, implant battery life, and other concerns, it is not plausible to expect
+a learning algorithm to have access to unlimited amounts of training data. As a result,
+it is necessary to make efficient use of the data we can acquire.
+
+One clear advantage of our co-processor approach is that, as in Bernal \cite{bernal.sim},
+we can train our stimulation agent to a great degree
+
+\subsection{Future application}
+(on a lower dimensional problem presumably)
+
 \section{Acknowledgements}
-Ganguly, Priya, Anca, Justin, Luciano
+TODO: funding sources
+
+The author would like to thank the following for discussions and insights related to
+the present topic:
+\begin{itemize}
+	\item Karunesh Ganguly, UC San Francisco
+	\item Priya Khanna, UC San Francisco
+	\item Anca Dragan, UC Berkeley
+	\item Justin Ong, University of Washington
+	\item Luciano De La Iglesia, University of Washington
+\end{itemize}
 
 \section{Ethical Statement}
-asdf
+TODO.
 

### 2021-11-14 cpn.py

@@ -147,7 +147,7 @@ class CPN_EN_CoProc(experiment.CoProc):
         done = False
         checkpoint_eidx = 0
         en_is_ready = False
-        while not en_is_ready and checkpoint_eidx < 30:
+        while not en_is_ready and checkpoint_eidx < 10:
             for bidx in range(self.recycle_thresh):
                 actuals, targets, trial_end, stims, brain_data = self.saved_data[bidx]
 

### 2021-11-14 paper/article.tex
Paper progress

@@
 grasping circuit, presented previously by Michaels et al. \cite{michaels.mrnn}. We
-explore:
-\begin{itemize}
-	\item The properties of the artificial neural networks (ANNs) that are needed to
+explore the properties of the artificial neural networks (ANNs) that are needed to
 successfully adapt to the long-running dynamics of a stimulated neural circuit,
-	      as well as to adapt to that circuit's ongoing changes.
-	\item Training methods for our stimulation model, to make it effective in later
-	      training our stimulation agent.
-	\item Data reuse methods, to improve learning efficiency, to better ensure we
-	      can train them with a biologically-realistic amount of data.
-\end{itemize}
+as well as to adapt to that circuit's ongoing connectivity changes.
+We additionally present a training method our co-processor, allowing it to learn
+an optimal stimulation pattern that drives improvements in external task performance,
+while also adapting to the non-stationarity of the brain.
 
@@
 
-Note that the EN effectively constitutes a stimulation model. We initially attempted
-an EN architecture resembling the common linear time-invariant state space model of
-stimulation, as in e.g. Yang et al. \cite{shanechi.stimmodel}, which is a helpfully
-simple (linear) approach. That approach seeks to model neural network dynamics in
-terms of a stationary linear model, which, due in part to its simplicity and the
-number of techniques developed for it, lends itself to useful interpretations and
-analyses of those neural dynamics. We found, however, that the linear approach,
-as well as the related vanilla RNN approach with nonlinearities added, was not
-sufficiently expressive of long-term dependencies, such that the co-processor
-was capable of learning well.
+Note that the EN effectively constitutes a stimulation model. However, in the case that
+a co-processor attempts to optimize for task performance, as opposed to neural
+activity directly, the stimulation model predicts the effect of stimulation and neural
+dynamics on task performance, rather than modeling and predicting the neural response
+to stimulation alone. This provides the key functionality needed to train the CPN.
+
+We initially attempted an EN architecture using vanilla RNN cells, with a nonlineaerity.
+This approach resembles the common linear time-invariant state space model of
+stimulation, as in e.g. Yang et al. \cite{shanechi.stimmodel}, though it additionally
+incorporates a nonlinearity. That approach has the benefit of being simple, and familiar.
+We found, however, that the linear approach, as well as the related vanilla RNN
+approach with nonlinearities added, was not sufficiently expressive of long-term
+dependencies, such that the co-processor was capable of learning well.
 
@@
 LSTM cells are designed to better capture the relevant aspects of
-long-running network dynamics, and in our case we found that they were crucial
-for that reason. As a result, both our EN and CPN models leverage LSTM cells.
+long-running network dynamics. We found that LSTM cells were crucial for
+enabling our co-processor to learn the long-running dynamics of the network
+it was stimulating. As a result, both our EN and CPN models leverage LSTM cells.
 
@@
 of stimulation the co-processor can learn to apply. There exists a clear mismatch between
-the dimensionality of the underlying problem, electrode technologies, and the amount of training
+the dimensionality of the underlying problem, sensor and stimulator technologies, and the amount of training
 data which can reasonably be collected to train a closed-loop neural controller. We argue
@@
 that nevertheless the insights we are generating through this simulation are likely to be
-applicable. As we show next, the simulation exhibits a number of properties that reflect
-a real world application, which our co-processor design must contend with. In the Discussion
-section below, we will explore how those insights are likely to lead to initial demonstrations
-of the co-processor approach on lower dimensional problems.
+applicable. In the discussion below, we will explore how those insights are likely to lead
+to initial demonstrations of the co-processor approach on lower dimensional problems.
+As we show next, the simulation exhibits a number of properties that reflect
+a real world application, which our co-processor design must contend with.
 
@@
 \subsubsection{Simulating recovery prior to co-processor use}
-We sought to additonally pre-train the mRNN after applying the lesion in
+We sought to additonally re-train the mRNN after applying the lesion in
 order to simulate stroke recovery. For simulated lesions which zero the outputs
@@
 degrees of freedom that a nearly full recovery can occur, unless so many
-neurons are lesioned that no stimulation could be effective. We go into
-this further in the Discussion section below.
+neurons are lesioned that no stimulation could be effective.
 
@@
 
-Addiontally, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
+Addtionaly, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
 With that, we use a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion}
+\label{sec:discussion}
+
+We present here a first-of-its-kind demonstration of a novel design for neural co-processors,
+allowing for an AI agent to learn stimulation policies that improve a user's task performance.
+The central notion of the design is to leverage a stimulation model, whose purpose is to provide
+training feedback to an AI agent responsible for generating the stimulation. We present here
+a first demonstration of the learning algorithms and key properties
+
+We proposed X
+It did Y
+It was found that Z
+It has potential to A
+
 \begin{itemize}
@@
 
+\section{Conclusion}
+...
 \subsection{Future application}
@@
 
-\section{Ethical Statement}
-TODO.
-
 \section{References}

### 2021-11-15 experiment/experiment.py
Fix coadapt :(

@@ -212,7 +212,7 @@ class Experiment:
 
         if cfg.recover_after_lesion:
             model_path = michaels_load.get_path(fname_override="recovered.model")
-            self.mike.load_state_dict(torch.load(model_path))
+            self.mike.load_weights_from_file(model_path)
             self.opt_mike = AdamW(self.mike.parameters(), lr=1e-7)
 
     @property
@@ -368,7 +368,10 @@ class Experiment:
             if self.cfg.coadapt and (self.loss_history.eidx % 5) == 0:
                 loss = torch.nn.MSELoss()(actuals, dout[:, 1:, :])
                 loss.backward(inputs=list(self.mike.parameters()))
-                self.mike.set_coadap_grads()
+                if self.cfg.lesion_instance.application == "connection":
+                    self.mike.set_connection_coadap_grads()
+                else:
+                    self.mike.set_end_to_end_coadap_grads()
                 self.opt_mike.step()
 
             self.loss_history.report_user_data(user_data)

### 2021-11-15 experiment/mRNN.py

@@ -83,7 +83,6 @@ class MichaelsRNN(nn.Module):
 
             self.J_zero_grad_mask = None
             self.I_zero_grad_mask = None
-            self.recalc_masks_from_weights()
         else:
             # Masks ---------------------------
             # (num_neurons, num_neurons)
@@ -188,6 +187,7 @@ class MichaelsRNN(nn.Module):
                 self.x0 = nn.Parameter(torch.empty(self.num_neurons))
                 self.fc = nn.Linear(num_neurons_per_module, output_dim)
 
+        self.recalc_masks_from_weights()
         self.lesion = lesion
         self.stimulus = stimulus
 
@@ -209,18 +209,32 @@ class MichaelsRNN(nn.Module):
         I_zero_grad_mask[npm:, :] = 1.0
         self.I_zero_grad_mask = I_zero_grad_mask
 
-        self.I_coadap_zero_grad_mask = torch.zeros(self.num_neurons, self.num_input_features)
-        self.I_coadap_zero_grad_mask[:npm, :] = 1.0
+        # The hold signal response can change, but not input response
+        self.I_connection_coadap_zero_grad_mask = torch.zeros(self.num_neurons, self.num_input_features)
 
+        # Only M1 internal mappings
+        self.J_connection_coadap_zero_grad_mask = torch.zeros(self.num_neurons, self.num_neurons)
+        self.J_connection_coadap_zero_grad_mask[:npm, :npm] = 1.0
 
-        self.J_coadap_zero_grad_mask = torch.zeros(self.num_neurons, self.num_neurons)
-        self.J_coadap_zero_grad_mask[:npm, :npm] = 1.0
+        # Only M1
+        self.S_connection_coadap_zero_grad_mask = torch.zeros(self.num_neurons, 1)
+        self.S_connection_coadap_zero_grad_mask[:npm] = 1.0
+
+        # Only M1
+        self.B_connection_coadap_zero_grad_mask = torch.zeros(self.num_neurons, 1)
+        self.B_connection_coadap_zero_grad_mask[:npm] = 1.0
 
         if self._cuda is not None:
             self.J_zero_grad_mask = self.J_zero_grad_mask.cuda(self._cuda)
             self.I_zero_grad_mask = self.I_zero_grad_mask.cuda(self._cuda)
-            self.I_coadap_zero_grad_mask = self.I_coadap_zero_grad_mask.cuda(self._cuda)
[…]

### 2021-11-16 cpn.py
Update reuse pattern

@@ -135,7 +135,7 @@ class CPN_EN_CoProc(experiment.CoProc):
 
         return update_task_loss
 
-    def train_en_closed_loop(self, loss_history, next_is_validation):
+    def train_en_closed_loop(self, loss_history, user_data, next_is_validation):
         # Unpack, aka form a single 'actuals', 'targets', 'trial_end',
         #  and pack a list of concatenated stims.
 
@@ -143,6 +143,17 @@ class CPN_EN_CoProc(experiment.CoProc):
         trial_len = len(self.saved_data[0][4])
 
         is_validation = next_is_validation
+
+
+
+        last_pred_loss = 1
+        if user_data is not None:
+            if user_data.pred_loss == user_data.pred_loss:
+                last_pred_loss = user_data.pred_loss
+        if last_pred_loss > 0.005:
+            self.saved_data = []
+            return False
+
 
         done = False
         checkpoint_eidx = 0
@@ -228,7 +239,7 @@ class CPN_EN_CoProc(experiment.CoProc):
                 self.epoch_type = EpochType.EN
 
             if len(self.saved_data) == self.recycle_thresh and not en_is_ready:
-                en_is_ready = self.train_en_closed_loop(loss_history,
+                en_is_ready = self.train_en_closed_loop(loss_history, user_data,
                     next_is_validation)
 
                 if en_is_ready:

### 2021-11-18 paper/article.tex
Paper progress.

@@
 \title[Modeling a Neural Co-Processor]
-{Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept}
+{Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept}
 
@@
 
-An example of such a simulation approach is Bernal et al. \cite{bernal.sim}. In this work, the
+An example of such a simulation approach is Dura-Bernal et al. \cite{bernal.sim}. In this work, the
 authors use a simulated spiking neural network to train a stimulation agent, much
@@
 			\item 90\% on connection lesion, both with and without co-adaptation
-			\item M1 50\%: 81\% recovery on first attempt with co-adaptation
-			\item AIP 50\%: 35\% recovery w/co-adaptation. Presumably, this is because we've lost encoding functionality, and co-adaptation is very slow. If we ran the simulation for 1 month, who knows?
+			\item M1 50\%: 81\% recovery on first attempt with co-adaptation. This is likely to go higher, since I found a bug.
+			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running.
+The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow.
+Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
 			\item Waiting on all other results.
@@
 We present here a first-of-its-kind demonstration of a novel design for neural co-processors,
-allowing for an AI agent to learn stimulation policies that improve a user's task performance.
-The central notion of the design is to leverage a stimulation model, whose purpose is to provide
-training feedback to an AI agent responsible for generating the stimulation. We present here
-a first demonstration of the learning algorithms and key properties
+allowing for an AI agent to learn neural stimulation policies that improve a user's performance
+of an external task. The design revolves around the use of a stimulation model for training the
+agent, providing a proxy to the true function mapping stimulation and neural activity
+to task performance. We base our demonstration on a simulation of a neural circuit engaged in
+an external grasping task.
 
-We proposed X
-It did Y
-It was found that Z
-It has potential to A
+Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
+50-90\% across our various experiments. To work this well, the co-processor needed to adapt to
+the long-running dynamics of the stimulated network, as well as the long-running effects of
+stimulation. In some experiments we required it to additionally adapt to non-stationarity in
+the neural circuit, which was actively changing at the same time the co-processor was learning.
+It achieved that as well, co-adapting with the simulated brain to improve external task
+performance.  TODO: pre-recovery result.
+
+Clearly, we need to examine carefully what can and cannot be inferred from simulation studies
+such as these. As in Dura-Bernal \cite{bernal.sim}, we note that our simulation differs
+drastically from an organic brain, both in scale and structure. Today, predicting long-running
+neural responses to stimulation remains a difficult problem, implying targeted neural control
+is also difficult. Even moreso, identifying the neural correlates of complex task success
+remains largely out of reach \cite{khanna.openloop}. What, then, does our simulation provide us?
+
+Abstractly, the co-processor lends itself to any closed-loop neural stimulation problem
+where a relevant stimulation model can be identified. There are many lower dimensional problems
+today which we address with far simpler stimulation regimes than would be necessary for
+e.g. remediating the symptoms of a stroke. For example, deep brain stimulators for
+treating the symptoms of Parkinson's Disease can now be controlled in a closed loop
+fashion to also minimize stimulation side-effects and maximize implant battery life.
+TODO: citation. Our simulation clearly does not imply that our co-processor design can
+translate immediately to stroke applications. However, it does demonstrate that our design
+can adapt to long-running effects of stimulation, as well as to the non-stationarity of
+the neural circuit to which it is applying the stimulation. These properties together
+suggest it could be useful in lower dimensional problems where a stimulation model can
+be estimated, and stimulation parameters can be tied to the target objective. TODO: more examples?
+
+On the whole, the co-processor quickly improved task performance, but required orders of magnitude
+more training examples to achieve its highest levels of performance. This training inefficiency remains
+a problem...
 
@@
 
-One clear advantage of our co-processor approach is that, as in Bernal \cite{bernal.sim},
+One clear advantage of our co-processor approach is that, as in Dura-Bernal \cite{bernal.sim},
 we can train our stimulation agent to a great degree

### 2021-11-21 experiment/experiment.py
Pre-recov machinery

@@ -1,3 +1,4 @@
+import errno
 import logging
 import os
 import time
@@ -19,6 +20,9 @@ from . import utils
 
 
 g_logger = logging.getLogger("experiment")
+
+# Rel path pointing to a pre-recovered mRNN
+RECOV_MODEL_PATH = "experiment/recovered"
 
 
 @attr.s(auto_attribs=True)
@@ -211,8 +215,17 @@ class Experiment:
             self.obs_drop_module_idx = None
 
         if cfg.recover_after_lesion:
-            model_path = michaels_load.get_path(fname_override="recovered.model")
-            self.mike.load_weights_from_file(model_path)
+            model_path = os.path.join(RECOV_MODEL_PATH, "recovered_mrnn_%s.model" %
+                str(lesion_instance))
+
+            try:
+                self.mike.load_weights_from_file(model_path)
+            except OSError as e:
+                if e.errno = errno.ENOENT:
+                    raise ValueError("No pre-generated recovered mRNN available "
+                        "for a lesion of type %s" % str(cfg.lesion_instance))
+                raise
+
             self.opt_mike = AdamW(self.mike.parameters(), lr=1e-7)
 
     @property

### 2021-11-21 experiment/config.py
Clean up M1

@@ -85,6 +85,7 @@ DEFAULT_OBS_OUT_DIM = 20
 DEFAULT_OBS_SIGMA = 1.75
 DEFAULT_NUM_STIM_CHANNELS = 16
 DEFAULT_STIM_SIGMA = 2.175
+DEFAULT_STIM_PAD_LEFT_NEURONS = 0
 DEFAULT_STIM_PAD_RIGHT_NEURONS = 200
 DEFAULT_OUT_DIM = 50
 DEFAULT_HOLDOUT_PCT = 0.2
@@ -149,6 +150,7 @@ def get(
     num_stim_channels=DEFAULT_NUM_STIM_CHANNELS,
     num_stim_neurons=None,
     stim_sigma=DEFAULT_STIM_SIGMA,
+    stim_pad_left_neurons=DEFAULT_STIM_PAD_LEFT_NEURONS,
     stim_pad_right_neurons=DEFAULT_STIM_PAD_RIGHT_NEURONS,
     stim_retain_grad=False,
     obs_out_dim=DEFAULT_OBS_OUT_DIM,
@@ -174,6 +176,9 @@ def get(
         num_stim_neurons = num_neurons_per_module
 
     if stimulation_type is stim.StimulationType.one_to_one:
+        if pad_left_neurons != 0:
+            raise NotImplementedError()
+
         stimulus = stimulation_type.value(
             num_stim_neurons,
             num_stim_neurons,
@@ -181,6 +186,9 @@ def get(
         )
     elif stimulation_type is stim.StimulationType.gaussian_alpha:
         if cuda:
+            raise NotImplementedError()
+
+        if pad_left_neurons != 0:
             raise NotImplementedError()
 
         # NOTE: can add the num_stim_channels and sigma arg above
@@ -191,12 +199,16 @@ def get(
             sigma=stim_sigma,
         )
     elif stimulation_type is stim.StimulationType.gaussian_exp:
+        assert (num_stim_neurons + stim_pad_left_neurons + stim_pad_right_neurons) == \
+                (3 * num_neurons_per_module)
+
         stimulus = stimulation_type.value(
             num_stim_channels,
             num_stim_neurons,
             batch_size=1,  # Will be reset before use
             sigma=stim_sigma,
             retain_grad=stim_retain_grad,
+            pad_left_neurons=stim_pad_left_neurons,
             pad_right_neurons=stim_pad_right_neurons,
             cuda=cuda,
         )

### 2021-11-21 experiment/experiment.py

@@ -46,7 +46,7 @@ def get_coadapt_config(cuda=None, **kwargs):
 def get_m1_lesion_config(cuda=None, coadapt=True, **kwargs):
     lesion_type = lesion.LesionType.outputs
     # Lesion the neurons with idxs 50-100, which are in M1
-    lesion_args = (50, 100)
+    lesion_args = (0, 50)
     return get_config(
         cuda=cuda,
         coadapt=coadapt,
@@ -54,7 +54,8 @@ def get_m1_lesion_config(cuda=None, coadapt=True, **kwargs):
         lesion_type=lesion_type,
         lesion_args=lesion_args,
         num_stim_neurons=50,
-        stim_pad_right_neurons=250,
+        stim_pad_left_neurons=50,
+        stim_pad_right_neurons=200,
         **kwargs,
     )
 

### 2021-11-21 experiment/stim.py

@@ -9,10 +9,12 @@ from . import utils
 
 class Stimulus(object):
     def __init__(
-        self, num_stim_channels, num_neurons, pad_right_neurons=200, batch_size=1
+        self, num_stim_channels, num_neurons, pad_left_neurons=0,
+        pad_right_neurons=200, batch_size=1
     ):
         self._num_stim_channels = num_stim_channels
         self._num_neurons = num_neurons
+        self._pad_left_neurons = pad_left_neurons
         self._pad_right_neurons = pad_right_neurons
         self._batch_size = batch_size
 
@@ -31,6 +33,10 @@ class Stimulus(object):
     @property
     def pad_right_neurons(self):
         return self._pad_right_neurons
+
+    @property
+    def pad_left_neurons(self):
+        return self._pad_left_neurons
 
     def add(self, params):
         raise NotImplementedError()
@@ -177,6 +183,7 @@ class StimulusGaussianExp(Stimulus):
         self,
         num_stim_channels,
         num_neurons,
+        pad_left_neurons=0,
         pad_right_neurons=200,
         sigma=2.5,
         decay=0.3,
@@ -185,11 +192,13 @@ class StimulusGaussianExp(Stimulus):
         cuda=None,
     ):
         super(StimulusGaussianExp, self).__init__(
-            num_stim_channels, num_neurons, pad_right_neurons, batch_size=batch_size
+            num_stim_channels, num_neurons, pad_left_neurons, pad_right_neurons,
+            batch_size=batch_size
         )
 
         self._sigma = sigma
-        self._vals = torch.zeros((batch_size, num_neurons + pad_right_neurons))
+        self._vals = torch.zeros((batch_size, pad_left_neurons + num_neurons +
+            pad_right_neurons))
         self._decay = decay
         self._norm = norm(0, self._sigma)
         self._retain_grad = retain_grad
@@ -237,7 +246,8 @@ class StimulusGaussianExp(Stimulus):
     def reset(self, batch_size=None):
         super(StimulusGaussianExp, self).reset(batch_size=batch_size)
         self._vals = torch.zeros(
-            (self.batch_size, self.num_neurons + self.pad_right_neurons)
+            (self.batch_size, self.pad_left_neurons + self.num_neurons +
+                self.pad_right_neurons)
         )
 
         if self._retain_grad:
@@ -266,7 +276,9 @@ class StimulusGaussianExp(Stimulus):
[…]

### 2021-11-21 cpn.py
Fix bug in pre-recov experiment.py logic

@@ -58,6 +58,8 @@ class CPN_EN_CoProc(experiment.CoProc):
         if log_dir is not None:
             self.log_dir = os.path.join(log_dir, cfg.cfg_str, str(self.uuid))
             os.makedirs(self.log_dir)
+
+            print("Log dir will be:", self.log_dir)
         else:
             self.log_dir = None
 

### 2021-11-21 experiment/experiment.py

@@ -217,14 +217,15 @@ class Experiment:
 
         if cfg.recover_after_lesion:
             model_path = os.path.join(RECOV_MODEL_PATH, "recovered_mrnn_%s.model" %
-                str(lesion_instance))
+                str(cfg.lesion_instance))
 
             try:
                 self.mike.load_weights_from_file(model_path)
             except OSError as e:
-                if e.errno = errno.ENOENT:
+                if e.errno == errno.ENOENT:
                     raise ValueError("No pre-generated recovered mRNN available "
-                        "for a lesion of type %s" % str(cfg.lesion_instance))
+                        "for a lesion of type %s; looked for path %s" % (
+                        str(cfg.lesion_instance), model_path))
                 raise
 
             self.opt_mike = AdamW(self.mike.parameters(), lr=1e-7)

### 2021-11-21 paper/article.tex
Paper progress

@@
 
-\section{Discussion}
+\section{Discussion and Future Work}
 \label{sec:discussion}
@@
 
-Clearly, we need to examine carefully what can and cannot be inferred from simulation studies
+Clearly, we need to discuss carefully what can and cannot be inferred from simulation studies
 such as these. As in Dura-Bernal \cite{bernal.sim}, we note that our simulation differs
@@
 treating the symptoms of Parkinson's Disease can now be controlled in a closed loop
-fashion to also minimize stimulation side-effects and maximize implant battery life.
-TODO: citation. Our simulation clearly does not imply that our co-processor design can
-translate immediately to stroke applications. However, it does demonstrate that our design
-can adapt to long-running effects of stimulation, as well as to the non-stationarity of
-the neural circuit to which it is applying the stimulation. These properties together
-suggest it could be useful in lower dimensional problems where a stimulation model can
-be estimated, and stimulation parameters can be tied to the target objective. TODO: more examples?
+fashion. That allows them to minimize stimulation side-effects and maximize implant
+battery life, while also reducing the patient's primary symptoms. TODO: citation. Our
+simulation does not imply that our co-processor design can translate immediately to
+stroke applications. However, it does demonstrate that our design can adapt to
+long-running effects of stimulation, as well as to the non-stationarity of the neural
+circuit to which it is applying the stimulation. These properties together suggest it may be
+useful in lower dimensional problems where a stimulation model can be estimated, and
+stimulation parameters can be tied to the target objective. TODO: more examples?
 
-On the whole, the co-processor quickly improved task performance, but required orders of magnitude
-more training examples to achieve its highest levels of performance. This training inefficiency remains
-a problem...
+On the whole, the co-processor quickly improved task performance, but required orders of
+magnitude more training examples to achieve its highest levels of performance. Even
+moderate amounts of recovery may be valuable to a user, but nevertheless we consider
+training efficiency to remain a problem. Due to limits on patient fatigue, time, implant
+battery life, and other concerns, it is not plausible to expect a learning algorithm
+to have access to unlimited amounts of training data. As a result, it is necessary to
+make efficient use of the data we can acquire. We believe there exist at least three
+mitigations for that:
 
 \begin{itemize}
-	\item Co-processors can be trained, and they can adapt to a difficult problem
-	\item One central issue that still needs work is data/training efficiency
-	\item Pitch a couple ideas to help solve it:
-		\begin{itemize}
-			\item Continuous retraining of EN, rather than throwing it away
-			\item Retain recent data and reuse it during training
-		\end{itemize}
-	\item Toward in vivo application: example of lower dimensional problems this may help with
+	\item Retraining an existing EN, rather than regularly creating a new one. As
+	      mentioned above, we encountered difficulty with this approach, but it remains
+	      to be seen if this is a fundamental problem with our training method, rather
+	      than a peculiarity of our simulation. Also, it remains unclear if a solution
+	      for our simulation exists as well, which we've simply yet to identify. This
+	      remains a future area of inquiry.
+	\item Making better use of what data we acquire. In this initial simulation we do
+	      not retain data beyond the present training epoch. In practice, we can likely
+	      retain training data for some time window. Non-stationarity requires us to
+	      regularly discard or discount data as it ages. However, in practice, training
+	      epochs operate on the order of seconds, suggesting that data can be retained
+	      and reused for a period of time. The `speed' of non-stationarity compared to
+	      the co-processor's learning therefore must be an area of future research.
+	\item Matching the dimensionality of the stimulation parameters and observation
+	      data to the amount of available data. If one chooses a stimulation regime with
+	      a small number of controllable parameters, where nevertheless task improvement
+	      from the stimulation is possible, that allows for a simpler co-processor, which
+	      may require less training data. We discuss this next.
 \end{itemize}
 
-\subsection{Learning efficiency}
-Learning efficiency is a key concern for our design. Due to limits on patient fatigue,
-limited time, implant battery life, and other concerns, it is not plausible to expect
-a learning algorithm to have access to unlimited amounts of training data. As a result,
-it is necessary to make efficient use of the data we can acquire.
-
-One clear advantage of our co-processor approach is that, as in Dura-Bernal \cite{bernal.sim},
-we can train our stimulation agent to a great degree
-
-\section{Conclusion}
-...
-\subsection{Future application}
-(on a lower dimensional problem presumably)
+TODO: future applications
 

### 2021-11-22 paper/article.tex
Paper progress

@@
 (a) specific regimes of neural activity, or (b) external task performance.
-For example, it may seek to learn to stimulate the motor cortex of a stroke patient,
+For example, one may seek to learn to stimulate the motor cortex of a stroke patient,
 conditioning the stimulation on upstream visual information, aiding the patient's
@@
 BCIs (BBCIs) allow stimulation to be conditioned by decoded brain activity as well as external
-sensor data (e.g. camera), which can allow for the application of real-time, fine-grained control of
+sensor (e.g. camera) data, which can allow for the application of real-time, fine-grained control of
 neural circuits and prosthetic devices, e.g. Nicolelis et al. \cite{nicolelis.bmbi}. These may lead,
@@
 enough to optimize not only for particular neural activities, but also for tasks external to the
-subject. For example, they may be able to aid a stroke victim by finding a stimulation pattern of
+subject. For example, they may be able to aid a stroke patient by finding a stimulation pattern of
 the motor cortex which helps restore lost limb function. Likewise, the framework generalizes enough
-to condition stimulation on both brain activity, and external sensors, e.g. cameras or LIDAR, in
-order to incorporate feedback for realtime control.
+to condition stimulation on both brain activity, and external sensors, e.g. cameras or light detection
+and ranging (LIDAR) sensors, in order to incorporate feedback for realtime control.
 
@@
 the circuit, i.e. brain, by updating its stimulation regime, while at the same time the
-brain is updating its response to the stimulation, and changing due to natural plasticity,
-aging, etc. This allows the co-processor to continually optimize for the intended cost
+brain is updating its response to the stimulation, and changing due to natural plasticity
+and aging. This allows the co-processor to continually optimize for the intended cost
 function, despite the signficant non-stationarity of the target circuit.
@@
 Here we provide a proof-of-concept in simulation for a co-processor that restores
-movement to a limb, after a subject has suffered a stroke affecting its ability to
+movement to a limb, after a subject has suffered a stroke affecting their ability to
 use that limb. It combines:
@@
 
+\section{Background}
 Significant advances have been made in modeling the effects of electrical stimulation
@@
 treating a variety of disorders. Open loop stimulation has been effective in
-treating Parkinson's Disease \cite{benabid.parkinsons}, as well as various
+treating Parkinson's Disease (PD) \cite{benabid.parkinsons}, as well as various
 psychiatric disorders \cite{holtzheimer.psy, kisely.psy, fraint.psy}.
@@
 occurs affecting a primate's motor cortex. The authors demonstrate that
-the use of low-frequency alternating current, applied epidurally and set to
-certain phases, can improve grasp performance.
+the use of low-frequency alternating current, applied epidurally,
+can improve grasp performance.
 
@@
 to better regulate the energy use of an implanted stimulator, allowing it to
-intelligently regulate when to apply stimulation, and to balance implant
-battery life and stimulation effect. Closed-loop stimulation also offers an
-opportunity minimize side-effects of stimulation, through real time regulation
-of the stimulation parameters, such as in the use of deep brain stimulation in
-Parkinson's patients.
+intelligently regulate when to apply stimulation, in order to presever implant
+battery life. Closed-loop stimulation also offers an opportunity minimize
+side-effects of stimulation, through real time regulation
+of the stimulation parameters, such as in the use of deep brain stimulation (DBS) in
+PD patients \cite{little.park}.
 
 Closed-loop stimulation conditions stimulation on observations of brain activity,
-possibly allowing it to shape the neural activity more precisely, and to adapt to changes
-in the circuit over time. This opens the door to real-time, targeted control of the neural
-circuit. It has been used to aid in learning new memories after some impairment
-\cite{berger.closedloop, kahana.biomarker}, to replay visually-invoked
-activations \cite{tafazoli.acls}, and for optogenetic control of
-a thalamocortical circuit \cite{bolus.opto}, among others.
+allowing it to shape the neural activity more precisely, and in response to changes
+in a patient's clinical condition. This may also allow an implant to adapt
+in real time to changes in the target circuit. It has been used to aid in learning
+new memories after some impairment \cite{berger.closedloop, kahana.biomarker},
+to replay visually-invoked activations \cite{tafazoli.acls}, and for optogenetic
+control of a thalamocortical circuit \cite{bolus.opto}, among others.
 
@@
 fundamental challenges in using neural stimulation to improve external task performance.
-First, neural networks exhibit long-running and nonlinear dependencies, necessitating
+First, neural networks exhibit long-running and nonlinear dynamics, necessitating
 the need for a stimulation agent to account for far-distant effects of the stimulation it
@@
 	\item A stimulation and neural dynamics model, known as an ``Emulator Network'' (EN). It models the relationship
-	      between the stimulation, neural dynamics, and external task. It's purpose is for training the second network.
+	      between the stimulation, neural dynamics, and external task. Its purpose is for training the second network.
 	\item A stimulation agent, known as the ``co-processor network'' (CPN), which maps neural activity, and possibly
@@
 
-We initially attempted an EN architecture using vanilla RNN cells, with a nonlineaerity.
+We initially attempted an EN architecture using vanilla RNN cells, with a nonlinearity.
 This approach resembles the common linear time-invariant state space model of
@@
 enabling our co-processor to learn the long-running dynamics of the network
-it was stimulating. As a result, both our EN and CPN models leverage LSTM cells.
+it was stimulating, and we use them as a result.
 
@@
 cheaply iterate on our design, prior to any \textit{in vivo} experiments.
-In order for the simulation to be admissible, we need to ensure it has some
+For the simulation to be admissible, we need to ensure it has some
 properties that allow it to strongly indicate if our design is improving in
@@
 design may be adapting to the pecularities of the simulation, without
-becoming more useful for the real world.
+becoming more useful.
 
 An example of such a simulation approach is Dura-Bernal et al. \cite{bernal.sim}. In this work, the
-authors use a simulated spiking neural network to train a stimulation agent, much
-like the work we present here. Their stimulation agent sought to restore the network's
-control of a simulated arm, to reach a target, after a simulated lesion was applied.
-As they point out, we have a limited ability to probe a neural circuit \textit{in vivo}
-in order to perform learning. As a result, we first need to design our approach through
-the use of an admissible simulation. The authors in this case simulated lesions by
-effectively removing parts of their simulated network, or by cutting connections between
-parts of the network. Likewise here, we use an established design for an artificial
-neural circuit which performs a grasping task, whose architecture and training methods
-were designed specifically to result in naturalistic dynamics.
+authors use a simulated spiking neural network to train a stimulation agent.
+Their stimulation agent sought to restore the network's
+control of a simulated arm, to reach a target, after a simulated lesion was applied,
+much like the work here.  As the authors point out, we have a limited ability to
+probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
+we first need to design our approach through the use of an admissible simulation.
+The authors in this case simulated lesions by effectively removing parts of their
+simulated network, or by cutting connections between parts of the network.
+Here, we use an established design for an artificial neural circuit which
+performs a grasping task, whose architecture and training methods were designed
+specifically to result in naturalistic dynamics.
 
@@
 were supplied to us by the lead author Jonathan Michaels, for the purpose of our present
-simulation. We re-implemented his models' logic in PyTorch, and loaded his trained parameters
+simulation. We re-implemented his model's logic in PyTorch, and loaded his trained parameters
 into it, for one of his subjects, arbitrarily chosen. See Fig. \ref{fig:michaels}.
@@
 Clearly, training and deploying a neural co-processor in a true brain involves challenges we
-don't capture with this simulation. We don't intend this simulation to constitute evidence
+don't capture with this simulation. This simulation does not constitute evidence
 that our method will, for example, immediately translate to true restoration of fine-grained
-control for individual fingers of a stroke patient. One clear challenge with doing so would be the
-sheer dimensionality of natural brain dynamics, and how that compares to the resolution
+control for individual fingers of a stroke patient. One clear challenge with doing so would
+be the sheer dimensionality of natural brain dynamics, and how that compares to the resolution
 of stimulation the co-processor can learn to apply. There exists a clear mismatch between
-the dimensionality of the underlying problem, sensor and stimulator technologies, and the amount of training
-data which can reasonably be collected to train a closed-loop neural controller. We argue
-that while this fact clearly creates a challenge for learning-based closed-loop stimulation,
-that nevertheless the insights we are generating through this simulation are likely to be
-applicable. In the discussion below, we will explore how those insights are likely to lead
-to initial demonstrations of the co-processor approach on lower dimensional problems.
-As we show next, the simulation exhibits a number of properties that reflect
-a real world application, which our co-processor design must contend with.
+the dimensionality of the underlying problem, sensor and stimulator technologies, and the
+amount of training data which can reasonably be collected to train a closed-loop neural
+controller. We argue that while this fact clearly creates a challenge for learning-based
+closed-loop stimulation, that nevertheless the insights we generate through this
+simulation are likely to be applicable. In the Discussion below, we will explore how
+those insights are likely to lead to initial demonstrations of the co-processor approach
+on lower dimensional problems. As we show next, the simulation exhibits a number of
+properties that reflect a real world application, which our co-processor design must
+contend with.
 
@@
 Simulating a brain lesion in terms of this artificial network results in error
-modes that resemble a natural lesion of certain regions of a primate brain. For example, if we
+modes that resemble a some natural lesions of a primate brain. For example, if we
 alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
@@
 it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
-simulated lesion closely resembles natural stroke symptoms, though the simulation is not
+simulated lesion closely resembles natural stroke symptoms. However, the simulation is not
 intended to constitute a physical model of a lesion, i.e. to directly explain the
-connection between hand spasticity and the lesion. Conversely, if we
-lesion of a portion of the output module of the Michaels mRNN, roughly
+connection between hand spasticity and the lesion.
+
+Conversely, if we lesion of a portion of the output module of the Michaels mRNN, roughly
 corresponding to M1, we see a more wholesale loss of movement, affecting even the
@@
 
-The co-processor's task in our simulation will be to identify the appropriate information,
+The co-processor's task in our simulation will be identifying the appropriate information,
 read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
-to effectively bridge across the lesion, forward propagating the object shape information,
-indirectly, through the use of stimulation.
+to effectively bridge across the lesion, forward propagating the object shape and task
+structyure information, indirectly, through the use of stimulation.
 
@@
 	      neurons to zero, causing a more holisitic loss of movement. Here, the lesion
-	      may make it impossible for the co-processor to find a solution, unless the
-	      brain co-adapts with it to re-map the M1 neurons onto the muscle space.
+	      may make it impossible for the co-processor to find a solution, but some
+	      recovery may be possible.
 \end{itemize}
@@
 simulated space, and to simulate how that may change the effects of stimulation.
-Instead, our stimulation model is merely attempting to make the co-processors task
-more realistic: it cannot precisely set the hidden state of every neuron,
-instantaneously and in each time step. Any stimulation it applies will diffuse
+Instead, our stimulation model is merely attempting to make the co-processor's task
+more realistic, in the sense that it cannot precisely set the hidden state of every
+neuron, instantaneously and in each time step. Any stimulation it applies will diffuse
 across ``space'' and time.
@@
 it outputs a modifier to the internal states of all neurons in the output module,
-based on the current state of memory in the function object. That memory allows
+based on the current state of memory. That memory allows
 the stimulation function to perform temporal smoothing. Specifically, we used a
@@
 is inappropriately low resolution for our task, for example epidural stimulation.
-We experimented with this value, and found we could, unsurprisingly, set it to a high
-enough or low enough value that the co-processor cannot succeed, though the co-processor
-could succeed over a broad range of values. However, we could not conclude much from
-those experiments, because it isn't clear what those experiments ultimately indicate.
-As a result, we don't have a principled way of choosing an appropriate $\sigma$ value;
-we set it to a moderate, balanced value, and don't vary it.
+However, we cannot conclude much from these experiments, because this simulation does
+not constitute a biophysical model, where we can identify some realistic value. The
+purpose of this smoothing technique is only to make the co-processor's job harder in
+the simple sense that it cannot target neurons individually.
+As a result, we set it to a moderate, balanced value, and don't vary it.
 
@@
 \begin{equation}
-\alpha_{t} = \tau\alpha_{t-1} + \theta_{t}
+\alpha_{t} = \tau\alpha_{t-1} + \theta_{t-1}
 \end{equation}
@@
 To demonstrate the co-processor's ability to adapt to the non-stationarity
-of the brain, we also change it throughout the co-processor's training.
+of the brain, we also change the brain throughout the co-processor's training.
 Specifically, we simulate the brain's co-adaptation with the stimulation
@@
 calculated and backpropagated into the mRNN, allowing it to adapt to
-its inputs, one of which is the stimulation being applied. The learning
-rate for the optimizer is set arbitrarily to a relatively low rate of
-$1e-7$. The learning rate was not chosen in a principled way, since
-we are not aware of a principled way of choosing it. However, we reason
-that it should be lower than the co-processor by orders of magnitude due
-to the rapidity by which an ANN can be retrained.
+the fact it receives stimulation. The learning rate for the optimizer
+is set arbitrarily to a relatively low rate of $1e-7$. The learning
+rate was not chosen in a principled way, since we are not aware of a
+principled way of choosing it. However, we reason that it should be
+lower than the co-processor by orders of magnitude due
+to the rapidity with which an ANN can be retrained.
 
@@
 
-After some experimentation, we composed each batch of EN training data in the following way:
-
-\begin{table}[h]
+After some experimentation, we composed each batch of EN training data in the following way: \\
+
 \begin{tabular}{|l|l|}
@@
 White noise stimulation                                                                                                        & 30\%                           \\ \hline
-\end{tabular}
-\end{table}
-
-Addtionaly, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
+\end{tabular} \\
+
+Additionally, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
 With that, we use a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
@@
 \subsubsection{CPN training}
-Once an EN is properly trainedx, CPN training is straightforward. For this phase,
+Once an EN is properly trained, CPN training is straightforward. For this phase,
 we generate training examples using the current CPN alone. The EN generates predictions of the
@@
 
-As with the EN, the choice of learning rate schedule for training the CPN is somewhat important.
+The choice of learning rate schedule for training the CPN is somewhat important.
 The CPN appears to train in two phases. In the first phase, it is largely learning the structure
@@
 \subsubsection{Interleaved CPN/EN training, and adapting to non-stationarity}
-Having defined the training epochs for the EN and CPN, we can define a training
+Having defined the training periods for the EN and CPN, we can define a training
 algorithm on the basis of those two. We train the two in alternation, creating
@@
 We solve this challenge through a simple set of metrics, which, when satisfied, indicate
-we need to transition to an EN training epoch. Our first metric simply captures the EN's
+we need to transition to an EN training period. Our first metric simply captures the EN's
 declining predictive power. When the EN's prediction error reaches a point which is
@@
 of the three lesion types described above. We additionally vary whether we simulate
-brain co-adaptation, in order to better understand what the co-processor is learning.
+brain co-adaptation, in order to understand its impact on the co-processor training.
 Finally, we perform one experiment where we attempt to simulate natural recovery from
-the lesion, prior to training the co-processor. In sum, our experiments are:
-
-\begin{table}[]
+the lesion, prior to training the co-processor. In sum, our experiments are: \\
+
 \begin{tabular}{|l|l|c|c|}
@@
 7              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 & X                                                                                               & X                                                                                              \\ \hline
-\end{tabular}
-\end{table}
+\end{tabular} \\
 
@@
 			\item 90\% on connection lesion, both with and without co-adaptation
-			\item M1 50\%: 81\% recovery on first attempt with co-adaptation. This is likely to go higher, since I found a bug.
+			\item M1 50\%: 81\% recovery with and without co-adpation.
+The co-processor is having trouble finding a better solution than this, due to loss in brain functionality.
+Increasing the rate of brain co-adaptation may show a better solution.
 			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running.
@@
 Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
-			\item Waiting on all other results.
+			\item Waiting on pre-recovery result. Unclear if the co-processor will help or not.
 		\end{itemize}
@@
 	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
+	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
 	\item Can co-adapt with the brain, as it changes
-	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
 	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
-	\item Show pre-recovery results; co-processor will improve beyond what recovery can achieve. (will run results this week)
+	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve. (Waiting on results; running now)
 \end{itemize}
@@
 today which we address with far simpler stimulation regimes than would be necessary for
-e.g. remediating the symptoms of a stroke. For example, deep brain stimulators for
-treating the symptoms of Parkinson's Disease can now be controlled in a closed loop
-fashion. That allows them to minimize stimulation side-effects and maximize implant
-battery life, while also reducing the patient's primary symptoms. TODO: citation. Our
-simulation does not imply that our co-processor design can translate immediately to
-stroke applications. However, it does demonstrate that our design can adapt to
-long-running effects of stimulation, as well as to the non-stationarity of the neural
-circuit to which it is applying the stimulation. These properties together suggest it may be
-useful in lower dimensional problems where a stimulation model can be estimated, and
-stimulation parameters can be tied to the target objective. TODO: more examples?
-
+e.g. remediating the symptoms of a stroke. In some applications, such as PD symptom relief,
+stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
+voltage parameters, where the key challenge is learning when to apply the stimulation, and
+its magnitude. Our simulation demonstrates the ability of our training method to adapt to
+higher dimensional problems than these, where additionally the long-running effects of
+stimulation on the intended figure-of-merit must be modeled. These properties together
+suggest it may be useful in lower dimensional problems where a stimulation model can be
+estimated, and stimulation parameters can be tied to the target objective.
+
+Our future work will include a reinforcement learning (RL) approach, where the co-processor
+explicitly learns when to apply stimulation, in addition to parameterizing it. Its target
+objective involves not only symptom relief, but also minimizing energy use. In this approach,
+our CPN and EN may correspond to the Actor and Critic portions of an Actor-Critic model
+for example. Here, the Critic would learn to value symptom relief, while attaching a cost
+to stimulation, thus incentivizing an energy efficient approach.
+
+One challenge remaining with our current co-processor design is training effiency.
 On the whole, the co-processor quickly improved task performance, but required orders of
@@
 	      epochs operate on the order of seconds, suggesting that data can be retained
-	      and reused for a period of time. The `speed' of non-stationarity compared to
+	      and reused for multiple epochs. The `speed' of non-stationarity compared to
 	      the co-processor's learning therefore must be an area of future research.
-	\item Matching the dimensionality of the stimulation parameters and observation
-	      data to the amount of available data. If one chooses a stimulation regime with
+	\item Matching the dimensionality of the stimulation and observations
+	      to the amount of available data. If one chooses a stimulation regime with
 	      a small number of controllable parameters, where nevertheless task improvement
-	      from the stimulation is possible, that allows for a simpler co-processor, which
-	      may require less training data. We discuss this next.
+	      from the stimulation is possible, that allows for a simpler co-processor which
+	      will likely require less training data.
 \end{itemize}
 
-TODO: future applications
+Having now demonstrated our method on a higher dimensional problem, we will next demonstrate
+it on a lower dimensional problem, one which is likely to transfer more readily to
+an \textit{in vivo} trial. We hypothesize that this co-processor model will transition well
+to applications in PD, or essential tremor (ET). In Castaño-Candamil et al. \cite{castano.pd},
+we see an ML-based approach to closed-loop stimulation for ET relief. Here, a simple learning
+model extracts neural markers (NMs) from a human patient's M1 cortex. It uses those to
+condition DBS stimulation parameters. The stimulation controller adapts to non-stationarity
+of the mapping between NMs and optimal control, which is necessary due to the sensitivity
+of NMs to context (e.g. sitting or standing). The co-processor model is likely to
+transition well to such an application, where it may provide additional energy savings
+or reduction of stimulation side-effects. Here, it will attempt to extract neural
+markers appropriate for each context, to detect transitions in context, and to apply
+optimal stimulation.
+
+\section{Conclusion}
+We have demonstrated a general framework for training a closed-loop neural stimulator,
+called a ``neural co-processor''. We outlined one specific co-processor design, and
+outlined an approach to training it using supervised learning. In simulation it
+successfully learned to alleviate some of the effects of a brain lesion. To do so, it
+needed to co-adapt to the simulated brain, which exhibited non-stationarity. Additionally,
+it needed to learn long-running effects of the stimulation it learned to apply. Because of
+these properites, the co-processor approach likely generalizes well to a wide range of
+clinical applications where closed-loop neural stimulation is appropriate.
 

### 2021-12-05 paper/article.tex
Paper progress

@@
 	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
-	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve. (Waiting on results; running now)
+	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve.
+Initial result: the co-adaptive case found a better solution.
+The co-processor couldn't improve yet on the pre-recovery case.
+The solution it found was (basically) to ignore the inputs.
 \end{itemize}
@@
 stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
-voltage parameters, where the key challenge is learning when to apply the stimulation, and
-its magnitude. Our simulation demonstrates the ability of our training method to adapt to
+parameters, where the key challenge is learning when to apply the stimulation, and
+its shape and power. Our simulation demonstrates the ability of our training method to adapt to
 higher dimensional problems than these, where additionally the long-running effects of
@@
 transition well to such an application, where it may provide additional energy savings
-or reduction of stimulation side-effects. Here, it will attempt to extract neural
+or reduction of stimulation side-effects. In that case, it will attempt to extract neural
 markers appropriate for each context, to detect transitions in context, and to apply

### 2021-12-11 paper/article.tex
Paper progress

@@
 
+TODO: table of raw healthy vs lesioned vs recovered vs pct recovered
+
+With each experiment, we track two metrics to characterize the learning progress. First, we
+track the task loss, which is an MSE loss that measures the co-processor's ability to restore
+movement towards the target trajectory. We characterize that measure in terms of its percent
+difference from lesioned and healthy performances.
+
+Second, we measure the degree to which the co-processor's solution successfully
+differentiates object classes. We measure the ratio of within-class and total variation,
+and compare that to the healthy network. Our separation metric $S$ is defined as:
+
+\begin{equation}
+	S = \frac{\sigma_{a}}{\sigma_{w}} - \frac{\sigma_{a,h}}{\sigma_{w,h}}
+\end{equation}
+
+where $\sigma$ measures the mean variation, across time, and across all data points in
+a given sample. $\sigma_{a}$ is the total variation of a dataset, and $\sigma_{w}$
+is the average of the within-class variations. The $h$ indicates the healthy
+metric. An $S$ value of 0.0 indicates that the grasps for the various object classes
+vary to the same degree as the healthy network.
+
+This metric $S$ allows us to differentiate overall task performance improvement
+and the ability of the co-processor to condition the grasp based on visual information.
+The structure of the delayed reach-to-grasp task can be learned by the co-processor,
+i.e. to hold muscle velocities to 0.0 for the initial part of each trial, without it
+also learning to differentiate the various object shapes from the brain observations.
+$S$ is our metric for differentiating those two co-processor behaviors.
+
+\subsection{Experiment 1: AIP 75\% lesion}
+This experiment measures the effect of a loss of task representation neural
+machinery on the ability of the co-processor to succeed. In the non-coadaptive
+version of this experiment, the loss of neurons involved in encoding the
+visual inputs results in the simulated brain being unable to condition its
+grasp for the object shape. Likewise, the co-processor doesn't have sufficient
+information to improve performance. The lesioned circuit largely
+
+In the co-adaptive case,
+
+See Fig. \ref{fig:results}.
+
+\subsection{Experiment 2: M1 75\% lesion}
+\subsection{Experiment 3: F5 and M1 connectivity lesion}
+\subsection{Experiment 4: M1 lesion with pre-recovery}
+
 TODO: summary of conclusions that will go here:

### 2021-12-12 paper/article.tex
Paper progress

@@
 gradients which train the CPN to output more useful stimulation. We discovered that
-this latter property does not necessarily occur simply by virtue of the former.
-We have found that an EN can be trained to extreme levels of predictive power, even
+this latter property does not occur simply by virtue of the former.
+We have found that an EN can be trained to high levels of predictive power, even
 on random stimulation, and to orders of magnitude lower loss than the task loss,
@@
 
-\subsubsection{Improving learning efficiency with data reuse}
-TODO. May drop this section if I can't get learning stabilized with this approach.
-
 \subsection{Experiments}
@@
 \begin{figure}[h]
-\includegraphics[width=\textwidth]{results_standin.png}
+\includegraphics[width=\textwidth]{training_results.png}
 \caption{Loss and class separation vs training epoch}
@@
 TODO: summary of conclusions that will go here:
+
+TODO:
+* AIP + coadaptation shows that we can adapt to changes in the input mappings.
+* M1 + coadaptation shows we can adapt to changes in the output mappings
+* Connection lesion shows we can act as a bridge
+
 \begin{itemize}
@@
 It achieved that as well, co-adapting with the simulated brain to improve external task
-performance.  TODO: pre-recovery result.
+performance. Finally, it adapted to a simulated brain which had already undergone some amount of
+lesion recovery. In that case, the co-processor successfully identified the information upstream
+from the lesion which was necessary to stimulate the motor cortex downstream.
 

### 2021-12-12 cpn.py
Pre-recov is working now.

@@ -30,7 +30,11 @@ class CPN_EN_CoProc(experiment.CoProc):
             activation_func=cfg.cpn_activation,
             cuda=cuda,
         )
-        self.opt_cpn = AdamW(self.cpn.parameters(), lr=1e-3)
+
+        if cfg.recover_after_lesion:
+            self.opt_cpn = AdamW(self.cpn.parameters(), lr=4e-3)
+        else:
+            self.opt_cpn = AdamW(self.cpn.parameters(), lr=1e-3)
 
         en_in_dim = cfg.observer_instance.out_dim + stim_dim + 1
         self.en, self.opt_en = stim_model.get_stim_model(

### 2021-12-12 cpn_epoch_cpn.py

@@ -33,6 +33,11 @@ class CPNEpochCPN:
         self.reg_stim_weight = None
         self.stims = []
 
+        if cfg.recover_after_lesion:
+            self.set_opt_lr = self.lr_sched_aggressive_refine3
+        else:
+            self.set_opt_lr = self.lr_sched_standard
+
         self.reset()
 
     def reset(self):
@@ -102,9 +107,71 @@ class CPNEpochCPN:
 
         self.recent_pred_loss = pred_loss.item()
 
-    def finish(self, loss_history, is_validation):
+    def lr_sched_aggressive_refine(self, rtl, eidx):
+        """
+        Args:
+            rtl - recent training loss, which we use to determine the learning rate
+        """
+        if rtl is None or rtl >= 0.005:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 1e-3
+        elif rtl >= 0.004:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 1e-4
+        elif rtl >= 0.002:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 2e-6
+        else:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 1e-6
 
-        rtl = self.recent_train_loss
+    def lr_sched_aggressive_refine2(self, rtl, eidx):
+        """
+        Args:
+            rtl - recent training loss, which we use to determine the learning rate
+        """
+        if rtl is None or rtl >= 0.004:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 5e-4
+        elif rtl >= 0.003:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 1e-5
+        elif rtl >= 0.002:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 2e-6
+        else:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 1e-6
+
+    def lr_sched_aggressive_refine3(self, rtl, eidx):
+        """
+        Args:
+            rtl - recent training loss, which we use to determine the learning rate
+        """
+        if rtl is None or eidx < 4000:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 1e-3
+        elif rtl >= 0.008:
+            for p in self.opt_cpn.param_groups:
+                p["lr"] = 1e-3
+        elif rtl >= 0.006:
+            for p in self.opt_cpn.param_groups:
[…]

### 2021-12-12 experiment/experiment.py

@@ -45,7 +45,7 @@ def get_coadapt_config(cuda=None, **kwargs):
 
 def get_m1_lesion_config(cuda=None, coadapt=True, **kwargs):
     lesion_type = lesion.LesionType.outputs
-    # Lesion the neurons with idxs 50-100, which are in M1
+    # Lesion the neurons with idxs 0-50, which are in M1
     lesion_args = (0, 50)
     return get_config(
         cuda=cuda,

### 2021-12-12 experiment/stats.py
Print healthy/lesion loss at start of run

@@ -1,5 +1,6 @@
 import copy
 import json
+import logging
 
 import attr
 import torch.nn
@@ -9,6 +10,7 @@ from .utils import render_none_or_float
 
 _G_MSELOSS = torch.nn.MSELoss()
 
+g_logger = logging.getLogger("stats")
 
 def calc_task_loss(actuals, targets):
     return _G_MSELOSS(actuals, targets[:, 1:, :])
@@ -46,6 +48,8 @@ class LossHistory:
 
         self.lesioned_loss = lesioned_loss
         self.healthy_loss = healthy_loss
+
+        g_logger.info("Healthy loss: %0.8f, Lesioned loss: %0.8f", lesioned_loss, healthy_loss)
 
         self.eidx = 0
 

### 2021-12-12 experiment/experiment.py
Print the healthy/lesioned raw losses, for clarity

@@ -194,6 +194,21 @@ class Experiment:
             for param in self.mike.parameters():
                 param.requires_grad = False
 
+        if cfg.recover_after_lesion:
+            model_path = os.path.join(RECOV_MODEL_PATH, "recovered_mrnn_%s.model" %
+                str(cfg.lesion_instance))
+
+            try:
+                self.mike.load_weights_from_file(model_path)
+            except OSError as e:
+                if e.errno == errno.ENOENT:
+                    raise ValueError("No pre-generated recovered mRNN available "
+                        "for a lesion of type %s; looked for path %s" % (
+                        str(cfg.lesion_instance), model_path))
+                raise
+
+            self.opt_mike = AdamW(self.mike.parameters(), lr=1e-7)
+
         self.observer = cfg.observer_instance
 
         (
@@ -214,21 +229,6 @@ class Experiment:
             self.obs_drop_module_idx = 0
         else:
             self.obs_drop_module_idx = None
-
-        if cfg.recover_after_lesion:
-            model_path = os.path.join(RECOV_MODEL_PATH, "recovered_mrnn_%s.model" %
-                str(cfg.lesion_instance))
-
-            try:
-                self.mike.load_weights_from_file(model_path)
-            except OSError as e:
-                if e.errno == errno.ENOENT:
-                    raise ValueError("No pre-generated recovered mRNN available "
-                        "for a lesion of type %s; looked for path %s" % (
-                        str(cfg.lesion_instance), model_path))
-                raise
-
-            self.opt_mike = AdamW(self.mike.parameters(), lr=1e-7)
 
     @property
     def cfg(self):

### 2021-12-12 experiment/stats.py

@@ -12,6 +12,7 @@ _G_MSELOSS = torch.nn.MSELoss()
 
 g_logger = logging.getLogger("stats")
 
+
 def calc_task_loss(actuals, targets):
     return _G_MSELOSS(actuals, targets[:, 1:, :])
 
@@ -49,7 +50,9 @@ class LossHistory:
         self.lesioned_loss = lesioned_loss
         self.healthy_loss = healthy_loss
 
-        g_logger.info("Healthy loss: %0.8f, Lesioned loss: %0.8f", lesioned_loss, healthy_loss)
+        g_logger.info(
+            "Healthy loss: %0.8f, Lesioned loss: %0.8f", healthy_loss, lesioned_loss
+        )
 
         self.eidx = 0
 
@@ -85,9 +88,7 @@ class LossHistory:
             "task_loss": render_none_or_float(rec.task_loss),
             "task_val_loss": render_none_or_float(rec.task_val_loss),
             "pct_recov": render_none_or_float(rec.pct_recov, fmt="%0.3f"),
-            "class_separation": render_none_or_float(
-                rec.class_separation, fmt="%0.3f"
-            ),
+            "class_separation": render_none_or_float(rec.class_separation, fmt="%0.3f"),
         }
 
         if rec.user_data is not None:
@@ -154,7 +155,6 @@ class LossHistory:
         self._recs.append(rec)
         self.eidx += 1
 
-
     def report_by_result(
         self, actuals, dout, labels, update_task_loss=True, user_data=None
     ):
@@ -172,7 +172,8 @@ class LossHistory:
     ):
         if update_task_loss:
             task_val_loss = calc_task_loss(actuals, dout)
-        else: task_val_loss = None
+        else:
+            task_val_loss = None
 
         self.report(None, task_val_loss=task_val_loss, user_data=user_data)
 

### 2021-12-12 paper/article.tex
Paper progress

@@
 
-TODO: table of raw healthy vs lesioned vs recovered vs pct recovered
+TODO: label lines/colors
+
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
+AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
+AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
+M1 No-coadapt  & 0.021136                                                          &                                                                  &                                                                      & 102.75\%           & 102.75\%                                                         \\ \hline
+M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
+Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
+Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
+Pre-recov*      & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
+\end{tabular}
+\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
+\end{table}
 
@@
 
+TODO: describe stopping conditions
+
 \subsection{Experiment 1: AIP 75\% lesion}
-This experiment measures the effect of a loss of task representation neural
-machinery on the ability of the co-processor to succeed. In the non-coadaptive
-version of this experiment, the loss of neurons involved in encoding the
-visual inputs results in the simulated brain being unable to condition its
-grasp for the object shape. Likewise, the co-processor doesn't have sufficient
-information to improve performance. The lesioned circuit largely
-
-In the co-adaptive case,
-
-See Fig. \ref{fig:results}.
+This experiment demonstrates the ability for the co-processor to succeed after the
+patient has lost some of the neural machinery responsible for representing the task.
+In the non-coadaptive version of this experiment, the loss of AIP neurons involved
+in encoding the visual inputs results in the simulated brain being unable to
+condition its grasp for the object shape, though reaching largely still succeeds.
+Likewise, the co-processor doesn't have sufficient information to improve performance.
+In this case, the co-processor is effectively unable to improve on existing performance,
+for lack of that information.
+
+In the co-adaptive case, the brain and co-processor find a solution that largely recovers
+performance. This demonstrates that the co-processor can adapt to the non-stationarity
+of the mapping between the observed brain activity and the optimal stimulation. Note,
+though, that the simulation is not designed to indicate if the co-processor speeds up
+recovery, or results in greater results than natural recovery, since the simulation
+is not designed to answer that question.
+
+Additionally, in Fig. \ref{fig:results} we see that initially the co-processor and
+brain did not strongly separate the object classes, but that they learned to as the
+training proceeded. This is again due to the co-processor and brain learning how
+to map the input visual information on to the task.
 
 \subsection{Experiment 2: M1 75\% lesion}
-\subsection{Experiment 3: F5 and M1 connectivity lesion}
-\subsection{Experiment 4: M1 lesion with pre-recovery}
-
-TODO: summary of conclusions that will go here:
-
-TODO:
-* AIP + coadaptation shows that we can adapt to changes in the input mappings.
+TODO: after final results
 * M1 + coadaptation shows we can adapt to changes in the output mappings
-* Connection lesion shows we can act as a bridge
-
-\begin{itemize}
-	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
-		\begin{itemize}
-			\item 90\% on connection lesion, both with and without co-adaptation
-			\item M1 50\%: 81\% recovery with and without co-adpation.
-The co-processor is having trouble finding a better solution than this, due to loss in brain functionality.
-Increasing the rate of brain co-adaptation may show a better solution.
-			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running.
-The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow.
-Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
-			\item Waiting on pre-recovery result. Unclear if the co-processor will help or not.
-		\end{itemize}
-	\item Initially, performance improves orders of magnitude more quickly than later.
+* Initially, performance improves orders of magnitude more quickly than later.
 50-75\% recovery during this phase.
@@
 not to move the hand until after the hold signal is lifted).
-	\item Fine tuning takes a far longer time
-	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
-	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
-	\item Can co-adapt with the brain, as it changes
-	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
-	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve.
-Initial result: the co-adaptive case found a better solution.
-The co-processor couldn't improve yet on the pre-recovery case.
-The solution it found was (basically) to ignore the inputs.
-\end{itemize}
+* Later: fine tuning and object class separation. Learning task structure.
+
+\subsection{Experiment 3: F5 and M1 connectivity lesion}
+As in the other cases, the co-processor's loss in this experiment improved
+quickly, initially, but takes a long time to refine. During that refinement,
+object differentiation occurs. Note that in this experiment, no information
+can propagate forward to the M1 module, since we have completely severed the
+input connections to that module. It is the sole responsibility of the
+co-processor to perform that forward-propagation. As a result, co-adaptation
+only affects the M1 module. The co-processor's learning algorithm successfully
+adapted to this case, where the observed brain activity was stationary, but
+the mapping between stimulation and task performance was not.
+
+\subsection{Experiment 4: M1 lesion with pre-recovery}
+In this case, we simulate some amount of recovery prior to the co-processor
 

### 2021-12-13 paper/article.tex
Paper progress

@@
 We sought to additonally re-train the mRNN after applying the lesion in
-order to simulate stroke recovery. For simulated lesions which zero the outputs
-of neurons, the mRNN is, unfortunately, not an appropriate model.
+order to simulate stroke recovery. For our simulated lesions which zero
+the outputs of neurons, the mRNN cannot supply an appropriate pre-recovery model.
 We know that strokes can cause damage from which the subject, in many cases,
@@
 
-We can, however, simulate pre-recovery in the case of a lesion that excludes
+We can, however, simulate pre-recovery in the case of a lesion that prevents
 communication between the F5 and M1 modules. In that case, object shape information
@@
 
+TODO: describe stopping conditions
+
 \section{Results}
@@
 \end{figure}
-
-TODO: label lines/colors
 
@@
 Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
-Pre-recov*      & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
+Pre-recovery*  & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
 \end{tabular}
@@
 $S$ is our metric for differentiating those two co-processor behaviors.
-
-TODO: describe stopping conditions
 
@@
 
-\subsection{Experiment 4: M1 lesion with pre-recovery}
+\subsection{Experiment 4: Connectivity lesion with pre-recovery}
 In this case, we simulate some amount of recovery prior to the co-processor
+learning. Complete recovery is impossible, because the output module's
+connections to F5 (both forward and backward) are masked. As a result,
+the pre-recovery causes the simulated brain to learn an object-agnostic
+stereotyped grasp. The co-processor's job is to forward-propagate the
+information necessary to condition the grasp on the object shape.
+
+The co-processor in this scenario successfully improved task performance.
+Note that the object separation class separation metric $S$ returned to
+a healthy value towards the end of the training, indicating that the
+learned solution did not simply refine the stereotyped grasp. That in
+turn demonstrates the co-processor's ability to propagate information
+``around'' the simulation lesion.
 

### 2021-12-15 paper/article.tex
Paper progress

@@
 
-TODO: describe stopping conditions
+\subsection{Stopping condition}
+For the purpose of these experiments, we train the co-processor until one of two
+conditions is satisfied:
+
+\begin{itemize}
+	\item The percent change in task loss between two consecutive 500 epoch windows is below a certain threshold: 0.01\%. This
+	      stopping criteria ensures we stop training when the economy of additional training is low.
+	\item The epoch count is greater than 250k. This criteria provides a backstop against very slow training running without end.
+\end{itemize}
 
@@
 AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
-M1 No-coadapt  & 0.021136                                                          &                                                                  &                                                                      & 102.75\%           & 102.75\%                                                         \\ \hline
+M1 No-coadapt  & 0.021136                                                          & 0.005765                                                         & 0.005604                                                             & 74.73\%            & 75.51\%                                                          \\ \hline
 M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
@@
 This experiment demonstrates the ability for the co-processor to succeed after the
-patient has lost some of the neural machinery responsible for representing the task.
+patient has lost some of the neural machinery responsible for representing the objects' shapes.
 In the non-coadaptive version of this experiment, the loss of AIP neurons involved
@@
 condition its grasp for the object shape, though reaching largely still succeeds.
-Likewise, the co-processor doesn't have sufficient information to improve performance.
-In this case, the co-processor is effectively unable to improve on existing performance,
+Likewise, the co-processor doesn't have sufficient information to improve performance,
 for lack of that information.
@@
 \subsection{Experiment 2: M1 75\% lesion}
-TODO: after final results
-* M1 + coadaptation shows we can adapt to changes in the output mappings
-* Initially, performance improves orders of magnitude more quickly than later.
-50-75\% recovery during this phase.
-This appears to be the co-processor learning the task structure (e.g.
-not to move the hand until after the hold signal is lifted).
-* Later: fine tuning and object class separation. Learning task structure.
+In this experiment, 75\% of the output module of the simulated network has been
+lesioned. In the coadaptive case, the co-processor needs to adapt to
+non-stationarity in the mapping between stimulation and task performance.
+In the non-coadaptive case, training quickly plateaus, and does not improve
+thereafter. That suggests we have lost some of our degrees of freedom to control
+the activations in the output layer, resulting in a reduced ability to recreate
+the target grasps.
+
+In the co-adapative case, performance likewise hits an inflexion point, but
+continues to improve slowly.  TODO: finish once we have the faster co-adaptation
+rate tested.
 
 \subsection{Experiment 3: F5 and M1 connectivity lesion}
-As in the other cases, the co-processor's loss in this experiment improved
-quickly, initially, but takes a long time to refine. During that refinement,
+As in the other cases, the co-processor's loss in this experiment improves
+quickly, initially, but taskes much longer to refine. During that refinement,
 object differentiation occurs. Note that in this experiment, no information
@@
 input connections to that module. It is the sole responsibility of the
-co-processor to perform that forward-propagation. As a result, co-adaptation
-only affects the M1 module. The co-processor's learning algorithm successfully
-adapted to this case, where the observed brain activity was stationary, but
-the mapping between stimulation and task performance was not.
+co-processor to perform that forward-propagation.
+
+In this case, co-adaptation only affects the M1 module, since we have nothing
+to use as a learning signal: there is no backpropagation to the AIP and
+F5 modules. As a result, this experiment primarily demonstrates that the
+co-processor's learning algorithm is capable of adapting to non-stationarity
+in the mapping between stimulation parameters and task performance. Contrast
+that with the results of Experiment 1, where the non-stationarity affected
+the mapping between the visual inputs and the brain observations.
 
@@
 The co-processor in this scenario successfully improved task performance.
-Note that the object separation class separation metric $S$ returned to
+Note that the object class separation metric $S$ returned to
 a healthy value towards the end of the training, indicating that the
-learned solution did not simply refine the stereotyped grasp. That in
-turn demonstrates the co-processor's ability to propagate information
-``around'' the simulation lesion.
+learned solution did not simply refine the stereotyped grasp.
 

### 2021-12-19 paper/article.tex
Paper progress

@@
 explore some of the core algorithms that may allow co-processors to successfully
-learn and to adapt to non-stationarity in both the brain and sensors.
+learn, and to adapt to non-stationarity in both the brain and sensors.
 \textit{Approach} We provide the first proof-of-concept of a neural co-processor that
@@
 as our approach to training it, leading to insights on how such a model may be
-developed for \textit{in vivo} use. We believe this co-processor design will allow for
-learning complex stimulation policies that help restore function to a stroke victim.
+developed for \textit{in vivo} use. We believe this co-processor design has potential to
+learn complex stimulation policies, leading to application in treating a variety of
+diseases.
 \end{abstract}
@@
 to better regulate the energy use of an implanted stimulator, allowing it to
-intelligently regulate when to apply stimulation, in order to presever implant
+intelligently regulate when to apply stimulation, in order to preserve implant
 battery life. Closed-loop stimulation also offers an opportunity minimize
@@
 it in a subject- and time-specific way. Our model seeks to accomplish that using
-deep learning, together with a data-efficient approach to training.
+deep learning.
 
@@
 grasping circuit, presented previously by Michaels et al. \cite{michaels.mrnn}. We
-explore the properties of the artificial neural networks (ANNs) that are needed to
+explore the properties of the co-processor design that are needed to
 successfully adapt to the long-running dynamics of a stimulated neural circuit,
 as well as to adapt to that circuit's ongoing connectivity changes.
-We additionally present a training method our co-processor, allowing it to learn
+We additionally present a training method for our co-processor, allowing it to learn
 an optimal stimulation pattern that drives improvements in external task performance,
@@
 A detailed simulation such as this allows us to explore some of the critical
-architectural details and training algorithms for a co-processor of this type.
-By first doing such exploration in simulation, we are able to rapidly and
-cheaply iterate on our design, prior to any \textit{in vivo} experiments.
-For the simulation to be admissible, we need to ensure it has some
+architectural details and training algorithms that will allow a co-processor
+to succeed.  By first doing such exploration in simulation, we are able to
+rapidly and cheaply iterate on our design, prior to any \textit{in vivo}
+experiments. For the simulation to be admissible, we need to ensure it has some
 properties that allow it to strongly indicate if our design is improving in
@@
 Here, we use an established design for an artificial neural circuit which
-performs a grasping task, whose architecture and training methods were designed
+performs a grasping task. Its architecture and training methods were designed
 specifically to result in naturalistic dynamics.
@@
 The emergent network dynamics show a number of other correspondences to the monkeys' natural
-brain activity as well, as detailed in the paper.
+brain activity as well, as detailed in the paper. For convenience, we will refer to these
+modules throughout this paper using the brain region they correspond to.
 
@@
 object shapes. That is - the visual information input to the network that differentiates
-one trial from another is leveraged by the circuit to properly condition
+one trial from another is leveraged by the circuit to successfully condition
 the hand shape trajectory for grasping the object of that trial's particular shape, size, and
@@
 machinery needed to translate or forward-propagate object shape information
-to movement-related cortex can result in a reduced ability to form the hand properly for
-the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
+to movement-related cortex can result in a reduced ability to pose the hand, though
+positioning of the hand may still roughly succeed. Muscle spasticity of
 the hand is also a common symptom of certain strokes in primates, and indeed many who
-suffer from it are still able to position their hand, even while being unable to form
+suffer from tem are still able to position their hand, even while being unable to form
 it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
-simulated lesion closely resembles natural stroke symptoms. However, the simulation is not
-intended to constitute a physical model of a lesion, i.e. to directly explain the
+simulated lesion closely resembles natural stroke symptoms. However, the simulation is
+not intended to constitute a physical model of a lesion, i.e. to directly explain the
 connection between hand spasticity and the lesion.
@@
 to effectively bridge across the lesion, forward propagating the object shape and task
-structyure information, indirectly, through the use of stimulation.
+structure information, indirectly, through the use of stimulation.
 
@@
 	      neurons to zero, effectively removing them from the network. This results in a
-	      loss of object shape information until the brain co-adapts to the lesion.
+	      loss of object shape information.
 	\item \textbf{Connection-based:} we prevent the propagation of
@@
 in time during a trial. If we repeat that experiment many times, we can see what the
-distribution of long- running effects tends to look like.
+distribution of long-running effects tends to look like on the output muscle velocities.
 
@@
 Instead, our stimulation model is merely attempting to make the co-processor's task
-more realistic, in the sense that it cannot precisely set the hidden state of every
-neuron, instantaneously and in each time step. Any stimulation it applies will diffuse
-across ``space'' and time.
+more realistically difficult, in the sense that it cannot precisely set the hidden
+state of every neuron, instantaneously and in each time step. Any stimulation it
+applies will diffuse across ``space'' and time.
 
@@
 is inappropriately low resolution for our task, for example epidural stimulation.
-However, we cannot conclude much from these experiments, because this simulation does
+We could not conclude much from such an experiment, because this simulation does
 not constitute a biophysical model, where we can identify some realistic value. The
 purpose of this smoothing technique is only to make the co-processor's job harder in
-the simple sense that it cannot target neurons individually.
-As a result, we set it to a moderate, balanced value, and don't vary it.
+the simple sense that it cannot target neurons individually. As a result, we set it
+to a moderate, balanced value, and don't vary it.
 
@@
 supplied by a highly trained co-processor, where task performance has been
-largely restored. We explain this experiment in more detail below.
+significantly restored. We explain this experiment in more detail below.
 
@@
 lower than the co-processor by orders of magnitude due
-to the rapidity with which an ANN can be retrained.
+to the rapidity with which an ANN can be retrained in practice.
 
@@
 cannot recover. The mRNN model has sufficient redundancy built into it that
-lesioning the outputs of large numbers of neurons leaves enough remaining
+lesioning the outputs of large numbers of neurons often leaves enough remaining
 degrees of freedom that a nearly full recovery can occur, unless so many
@@
 this latter property does not occur simply by virtue of the former.
-We have found that an EN can be trained to high levels of predictive power, even
+An EN of this design can be trained to high levels of predictive power, even
 on random stimulation, and to orders of magnitude lower loss than the task loss,
@@
 orders of magnitude lower than in the first phase. We will cover this in some more detail in
-Section \ref{sec:results}.
+the Results section (Section \ref{sec:results}).
 
@@
 though, that the simulation is not designed to indicate if the co-processor speeds up
-recovery, or results in greater results than natural recovery, since the simulation
-is not designed to answer that question.
+recovery, or results in greater results than natural recovery>
 
@@
 In this experiment, 75\% of the output module of the simulated network has been
-lesioned. In the coadaptive case, the co-processor needs to adapt to
+lesioned. In the co-adaptive case, the co-processor needs to adapt to
 non-stationarity in the mapping between stimulation and task performance.
@@
 In the co-adapative case, performance likewise hits an inflexion point, but
-continues to improve slowly.  TODO: finish once we have the faster co-adaptation
-rate tested.
+continues to improve slowly. The ability to improve further is tied to the
+simulated recovery restoring the degrees of freedom needed for success. The
+co-processor in this later phase of training is tasked with adapting to the
+non-stationarity that results from the recovery process, which it does
+successfully.
 
@@
 object differentiation occurs. Note that in this experiment, no information
-can propagate forward to the M1 module, since we have completely severed the
-input connections to that module. It is the sole responsibility of the
-co-processor to perform that forward-propagation.
+can propagate forward to the M1 module, or backwards from it, since we have
+completely severed the inter-module connections involving that module. It is
+the sole responsibility of the co-processor to perform that
+forward-propagation.
 
@@
 to use as a learning signal: there is no backpropagation to the AIP and
-F5 modules. As a result, this experiment primarily demonstrates that the
+F5 modules. As a result, this experiment demonstrates that the
 co-processor's learning algorithm is capable of adapting to non-stationarity
@@
 
-The co-processor in this scenario successfully improved task performance.
-Note that the object class separation metric $S$ returned to
-a healthy value towards the end of the training, indicating that the
-learned solution did not simply refine the stereotyped grasp.
+The co-processor in this scenario successfully improved task performance
+beyond the recovery point. Note that the object class separation metric
+$S$ returned to a healthy value towards the end of the training,
+indicating that the learned solution did not simply refine the
+stereotyped grasp.
 
@@
 Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
-50-90\% across our various experiments. To work this well, the co-processor needed to adapt to
-the long-running dynamics of the stimulated network, as well as the long-running effects of
+75-90\% across our various experiments. To work this well, the co-processor needed to adapt to
+the long-running dynamics of the simulated network, as well as the long-running effects of
 stimulation. In some experiments we required it to additionally adapt to non-stationarity in
@@
 today which we address with far simpler stimulation regimes than would be necessary for
-e.g. remediating the symptoms of a stroke. In some applications, such as PD symptom relief,
+e.g. fine-grained finger control. In some applications, such as PD symptom relief,
 stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
 parameters, where the key challenge is learning when to apply the stimulation, and
-its shape and power. Our simulation demonstrates the ability of our training method to adapt to
-higher dimensional problems than these, where additionally the long-running effects of
-stimulation on the intended figure-of-merit must be modeled. These properties together
-suggest it may be useful in lower dimensional problems where a stimulation model can be
-estimated, and stimulation parameters can be tied to the target objective.
+its shape and power. Our simulation demonstrates the ability of our training method to
+adapt to higher dimensional problems than these, where additionally the long-running
+effects of stimulation on the intended figure-of-merit must be modeled. These properties
+together suggest it may be useful in lower dimensional problems where a stimulation model
+can be estimated, and stimulation parameters can be tied to the target objective.
 
@@
 successfully learned to alleviate some of the effects of a brain lesion. To do so, it
-needed to co-adapt to the simulated brain, which exhibited non-stationarity. Additionally,
+needed to co-adapt with the simulated brain, which exhibited non-stationarity. Additionally,
 it needed to learn long-running effects of the stimulation it learned to apply. Because of

### 2022-09-22 paper/article.tex
Paper progress. .tex not sync'd with overleaf yet, but figs are up to date.

@@
 \usepackage{subcaption}
+\usepackage{lscape}
 \graphicspath{ {./figs/} }
@@
 Aided in part by application of advanced AI techniques, brain-computer interfaces (BCIs) have made
-advancements over the last several decads, allowing for decoded brain signals to be used for
+advancements over the last several decades, allowing for decoded brain signals to be used for
 control of a wide variety of virtual and physical prostheses \cite{rao.bcibook, wolpaw.bcibook,
@@
 the effect of stimulation, we additionally train it to be useful for then learning an
-optimal stimulation policy, which is a property somewhat distinct from predictive
-power alone.
+optimal stimulation policy, which we will show is a property somewhat distinct from
+predictive power alone.
 
@@
 the use of open loop stimulation in restoring dexterity after a lesion
-occurs affecting a primate's motor cortex. The authors demonstrate that
+of a primate's motor cortex. The authors demonstrate that
 the use of low-frequency alternating current, applied epidurally,
@@
 as a result of differing inputs and ongoing activity. Stimulation
-therefore needs to be proactively adapted in response. This need is even
-greater over longer time scales as the effects of plasticity and ageing change
+therefore needs to be proactively adapted in response \cite{shanechi.stimmodel, castano.pd}.
+This need is even greater over longer time scales as the effects of plasticity and ageing change
 the connectivity of the brain. Closed-loop stimulation may also provide means
@@
 For a neurologically complex task such as grasping, we cannot identify
-\textit{a priori} a real time controller of the neural circuit. That is
-due in large part to the variability of circuits from subject-to-subject,
-as well as variations in the placement of sensors and stimulators in the
-brain. The only plausible path to such a real time controller is to parameterize
-it in a subject- and time-specific way. Our model seeks to accomplish that using
-deep learning.
+\textit{a priori} a real time stimulation regime of the neural circuit which would
+result in precise task-relevant control. That is due in large part to the variability
+of circuits from subject-to-subject, as well as variations in the placement of sensors
+and stimulators in the brain. The only plausible path to such a real time controller
+is to parameterize it in a subject- and time-specific way. Our model seeks to
+accomplish that using deep learning.
 
@@
 	\item A stimulation and neural dynamics model, known as an ``Emulator Network'' (EN). It models the relationship
-	      between the stimulation, neural dynamics, and external task. Its purpose is for training the second network.
+	      between the stimulation, neural dynamics, and external task. Its purpose is for training the second network,
+	      by accurately modeling the effect of stimulation.
 	\item A stimulation agent, known as the ``co-processor network'' (CPN), which maps neural activity, and possibly
@@
 We co-train the EN and CPN, with the goal of training a CPN whose output stimulation parameters
-improve task performance. By continually training both, they adapt to the brain as it changes,
-effectively allowing for brain-stimulator co-adaptation. The EN is a tool for training the CPN,
-giving us a way to backpropagate task error to the CPN. It outputs task-relevant
-metrics - a prediction of muscle velocities in our case - given measurements of neural
+improve task performance. By iteratively training and retraining both over time, they adapt to
+the brain as it changes, effectively allowing for brain-stimulator co-adaptation. The EN is a
+tool for training the CPN, giving us a way to backpropagate task error to the CPN. It outputs
+task-relevant metrics - a prediction of muscle velocities in our case - given measurements of neural
 activity, and the stimulation parameters output from the CPN. If the EN is trained in a
@@
 the CPN with it. When training the CPN, we in-effect treat the EN's output as the true
-task performance, or a related metric, and then backpropagate the loss defined in terms
-of that metric in order to train the CPN. See Fig. \ref{fig:weill}. We will illustrate the details
-of the training algorithm below.
+value of the task-relevant metric, or a related metric, and then backpropagate the loss defined
+in terms of that metric in order to train the CPN. See Fig. \ref{fig:weill}. We will illustrate
+the details of the training algorithm below.
+
+Put another way: we define our loss functions for both the EN and CPN based on an L2 error over
+a vector of muscle length velocities. The EN's error measures the EN's prediction error for some
+stimulation and measured neural activity. The CPN's error measures the difference between actual
+and target muscle length velocities. In our case, those targets come from the simulated task;
+see below.
 
@@
 
-An example of such a simulation approach is Dura-Bernal et al. \cite{bernal.sim}. In this work, the
-authors use a simulated spiking neural network to train a stimulation agent.
-Their stimulation agent sought to restore the network's
-control of a simulated arm, to reach a target, after a simulated lesion was applied,
-much like the work here.  As the authors point out, we have a limited ability to
-probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
-we first need to design our approach through the use of an admissible simulation.
-The authors in this case simulated lesions by effectively removing parts of their
-simulated network, or by cutting connections between parts of the network.
-Here, we use an established design for an artificial neural circuit which
-performs a grasping task. Its architecture and training methods were designed
-specifically to result in naturalistic dynamics.
-
 The simulated circuit, from Michaels et al. \cite{michaels.mrnn}, was trained
-to resemble the grasping circuits of monkey subjects engaged in a delayed reach-to-grasp task.
+to resemble the grasping circuits of primate subjects engaged in a delayed reach-to-grasp task.
 Its design draws on a body of literature focused on architectures and training methods for RNNs which
@@
 and are connected to each other sparsely (10\% connectivity). The inputs are visual features
-intended to capture the view the monkeys had during the task, specifically VGGNet features \cite{simonyan.vgg},
-from 3D renderings of the same objects which the monkeys grasped. The outputs are muscle length
-velocities for the shoulder, arm, and hand of the monkey during the trial. The natural
+intended to capture the view the primates had during the task, specifically VGGNet features \cite{simonyan.vgg},
+from 3D renderings of the same objects which the primates grasped. The outputs are muscle length
+velocities for the shoulder, arm, and hand of the primate during the trial. The natural
 velocities were captured with a motion capturing glove, and the artificial neural network
@@
 		\includegraphics[width=\textwidth]{michaels.eps}
-		\caption{Michaels Modular RNN (mRNN), a simulated grasping circuit. The emergent dynamics of the
-		three modules correspond well to natural neural activity measured from primate AIP, F5, M1 regions,
-		respectively, from the same task. Visual (VGGNet) features forward propagate through the
-		network, conditioning the grasp for the object of the particular size, shape, and location.}
+		\caption{Michaels Modular RNN (mRNN), a simulated grasping circuit \cite{michaels.mrnn}.
+		The emergent dynamics of the three modules correspond well to natural neural activity
+		measured from primate AIP, F5, M1 regions, respectively, from the same task. Visual
+		(VGGNet) features forward propagate through the network, conditioning the grasp for the
+		object of the particular size, shape, and location.}
 	\end{subfigure}
@@
 emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
-to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
+to the AIP, F5, and M1 portions of the primate subjects' brains. `Correspond' here, notably,
 refers to the fact that the activity of the module receiving the visual inputs resembles
-the AIP activity of the monkey subject from the same trials. Likewise, the second and
-third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
-The emergent network dynamics show a number of other correspondences to the monkeys' natural
+the AIP activity of the primate subject from the same trials. Likewise, the second and
+third modules resemble the natural activity from the primate's F5 and M1 regions, respectively.
+The emergent network dynamics show a number of other correspondences to the primates' natural
 brain activity as well, as detailed in the paper. For convenience, we will refer to these
@@
 stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
-grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
-waiting period, but performance suffers drastically.
+grasps the given primate performed, due to all trials being a reach-to-grasp preceded by a
+holding period, but performance suffers drastically.
 
@@
 the hand is also a common symptom of certain strokes in primates, and indeed many who
-suffer from tem are still able to position their hand, even while being unable to form
-it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
-simulated lesion closely resembles natural stroke symptoms. However, the simulation is
-not intended to constitute a physical model of a lesion, i.e. to directly explain the
-connection between hand spasticity and the lesion.
-
-Conversely, if we lesion of a portion of the output module of the Michaels mRNN, roughly
-corresponding to M1, we see a more wholesale loss of movement, affecting even the
-ability to reach for the grasp. Finally, if we ``disconnect'' communications between the
-F5 and M1 modules, we see a failure similar to an AIP lesion: movement is generally
-achieved, but we see a disproportionate impact on hand pose.
+suffer from them are still able to position their hand, even while being unable to form
+it properly for a grasp \cite{khanna.openloop,puthenveettil.hand}. In that sense, the
+error mode of this simulated lesion closely resembles natural stroke symptoms. However,
+the simulation is not intended to constitute a physical model of a lesion, i.e. to
+directly explain the connection between hand spasticity and the lesion.
+
+If we instead ``disconnect'' communications between the F5 and M1 modules, we see a
+failure similar to the AIP lesion: movement is generally achieved, but we see a
+disproportionate impact on hand pose. Finally, if we lesion of a portion of the output
+module of the Michaels mRNN, roughly corresponding to M1, we see a more wholesale loss
+of movement, affecting even the ability to reach for the grasp.
 See Fig. \ref{fig:lesion} for examples.
@@
 	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
-	    \caption{Example muscle trajectories during a single trial, for one hand and shoulder muscle}
+	    \caption{Example muscle trajectories during a single trial, for one hand and shoulder muscle.
+	             Orange: original trajectory. Blue: with lesion.}
 	\end{subfigure}
@@
 	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
-	    \caption{Ratio of hand muscle MSE losses vs. overall MSE losses, for different lesion types}
+	    \caption{Ratio of hand muscle MSE losses vs. overall MSE losses, for different lesion types.
+	             Average of all trials, with +/- 1 stdev shown.}
 	\end{subfigure}
@@
 \begin{itemize}
-	\item \textbf{Output-based (AIP):} we force the output of some proportion of ``AIP''
+	\item \textbf{Output-based AIP:} we force the output of some proportion of ``AIP''
 	      neurons to zero, effectively removing them from the network. This results in a
@@
 	      between them run both directions, and we are lesioning both.
-	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
+	\item \textbf{Output-based M1:} we force the output of some proportion of ``M1''
 	      neurons to zero, causing a more holisitic loss of movement. Here, the lesion
@@
 	    \includegraphics[width=\textwidth]{perturbe_dist.pdf}
-	    \caption{Distribution of effects across n=1000 samples}
+	    \caption{Distribution of effects across n=1000 samples. Greyscale
+	             field represents variance across samples.}
 	\end{subfigure}
@@
 
-In order to inject another aspect of realism into our simulation, we subject
+In order to include another aspect of realism in our simulation, we subject
 the co-processor to a stimulation model, rather than allowing it to directly
@@
 it affects others, according to each neuron's Mahalanobis distance from it. The neurons
-are aligned along that dimension arbitrarily. We fix the width of the Gaussians
+are aligned along that dimension arbitrarily. We fix the width of the Gaussian distributions
 arbitrarily to $\sigma=1.75$. We are not aware of a principled way of picking this
@@
 	\item $I$: the input response matrix
+	\item $v_{t}$: the visual input vector
 	\item $b$: the neuron activation bias
@@
 the fact it receives stimulation. The learning rate for the optimizer
-is set arbitrarily to a relatively low rate of $1e-7$. The learning
+is set arbitrarily to a relatively low rate of $1e^{-7}$. The learning
 rate was not chosen in a principled way, since we are not aware of a
@@
 \subsubsection{Simulating recovery prior to co-processor use}
-We sought to additonally re-train the mRNN after applying the lesion in
+We sought to additionally re-train the mRNN after applying the lesion in
 order to simulate stroke recovery. For our simulated lesions which zero
@@
 cannot propagate forward in the network, in order to condition the hand for
-grasping. The mRNN can, in pre-recovery, learn a stereotyped grasp, at which
-point the co-processor's job will be to forward-propagate that information.
-We explore this in one of our experiments, explained below.
+grasping. The mRNN can, in pre-recovery, learn a stereotyped grasp. After that point,
+the co-processor's job will be to forward-propagate object shape information
+to specialize that grasp for the individual objects. We explore this in one of our
+experiments, explained below.
+
+\subsubsection{Simulating a non-stationary observation function, e.g. sensor drift}
+We also demonstrate the co-processor's ability to adapt to non-stationarity in the observation
+function. Over time, implanted sensors may drift from true readings. To test the co-processor's
+ability to adapt to that drift, we perform an experiment where the observation function
+has a bias term which changes over time, according to a random process. Each epoch, we
+add a random value to each element of the bias, drawn from a mean-0 Gaussian distribution. That
+causes the ``sensors'' to have a drifting zero point over time.
+See Supplement \ref{sup:drift} for further details.
 
@@
 EN and CPN training epochs. EN training epochs concentrate on updating the EN,
-based on observations of the effect of stimulation on the simulated network's output.
-The loss function when training the EN is the mean squared error (MSE) loss between the
-actual and predicted output of the mRNN (i.e. muscle velocities). We then use the EN
-to train the CPN during the CPN training epochs. During that time, we backpropagate
-the MSE loss between the EN's predicted output, and target output, to the CPN. In
-other words: we treat the EN output as a proxy to the true brain, allowing us
-to train the CPN, since we have no way to ``backpropagate through the brain'' itself.
+based on observations of the effect of stimulation on the simulated network's output,
+and comparing it to the EN's predictied output. The loss function when training the
+EN is the mean squared error (MSE) loss between the actual and predicted output of
+the mRNN (i.e. muscle velocities). We then use the EN to train the CPN during the
+CPN training epochs. During that time, we backpropagate the MSE loss between the
+EN's predicted output, and target output, to the CPN. In other words: we treat the
+EN's output as a proxy to the true brain, allowing us to train the CPN, since we
+have no way to ``backpropagate through the brain'' itself.
 See Fig. \ref{fig:training}.
@@
 the object shape information observed from the simulated brain, and how that should map onto
-stimulation. That phase of the training takes much longer, and requires a learning rate 2-3
-orders of magnitude lower than in the first phase. We will cover this in some more detail in
-the Results section (Section \ref{sec:results}).
+stimulation. That phase of the training requires a learning rate 2-3 orders of magnitude lower than
+in the first phase. We will cover this in some more detail in the Results section
+(Section \ref{sec:results}).
 
 \subsubsection{Interleaved CPN/EN training, and adapting to non-stationarity}
-Having defined the training periods for the EN and CPN, we can define a training
-algorithm on the basis of those two. We train the two in alternation, creating
-a new EN each time we enter an EN training period. We explored the possibility
+Having defined the training procedures for the EN and CPN, we can define a training
+algorithm combining those two. We train the two in alternation, creating
+a new EN each time we enter an EN training phase. We explored the possibility
 of reusing an existing EN by retraining it, for the purpose of training efficiency.
@@
 
-The key remaining challenge is determining when an EN is no longer suitable for training the
-current CPN. Expiring an EN at the right time accomplishes two things. First, it ensures
-that our current EN is trained properly for the current CPN; without that, the CPN training
-will become unstable. Second, expiring the EN is how we adapt our learning to the brain's
-non-stationarity. At some point, the brain will have changed sufficiently for our EN to be
+The key remaining challenge at that point is determining when an EN is no longer
+suitable for training the current CPN. Expiring an EN at the right time
+accomplishes two things. First, it ensures that our current EN is trained properly
+for the current CPN; without that, the CPN training will become unstable. Second,
+expiring the EN is how we adapt our learning to the brain's non-stationarity.
+At some point, the brain will have changed sufficiently for our EN to be
 out-of-date, and therefore requiring replacement.
@@
 these two metrics appear effective in ensuring we expire the EN at the right time.
+See Supplement \ref{sup:encpninter} for additional details.
 
@@
 brain co-adaptation, in order to understand its impact on the co-processor training.
-Finally, we perform one experiment where we attempt to simulate natural recovery from
-the lesion, prior to training the co-processor. In sum, our experiments are: \\
-
-\begin{tabular}{|l|l|c|c|}
+Additionally, we perform one experiment where we attempt to simulate natural recovery
+from the lesion, prior to training the co-processor. Our final experiment includes
+sensor drift, as well as co-adaptation. In sum, our experiments are: \\
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|c|c|c|}
 \hline
-\textbf{Index} & \textbf{Lesion}                                           & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ co-adaptation?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ pre-recovery?\end{tabular}}} \\ \hline
-1              & 50\% AIP loss                                             &                                                                                                 &                                                                                                \\ \hline
-2              & 50\% AIP loss                                             & X                                                                                               &                                                                                                \\ \hline
-3              & 50\% M1 loss                                              &                                                                                                 &                                                                                                \\ \hline
-4              & 50\% M1 loss                                              & X                                                                                               &                                                                                                \\ \hline
-5              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 &                                                                                                 &                                                                                                \\ \hline
-6              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 & X                                                                                               &                                                                                                \\ \hline
-7              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 & X                                                                                               & X                                                                                              \\ \hline
-\end{tabular} \\
+ & \textbf{Lesion}                                           & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ co-adaptation?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ pre-recovery?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ sensor\\ drift?\end{tabular}}}\\ \hline
+1              & 50\% AIP loss                                             &                                                                                                 &                                                                                                &                                                                                                 \\ \hline
+2              & 50\% AIP loss                                             & X                                                                                               &                                                                                                &                                                                                                 \\ \hline
+3              & 50\% M1 loss                                              &                                                                                                 &                                                                                                &                                                                                                 \\ \hline
+4              & 50\% M1 loss                                              & X                                                                                               &                                                                                                &                                                                                                 \\ \hline
+5              & 100\% connection loss F5\textless{}-\textgreater{}M1 &                                                                                                 &                                                                                                &                                                                                                 \\ \hline
+6              & 100\% connection loss F5\textless{}-\textgreater{}M1 & X                                                                                               &                                                                                                &                                                                                                 \\ \hline
+7              & 100\% connection loss F5\textless{}-\textgreater{}M1 & X                                                                                               & X                                                                                              &                                                                                                 \\ \hline
+8              & 100\% connection loss F5\textless{}-\textgreater{}M1 & X                                                                                               &                                                                                                & X                                                                                               \\ \hline
+\end{tabular}
+\end{table}
 
@@
 \begin{itemize}
-	\item The percent change in task loss between two consecutive 500 epoch windows is below a certain threshold: 0.01\%. This
+	\item The percent change in task loss between two consecutive 500 epochs is below a certain threshold: 0.01\%. This
 	      stopping criteria ensures we stop training when the economy of additional training is low.
@@
 
+\begin{landscape}
+\begin{figure}[hp]
+\centering
+\includegraphics[scale=1.0]{training_results.png}
+\caption{Loss and class separation vs training epoch}
+\label{fig:results}
+\end{figure}
+\end{landscape}
+
 \section{Results}
@@
 
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{training_results.png}
-\caption{Loss and class separation vs training epoch}
-\centering
-\label{fig:results}
-\end{figure}
-
-\begin{table}[]
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
-AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
-AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
-M1 No-coadapt  & 0.021136                                                          & 0.005765                                                         & 0.005604                                                             & 74.73\%            & 75.51\%                                                          \\ \hline
-M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
-Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
-Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
-Pre-recovery*  & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
-\end{tabular}
-\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
-\end{table}
-
 With each experiment, we track two metrics to characterize the learning progress. First, we
 track the task loss, which is an MSE loss that measures the co-processor's ability to restore
-movement towards the target trajectory. We characterize that measure in terms of its percent
-difference from lesioned and healthy performances.
+movement towards the target trajectory. We define it over the muscle length trajectory vector
+output from the simulated network, compared to the ground truth muscle length data captured
+during the Michaels experiment. In graphs, we characterize this loss measure in terms of its
+percent recovery - i.e. difference from lesioned and healthy performances. See Fig.
+\ref{fig:results}, and Table \ref{tab:results} for detailed results.
 
@@
 and the ability of the co-processor to condition the grasp based on visual information.
+To successfully grasp in the real world, the hand must be preformed appropriately for
+the shape of the object being grasped, and must close around it appropriately.
 The structure of the delayed reach-to-grasp task can be learned by the co-processor,
-i.e. to hold muscle velocities to 0.0 for the initial part of each trial, without it
+e.g. to hold muscle velocities to 0.0 for the initial part of each trial, without it
 also learning to differentiate the various object shapes from the brain observations.
-$S$ is our metric for differentiating those two co-processor behaviors.
+Roughly speaking, $S$ is our metric for differentiating overall decrease of task loss,
+and improvement in object differentiation.
 
@@
 condition its grasp for the object shape, though reaching largely still succeeds.
-Likewise, the co-processor doesn't have sufficient information to improve performance,
-for lack of that information.
+Note in Supplement \ref{sup:results} that lesioned loss is lower in this experiment than
+in others: the user continues to reach successfully, but not to form the hand properly, as
+explore above. Likewise, the co-processor doesn't have sufficient information to improve
+performance, for lack of that information.
 
@@
 though, that the simulation is not designed to indicate if the co-processor speeds up
-recovery, or results in greater results than natural recovery>
+recovery, or provides better results than natural recovery.
 
@@
 
-\subsection{Experiment 2: M1 75\% lesion}
-In this experiment, 75\% of the output module of the simulated network has been
+\subsection{Experiment 2: M1 50\% lesion}
+In this experiment, 50\% of the output module of the simulated network has been
 lesioned. In the co-adaptive case, the co-processor needs to adapt to
@@
 simulated recovery restoring the degrees of freedom needed for success. The
-co-processor in this later phase of training is tasked with adapting to the
-non-stationarity that results from the recovery process, which it does
+co-processor in this later phase of training must adapt to the
+non-stationarity that results from that recovery process, which it does
 successfully.
 
-\subsection{Experiment 3: F5 and M1 connectivity lesion}
+\subsection{Experiment 3: F5 and M1 connection lesion}
 As in the other cases, the co-processor's loss in this experiment improves
-quickly, initially, but taskes much longer to refine. During that refinement,
+quickly, initially, but takes much longer to refine. During that refinement,
 object differentiation occurs. Note that in this experiment, no information
@@
 In this case, co-adaptation only affects the M1 module, since we have nothing
-to use as a learning signal: there is no backpropagation to the AIP and
+to use as a learning signal: there is no ``backpropagation'' to the AIP and
 F5 modules. As a result, this experiment demonstrates that the
@@
 
-\subsection{Experiment 4: Connectivity lesion with pre-recovery}
+Object separation is initially low and unstable, as the co-processor begins
+to learn the task. Afterwards, it stabilizes as the co-processor begins to
+leverage observed brain activity to differentiate object shapes. In that
+phase of training, the object separation eventually begins convergence towards
+healthy activity.
+
+\subsection{Experiment 4: Connection lesion with pre-recovery}
 In this case, we simulate some amount of recovery prior to the co-processor
@@
 
+\subsection{Experiment 5: Connection lesion with sensor drift}
+Here we simulate sensor drift, using the same connection lesion above, and
+co-adaptation. As in the other connection lesion experiments, the co-processor
+learns the reach-to-grasp task, and forward-propagates visual information to
+condition the grasp. Notably, because the observation function is non-stationary
+in this experiment, the class separation exhibits far higher epoch-to-epoch
+variance in the later epochs. The co-processor presumably becomes more
+reliant on upstream visual information over time, as it learns how to map it
+on to the task. As a result: it becomes more sensitive to upstream
+non-stationarity as it learns. That is - earlier training epochs involve
+learning the task structure - i.e. delayed reach-to-grasp - but in a
+largely object-agnostic sense. Later, it further leverages the observations
+in order to differentiate object shapes. Note however that it continues to
+perform well in later epochs in terms of task loss.
+
+We present the results of this experiment without our stopping condition
+applied, due to the noise in the training causing stoppage while results
+are still improving. For this experiment, we opted instead to stop training
+after running the experiment long enough to get results near to the version
+of this experiment without sensor drift.
+
+Note also that training efficiency was, unsurprisingly, decreased relative to
+the experiment without sensor drift. For example, the sensor drift version
+required $282k$ training epochs to reach $85\%$ recovery, whereas the version
+without sensor drift required only $93k$.
+
 \section{Discussion and Future Work}
@@
 It achieved that as well, co-adapting with the simulated brain to improve external task
-performance. Finally, it adapted to a simulated brain which had already undergone some amount of
+performance. It also adapted to a simulated brain which had already undergone some amount of
 lesion recovery. In that case, the co-processor successfully identified the information upstream
-from the lesion which was necessary to stimulate the motor cortex downstream.
+from the lesion which was necessary to stimulate the motor cortex downstream. Finally, it was
+able to adapt to simulated sensor drift, though with a penalty to training efficiency.
 
 Clearly, we need to discuss carefully what can and cannot be inferred from simulation studies
-such as these. As in Dura-Bernal \cite{bernal.sim}, we note that our simulation differs
-drastically from an organic brain, both in scale and structure. Today, predicting long-running
+such as these. An example of such a simulation approach is Dura-Bernal et al.
+\cite{bernal.sim}. In this work, the authors use a simulated spiking neural network to train
+a stimulation agent. Their stimulation agent sought to restore the network's
+control of a simulated arm, to reach a target, after a simulated lesion was applied,
+much like the work here. As the authors point out, we have a limited ability to
+probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
+we first need to design our approach through the use of an admissible simulation.
+The authors in this case simulated lesions by effectively removing parts of their
+simulated network, or by cutting connections between parts of the network.
+Here, we use an established design for an artificial neural circuit which
+performs a grasping task. Its architecture and training methods were designed
+specifically to result in naturalistic dynamics.
+
+As in Dura-Bernal \cite{bernal.sim}, we note that our simulation differs drastically from an
+organic brain, both in scale and structure. Today, predicting long-running
 neural responses to stimulation remains a difficult problem, implying targeted neural control
@@
 \section{Acknowledgements}
-TODO: funding sources
-
-The author would like to thank the following for discussions and insights related to
+This work was supported by a Weill Neurohub Investigator grant, a CJ and Elizabeth Hwang
+endowed professorship (RPNR), and National Science Foundation grant no. EEC-1028725.
+
+The authors would like to thank the following for discussions and insights related to
 the present topic:
@@
 	\item Karunesh Ganguly, UC San Francisco
-	\item Priya Khanna, UC San Francisco
+	\item Priya Khanna, UC Berkeley
 	\item Anca Dragan, UC Berkeley
@@
 
+\newpage
+
 \section{References}
@@
 \bibliography{refs}
+
+\newpage
+
+\section{Supplement}
+\subsection{EN/CPN Interleaving}
+\label{sup:encpninter}
+We train a CPN until an EN expires. At that point we train a new EN. Expiration
+occurs when one of a number of predicates are true:
+
+\begin{itemize}
+	\item The EN prediction loss is greater than $min(6e^{-4}, L/10)$, where $L$ is the most recent
+	      task loss. EN prediction loss is an MSE loss between the EN's prediction of muscle
+	      trajectories, and the actual muscle trajectories output by the mRNN.
+	\item Task loss increased across 15 of the most previous 30 training epochs.
+	\item 100 CPN training epochs have elapsed.
+\end{itemize}
+
+\subsection{Sensor drift}
+\label{sup:drift}
+To simulate sensor drift, we add a vector to a bias term introduced to the observation function.
+We draw the elements of the vector from a mean-0 Gaussian distribution, with variance based
+on the mean value of the observation function from prior connection-lesion experiments. That
+allows us to put it into a reasonable range, where it is effective but not extreme. We attempted
+the experiment with several values of the variance, and found the results to be principally the same:
+the co-processor eventually learned, but at a rate slower than otherwise. We present results for
+a variance of $1.5e^{-3}$, which is $2\%$ of the mean observation value.
+
+\subsection{Results detail}
+\label{sup:results}
+
+\begin{table}[h]
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
+AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
+AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
+M1 No-coadapt  & 0.021136                                                          & 0.005765                                                         & 0.005604                                                             & 74.73\%            & 75.51\%                                                          \\ \hline
+M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
+Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
+Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
+Pre-recovery*  & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
+Sensor drift   & 0.020719                                                          & 0.002966                                                        & 0.002825                                                              & 88.10\%            & 88.83\%                                                          \\ \hline
+\end{tabular}
+\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
+\end{table}
+
 \end{document}

### 2022-10-22 paper/article.tex
JNE paper

--- 2021-12-19 paper/article.tex
@@
 \usepackage{subcaption}
+\usepackage{lscape}
+\usepackage{hyperref}
 \graphicspath{ {./figs/} }
@@
 
-\title[Modeling a Neural Co-Processor]
-{Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept}
-
+\title[Neural Co-Processors]
+{Neural Co-Processors for Restoring Brain Function:\\Results from a Cortical Model of Grasping}
+
+% Authorship temporarily removed for double-anonymous review
 \author{Matthew J Bryan$^{1}$, Linxing Preston Jiang$^{1}$, Rajesh P N Rao$^{1}$}
 
-\address{$^{1}$ Neural Systems Laboratory, Department of Computer
-Science and Engineering, University of Washington, Box 352350,
+\address{$^{1}$ Neural Systems Laboratory, Paul G. Allen School of Computer
+Science \& Engineering, University of Washington, Box 352350,
 Seattle, WA 98105, USA}
 
-\ead{matthew.bryan@dell.com}
+\ead{\{mmattb,prestonj,rao\}@cs.washington.edu}
 \vspace{10pt}
 \begin{indented}
-\item[]September 2021
+\item[]November 2022
 \end{indented}
@@
 \begin{abstract}
-\textit{Objective} Brain co-processors \cite{rao.coproc} are devices which use artificial
-intelligence (AI) for closed-loop neurostimulation, to shape neural activity and to bridge
-injured neural circuits for targeted repair and rehabilitation. The co-processor framework
-offers a flexible approach to learning closed-loop stimulation policies that optimize for
-(a) specific regimes of neural activity, or (b) external task performance.
-For example, one may seek to learn to stimulate the motor cortex of a stroke patient,
-conditioning the stimulation on upstream visual information, aiding the patient's
-attempt to grasp an object. Through the use of artificial neural
-networks (ANNs) and deep learning, the co-processor co-adapts with the
-neural circuit, allowing it to seek optimal stimulation policies, and adapt them
-as the neural circuit changes. The results presented here demonstrate a
-neural co-processor for the first time, through the use of a simulation. We
-explore some of the core algorithms that may allow co-processors to successfully
-learn, and to adapt to non-stationarity in both the brain and sensors.
-\textit{Approach} We provide the first proof-of-concept of a neural co-processor that
-leverages deep learning, through the use of a simulated neural circuit.
-That circuit performs a reach-to-grasp task, based on visual input, and is designed to
-closely resemble a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate
-a variety of lesions by altering the model, and then demonstrate our co-processor's ability
-to restore lost function through ``stimulation'' of that model. We further test the
-ability of our co-processor to adapt its stimulation as the simulated brain undergoes changes.
-\textit{Main results} Our simulated co-processor successfully co-adapts with the neural
-circuit to accomplish the external reaching task. The co-processor framework
-demonstrated here adapts to a variety of lesion types, and to ongoing changes in the
-simulated brain.
-\textit{Significance} The proof-of-concept here outlines a co-processor model, as well
-as our approach to training it, leading to insights on how such a model may be
-developed for \textit{in vivo} use. We believe this co-processor design has potential to
-learn complex stimulation policies, leading to application in treating a variety of
-diseases.
+\textit{Objective} A major challenge in closed-loop brain-computer interfaces (BCIs)
+is finding optimal stimulation patterns as a function of ongoing neural activity for
+different subjects and for different objectives. Traditional approaches, such as
+those currently used for deep brain stimulation (DBS), have largely followed a manual
+trial-and-error strategy to search for effective open-loop stimulation parameters,
+a strategy that is inefficient and does not generalize to closed-loop
+activity-dependent stimulation.
+\textit{Approach} To achieve goal-directed closed-loop neurostimulation, we propose
+the use of brain co-processors, devices which exploit artificial intelligence (AI)
+to shape neural activity and bridge injured neural circuits for targeted repair and
+rehabilitation.  Here we investigate a specific type of co-processor called a ``neural
+co-processor'' which uses artificial neural networks (ANNs) and deep learning to learn optimal
+closed-loop stimulation policies. The co-processor adapts the stimulation policy as
+the biological circuit itself adapts to the stimulation, achieving a form of brain-device
+co-adaptation. We tested the neural co-processor's ability to restore function after
+stroke by simulating a variety of lesions in a previously published cortical model
+of grasping.  We additionally tested the ability of our co-processor
+to adapt its stimulation as the simulated cortical network and simulated sensors
+underwent changes.
+\textit{Main results} Our results show that a neural co-processor can restore
+reaching and grasping function after a simulated stroke in a cortical model, achieving
+recovery towards healthy function in the range 75-90\%. Our co-processor successfully
+co-adapted to accomplish the reach-and-grasp task after a variety of lesions to the
+simulated cortical circuit.
+\textit{Significance} Our results provide the first proof-of-concept demonstration,
+using computer simulations, of a neural co-processor for activity-dependent closed-loop
+neurosimulation for optimizing a rehabilitation goal after injury. Our results provide
+insights on how such co-processors may eventually be developed for \textit{in vivo} use
+to learn complex adaptive stimulation policies for a variety of neural rehabilitation
+and neuroprosthetic applications. 
 \end{abstract}
@@
 \vspace{2pc}
-\noindent{\it Keywords}: brain-computer interface, neural co-processor, ai, machine learning, stimulation
+\noindent{\it Keywords}: brain-computer interface, neural co-processor, AI, machine learning, deep learning, neural networks, stimulation, computational models
 %
@@
 \section{Introduction}
-Aided in part by application of advanced AI techniques, brain-computer interfaces (BCIs) have made
-advancements over the last several decads, allowing for decoded brain signals to be used for
-control of a wide variety of virtual and physical prostheses \cite{rao.bcibook, wolpaw.bcibook,
-moritz.neuro, lebedev.bmi}. Separately: advances in stimulation techniques and modeling have allowed
+Brain-computer interfaces (BCIs) have made significant advances over the last several decades, leading
+to the control of a wide variety of virtual and physical prostheses through neural signal decoding
+\cite{rao.bcibook, wolpaw.bcibook, moritz.neuro, lebedev.bmi}. Separately, advances in stimulation techniques and modeling have allowed
 us to probe neural circuit dynamics (e.g. \cite{walker.inception}) and learn to better drive neural
-circuits towards target dynamics, by encoding and delivering information through
+circuits towards desired target dynamics by encoding and delivering information through
 stimulation \cite{niparko.cochlear, weiland.retinal, tomlinson.propr, tabot.tact, tyler.tact,
-dadarlat.tact, sharlene.tact, cronin.tact}.
-Recently, there has been increasing interest in building on these advances to combine decoding
-and encoding in a single system, for closed-loop stimulation of a neural circuit. Bi-directional
-BCIs (BBCIs) allow stimulation to be conditioned by decoded brain activity as well as external
-sensor (e.g. camera) data, which can allow for the application of real-time, fine-grained control of
-neural circuits and prosthetic devices, e.g. Nicolelis et al. \cite{nicolelis.bmbi}. These may lead,
-for example, to neuro-prostheses that are capable of restoring movement which was lost due to
-traumatic brain injury (TBI), to a degree not previously possible.
-
-Motivated by that progress, we demonstrate here a flexible framework for combining encoding
-and decoding, which we term ``neural co-processors'' \cite{rao.coproc}. Neural co-processors leverage
-AI and deep learning to identify optimal, closed-loop stimulation patterns. The approach is flexible
-enough to optimize not only for particular neural activities, but also for tasks external to the
-subject. For example, they may be able to aid a stroke patient by finding a stimulation pattern of
-the motor cortex which helps restore lost limb function. Likewise, the framework generalizes enough
-to condition stimulation on both brain activity, and external sensors, e.g. cameras or light detection
-and ranging (LIDAR) sensors, in order to incorporate feedback for realtime control.
-
-Additionally, the co-processor framework allows a neuro-prosthesis to actively adapt to
-a neural circuit as it changes with time. This framework is capable of co-adapting with
-the circuit, i.e. brain, by updating its stimulation regime, while at the same time the
-brain is updating its response to the stimulation, and changing due to natural plasticity
-and aging. This allows the co-processor to continually optimize for the intended cost
-function, despite the signficant non-stationarity of the target circuit.
-
-Here we provide a proof-of-concept in simulation for a co-processor that restores
-movement to a limb, after a subject has suffered a stroke affecting their ability to
-use that limb. It combines:
+dadarlat.tact, sharlene.tact, cronin.tact}. Bi-directional BCIs (BBCIs) allow stimulation to be
+conditioned on decoded brain activity and encoded sensor data for applications such as real-time,
+fine-grained control of neural circuits and prosthetic devices (e.g., \cite{nicolelis.bmbi}).
+
+Motivated by these advances, we investigate here a flexible framework for combining encoding
+and decoding using ``neural co-processors'' \cite{rao.coproc}, a type of brain co-processor \cite{rao.braincoproc}. Neural co-processors leverage artificial neural networks (ANNs) and deep
+learning to compute optimal closed-loop stimulation patterns. The approach can be used to not only drive neural activity
+toward desired activity regimes, but also to achieve task goals external to the subject, such as finding closed-loop
+stimulation patterns for motor cortical neurons for restoring the ability to reach and grasp an object. Likewise, the
+framework generalizes to stimulation based on both brain activity and external sensor measurements, e.g., from cameras or
+light detection and ranging (LIDAR) sensors, in order to restore perception (e.g., cortical visual prosthesis) or
+incorporate feedback for real-time prosthetic control (see \cite{rao.braincoproc} for details).
+
+The co-processor framework also allows co-adaptation with biological circuits in the brain by updating its stimulation policy, while the brain updates its own response to the stimulation via adaptation and neural plasticity, or updates due to other reasons.
+This allows the co-processor to continually optimize its outputs for the desired optimization function in the presence of significant non-stationarity in the brain.
+
+Here we use computer simulations to demonstrate a neural co-processor that restores
+movement in a computational model of cortical networks controlling a limb, after a simulated stroke affects the ability to
+use that limb. Our demonstration combines:
 \begin{itemize}
-	\item A stimulation model, which models the relationship between decoded brain activity,
-	      stimulation, and task performance.
-	\item An AI agent which determines the stimulation to apply in a closed-loop fashion, in real time.
+	\item An emulation model based on ANNs, which models the relationship between 
+	      stimulation, decoded brain activity, and task performance.
+	\item An AI agent based on ANNs which determines the best stimulation to apply in a closed-loop fashion in real time.
 \end{itemize}
@@
 \section{Background}
-Significant advances have been made in modeling the effects of electrical stimulation
-of the brain, some of which can be leveraged for our co-processor design, as we outline
-below. Researchers have explored how information can be biomimetically or
+Significant advances have been made in understanding and modeling the effects of electrical stimulation
+on the brain. Researchers have explored how information can be biomimetically or
 artificially encoded and delivered via stimulation to neuronal networks in the brain and
@@
 \cite{tabot.tact, tyler.tact, dadarlat.tact, sharlene.tact, cronin.tact} perception.
-Advancements have also been made in modeling the effects of stimulation over large scale, multi-region
-networks, and across time \cite{shanechi.stimmodel}. Some have additionally designed models which
-can adapt to ongoing changes in the brain, including changes due to the stimulation itself
-\cite{tafazoli.acls}. In our proof-of-concept outlined below, we will use a stimulation
-model, not unlike those cited here, which seeks to account for both network dynamics
+Advances have also been made in modeling the effects of stimulation over large scale, multi-region
+networks, and across time \cite{shanechi.stimmodel}. Some models can additionally adapt to ongoing
+changes in the brain, including changes due to the stimulation itself
+\cite{tafazoli.acls}. For our simulations described below, we use a stimulation
+model, not unlike those cited above, which seeks to account for both network dynamics
 and non-stationarity. In addition to training the model to have a strong ability to predict
-the effect of stimulation, we additionally train it to be useful for then learning an
-optimal stimulation policy, which is a property somewhat distinct from predictive
+the effect of stimulation, we additionally adapt it to be useful for learning an
+optimal stimulation policy, a property distinct from predictive
 power alone.
 
-Advances have also been made in both open- and closed-loop stimulation for
+Researchers have also explored both open- and closed-loop stimulation protocols for
 treating a variety of disorders. Open loop stimulation has been effective in
@@
 psychiatric disorders \cite{holtzheimer.psy, kisely.psy, fraint.psy}.
-More directly related to this paper, we see in Khanna et al. \cite{khanna.openloop},
-the use of open loop stimulation in restoring dexterity after a lesion
-occurs affecting a primate's motor cortex. The authors demonstrate that
-the use of low-frequency alternating current, applied epidurally,
+In research more directly related to our work, Khanna et al. \cite{khanna.openloop}, investigated 
+the use of open loop stimulation in restoring dexterity after a lesion in a nonhuman primate's (NHP's) motor cortex.
+The authors demonstrate that the use of low-frequency alternating current, applied epidurally,
 can improve grasp performance.
 
-While open loop stimulation techniques have yielded clinically useful results,
-their results in many domains have been mixed, such as use in visual
-prostheses \cite{bosking.visual}, and use in invoking somatosensory feedback
-\cite{cronin.tact}. Likely this is due to the stimulation not being conditioned
-on the ongoing dynamics of the circuit being stimulated. Moment-to-moment and
-throughout the day, the circuit will respond differently to the same stimulus,
-as a result of differing inputs and ongoing activity. Stimulation
+While open loop stimulation techniques have yielded clinically useful results, results in other domains have been mixed, such as in visual
+prostheses \cite{bosking.visual}, and in invoking somatosensory feedback
+\cite{cronin.tact}. We believe this is due to the stimulation not being conditioned
+on the ongoing dynamics of the neural circuit being stimulated. From moment to moment and
+throughout the day, a neuronal circuit in the brain can be expected to respond differently even when the same stimulation parameters are used,
+due to the multitude of different external and internal inputs influencing the circuit's ongoing activity. Stimulation
 therefore needs to be proactively adapted in response. This need is even
-greater over longer time scales as the effects of plasticity and ageing change
-the connectivity of the brain. Closed-loop stimulation may also provide means
+greater over longer time scales as the effects of plasticity, changes in clinical conditions, and ageing change
+the dynamics and connectivity of the brain. Closed-loop stimulation may also provide means
 to better regulate the energy use of an implanted stimulator, allowing it to
 intelligently regulate when to apply stimulation, in order to preserve implant
-battery life. Closed-loop stimulation also offers an opportunity minimize
+battery life. Another benefit is that closed-loop stimulation offers an opportunity to minimize the 
 side-effects of stimulation, through real time regulation
 of the stimulation parameters, such as in the use of deep brain stimulation (DBS) in
-PD patients \cite{little.park}.
-
-Closed-loop stimulation conditions stimulation on observations of brain activity,
-allowing it to shape the neural activity more precisely, and in response to changes
-in a patient's clinical condition. This may also allow an implant to adapt
-in real time to changes in the target circuit. It has been used to aid in learning
+PD patients \cite{little.park}. In recent years, closed-loop stimulation has been used to aid in learning
 new memories after some impairment \cite{berger.closedloop, kahana.biomarker},
@@
 
-Something that remains unclear is how to leverage closed-loop control for real-time
-co-adaption with the brain to accomplish an external task. ``Co-adaption'' here refers
-to the ability of a neuro-prosthesis to adapt its stimulation regime to the ongoing
+A major open question is: how does one leverage closed-loop control for real-time
+co-adaptation with the brain to accomplish an external task? ``Co-adaptation'' here refers
+to the ability of a BCI to adapt its stimulation regime to the ongoing
 changes in the circuit it is stimulating, and to adapt with that circuit to accomplish
-the external task, such as grasping. The neural co-processor we present here provides one
-potential model for accomplishing that. Through the use of deep learning, the
-co-processor model we present co-adapts an AI agent, which governs the
-stimulation, with both a stimulation model, and the neural circuit being stimulated.
-
-For a neurologically complex task such as grasping, we cannot identify
-\textit{a priori} a real time controller of the neural circuit. That is
+the external task (e.g., grasping). The neural co-processor we present here provides one
+potential model for accomplishing that. Through the use of deep learning, a neural 
+co-processor model co-adapts its AI, which governs the
+stimulation, with the neural circuit being stimulated.
+
+For a neurologically complex task such as grasping, it is not possible to design
+\textit{a priori} a fixed real-time controller for the (potentially impaired) neural circuits involved in the task. That is
 due in large part to the variability of circuits from subject-to-subject,
 as well as variations in the placement of sensors and stimulators in the
-brain. The only plausible path to such a real time controller is to parameterize
-it in a subject- and time-specific way. Our model seeks to accomplish that using
-deep learning.
-
-Before attempting \textit{in vivo} experiments using such a model, we first demonstrate
-here a number of crucial design elements of it, through the use of a simulated
-grasping circuit, presented previously by Michaels et al. \cite{michaels.mrnn}. We
-explore the properties of the co-processor design that are needed to
-successfully adapt to the long-running dynamics of a stimulated neural circuit,
-as well as to adapt to that circuit's ongoing connectivity changes.
-We additionally present a training method for our co-processor, allowing it to learn
-an optimal stimulation pattern that drives improvements in external task performance,
+brain. The only plausible path to implementing a real-time controller is therefore to allow the device to adapt to the subject,
+long-term changes in their brain activities, and variability in the sensors, stimulators and hardware. Our proposed neural
+co-processors seek to accomplish such adaptation through ANNs and deep learning.
+
+To gain insights into neural co-processors before testing them in \textit{in vivo} experiments, we investigated a number of crucial design elements through the use of a neural network model of the cortical areas involved in
+grasping, presented previously by Michaels et al. \cite{michaels.mrnn}. We
+explored what properties of the co-processor allow
+successful adaptation to the short-term dynamics of the cortical model as it is being stimulated,
+as well as adaptation to longer-term connectivity changes in the cortical model.
+We present a training method for neural co-processors for learning optimal stimulation patterns that drive improvements in external task performance,
 while also adapting to the non-stationarity of the brain.
 
-\section{Method}
+\section{Methods}
 
 \begin{figure}
+    \centering
+    \begin{subfigure}[c]{0.75\textwidth}
+		\centering
 		\includegraphics[width=\textwidth]{weill_arch.png}
-\caption{Using a co-processor to drive external task performance after a traumatic brain injury}
-\centering
-\label{fig:weill}
+		\caption{}
+	\end{subfigure}
+	\hfill
+    \begin{subfigure}[c]{0.75\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.png}
+		\caption{}
+	\end{subfigure}
+	\hfill
+    \caption{{\bf Neural co-processor for restoring function after a brain injury}.
+    An artificial neural network called the “Co-Processor Network” (CPN) is used to
+    map input neural activity patterns in an area A to output stimulation patterns
+    in same or other areas B in order to achieve a neural or behavioral goal using
+    another ANN, an “Emulator Network” (EN) - see text for details. (a) The example
+    here shows the CPN creating a new information processing pathway between prefrontal
+    cortex and motor cortex, bypassing an intermediate area affected by brain injury
+    (e.g., stroke). Adapted from \cite{rao.coproc}, (b) Our current study involves a
+    simulated cortical grasping circuit. The CPN and EN both receive simulated brain
+    recordings, which are created according to a recording model (see Methods).
+    The CPN outputs stimulation parameters, which are applied to the simulated
+    grasping circuit, according to a stimulation model. The EN models the relationship between the
+    stimulation, brain recordings, and external task.}
+    \label{fig:arch}
 \end{figure}
@@
 \subsection{Architecture Overview}
-First, we present the architecture of our co-processor design. This design aims to solve two
+First, we present the architecture of our neural co-processor design. This design aims to solve two
 fundamental challenges in using neural stimulation to improve external task performance.
-First, neural networks exhibit long-running and nonlinear dynamics, necessitating
-the need for a stimulation agent to account for far-distant effects of the stimulation it
-applies. Second, with neurologically complex tasks, such as grasping, the mapping between
-neural activity as-measured and the external task cannot be determined \textit{a priori}.
-As a result, the co-processor must somehow learn what stimulation is appropriate for
-aiding the user in the external task they are attempting to perform.
-
-Our co-processor attempts to solve these with a pair of artificial neural networks:
+First, to restore function for complex tasks, such as grasping, the mapping between
+neural activity, such as the activity encoding the intention to grasp, and the stimulation patterns
+to be applied to a downstream motor area to enable grasping cannot be pre-determined.
+As a result, the co-processor must learn what stimulation pattern is appropriate for achieving the
+external task. Unfortunately, in most cases, we do not know the optimal stimulation patterns to train
+the co-processor. This is because stimulation shapes the nonlinear dynamics of circuits in the brain
+in complex ways, leading to complex effects in external behavior. It is therefore not obvious what
+stimulation patterns will produce a desired behavior.
+
+A neural co-processor attempts to solve these problems with a pair of artificial neural networks (ANNs)
+(Fig.~\ref{fig:arch}):
 \begin{itemize}
-	\item A stimulation and neural dynamics model, known as an ``Emulator Network'' (EN). It models the relationship
-	      between the stimulation, neural dynamics, and external task. Its purpose is for training the second network.
-	\item A stimulation agent, known as the ``co-processor network'' (CPN), which maps neural activity, and possibly
-	      data from external sensors, to stimulation parameters.
+    \item a ``Co-Processor Network'' (CPN), which maps neural activity, and possibly
+	      data from external sensors, to appropriate stimulation parameters.
+    \item an ``Emulator Network'' (EN) which models the effect of stimulation on neural dynamics and
+          behavior for the external task.
 \end{itemize}
 
-We co-train the EN and CPN, with the goal of training a CPN whose output stimulation parameters
-improve task performance. By continually training both, they adapt to the brain as it changes,
-effectively allowing for brain-stimulator co-adaptation. The EN is a tool for training the CPN,
-giving us a way to backpropagate task error to the CPN. It outputs task-relevant
-metrics - a prediction of muscle velocities in our case - given measurements of neural
-activity, and the stimulation parameters output from the CPN. If the EN is trained in a
-particular way, and to a sufficient level of precision, it can be used as a function
-approximator relating stimulation and a task, and do so in a way that allows us to train
-the CPN with it. When training the CPN, we in-effect treat the EN's output as the true
-task performance, or a related metric, and then backpropagate the loss defined in terms
-of that metric in order to train the CPN. See Fig. \ref{fig:weill}. We will illustrate the details
-of the training algorithm below.
-
-In our present demonstration, the EN is constituted as a single layer, fully connected,
-long short-term memory (LSTM) recurrent neural network (RNN), with hyperbolic tangent ($tanh$)
-activations, and a linear readout. The general notion of a co-processor does
-not require this precise architecture, but for our example, we found that the LSTM
-approach allows the network to continuously adapt to long-running dependencies in
-the simulated neural dynamics, far better than a vanilla RNN. The CPN is constituted
-as an almost identical network, though with a different dimensionality, as we explain
-below. There is no strict requirement for the EN and CPN to be so similar, but we
-found this simple architecture to work well in our example.
-
-Note that the EN effectively constitutes a stimulation model. However, in the case that
-a co-processor attempts to optimize for task performance, as opposed to neural
-activity directly, the stimulation model predicts the effect of stimulation and neural
-dynamics on task performance, rather than modeling and predicting the neural response
-to stimulation alone. This provides the key functionality needed to train the CPN.
-
-We initially attempted an EN architecture using vanilla RNN cells, with a nonlinearity.
-This approach resembles the common linear time-invariant state space model of
-stimulation, as in e.g. Yang et al. \cite{shanechi.stimmodel}, though it additionally
-incorporates a nonlinearity. That approach has the benefit of being simple, and familiar.
-We found, however, that the linear approach, as well as the related vanilla RNN
-approach with nonlinearities added, was not sufficiently expressive of long-term
-dependencies, such that the co-processor was capable of learning well.
-
-Separately, nonlinear LSTM models have also been applied to modeling natural
-neural networks, such as use in predicting local field potentials
-\cite{kim.lstm}, and for stimulation modeling \cite{guclu.lstm}.
-LSTM cells are designed to better capture the relevant aspects of
-long-running network dynamics. We found that LSTM cells were crucial for
-enabling our co-processor to learn the long-running dynamics of the network
-it was stimulating, and we use them as a result.
+The CPN can be trained using the backpropagation algorithm, the workhorse of deep learning for training ANNs.
+However, backpropagation requires the error between the output of the CPN and a desired output, and as
+discussed above, we do not have the desired output stimulation pattern. We do however know what the desired
+output behavior in a task should be, e.g., a particular kind of grasp for a particular object or a particular
+type of neural activity in a brain area corresponding to healthy activity. We can therefore compute the error
+between this desired behavior and the actual behavior caused by the CPN due to stimulation. How do we
+backpropagate this error, which is external to the CPN, to update the parameters of the CPN?
+
+We use a trained EN to backpropagate the external task error to the CPN. We train the EN to predict
+task-relevant parameters - a prediction of muscle velocities in our case - given the stimulation
+parameters output by the CPN and any measured neural activity. If the EN is trained to a sufficiently
+high level of precision, it can be used as a function approximator mapping stimulation to task output.
+When training the CPN, we treat the EN's output as the actual task output (e.g., actual grasp
+parameters).  We backpropagate through the EN (without changing its weights) the error between
+the EN's output and the desired task output and use this backpropagated error to train the
+CPN (see Fig. \ref{fig:arch}).
+
+In our experiments, the EN was a single layer fully connected long short-term memory (LSTM) recurrent
+neural network (RNN), with hyperbolic tangent ($tanh$) activations, and a linear readout. It has 87 LSTM neurons,
+which was chosen somewhat arbitrarily as a function of the input and output vector sizes. We found varying this neuron
+count did not drastically change results. Although other architectures could also be used, we found that this
+LSTM architecture allows the EN to continuously adapt to long-running dependencies in the simulated neural dynamics, 
+ar better than a vanilla RNN. The CPN had an almost identical architecture, but with 61 LSTM neurons. As with the CPN,
+this neuron count was chosen as a function of the input and output vector sizes, but varying it upward had little effect on results.
+There is no requirement for the EN and CPN to have similar network architectures, but we found that these choices worked
+well in our experiments.
+
+Note that the EN is more general than traditional models of neurostimulation which attempt to predict the effects of
+stimulation on \em{neural activity}. The EN in a neural co-processor predicts the effects of stimulation (taking into
+account ongoing neural dynamics) on \em{task performance}. This provides the key functionality needed to train the
+CPN. In the special case where the task involves driving neural circuits in the brain to desired neural activities,
+the EN reduces to more traditional models of stimulation.
+
+For comparison, consider an EN architecture based on a traditional RNN with a nonlinearity. This is a nonlinear version
+of the common linear time-invariant state space model of stimulation (see, e.g.,  \cite{shanechi.stimmodel}). 
+We found, however, that compared to LSTM-based EN, the linear model and the traditional nonlinear RNN are both not
+sufficiently powerful to capture the long-term dependencies in stimulation effects needed for the CPN to learn well.
+Our use of an LSTM-based EN builds on previous work modeling biological neural networks for predicting local field potentials
+\cite{kim.lstm}, and modeling stimulation \cite{guclu.lstm}.
 
 \subsection{Simulation Overview}
-We demonstrate our approach here using a simulated grasping circuit.
-A detailed simulation such as this allows us to explore some of the critical
-architectural details and training algorithms that will allow a co-processor
-to succeed.  By first doing such exploration in simulation, we are able to
-rapidly and cheaply iterate on our design, prior to any \textit{in vivo}
-experiments. For the simulation to be admissible, we need to ensure it has some
-properties that allow it to strongly indicate if our design is improving in
-a direction that will later allow for real deployments. Otherwise, our
-design may be adapting to the pecularities of the simulation, without
-becoming more useful.
-
-An example of such a simulation approach is Dura-Bernal et al. \cite{bernal.sim}. In this work, the
-authors use a simulated spiking neural network to train a stimulation agent.
-Their stimulation agent sought to restore the network's
-control of a simulated arm, to reach a target, after a simulated lesion was applied,
-much like the work here.  As the authors point out, we have a limited ability to
-probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
-we first need to design our approach through the use of an admissible simulation.
-The authors in this case simulated lesions by effectively removing parts of their
-simulated network, or by cutting connections between parts of the network.
-Here, we use an established design for an artificial neural circuit which
-performs a grasping task. Its architecture and training methods were designed
-specifically to result in naturalistic dynamics.
-
-The simulated circuit, from Michaels et al. \cite{michaels.mrnn}, was trained
-to resemble the grasping circuits of monkey subjects engaged in a delayed reach-to-grasp task.
-Its design draws on a body of literature focused on architectures and training methods for RNNs which
-seek to create artificial neural circuits that have activation dynamics similar to natural circuits,
-including delayed grasping tasks \cite{susillo.mrnn}. The Michaels circuit consists of a
-``modular'' vanilla RNN (mRNN), and a linear readout layer. Each ``module'' consists of 100 vanilla
+To test the feasibility of the neural co-processor approach, we used a previously published model of a
+grasping circuit in a nonhuman primate (NHP) brain. Using such a model allows us to explore some of the critical
+architectural choices and training algorithms, enabling us to rapidly and cheaply iterate on our
+design, prior to any \textit{in vivo} experiments.
+
+In the present study, we use the grasping network model of Michaels et al. \cite{michaels.mrnn}
+(Fig.~\ref{fig:michaels}a). The network, representing multiple interconnected cortical areas, was trained
+to mimic the grasping circuits of NHP subjects engaged in a delayed reach-to-grasp task.
+The model's design draws on a body of literature focused on architectures and training methods for RNNs which
+seek to create artificial neural networks with activation dynamics similar to biological circuits,
+including circuits for delayed grasping tasks \cite{susillo.mrnn}. The model consists of three
+``modular'' vanilla RNNs (mRNNs), representing the cortical areas AIP, F5 and M1 respectively, and a linear
+readout layer (Fig.~\ref{fig:michaels}a). Each ``module'' consists of 100 vanilla
 RNN neurons, with a nonlinearity applied on the outputs. The modules are internally fully connected,
-and are connected to each other sparsely (10\% connectivity). The inputs are visual features
-intended to capture the view the monkeys had during the task, specifically VGGNet features \cite{simonyan.vgg},
-from 3D renderings of the same objects which the monkeys grasped. The outputs are muscle length
-velocities for the shoulder, arm, and hand of the monkey during the trial. The natural
-velocities were captured with a motion capturing glove, and the artificial neural network
-was trained to recapitulate those grasping motions. Data and trained models from this work
-were supplied to us by the lead author Jonathan Michaels, for the purpose of our present
-simulation. We re-implemented his model's logic in PyTorch, and loaded his trained parameters
-into it, for one of his subjects, arbitrarily chosen. See Fig. \ref{fig:michaels}.
+and are connected to each other sparsely (10\% connectivity in our case). The inputs to the network are visual features
+representing the object to be grasped, as well as a hold signal. The visual features
+intend to capture the features represented in the subject's visual cortex. They were extracted using
+VGGNet \cite{simonyan.vgg} from 3D renderings of the same objects which the subjects grasped. The hold signal is
+a boolean which encodes the point in the experiment when the subject began to reach. The outputs of the
+network are muscle length velocities for the shoulder, arm, and hand of the subject. The actual
+velocities were captured with a motion capture glove, and the Michaels et al. network model 
+was trained to recapitulate those grasping motions.\footnote{Data and trained models from this work
+were supplied to us by the lead author of \cite{michaels.mrnn}. We re-implemented their model in PyTorch,
+and used their trained parameters for an arbitrarily chosen subject.}
 
@@
 		\centering
-		\includegraphics[width=\textwidth]{michaels.eps}
-		\caption{Michaels Modular RNN (mRNN), a simulated grasping circuit. The emergent dynamics of the
-		three modules correspond well to natural neural activity measured from primate AIP, F5, M1 regions,
-		respectively, from the same task. Visual (VGGNet) features forward propagate through the
-		network, conditioning the grasp for the object of the particular size, shape, and location.}
+		\includegraphics[width=\textwidth]{michaels.png}
+		\caption{}
 	\end{subfigure}
@@
 		\includegraphics[width=\textwidth]{mRNN_J.png}
-		\caption{Connectivity matrix $J$. Note the within-module connections
-		(on the diagonal) are fully connected, and connections to adjacent
-		modules are sparse $10\%$.}
+		\caption{}
 	\end{subfigure}
 	\hfill
-\caption{Architecture of the Michaels mRNN}
+\caption{\textbf{Architecture of the Brain Model.} (a) Modular RNN (mRNN) used by Michaels et al.
+        \cite{michaels.mrnn} to model cortical circuits involved in grasping objects.
+        The emergent dynamics of the three modules correspond well to neural activity in
+        primate cortical areas AIP, F5, M1, respectively, for the same task. Visual
+        features of an object (derived from VGGNet) propagate forward through the
+		network, conditioning the grasp on the object's size and shape. (b)
+		Connectivity matrix $J$. Connected neuron pairs are indicated in black, though the
+		actual network weights are floating point values. White indicates non-connections.
+		Note the within-module connections (black squares along	the diagonal) are fully
+		connected while connections to adjacent modules are sparse
+		($10\%$ of possible connections).}
 \label{fig:michaels}
@@
 
-The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
-representing the visual processing needed to reach the hand to the appropriate position for
-the grasp, and to form the hand properly for grasping the particular shape of object. The
-emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
-to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
-refers to the fact that the activity of the module receiving the visual inputs resembles
-the AIP activity of the monkey subject from the same trials. Likewise, the second and
-third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
-The emergent network dynamics show a number of other correspondences to the monkeys' natural
-brain activity as well, as detailed in the paper. For convenience, we will refer to these
-modules throughout this paper using the brain region they correspond to.
-
-Importantly, the simulated circuit's activity shows a relatively clear separation of the
-object shapes. That is - the visual information input to the network that differentiates
-one trial from another is leveraged by the circuit to successfully condition
-the hand shape trajectory for grasping the object of that trial's particular shape, size, and
-location. The visual information forward-propagates through the network, through successive
-processing steps, where it is eventually leveraged to recapitulate the grasp appropriate
-for the object represented by the input visual features. It follows, then, that the classes
-are separable in some way by observing the distinct ways that the mRNN is activated by the
-classes' corresponding visual inputs. As we will note below: in the absense of this
-visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
-stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
-grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
-waiting period, but performance suffers drastically.
-
-Clearly, training and deploying a neural co-processor in a true brain involves challenges we
-don't capture with this simulation. This simulation does not constitute evidence
-that our method will, for example, immediately translate to true restoration of fine-grained
-control for individual fingers of a stroke patient. One clear challenge with doing so would
-be the sheer dimensionality of natural brain dynamics, and how that compares to the resolution
-of stimulation the co-processor can learn to apply. There exists a clear mismatch between
-the dimensionality of the underlying problem, sensor and stimulator technologies, and the
-amount of training data which can reasonably be collected to train a closed-loop neural
-controller. We argue that while this fact clearly creates a challenge for learning-based
-closed-loop stimulation, that nevertheless the insights we generate through this
-simulation are likely to be applicable. In the Discussion below, we will explore how
-those insights are likely to lead to initial demonstrations of the co-processor approach
-on lower dimensional problems. As we show next, the simulation exhibits a number of
-properties that reflect a real world application, which our co-processor design must
-contend with.
-
-\subsubsection{Simulated lesions cause real world failure modes}
-Simulating a brain lesion in terms of this artificial network results in error
-modes that resemble a some natural lesions of a primate brain. For example, if we
-alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
-the reaching motion generally succeeds, but the finger muscle velocities show a high degree
-of error - effectively implying that the subject can somewhat reach to grasp, but cannot form a
-grasp appropriate for the object. That suggests that losing a portion of the cortical
-machinery needed to translate or forward-propagate object shape information
-to movement-related cortex can result in a reduced ability to pose the hand, though
-positioning of the hand may still roughly succeed. Muscle spasticity of
-the hand is also a common symptom of certain strokes in primates, and indeed many who
-suffer from tem are still able to position their hand, even while being unable to form
-it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
-simulated lesion closely resembles natural stroke symptoms. However, the simulation is
-not intended to constitute a physical model of a lesion, i.e. to directly explain the
-connection between hand spasticity and the lesion.
-
-Conversely, if we lesion of a portion of the output module of the Michaels mRNN, roughly
-corresponding to M1, we see a more wholesale loss of movement, affecting even the
-ability to reach for the grasp. Finally, if we ``disconnect'' communications between the
-F5 and M1 modules, we see a failure similar to an AIP lesion: movement is generally
-achieved, but we see a disproportionate impact on hand pose.
-See Fig. \ref{fig:lesion} for examples.
+The grasping model we use \cite{michaels.mrnn} implements a vision-to-grasp pipeline, from the visual
+processing needed to move the hand to the appropriate position for the grasp to shaping the hand for
+grasping an object of a particular shape. The emergent dynamics of the model's ``modules'', once trained,
+correspond roughly to the responses in the cortical areas AIP, F5, and M1 in a NHP subject's brain.
+Specifically, the activity of the first module (Fig.~\ref{fig:michaels}a), receiving the visual inputs in a trial resembles
+the activity in area AIP of the NHP subject from the same trial. Likewise, the second and
+third modules resemble the activity from the subjects's F5 and M1 cortical areas, respectively.
+The emergent network dynamics show a number of other correspondences to the subjects' 
+brain activity as well (see \cite{michaels.mrnn} for details). For convenience, we will refer to the three
+modules in the model using the cortical areas (AIP, F5, M1) they correspond to. For details on how this network
+was trained, refer to Supplementary Materials \ref{sup:michaelstraining}. For additional details on the task structure, see
+Supplementary Materials \ref{sup:michaelstask}.
+
+An important attribute of the grasping network model is that  the simulated circuit's activity shows a relatively
+clear separation for different object shapes, i.e. the visual input to the network is leveraged by the model to
+successfully generate hand shape trajectories for grasping objects of particular shapes, and sizes. As we will note
+below, if the visual information is prevented from propagating through the modules, the model can at best learn a
+grasp stereotyped across all object sizes and shapes.
+
+\subsubsection{Lesioning the model causes real world failure modes.}
+Simulating a brain lesion in the grasping model described above results in error
+modes which resemble some natural lesions of the primate brain. For example, a ``lesion''
+involving zeroing some of the outputs of the first (AIP) module 
+leads to a reaching motion generally succeeding, but finger muscle velocities show a high
+degree of error - effectively implying that the subject can reach to grasp, but cannot
+tailor the grasp to the current object. This is likely due to object shape information not
+being fully conveyed to the motor cortex (M1) module. Such a form of hand muscle spasticity
+is also a common symptom in certain strokes where the subject is able to position their
+hand, but is unable to properly form the appropriate grasp \cite{khanna.openloop,
+puthenveettil.hand}.
+
+On the other hand, if we lesion of a portion of the motor cortex (M1) module in the model,
+we see a more complete loss of movement, affecting even the ability to reach for the grasp.
+Finally, if we ``disconnect'' communication between the F5 and M1 modules, we see a
+failure similar to an AIP lesion: the reaching movement is generally achieved, but we see
+a disproportionate impact on forming the appropriate hand grasp. See Fig. \ref{fig:lesion}
+for examples.
 
@@
 	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
-	    \caption{Example muscle trajectories during a single trial, for one hand and shoulder muscle}
+	    \caption{}
 	\end{subfigure}
@@
 	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
-	    \caption{Ratio of hand muscle MSE losses vs. overall MSE losses, for different lesion types}
+	    \caption{}
 	\end{subfigure}
 	\hfill
-	\caption{Lesion designs which prevent forward propagation of object shape information
-	         differentially impact hand pose. Loss here is an L2 distance
-		 measured relative to the circuit's trajectories prior to the lesion.}
+	\caption{\textbf{Simulating strokes by lesioning different parts of the model 
+	         lead to differential impact on grasping behavior.}
+             (a) Example muscle trajectories before (orange) and after (blue)
+             simulated stroke due to AIP, F5-M1 and M1 lesions for a shoulder
+             muscle (top row) and a hand muscle (bottom row). Lesions which prevent
+             a forward propagation of visual information tend to have a larger effect
+             on hand pose (example EIP muscle shown) than on the shoulder, which is more
+             involved in the reach motion (example TMAJ muscle shown). Lesions of M1
+             tend to cause a significant task loss on both.
+             (b) Ratio of hand muscle mean squared error (MSE) losses vs.
+             overall MSE losses, for different lesion types. The loss is measured
+             relative to the model's movement trajectories prior to the lesion.
+             Average of all trials, with +/- 1 stdev shown.}
 	\label{fig:lesion}
@@
 
-The co-processor's task in our simulation will be identifying the appropriate information,
-read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
-to effectively bridge across the lesion, forward propagating the object shape and task
-structure information, indirectly, through the use of stimulation.
-
-In our experiments below, we demonstrate our co-processor design on three types of
+Given a particular type of lesion in the model, the co-processor's task is to
+generate the appropriate stimulation pattern given mRNN activities. The
+co-processor thus seeks to effectively bridge across the lesion and deliver
+stimulation to enable grasping behavior tailored to the current object's shape.
+
+In our experiments, we studied our co-processor's performance on three types of
 simulated lesions:
 \begin{itemize}
-	\item \textbf{Output-based (AIP):} we force the output of some proportion of ``AIP''
-	      neurons to zero, effectively removing them from the network. This results in a
+	\item \textbf{Loss of AIP Neurons:} We force the output of some proportion of ``AIP''
+	      neurons to zero, effectively removing them from the network. This results in some amount of 
 	      loss of object shape information.
-	\item \textbf{Connection-based:} we prevent the propagation of
+	\item \textbf{Loss of F5-M1 Connections:} We prevent the propagation of
 	      information between the ``F5'' and ``M1'' modules, effectively representing a
-	      severing of the connections between the two. Note that the sparse connections
-	      between them run both directions, and we are lesioning both.
-	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
-	      neurons to zero, causing a more holisitic loss of movement. Here, the lesion
-	      may make it impossible for the co-processor to find a solution, but some
-	      recovery may be possible.
+	      severing of the connections between the two modules. Note that the connections are sparse
+	      and run in both directions, and we lesion the connections in both directions.
+	\item \textbf{Loss of M1 Neurons:} We force the output of some proportion of ``M1''
+	      neurons to zero. Here, the lesion may make it impossible for the co-processor
+	      to find a perfect solution, since the loss of M1 neurons may make it impossible to
+	      activate muscles in the same ways. However, in Results we will see that some recovery
+	      is possible.
 \end{itemize}
 
-\subsubsection{Simulated network exhibits long running dynamics}
-Just as the co-processor must adapt to the lesion, it must also adapt to the dynamics of
-the network it is stimulating. Natural neural networks as well as our simulated network
-exhibit long running dynamics, which our CPN and EN must account for in their learning.
-A perturbation of a network (i.e. due to stimulation) will cause changes in neuron
-activations long after the stimulation is applied, sometimes far from the site of
-stimulation. Our simulated grasping circuit exhibits the same behavior.
-
-To illustrate this, suppose we applied a small, one-time, instantaneous perturbation
-of the hidden state of 10 randomly chosen neurons in the output (M1) module at some point
-in time during a trial. If we repeat that experiment many times, we can see what the
+\subsubsection{Simulated network exhibits long running and stable dynamics.}
+To compute stimulation patterns that optimize for a task, the co-processor must
+learn to adapt to the dynamics of the network it is stimulating. Biological neural
+networks as well as our simulated grasping network exhibit long range changes in dynamics
+due to stimulation. A perturbation of the network (i.e. due to stimulation) will cause
+changes in neuronal activations long after the stimulation has been applied, sometimes far
+from the site of stimulation. Our model of the cortical grasping network exhibits the same
+behavior.
+
+To illustrate this, suppose we apply a small, one-time perturbation
+to the hidden states of 10 randomly chosen neurons in the output (M1) module at some point
+in time during a trial. By repeating this experiment many times, we can understand what the
 distribution of long-running effects tends to look like on the output muscle velocities.
 
-In Fig. \ref{fig:dynamics} we can see that even a single, one-time perturbation in
-the network has effects dozens of time steps later. Our co-processor will need to
-account for these dynamics, since stimulation is intended to cause perturbations
-in a network. As we will show in the next subsection, the problem the co-processor faces
-is in-fact even harder than this, due to a stimulation model that includes both
-spatial and temporal smoothing.
+In Fig.~\ref{fig:dynamics} we see that even a single, one-time perturbation in
+the network has effects dozens of time steps later. Our co-processor will need to learn to take
+these dynamics into account. As we will show in the next subsection, the problem our co-processor
+faces in our simulations is in fact even harder than this since we also incorporate
+a model of stimulation effects on the neural circuit which includes both spatial and temporal
+smoothing.
 
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{perturbe_single.pdf}
-	    \caption{Example trial}
+	    \includegraphics[width=\textwidth]{perturbe_single_tmaj.pdf}
+	    \caption{}
 	\end{subfigure}
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{perturbe_dist.pdf}
-	    \caption{Distribution of effects across n=1000 samples}
+	    \includegraphics[width=\textwidth]{perturbe_dist_tmaj.pdf}
+	    \caption{}
 	\end{subfigure}
 	\hfill
-	\caption{Instantaneous perturbations of a random sample (n=10) of M1
-	         neurons at time t=200 results in long-running effects on output.}
-	\label{fig:dynamics}
-\end{figure}
-
-\subsubsection{Stimulation model}
-
-\begin{figure}[h]
-	\begin{subfigure}[c]{0.45\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{stim_single.png}
-		\caption{$\sigma=1.875$}
+	\begin{subfigure}[c]{0.48\textwidth}
+	    \centering
+	    \includegraphics[width=\textwidth]{perturbe_single_eip.pdf}
+	    \caption{}
 	\end{subfigure}
 	\hfill
-	\begin{subfigure}[c]{0.45\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{stim_single2.png}
-		\caption{$\sigma=2.875$}
+	\begin{subfigure}[c]{0.48\textwidth}
+	    \centering
+	    \includegraphics[width=\textwidth]{perturbe_dist_eip.pdf}
+	    \caption{}
 	\end{subfigure}
-
-	\caption{Spatial and temporal smoothing of a single 16 dimensional,
-	         randomized $\theta$, onto 100 simulated neurons.
-		 Color values indicate the magnitude of the value summed into
-		 each neuron's hidden state in that time step. After $t=0$,
-		 $\theta$ is the zero vector.}
-	\label{fig:stim_single}
-\end{figure}
-
-In order to inject another aspect of realism into our simulation, we subject
-the co-processor to a stimulation model, rather than allowing it to directly
-influence the hidden state of the simulated network's neurons. Our stimulation
-model includes aspects of both spatial and temporal smoothing.
-
-The intent with this approach is not to create a realistic biophysical model. The
-mRNN model does not contain sufficient information to arrange the neurons in a
-simulated space, and to simulate how that may change the effects of stimulation.
-Instead, our stimulation model is merely attempting to make the co-processor's task
-more realistically difficult, in the sense that it cannot precisely set the hidden
-state of every neuron, instantaneously and in each time step. Any stimulation it
-applies will diffuse across ``space'' and time.
-
-In our experiments, we stimulate only the output module, since this co-processor's
+	\hfill
+	\caption{\textbf{Perturbations of the simulated cortical network (i.e. due to stimulation)
+	         have long-running effects on muscle activation.} Our results are consistent with
+	         the fixed point analysis of the network in \cite{michaels.mrnn}, showing that the network
+	         exhibits stability. It tends towards the original trajectory after perturbation.
+	         Instantaneous random perturbation of a random group of 10 M1
+	         neurons performed at time t=2s. (a, c) Single trial example for TMAJ, EIP muscles
+	         respectively. (b, d) Distribution of effects across n=1000 samples on TMAJ, EIP muscles
+	         respectively. Probability distribution shown to +/- 2 stdevs. Network exhibits similar
+	         behavior when perturbance is performed at other times.}
+	\label{fig:dynamics}
+\end{figure}
+
+\subsubsection{Stimulation model.}
+In our experiments, we stimulate neurons only in the output area (here M1) since the co-processor's
 purpose is to improve external task performance, which it is able to do with
-only stimulation of the output module of the network. It is conceivable that
-a co-processor could stimulate other areas of the brain to improve task performance
-downstream, or to probe the brain to better reveal the user's intent (i.e. object
-shape), but we did not explore those possibilities in this paper.
-
-The stimulation function $S$ receives as inputs the stimulation parameters $\theta$
-provided by the CPN, which it accumulates in an internal memory. Each time step,
-it outputs a modifier to the internal states of all neurons in the output module,
-based on the current state of memory. That memory allows
-the stimulation function to perform temporal smoothing. Specifically, we used a
-simple exponential decay model where the function reads its current state of memory,
-sums in the new parameters, and then decays each memory element towards $0.0$ at some
-rate. Roughly speaking, this design intends to approximate the notion of dissipation: the
-effect of stimulation is not instantaneous, but rather decays with time, as charge
-dissipates into the surrounding area.
-
-Likewise, the stimulation function maps the relatively low dimensional stimulation
-parameter vector $\theta$ onto the M1 cells which it stimulates. In our experiments,
-our stimulation parameters $\theta$ had 16 dimensions, which we found was
-sufficient to allow significant improvement in task performance. The stimulation
-function uses a simple Gaussian smoothing to map those parameters onto the simulated
-M1 neurons. Each parameter represents, in a sense, an electrode located along a single
-spatial dimension, whose stimulation affects the neurons in its vicinity more than
-it affects others, according to each neuron's Mahalanobis distance from it. The neurons
-are aligned along that dimension arbitrarily. We fix the width of the Gaussians
-arbitrarily to $\sigma=1.75$. We are not aware of a principled way of picking this
-value, and clearly if we set it to extreme values, it makes our simulation less useful.
-For example, if it is set to an extreme high value, the stimulation is effectively
-lower dimensional, resulting in a model corresponding to a stimulation technology that
-is inappropriately low resolution for our task, for example epidural stimulation.
-We could not conclude much from such an experiment, because this simulation does
-not constitute a biophysical model, where we can identify some realistic value. The
-purpose of this smoothing technique is only to make the co-processor's job harder in
-the simple sense that it cannot target neurons individually. As a result, we set it
-to a moderate, balanced value, and don't vary it.
-
-Thus, the governing equations of the stimulation become:
+appropriate stimulation of the M1 output region of the network model. In a more general setting, it is
+conceivable that a co-processor could stimulate other areas of the brain as well to improve task
+performance downstream, or to probe the brain to better reveal the user's intent (i.e. object
+shape), but we leave these directions to future work.
+
+To make our stimulation model more realistic, rather than allowing stimulation to directly change the
+output of single neurons, we simulated how stimulation may affect a network of neurons using a
+model that incorporates aspects of both spatial and temporal smoothing. Our intent was not to create a
+biophysical model of stimulation (our model does not arrange neurons in a volume to allow for such detailed
+simulation). Instead, the stimulation model approximates the effects of in-vivo biological stimulation (e.g.
+extracellular electrical stimulation) by diffusing the effects of stimulation across neural tissue space and time.
+Including this model implies that our EN must approximate this stimulation function, in addition to approximating
+the cortical dynamics of the brain and the mapping of those dynamics onto the grasping task. In Supplementary Materials
+\ref{sup:passthrough}, we compare results to the case with no stimulation or recording function.
+
+Specifically, we use a stimulation function $S$ which receives as input the stimulation parameter vector $\theta$
+from the CPN. $S$ performs temporal smoothing using a simple exponential decay model by adding its current input to
+an exponentially decaying sum of inputs that decays towards $0.0$ at some rate (see Equation ~\ref{eq:temporalstim} below).
+The effect of stimulation is thus not instantaneous, but rather decays with time. Likewise, to approximate how charge
+dissipates into a surrounding volume, we applied Gaussian smoothing to map $\theta$ (16-D in our experiments) to changes
+in the activations of a large number of the simulated M1 neurons (100 neurons in our experiments). If we assume each element
+of the $\theta$ represents the stimulation parameter for a single electrode, our Gaussian smoothing operation emulates how
+stimulation may affect the neurons in its vicinity more than it affects neurons further away. To accomplish that, we assume
+our model neurons are aligned along a spatial dimension arbitrarily and fix $\sigma=1.75$.
+
+The resulting equations defining our stimulation model are:
 
 \begin{equation}
-\alpha_{t} = \tau\alpha_{t-1} + \theta_{t-1}
+\alpha_{t} = \tau\alpha_{t-1} + \theta_{t-1} \label{eq:temporalstim}
 \end{equation}
 \begin{equation}
-s_{t} = C(\alpha_{t})
+s_{t} = C\alpha_{t}
 \end{equation}
@@
 \begin{itemize}
-	\item $\alpha$: the 16 dimensional internal activation, or memory
-        of our stimulation
-	\item $\tau$: our decay rate, which we set arbitrarily to $0.7$
-	\item $C$: a precalculated $100 x 16$ matrix which provides our
-	Gaussian smoothing
-	\item $s_{t}$: the stimulation we apply to each neuron at the
-	given time step.
+	\item $\alpha$: the 16 dimensional internal state 
+        of the stimulation function
+	\item $\tau$: our decay rate for temporal smoothing, which we set arbitrarily to $0.7$
+	\item $C$: a fixed $100 \times 16$ Gaussian smoothing matrix containing a single 1-D Gaussian
+	           in each column, to implement spatial smoothing and spread of stimulation to the
+	           stimulated neural population. 
+	\item $s_{t}$: the spatiotemporally smoothed stimulation vector whose elements denote how
+	               much stimulation is applied to each neuron at time step $t$ (see
+	               Equation~\ref{eq:neural-activation} below).
 \end{itemize}
 
-The governing equations of the simulated network then become:
+The governing equations of an mRNN with stimulation then become:
 
@@
 x_{t+1} = Jx_{t} + Iv_{t} + s_{t} + b
+\label{eq:neural-activation}
 \end{equation}
@@
 \begin{itemize}
-	\item $x$: the hidden state of each neuron
-	\item $J$: the recurrence weight matrix
-	\item $I$: the input response matrix
-	\item $b$: the neuron activation bias
-	\item $L$, $l$: the parameters of the linear readout
-	\item $y$: the output of the network
+	\item $x$: the hidden state of each model neuron
+	\item $J$: recurrent weight matrix
+	\item $I$: input weight matrix
+    \item $v_{t}$: input 
+	\item $b$: activation bias
+	\item $L$, $l$: parameters of the linear readout layer
+	\item $y_{t}$: the output of the network
 \end{itemize}
 
-See Fig. \ref{fig:stim_single} for a visual depiction of an example where
-$\theta$ is non-zero at $t=0$, and zero for all other times, in order to see how
-stimulation is applied across space and time. In Fig. \ref{fig:stim_and_obs},
-we see stimulation applied during the simulation, in a real trial. In this case,
-the trial is one where $50\%$ of M1 neurons are lesioned to have zero output.
-Observations of the network's hidden state (explained in the next section) are taken
-from the first two modules, and stimulation is applied to the last module. This fact
-is intended to capture the difficulties with observing the same region that you are
-stimulating, due to stimulation artifacts. The trial we show depicts stimulation
-supplied by a highly trained co-processor, where task performance has been
-significantly restored. We explain this experiment in more detail below.
-
 \begin{figure}[h]
+	\begin{subfigure}[c]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{stim_single.png}
+		\caption{$\sigma=1.875$}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[c]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{stim_single2.png}
+		\caption{$\sigma=2.875$}
+	\end{subfigure}
+
+	\caption{\textbf{Simulating the effects of stimulation on a recurrent network across space and time.} The stimulation model performs spatial and temporal smoothing of a 16-dimensional
+	         vector of stimulation parameters $\theta$ onto 100 neurons in a recurrent network.
+	         The 16 elements of the vector represent in-effect 16 electrodes located
+	         evenly along a spatial dimension, along which the 100 neurons are arranged
+	         arbitrarily. We show here the effects of two randomized $\theta$ with two respective smoothing
+	         parameters $\sigma$. After $t=0$,
+		     $\theta$ is the zero vector. Color values indicate the magnitude of the value summed into
+		     each neuron's hidden state at that time step. }
+	\label{fig:stim_single}
+\end{figure}
+
+Fig. \ref{fig:stim_single} depicts an example where
+$\theta$ is non-zero at $t=0$ and zero for all other times, illustrating the effects of 
+stimulation on the network across space and time.  Fig. \ref{fig:stim_and_obs} illustrates the
+effects of stimulation during a stroke simulation. In this case, the trial is one where $50\%$
+of M1 neurons have been  lesioned (their outputs are zero), the inputs to the CPN are from the
+AIP and F5 modules, and stimulation is applied to the M1 module.
+
+\begin{figure}[h]
 	\centering
 	\includegraphics[width=\textwidth]{stim_and_obs.png}
-	\caption{Example of stimulation and observation for a single trial. Here, M1
-	has been lesioned 50\%, as seen in the zero (light blue) hidden states. We
-	gather observations only from the AIP and F5 modules, as we explain in
-	Section \ref{sec:experiments}. Stimulation is applied to M1, to drive the
-	network output.}
+	\caption{\textbf{Example of stimulation and recording for a single trial.} Here, M1
+	has been lesioned 50\%, indicated by hidden states of M1 neurons with a value of zero (bottom
+	left panel, lowest part of the plot) . We record neural activities (recording model) only
+	from the AIP and F5 modules, as explained in Section \ref{sec:experiments}. Stimulation is applied
+	to M1 to drive the network's output. Stimulation is mostly constant until approximately t=275,
+	when the hold signal was lifted and reach began.}
 	\label{fig:stim_and_obs}
@@
 
-\subsubsection{Observation model}
-Our observation model likewise relies on a notion of electrodes spread along
-a single spatial dimension. As with stimulation, this is not intended to
-capture true physical relationships between neurons and electrodes, but
-rather to act as a dimensionality reduction that isn't designed to specifically
-favor task success, such as PCA. We treat the neurons of each module as
-occupying their own space, with 20 electrodes arrayed along each.
-
-We use a similar Gaussian-based approach as with stimulation, but in this
-case our approach looks much like a Gaussian convolution, where the Gaussian
-kernel is centered at each electrode position. Each `electrode' is read out
-as a weighted average of all neurons in the given module, with weights being
-provided by the Gaussian kernel. As above, refer to Fig. \ref{fig:stim_and_obs}
-for a visual example.
-
-\subsubsection{Brain co-adaptation}
-To demonstrate the co-processor's ability to adapt to the non-stationarity
-of the brain, we also change the brain throughout the co-processor's training.
-Specifically, we simulate the brain's co-adaptation with the stimulation
-to cooperatively solve the problem, as the stimulation is also changing
-through the CPN's training.
-
-We achieve this through a simple error backpropagation, using PyTorch's
-implementation of the $Adam$ optimizer. With each trial, task loss is
-calculated and backpropagated into the mRNN, allowing it to adapt to
-the fact it receives stimulation. The learning rate for the optimizer
-is set arbitrarily to a relatively low rate of $1e-7$. The learning
-rate was not chosen in a principled way, since we are not aware of a
-principled way of choosing it. However, we reason that it should be
-lower than the co-processor by orders of magnitude due
-to the rapidity with which an ANN can be retrained in practice.
+\subsubsection{Recording model.}
+To simulate recordings of our model neurons, we used a recording model that assumed the model neurons
+are laid out along a single spatial dimension with the given number of electrodes spread evenly apart.
+We then applied Gaussian convolution, where the Gaussian kernel is centered at each electrode position.
+Thus, the recording from each `electrode' is a Gaussian-weighted average of the activities of all neurons
+in the given module. Fig.~\ref{fig:stim_and_obs} (bottom panels) show an example of recordings obtained
+using this recording model; note that the Gaussian-averaging of activities provides only a low-pass
+filtered, potentially ambiguous view of the activities of the underlying neurons, making it more challenging
+for the co-processor to interpret the neural activity and produce an appropriate stimulation pattern. Our
+simulated recordings are similar in spirit, though not actually modeling, local field potentials recorded
+extracellularly in biological neural tissue. In Supplemental section
+\ref{sup:observability} we show that as we vary the observability of the
+brain via this recording model the results are largely the same, up to a
+point where objects can no longer be distinguished by the recording data.
+
+\subsubsection{Simulating co-adaptation by the brain.}
+To demonstrate the co-processor's ability to adapt with the brain as it adapts to the co-processor's
+stimulation, we modify the cortical grasping network's parameters (synaptic weights and biases) throughout
+the co-processor's training. We use the standard error backpropagation algorithm to adapt the grasping network
+(using PyTorch's implementation of the $AdamW$ optimizer). With each trial, task loss is calculated and
+backpropagated through the mRNNs as they receive stimulation. The learning rate was set arbitrarily
+to a relatively low rate of $1\mathrm{e}{-7}$ so that the network adapts more slowly than the co-processor.
 
 \subsubsection{Simulating recovery prior to co-processor use}
-We sought to additonally re-train the mRNN after applying the lesion in
-order to simulate stroke recovery. For our simulated lesions which zero
-the outputs of neurons, the mRNN cannot supply an appropriate pre-recovery model.
-We know that strokes can cause damage from which the subject, in many cases,
-cannot recover. The mRNN model has sufficient redundancy built into it that
-lesioning the outputs of large numbers of neurons often leaves enough remaining
-degrees of freedom that a nearly full recovery can occur, unless so many
-neurons are lesioned that no stimulation could be effective.
-
-We can, however, simulate pre-recovery in the case of a lesion that prevents
-communication between the F5 and M1 modules. In that case, object shape information
-cannot propagate forward in the network, in order to condition the hand for
-grasping. The mRNN can, in pre-recovery, learn a stereotyped grasp, at which
-point the co-processor's job will be to forward-propagate that information.
-We explore this in one of our experiments, explained below.
-
-\subsection{Training Algorithm}
-In all experiments, training the co-processor requires a careful interleaving of
-EN and CPN training epochs. EN training epochs concentrate on updating the EN,
-based on observations of the effect of stimulation on the simulated network's output.
-The loss function when training the EN is the mean squared error (MSE) loss between the
-actual and predicted output of the mRNN (i.e. muscle velocities). We then use the EN
-to train the CPN during the CPN training epochs. During that time, we backpropagate
-the MSE loss between the EN's predicted output, and target output, to the CPN. In
-other words: we treat the EN output as a proxy to the true brain, allowing us
-to train the CPN, since we have no way to ``backpropagate through the brain'' itself.
-See Fig. \ref{fig:training}.
+After a stroke, the human brain has the ability to learn and recover to some extent the behaviors
+affected by the stroke. We simulated this ability in our grasping network model by re-training the
+network for the grasping task after lesioning it. For our simulated lesions which zero
+the outputs of neurons, the mRNN often cannot supply a partial recovery model.
+This is because the mRNN model has sufficient redundancy built into it that lesioning it by
+inactivating large numbers of neurons often leaves enough remaining degrees of freedom that a
+nearly full recovery can occur, unless so many neurons are lesioned that no stimulation could be
+effective.
+
+In the case of a lesion that prevents communication between the F5 and M1 modules, information about
+the input object's shape cannot propagate forward in the network to allow shaping of the hand for
+grasping. However, the mRNN can learn a stereotyped grasp during the recovery period. After this recovery
+period, a co-processor can further boost grasp accuracy by forward-propagating object shape information
+from AIP and F5 to M1, acting as an artificial neural bridge. We explore this application of the
+co-processor in one of our experiments below.
+
+\subsubsection{Simulating a non-stationary recording function, e.g. sensor drift}
+We also demonstrate the co-processor's ability to adapt to non-stationarity in the recording
+function. Over time, implanted sensors may drift from true readings. To test the co-processor's
+ability to adapt to that drift, we perform an experiment where the observation function
+has a bias term which changes over time, according to a random process. Each epoch, we
+add a random value to each element of the bias, drawn from a mean-0 Gaussian distribution. That
+causes the ``sensors'' to have a drifting zero point over time.
+See Supplementary Materials \ref{sup:drift} for further details.
+
+\subsection{Training Paradigm}
+\label{sec:training}
+
+To train the CPN to generate appropriate stimulation patterns for grasping a given input object of a
+particular shape, we can use the backpropagation algorithm to minimize the errors between the current
+generated grasp and the desired target grasp. However, this error is in terms of grasp error instead of
+error in stimulation patterns (which is what the CPN needs), and we have no way of backpropagating grasp
+error through biological networks in the brain. This motivates our use of the EN as a model for the
+transformation from stimulation to muscle velocities: we backpropagate grasp error through the EN,
+while keeping its parameters fixed, and use this backpropagated error to change the parameters of the
+CPN towards minimizing grasp error.
+
+Using the EN to train the CPN requires careful interleaving of EN and CPN training epochs. EN training
+epochs concentrate on updating the EN, based on observations of the effect of stimulation on the grasping
+network model's output. The loss function when training the EN is the mean squared error (MSE) loss between
+the EN's prediction and the  actual output of the grasping network (i.e. muscle velocities). During the CPN
+training epochs, we keep the EN's parameters fixed and use the EN to train the CPN. Specifically, we
+backpropagate the MSE loss between the EN's predicted output, and the desired output (target muscle velocities
+for grasping the input object) through the CPN. Fig. \ref{fig:training} illustrates the EN and CPN training
+processes.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
-\caption{Experiment architecture overview}
-\centering
-\label{fig:exp_overview}
-\end{figure}
-
-\begin{figure}
 	\centering
@@
 		\centering
-		\includegraphics[width=\textwidth]{backprop_en.eps}
-		\caption{EN training phase: backpropagate actual vs predicted}
+		\includegraphics[width=\textwidth]{backprop_en.png}
+		\caption{}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[width=\textwidth]{backprop_cpn.eps}
-		\caption{CPN training phase: backpropagate actual vs target MSE loss, via EN, to CPN. This effectively treats the EN output as the actual, and thus the EN must be trained first.}
+		\includegraphics[width=\textwidth]{backprop_cpn.png}
+		\caption{}
 	\end{subfigure}
 	\hfill
-\caption{Two Phase Training Regime}
+\caption{\textbf{Training the neural co-processor.} (a) EN training phase: 
+         backpropagate actual vs predicted muscle velocity MSE loss.
+         (b) CPN training phase: backpropagate actual vs target
+         MSE loss, via EN, to CPN. This effectively treats the EN
+         output as the actual, and thus the EN must be trained first.}
 \label{fig:training}
@@
 
-\subsubsection{EN training}
-For the EN to be useful, it must accurately predict the effect of stimulation
-produced by the CPN. In addition to that, backpropagating through it must yield
-gradients which train the CPN to output more useful stimulation. We discovered that
-this latter property does not occur simply by virtue of the former.
-An EN of this design can be trained to high levels of predictive power, even
-on random stimulation, and to orders of magnitude lower loss than the task loss,
-while at the same time backpropagation through it yields gradients which are not
-useful for training the CPN. As a result, if the EN is not trained in specific ways,
-the CPN exhibits unstable training. We have thus far been unable to discover a way
-to measure and optimize for this second property directly, though pursuing it may be a
-fruitful subject of future applied ML research. Instead of attempting to train an EN
-directly to exhibit this property, we discovered a training regime that indirectly,
-but reliably, makes it occur.
-
-The central concept of our training regime is to provide a carefully chosen variety of
-training examples, and to use heavy weight regularization, in order to force
-the EN to generalize well to stimulation which it hasn't yet seen. We hypothesize that
-the crux of our EN training problem is one of over-fitting: that the EN may be trained
-to high predictive power on a set of stimulation examples, but that the effective
-dimensionality of the problem it is learning is comparatively quite high, due to both
-the dimensionality of our stimulation parameters $\theta$, and the dynamics of the
-network being stimulated. As a result, the hypersurface which it fits to the training
-examples may exhibit gradients not reflective of reality.
-
-First, and most significantly, we structure the training dataset for each EN training epoch
-in a very specific way. Each epoch, we include:
+\subsubsection{Training and testing data sets.}
+For training and evaluation of our model, we use the same data as in Michaels et al.
+\cite{michaels.mrnn}, which was provided to us by the lead author. The data consist of:
 \begin{itemize}
-	\item examples using stimulation from the current CPN
-	\item examples from a small (100s) collection of CPNs which are copies of the
-	      current CPN, with Gaussian-distributed mean-0 noise added to their parameters
-	\item examples of white noise stimulation
+    \item The input visual features and hold signal. We input these to the simulated
+          cortical circuit;
+    \item The object identity. We use this to calculate how well the co-processor
+          can differentiate objects; see Results section below;
+    \item The muscle velocity data for the trial, as extracted from the data glove
+          during the NHP trial, and as processed. This is our target output.
 \end{itemize}
-We developed this structure under the belief that we needed to cover a sufficient
-variety of examples to prevent overfit, and to do so in a way that emphasizes the neighborhood
-in CPN parameter space of the current CPN. We initially attempted to pad our dataset with white
-noise examples alone, but that was not sufficient to stabilize CPN training, despite the EN's
-ability to reach a high predictive power. Adding in the addition of samples from the parameter
-space near the current CPN appears to help ensure that the EN is localized enough to give
-useful gradients for that area. As we will explain below, this produces an additional problem
-that the EN is too specialized to the local area of CPN parameter space, but we then solve
-that problem using EN retraining, interleaved with the CPN training.
-
-After some experimentation, we composed each batch of EN training data in the following way: \\
+
+For each training session, we hold out a random sample of $20\%$ of the data to act as our validation data set; we
+use the other $80\%$ for training.
+
+\subsubsection{EN training.}
+For the EN to be useful for training the CPN, it must accurately predict the behavioral effects of stimulation
+produced by the CPN. In addition, backpropagating through the EN must yield gradients which are useful for
+changing the weights of the CPN in order to produce better stimulation patterns or minimizing the error.
+We discovered that this latter property does not occur simply by virtue of the former. Specifically, an EN
+can be trained to high levels of predictive power, even for random stimulation, and to orders of magnitude
+lower loss than the task loss, but backpropagation through the EN can still yield gradients which lead to
+unstable training of the CPN.
+
+We hypothesize that the crux of the EN training problem is one of over-fitting: the EN may be trained
+to achieve high predictive power on a data set of stimulation examples, but due to 
+the high dimensionality of our stimulation parameters $\theta$, and the dynamics of the
+network being stimulated, the function learned by the EN may be biased and provide gradients not suitable
+for the stimulation inputs generated by the CPN during its training.
+
+To address this problem, we first structure the training dataset for each EN training epoch
+in a specific way. Each epoch, we include:
+\begin{itemize}
+	\item training examples using stimulation inputs generated by the current CPN;
+	\item training examples using stimulation inputs generated by a small collection of CPNs obtained
+	      by adding zero-mean Gaussian noise to the current CPN's parameters; and 
+	\item training examples with white noise stimulation inputs.
+\end{itemize}
+
+Such a training paradigm is designed to cover a sufficient variety of examples to prevent overfitting,
+and to do so in a way that explores the neighborhood of the current CPN's parameter space. Augmenting
+a CPN-generated data set with white noise examples alone was not sufficient to stabilize CPN training,
+but including additional examples generated by CPNs in the neighborhood of the current CPN helped train
+the EN to perform well in that neighborhood and produce useful gradients for the CPN. ``Neighborhood''
+here refers to the region of CPN parameter space near the current CPN. This may cause the EN to be
+overfit, i.e. fit to the local area of CPN parameter space but exhibiting poor predictive power in
+other areas of the space. However, we solve this problem by retraining the EN, interleaved with CPN
+training. We show the effect of this dataset composition in one results section below
+(\ref{sec:results_con}).
+
+We composed each batch of EN training data as follows: \\
 
@@
 \textbf{Source}                                                                                                                            & \textbf{Proportion of dataset} \\ \hline
-Current CPN                                                                                                       & 10\%                           \\ \hline
-\begin{tabular}[c]{@{}l@{}}Current CPN with parameter noise\end{tabular} & 60\%                           \\ \hline
-White noise stimulation                                                                                                        & 30\%                           \\ \hline
+Data from current CPN                                                                                                       & 10\%                           \\ \hline
+\begin{tabular}[c]{@{}l@{}}Data from current CPN with parameter noise\end{tabular} & 60\%                           \\ \hline
+Data from white noise stimulation                                                                                                        & 30\%                           \\ \hline
 \end{tabular} \\
 
-Additionally, we use PyTorch's $AdamW$ optimizer, which includes an option for weight regularization.
-With that, we use a carefully chosen learning rate schedule. Our intuition is to leverage regularization as
-one of the standard mitigations for overfit. To converge, the learning rate schedule needs
-to be chosen carefully: a relatively high learning rate at first, dropping to a very low rate soon
-thereafter. Using too-low or too-high of a learning rate in any phase causes the EN learning
-to not converge. EN training proceeds until it reaches a threshold of prediction error, defined as
-a fraction of the current CPN's task loss.
+Additionally, we use decoupled weight decay regularization \cite{loshchilov2017decoupled} to
+mitigate overfitting, and a carefully chosen learning rate schedule. The schedule begins with
+a higher learning rate initially ($4\mathrm{e}{-3}$), ramping down to a lower rate ($1\mathrm{e}{-4}$).
+The learning schedule is based on the most recent prediction error on the validation portion of the
+data set. We found that using rates much lower or higher than these rates in any training phase
+may cause the EN learning to not converge. EN training proceeds until the network's output error
+exceed a threshold, defined as a fraction of the current CPN's task loss. See Section
+\ref{sec:interleaving} for further details on when we switch between training the EN and CPN.
 
 \subsubsection{CPN training}
-Once an EN is properly trained, CPN training is straightforward. For this phase,
-we generate training examples using the current CPN alone. The EN generates predictions of the
-effect of those stimulations, and we backpropagate through the EN to generate training gradients
-for the CPN. As with the EN, we constitute the dataset with a large random sample of trials from
-the original Michaels task. See Fig. \ref{fig:training}.
-
-The choice of learning rate schedule for training the CPN is somewhat important.
-The CPN appears to train in two phases. In the first phase, it is largely learning the structure
-of the reach-to-grasp task, e.g. that the muscles need to stay still until the reach begins. During
-that phase, the learning rate can be quite high. Later, the CPN begins to learn the mapping between
-the object shape information observed from the simulated brain, and how that should map onto
-stimulation. That phase of the training takes much longer, and requires a learning rate 2-3
-orders of magnitude lower than in the first phase. We will cover this in some more detail in
-the Results section (Section \ref{sec:results}).
+Once an EN is properly trained, we can use the EN as a surrogate for the part of the grasping model
+that transforms stimulation patterns and neural activity to grasping behavior. We use a large number
+of randomly sampled trials from the original Michaels et al. task \cite{michaels.mrnn} and generate
+stimulation patterns using the current CPN alone. These stimulation patterns are passed through the EN
+to generate predictions of output muscle velocities, which are compared to the desired target velocities
+to compute the error. We backpropagate this error through the EN (without modifying its parameters) to
+generate training gradients for the CPN. Fig.~\ref{fig:training} illustrates this training process.
+
+We found that the CPN appears to train in two phases. In the first phase, it is largely learning the structure
+of the reach-to-grasp task, e.g. that the muscles need to remain at a position during the hold period until the
+reach begins. During this early training phase, the learning rate can be quite high ($1\mathrm{e}{-3}$). Later, the CPN
+begins to learn the mapping between object shape information and the stimulation patterns needed to approximate
+the target grasp for that object. This later phase of training requires a learning rate 2-3 orders of magnitude
+lower than in the first phase ($1\mathrm{e}{-6} - 5\mathrm{e}{-5}$). See Section~\ref{sec:results} for more details.
 
 \subsubsection{Interleaved CPN/EN training, and adapting to non-stationarity}
-Having defined the training periods for the EN and CPN, we can define a training
-algorithm on the basis of those two. We train the two in alternation, creating
-a new EN each time we enter an EN training period. We explored the possibility
-of reusing an existing EN by retraining it, for the purpose of training efficiency.
-However, we found that retraining an EN was no more fast than training a new one,
-and in-fact usually took longer, despite significant experimentation with learning
-rates. It is unclear if that fact is an artifact of our simulation, or of the
-problem more generally, and so it needs more examination in the future, for the
-purpose of training efficiency.
-
-The key remaining challenge is determining when an EN is no longer suitable for training the
-current CPN. Expiring an EN at the right time accomplishes two things. First, it ensures
-that our current EN is trained properly for the current CPN; without that, the CPN training
-will become unstable. Second, expiring the EN is how we adapt our learning to the brain's
-non-stationarity. At some point, the brain will have changed sufficiently for our EN to be
-out-of-date, and therefore requiring replacement.
-
-We solve this challenge through a simple set of metrics, which, when satisfied, indicate
-we need to transition to an EN training period. Our first metric simply captures the EN's
-declining predictive power. When the EN's prediction error reaches a point which is
-sufficiently above some ratio to the CPN's task loss, we expire the EN immediately.
-Second, we expire the EN if CPN loss does not improve (or gets worse) across a sufficient
-number of recent training steps. This resembles common stopping conditions in
-iterative learning: one stops when learning is no longer improving the model. Together,
-these two metrics appear effective in ensuring we expire the EN at the right time.
+\label{sec:interleaving}
+Having defined the training procedures for the EN and CPN, we can define a training
+algorithm combining those two. We train the two in alternation, creating
+a new EN each time we enter an EN training phase. We explored the possibility
+of reusing an existing EN but found that retraining an EN was no more fast than training a new one.
+Training a new EN also allows us to adapt to the brain's non-stationarity. That requires us to
+determine when the current EN is no longer suitable for training the current CPN. 
+Our algorithm retires the current EN and trains a new one when one of two conditions
+is satisfied: (1) the EN's prediction error is sufficiently above some fraction of the CPN's task loss;
+or (2) if CPN loss does not improve (or gets worse) over a sufficient number of recent training steps 
+(this is a common stopping criterion in machine learning). Additional details can be found in the
+Supplementary Materials section ~\ref{sup:encpninter}.
 
@@
 
-Using the simulation described above, we perform several experiments to
-demonstrate the co-processor's ability to learn. In each experiment, we apply one
-of the three lesion types described above. We additionally vary whether we simulate
-brain co-adaptation, in order to understand its impact on the co-processor training.
-Finally, we perform one experiment where we attempt to simulate natural recovery from
-the lesion, prior to training the co-processor. In sum, our experiments are: \\
-
-\begin{tabular}{|l|l|c|c|}
+We performed 8 experiments to investigate the co-processor's ability to learn under a variety of
+conditions: we studied three types of lesions to the grasping network (simulated stroke) and for
+each lesion type, we either turned on or off brain co-adaptation, natural recovery before co-processor
+use, and non-stationarity due to sensor drift. The following table summarizes the experiments
+(a blank entry under a condition (e.g., Simulate co-adaptation?) means ``No'' while an X means ``Yes''): \\
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|c|c|c|}
 \hline
-\textbf{Index} & \textbf{Lesion}                                           & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ co-adaptation?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ pre-recovery?\end{tabular}}} \\ \hline
-1              & 50\% AIP loss                                             &                                                                                                 &                                                                                                \\ \hline
-2              & 50\% AIP loss                                             & X                                                                                               &                                                                                                \\ \hline
-3              & 50\% M1 loss                                              &                                                                                                 &                                                                                                \\ \hline
-4              & 50\% M1 loss                                              & X                                                                                               &                                                                                                \\ \hline
-5              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 &                                                                                                 &                                                                                                \\ \hline
-6              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 & X                                                                                               &                                                                                                \\ \hline
-7              & 100\% loss of connectivity F5\textless{}-\textgreater{}M1 & X                                                                                               & X                                                                                              \\ \hline
-\end{tabular} \\
-
-In each experiment, we use a single mRNN model, sourced directly from the results
-cited in Michaels \cite{michaels.mrnn}. We drive the experiment using the same set of
-input and output data that the mRNN was trained on. The dataset contains a total of
-502 trials, spread uniformally among 42 object classes. In each experiment, we separate
+ & \textbf{Lesion}                                           & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ co-adaptation?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ recovery?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ sensor\\ drift?\end{tabular}}}\\ \hline
+1              & 50\% AIP loss                                             &                                                                                                 &                                                                                                &                                                                                                 \\ \hline
+2              & 50\% AIP loss                                             & X                                                                                               &                                                                                                &                                                                                                 \\ \hline
+3              & 50\% M1 loss                                              &                                                                                                 &                                                                                                &                                                                                                 \\ \hline
+4              & 50\% M1 loss                                              & X                                                                                               &                                                                                                &                                                                                                 \\ \hline
+5              & 100\% connection loss F5\textless{}-\textgreater{}M1 &                                                                                                 &                                                                                                &                                                                                                 \\ \hline
+6              & 100\% connection loss F5\textless{}-\textgreater{}M1 & X                                                                                               &                                                                                                &                                                                                                 \\ \hline
+7              & 100\% connection loss F5\textless{}-\textgreater{}M1 & X                                                                                               & X                                                                                              &                                                                                                 \\ \hline
+8              & 100\% connection loss F5\textless{}-\textgreater{}M1 & X                                                                                               &                                                                                                & X                                                                                               \\ \hline
+\end{tabular}
+\end{table}
+
+For each experiment, we used the mRNN grasping model from Michaels et al.\cite{michaels.mrnn}, 
+and the same set of input and output data used in their work. The dataset contains a total of
+502 trials, sampled uniformly across 42 object classes. In each experiment, we held out 
 a random sample of 20\% of the dataset for validation.
@@
 \subsection{Stopping condition}
-For the purpose of these experiments, we train the co-processor until one of two
+In these experiments, we train the co-processor until one of two
 conditions is satisfied:
@@
 \begin{itemize}
-	\item The percent change in task loss between two consecutive 500 epoch windows is below a certain threshold: 0.01\%. This
-	      stopping criteria ensures we stop training when the economy of additional training is low.
-	\item The epoch count is greater than 250k. This criteria provides a backstop against very slow training running without end.
+	\item The percent change in average task loss between two consecutive ranges of 500 epochs
+	      is below a threshold of 0.01\%, indicating minimal benefits of further training. 
+	\item We have exceeded 250k epochs, which acts as a hard stopping criterion. 
 \end{itemize}
@@
 
+For each lesioning experiment, we track two metrics to characterize the learning progress. First, we
+track the task loss, which is an MSE loss that measures the co-processor's ability to restore
+movement towards the target trajectory. We define it over the muscle length trajectory vector
+output from the lesioned grasping network model, compared to the ground truth muscle length data captured
+during the Michaels et al. experiment \cite{michaels.mrnn}. In graphs, we characterize this loss in terms of 
+percent recovery - i.e. difference between lesioned and healthy performance. Figs.
+\ref{fig:results_aip}-\ref{fig:results_drift} and Table \ref{tab:results} provide the detailed results.
+
+Second, we measure the degree to which lesioned grasping network, when coupled with the
+co-processor, successfully differentiates object classes. We compared the ratio of within-class
+to total variation between the lesioned and the healthy network. This metric $S$ is defined as:
+
+\begin{equation}
+	S = \frac{\sigma_{a}}{\sigma_{w}} - \frac{\sigma_{a,h}}{\sigma_{w,h}}
+\end{equation}
+
+where $\sigma$ measures the mean variation, across time, and across all data points in
+a given sample, $\sigma_{a}$ is the total variation of a dataset, and $\sigma_{w}$
+is the average of the within-class variations ($h$ indicates healthy). An $S$ value of 0.0 indicates that the
+grasps for the different object classes vary to the same degree as the healthy network.
+
+The metric $S$ allows us to differentiate overall task performance improvement
+and the ability of the co-processor to condition the grasp on the input object based on visual information.
+To successfully grasp in the real world, the hand must be preformed appropriately for
+the shape of the object being grasped, and must close around it appropriately.
+Note that the structure of the delayed reach-to-grasp task can potentially be learned by the co-processor,
+(e.g. to hold muscle velocities to 0.0 for the initial part of each trial) without the co-processor
+also learning to differentiate the various object shapes from the brain observations.
+Thus, $S$ is useful for differentiating improvement in object differentiation as opposed only overall
+decrease in task loss.
+
+\subsection{Experiment 1: AIP 75\% lesion}
 \begin{figure}[h]
-\includegraphics[width=\textwidth]{training_results.png}
-\caption{Loss and class separation vs training epoch}
-\centering
-\label{fig:results}
-\end{figure}
-
-\begin{table}[]
+\centering
+\includegraphics[scale=1.3]{training_results_aip.png}
+\caption{Training Results: AIP 50\% Lesion}
+\label{fig:results_aip}
+\end{figure}
+
+Our first experiment tested the ability of the co-processor to compensate for the loss of a significant
+fraction of neurons in the area AIP of grasping model, which is responsible for encoding object shapres
+from visual inputs. Loss of AIP neurons results in the network model, without further adaptation, being unable to
+condition its grasp on the object shape, after largely succeeding in reaching toward the object.\footnote{Note
+in Supplementary Materials \ref{sup:results} that lesioned loss is lower in this experiment than
+in others: the user continues to reach successfully, but not to form the hand properly, as
+explore above. Likewise, the co-processor doesn't have sufficient information to improve
+performance, for lack of that information.} See Fig. \ref{fig:results_aip}a. Likewise, the co-processor cannot
+observe the object shape encoding in the brain, because that encoding is no longer happening. As a result,
+further recovery is not possible. Note in Fig. \ref{fig:results_aip}b that the co-processor never learns to
+separate object classes, which would have been indicated by an $S$ value near $0.0$.
+
+In the co-adaptive case, the brain and co-processor find a solution that reaches 76\% recovery towards non-lesioned
+performance. This demonstrates that the co-processor can adapt to the non-stationarity
+of the mapping between the observed brain activity and the CPN-delivered stimulation.
+As the brain adapts, it learns to encode the input visual information, which in turn allows the co-processor
+to observe that information through brain recordings. The co-processor can then leverage that information to condition
+stimulation. See Fig. \ref{fig:results_aip}c. Note, though, that the simulation is not designed to indicate if
+the co-processor speeds up recovery, or provides better results than natural recovery, since we aren't modeling the
+timeline of natural recovery. With this lesion design, simulated recovery would in-fact allow near-complete recovery
+of task performance. As a result, this experiment intends only to demonstrate co-adaptation.
+
+In Fig.~\ref{fig:results_aip}d, which plots the object separability metric $S$ discussed above, we see that initially
+the co-processor and brain do not strongly separate the object classes, resulting in negative $S$ values, but as the
+training proceeds, the metric exceeds 0.0, indicating separability. This can be attributed to the co-processor and the
+simulated ``brain'' learning how to map the input visual information about object shape to the appropriate grasp for the object.
+
+In Figs \ref{fig:results_aip}a and c, we see the co-processor initially causes worse performance, as
+indicated by the negative recovery rate. At this point the CPN is newly initialized, and has therefore
+not yet learned an interpretation of the recordings, or the problem structure.
+
+\subsection{Experiment 2: M1 50\% lesion}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=1.3]{training_results_m1.png}
+\caption{Training Results: M1 50\% Lesion}
+\label{fig:results_m1}
+\end{figure}
+
+In this experiment, we lesioned 50\% of the M1 module of the grasping network model. 
+In the non-coadaptive case, training quickly plateaus at 75\%, and does not improve
+thereafter. This suggests the lesion inactivated some of the degrees of freedom needed to
+control the output layer, resulting in a reduced ability to recreate the different target grasps
+for different objects. See Fig. \ref{fig:results_m1}a.
+
+% Note to self: we could probably quantify the loss of degrees-of-freedom. i.e. some "reachability
+% gamut" calculation, based on measuring the dynamic range of network outputs, varying only the ouputs
+% M1 neurons that weren't lesioned.
+% Add note on the relevance of this to the co-processor idea: if we don't have descending neural
+% neural connections, we can't recover. The co-processor needs a pathway to the muscles.
+
+In the co-adaptive case, performance likewise hits an inflexion point at 75\%, but
+continues to improve slowly. The ability to improve further is tied to the
+``natural'' recovery occurring in the co-adapting network model, which restores some of the
+degrees of freedom needed for improving grasping performance. The co-processor in this case adapts
+successfully to this non-stationarity due to the recovery process. See Fig. \ref{fig:results_m1}c.
+
+\subsection{Experiment 3: F5 and M1 connection lesion}
+\label{sec:results_con}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=1]{training_results_con.png}
+\caption{Training Results: F5-M1 Connection Lesion}
+\label{fig:results_con}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=1]{training_results_no_explore.png}
+\caption{Training Results: F5-M1 Connection Lesion. Use of exploratory
+         and randomized examples in the training data set lead to stabilized
+         and faster training. (a) Without exploratory examples in the
+         neighborhood of the CPN, we have unstable and slower training. In
+         this case, the data set is composed of $10\%$ examples drawn from the
+         CPN, and $90\%$ randomized examples (no exploration).
+         (b) Using only examples from the CPN and/or exploratory examples also
+         leads to unstable training. Here we additionally show experiments with
+         $90\%$ exploratory examples (no noise), and $100\%$ examples drawn from
+         the CPN (neither).}
+\label{fig:results_con_no_explore}
+\end{figure}
+
+In this experiment, we disconnected F5 from M1 completely to test whether the co-processor can
+act as a bridge between the two areas to appropriately convey information required for the grasping
+task from one area to the other. As in the other experiments, the co-processor's loss in this experiment
+improved quickly, but took much longer to refine the stimulation patterns to enable 
+object differentiation. See Fig. \ref{fig:results_con}a and c.
+
+Since feedback from the output cannot ``backpropagate'' to the F5 and AIP modules, co-adaptation and
+learning in this experiment only affects the M1 module. As a result, this experiment demonstrates
+that the co-processor's learning algorithm is capable of adapting to non-stationarity
+in the network model's mapping between stimulation parameters and the behavioral output (muscle velocities
+for grasping). These results contrast  with the results of Experiment 1, where the non-stationarity affected
+the mapping between the visual inputs and the AIP outputs.
+
+As seen in Fig.~\ref{fig:results_con}b and d, object separation is initially low and unstable, as the
+co-processor begins to learn the task. Afterwards, it stabilizes as the co-processor begins to
+leverage observed brain activity to differentiate object shapes, gradually converging towards the
+pre-lesion grasping performance.
+
+To illustrate the effect of EN training data set composition as described in \ref{sec:training}, we
+performed additional experiments where we vary the data set composition. We chose to base those
+experiments on the F5-M1 connection lesion arbitrarily. As outlined in Training (\ref{sec:training}),
+the training data set for the EN is composed of stimulation examples drawn from the current CPN,
+examples drawn from other CPNs with parameters in the neighborhood of the current CPN in parameter
+space (referred to as ``exploratory examples''), and randomized stimulation. In Fig.
+\ref{fig:results_con_no_explore} we see the effect of removing the latter two of those.
+Exploratory examples proved necessary to stabilize training, and to achieve higher levels of recovery,
+as in Fig. \ref{fig:results_con_no_explore}a. Removing the examples of randomized stimulation results in
+even more unstable training, where no improvement in task performance occurs.
+
+\subsection{Experiment 4: Connection lesion with recovery}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=1]{training_results_recov.png}
+\caption{Training Results: Connection Lesion with Recovery}
+\label{fig:results_recov}
+\end{figure}
+
+It is often the case that after a stroke, the brain recovers some of the function lost immediately after the
+stroke. We simulated this recovery in the grasping network model after lesioning the F5-M1 connections by
+training the lesioned network on the grasping task. Complete recovery is impossible because the M1 module's
+connections to F5 (both forward and backward) are disconnected. The grasping network model in this case learns
+an object-agnostic stereotypical grasp. We tested whether a co-processor can improve grasping performance by
+propagating to M1 the object-related information from F5 necessary to tailor the grasp to the current
+input object's shape.
+
+We found that the co-processor can successfully improve task performance beyond the ``natural''
+recovery after stroke (Fig.~\ref{fig:results_recov}a). The dashed line
+indicates the task loss after initial recovery. As in some experiments above, loss is
+initially higher at first, as the co-processor trains. Eventually however it learns
+to drive the task loss lower, as it learns to forward-propagate information from the
+``earlier'' parts of the simulated brain. Also, note the object class separation metric $S$
+returned to a healthy value towards the end of co-processor training
+(Fig.~\ref{fig:results_recov})b.
+
+\subsection{Experiment 5: Connection lesion with sensor drift}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=1]{training_results_drift.png}
+\caption{Training Results: Connection Lesion with Sensor Drift}
+\label{fig:results_drift}
+\end{figure}
+
+In our final experiment, we again lesioned the F5-M1 connections and allowed co-adaptation by the
+simulated model network but additionally, allowed the sensor readings to drift over time. This
+drift was modeled in the recording function, where the readings had a non-stationary zero point
+across epochs. We added a bias term which we updated between epochs according to a random process.
+See Supplementary Materials \ref{sup:drift} for details. Sensor-related non-stationarities are
+often seen in real-world neural recording systems due to a variety of factors, 
+from impedance changes due to sensor movement to biological scar tissue formation.
+
+As seen in Fig.~\ref{fig:results_drift}, the co-processor was able to quickly 
+recover the reach-to-grasp part of the task, and gradually learned to
+condition the grasp on the object information. Because the recording model is non-stationary
+in this experiment, the object class separation exhibits far greater epoch-to-epoch
+variance in the later epochs. The co-processor presumably becomes more
+reliant on upstream visual information over time, as it learns how to leverage this
+in order to differentiate object shapes.\footnote{For this experiment, we opted to stop training
+after running the experiment long enough to get results near to the version
+of this experiment without sensor drift.}
+
+As expected, training efficiency was decreased relative to
+the experiment without sensor drift: with the sensor drift,
+282K training epochs was needed to reach $85\%$ recovery while
+without sensor drift,  only 93K was required.
+
+\section{Discussion and Future Work}
+\label{sec:discussion}
+
+We present here a first-of-its-kind demonstration of a novel design for neural co-processors,
+allowing for an AI agent to learn neural stimulation policies that improve a user's performance
+of an external task. The design revolves around the use of a stimulation model for training the
+agent, providing a proxy to the true function mapping stimulation and neural activity
+to task performance. We base our demonstration on a simulation of a neural circuit engaged in
+an external grasping task.
+
+Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
+75-90\% across our various experiments. To work this well, the co-processor needed to adapt to
+the long-running dynamics of the simulated network, as well as the long-running effects of
+stimulation. In some experiments we required it to additionally adapt to non-stationarity in
+the neural circuit, which was actively changing at the same time the co-processor was learning, which it achieved as well.
+In one experiment, it also adapted to a simulated brain which had already undergone some amount of
+lesion recovery. In that case, the co-processor successfully identified the information upstream
+from the lesion which was necessary to stimulate the motor cortex (M1) module
+downstream. In the final experiment, it also successfully adapted to a non-stationary recording model.
+
+A previous example of such a simulation approach is the work of Dura-Bernal et al. \cite{bernal.sim}. In this work, the
+authors used a simulated spiking neural network to train a stimulation agent.
+Their stimulation agent sought to restore the network's
+control of a simulated arm to reach a target, after a simulated lesion was
+applied. As the authors point out, we have a limited ability to
+probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
+we first need to design our approach through the use of an admissible simulation.
+As we did, the authors in this case simulated lesions by effectively removing parts
+of their simulated network, or by cutting connections between parts of the network.
+
+Clearly, we need to discuss what can and cannot be inferred from simulation studies
+such as these. We note that our simulation differs drastically from an organic brain, both in scale and structure. Today, predicting long-running neural responses
+to stimulation remains a difficult problem, implying targeted neural control is also difficult. Even
+more so, identifying the neural correlates of complex task success remains largely out of reach
+\cite{khanna.openloop}. What, then, does our simulation provide us?
+
+Our simulation does not constitute evidence that our method will, for example, immediately translate
+to true restoration of fine-grained control for individual fingers of a stroke patient. One clear
+challenge with doing so would be the sheer dimensionality of natural brain dynamics, and how
+that compares to the resolution of stimulation the co-processor can learn to apply.
+There exists a  mismatch between the dimensionality of the underlying problem, sensor and
+stimulator technologies, and the amount of training data which can reasonably be collected to
+train a closed-loop neural controller. We argue that while this fact clearly creates a challenge
+for learning-based closed-loop stimulation, that nevertheless the insights we generate through this
+simulation are likely to be applicable.
+
+First, this simulation requires the co-processor to contend with issues of long-running
+dynamics, high dimensionality, and non-stationarity, which are likely to be key issues
+facing real world deployments. We designed our training algorithm to contend with dimensionality
+by training the CPN with a local stimulation model. The algorithm adapts to non-stationarity by
+regularly updating that stimulation model, allowing it to adapt to changes in the brain, in
+addition to ``following'' the CPN through stimulation parameter space.
+
+Second, the co-processor abstractly lends itself to any closed-loop neural stimulation problem
+where a relevant stimulation model can be identified. There are many lower dimensional problems
+today which we address with far simpler stimulation regimes than would be necessary for
+e.g. fine-grained finger control. In some applications, such as Parkinson's disease symptom relief,
+stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
+parameters, where the key challenge is learning when to apply the stimulation, and
+its shape and power. Our simulation demonstrates the ability of our training method to
+adapt to higher dimensional problems than these, where additionally the long-running
+effects of stimulation on the intended figure-of-merit must be modeled. These properties
+together suggest it may be useful in lower dimensional problems where a stimulation model
+can be estimated, and stimulation parameters can be tied to the target objective.
+
+Our future work will include a reinforcement learning (RL) approach, where the co-processor
+explicitly learns when to apply stimulation, in addition to parameterizing it. Its target
+objective involves not only symptom relief, but also minimizing energy use. In this approach,
+our CPN and EN may correspond to the Actor and Critic portions of an Actor-Critic model
+for example. Here, the Critic would learn to value symptom relief, while attaching a cost
+to stimulation, thus incentivizing an energy efficient approach.
+
+One challenge remaining with our current co-processor design is training efficiency.
+On the whole, the co-processor quickly improved task performance, but required orders of
+magnitude more training examples to achieve its highest levels of performance. Even
+moderate amounts of recovery may be valuable to a user, but nevertheless we consider
+training efficiency to remain a problem. Due to limits on patient fatigue, time, implant
+battery life, and other concerns, it is not plausible to expect a learning algorithm
+to have access to unlimited amounts of training data. As a result, it is necessary to
+make efficient use of the data we can acquire. We believe there exist at least three
+mitigations for that:
+
+\begin{itemize}
+	\item Retraining an existing EN, rather than regularly creating a new one. As
+	      mentioned above, we encountered difficulty with this approach, but it remains
+	      to be seen if this is a fundamental problem with our training method, rather
+	      than a peculiarity of our simulation. Also, it remains unclear if a solution
+	      for our simulation exists as well, which we've simply yet to identify. This
+	      remains a future area of inquiry.
+	\item Making better use of what data we acquire. In this initial simulation we do
+	      not retain data beyond the present training epoch because non-stationarity requires us to
+	      regularly discard or discount data as it ages. We found that in the presence
+	      of a rapidly learning CPN, and especially in our co-adaptive experiments,
+	      data retention in-fact caused training instability such that training
+	      eventually required more data collection than otherwise. However, in
+	      practice, training epochs operate on the order of seconds, suggesting
+	      that data can be retained and reused for multiple epochs. The `speed' of 
+	      non-stationarity compared to the co-processor's learning therefore is an
+	      area for future research.
+	\item Matching the dimensionality of the stimulation and recording parameters
+	      to the amount of available data. A stimulation regime with
+	      a small number of controllable parameters that nevertheless allows
+	      improvements in task performance  will likely require less training data.
+\end{itemize}
+
+Future work will test the neural co-processor framework on  lower dimensional
+problems that are likely to transfer more readily to \textit{in vivo} experiments
+and human trials. We hypothesize that the co-processor framework will transition well
+to treatments for conditions such as Parkinson's disease (PD), and essential tremor
+(ET). For example, Castaño-Candamil et al. \cite{castano.pd} used an ML-based approach
+to closed-loop stimulation for ET where a simple learning model extracted neural
+markers (NMs) from a human patient's M1 cortex and used the NMs to determine deep
+brain stimulation (DBS) stimulation parameters. Our co-processor model is
+likely to transition well to such an application by being able to extract neural
+markers appropriate for various contexts (e.g. sitting or standing) and applying 
+stimulation appropriate to the current context.
+
+\section{Conclusion}
+Using a simulated model of circuits in the primate brain involved in grasping, we
+demonstrated that a closed-loop neural stimulator, called a ``neural co-processor''
+can be trained to restore grasping function after different types of lesions to the
+model. We showed a neural co-processor can be  trained using supervised learning
+through backpropagation with the help of two networks: a co-processor network (CPN)
+that learns to generate stimulation patterns to optimize task performance, and an
+emulator network (EN) that learns to predict the effects of stimulation. Our results
+show that a neural co-processor can improve task performance after a lesion by
+co-adapting with the simulated brain model in the presence of nonstationarities in the
+brain and sensors. Given the generality of the framework, we expect neural
+co-processors  to be applicable to a wide range of clinical applications that
+required adaptive closed-loop neural stimulation.
+
+\section{Data availability statement}
+Code including analysis code used to
+generate figures can be found at
+\url{https://github.com/mmattb/coproc-poc}.
+Data is available upon reasonable request to the authors.
+
+\section{Conflict of interest statement}
+The authors of this work are not aware of any conflicts of interest
+related to it.
+
+\section{Acknowledgements}
+This work was supported by a Weill Neurohub Investigator grant, a CJ and Elizabeth Hwang
+endowed professorship (RPNR), and National Science Foundation grant no. EEC-1028725.
+The authors would like to thank Karunesh Ganguly (UC San Francisco), Priya Khanna
+(UC Berkeley), Anca Dragan (UC Berkeley), Justin Ong, and Luciano De La Iglesia
+for discussions and insights related to this work.
+
+\section{References}
+\bibliographystyle{iopart-num}
+\bibliography{refs}
+
+\newpage
+
+\section{Supplementary Materials}
+\subsection{Details of the simulated task.}
+\label{sup:michaelstask}
+
+The reach-to-grasp task on which the mRNN was trained involves multiple phases, which can be identified by the
+mRNN's inputs. Initially, the NHP subject sat in a dark room, with only a red cue visible. After a short period,
+an object was presented to the subject visually, after which the overhead light was turned off and once again only the red
+cue was visible. After another brief period, the red cue was turned off, which cued the subject to perform a grasp and hold of
+the object (still in the dark). This task is presented in more detail in the work of Susillo \cite{susillo.mrnn} and Michaels et al.
+\cite{michaels.mrnn}.
+
+The inputs to the mRNN network encode the task phase information. It consists of a twenty dimensional vector of visual
+features, which are presented only to the input (AIP) end of the network, and a 1D hold signal, which is presented to all
+neurons across the network. The authors of Michaels et al. \cite{michaels.mrnn} derived the hold signal
+from the grasp data, and it represents a point in time shortly before grasping began. As explained above, the visual features are
+drawn from VGGNet \cite{simonyan.vgg}, based on rendered images. The images consist of some combination of the red cue
+light and the object image. Note that the visual input vector and hold signal are not observed directly by the co-processor,
+but rather indirectly via brain recordings.
+
+The phases of the task are depicted in Table \ref{tab:phases}, and the timeline of a single trial is depicted in Fig. \ref{fig:trial_timeline}.
+
+\begin{table}[h!]
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{Description}                                                                                                               & \textbf{Cue} & \textbf{\begin{tabular}[c]{@{}l@{}}Object\\ image\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Hold\\ signal\end{tabular}} \\ \hline
+A rest period, during which a red cue light is presented                                                                           & On           & Off                                                             & On                                                             \\ \hline
+\begin{tabular}[c]{@{}l@{}}The object presentation period, during which the\\ image of the object and cue are visible\end{tabular} & On           & On                                                              & On                                                             \\ \hline
+A rest period                                                                                                                      & On           & Off                                                             & On                                                             \\ \hline
+Cue is turned off                                                                                                                  & Off          & Off                                                             & On                                                             \\ \hline
+Subject performs grasp                                                                                                             & On           & Off                                                             & Off                                                            \\ \hline
+\end{tabular}
+\caption{\label{tab:phases} Phases of the reach to grasp task. During each phase, a visual stimulus as well as a binary hold signal are presented.
+The visual stimulus encodes an image, consisting of a red cue light and/or an image of an object.}
+\end{table}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.7]{trial_timeline.png}
+\caption{Timeline of a Single Trial}
+\label{fig:trial_timeline}
+\end{figure}
+
+\subsection{Training the grasping model network.}
+\label{sup:michaelstraining}
+
+For our simulation, we leverage a pre-trained mRNN model, provided by the
+lead author of Michaels et al. \cite{michaels.mrnn}, and used with their permission.
+The training method for that model is based on prior work by the same authors, and
+others \cite{susillo.mrnn}.
+
+Training involved the use of Hessian-free optimization \cite{martens.hessianfree},
+rather than the more common first-order stochastic gradient descent. Additionally, an L2 firing
+rate regularization, and an L2 input and output model weight regularization were used.
+These regularizations are believed to result in more biologically plausible RNNs \cite{kao.sim}.
+
+The authors varied the model's nonlinearities, regularization weights, and
+inter-module sparsities. Among those, we chose one model arbitrarily for our study. Specifically,
+the model we chose was structured, and trained using:
+
+\begin{itemize}
+    \item The rectified hyperbolic tangent nonlinearity
+    \item $1\mathrm{e}{-1}$ Inter-module sparsity
+    \item $1\mathrm{e}{-3}$ L2 firing rate regularization
+    \item $1\mathrm{e}{-5}$ L2 weight regularization
+\end{itemize}
+
+\subsection{EN/CPN Interleaving.}
+\label{sup:encpninter}
+We train a CPN until an EN is no longer useful. At that point we train a new EN. We deem an EN
+as no longer useful if one of the following predicates becomes true:
+
+\begin{itemize}
+	\item The EN prediction loss is greater than $min(6\mathrm{e}{-4}, L/10)$, where $L$ is the most recent
+	      task loss. EN prediction loss is an MSE loss between the EN's prediction of muscle
+	      trajectories, and the actual muscle trajectories output by the grasping model network.
+	\item Task loss increased across 15 of the previous 30 training epochs.
+	\item 100 CPN training epochs have elapsed.
+\end{itemize}
+
+Likewise, the EN training period ends when its prediction loss $L$ on the validation data set drops
+below a threshold of $max(3\mathrm{e}{-4}, L/50)$.
+
+\subsection{Sensor drift.}
+\label{sup:drift}
+To simulate sensor drift, we add a vector to a bias term introduced to the recording function.
+We draw the elements of the vector from a zero-mean Gaussian distribution, with variance based
+on the mean value of the recording function from prior connection-lesion experiments. That
+allows us to put it into a reasonable range, where it is effective but not extreme. We attempted
+the experiment with several values of the variance, and found the results to be principally the same:
+the co-processor eventually learned, but at a rate slower than otherwise. We present results for
+a variance of $1.5\mathrm{e}{-3}$, which is $2\%$ of the mean recording value.
+
+\subsection{Passthrough reference model.}
+\label{sup:passthrough}
+
+To understand the effect of the recording and stimulation functions on the co-processor's performance,
+we repeated our experiment with the F5-M1 connection lesion and coadaptation, with passthrough recording
+and stimulation functions. ``Passthrough'' here means that the recording and stimulation functions have
+dimensionality equal to the number of neurons we are recording from or stimulating, respectively, and that
+there is no temporal or spatial smoothing used. That allows the co-processor to directly observe the hidden
+state of the AIP and F5 neurons in the simulated grasping circuit, and to directly influence the hidden state
+of the neurons in M1 it stimulates.
+
+For this passthrough version of the experiment, we additionally increased the number of artifical neurons
+making up the CPN and EN, to account for the additional dimensionality of the inputs and outputs. Specifically,
+we increased the CPN from 61 to 200 neurons, and the EN from 87 to 351. We measured whether the increase in
+neurons alone caused in the different results we see here, and found it did not; we omit those results for
+brevity.
+
+In Fig. \ref{fig:results_passthrough} we see that the co-processor performed significantly better with
+the passthrough functions. It learned faster, and to a higher recovery level: $90\%$ compared with $97\%$.
+Note: we truncate the results to the point we stopped the passthrough experiment. The reader can find the
+final results in Supplementary Materials \ref{sup:results}.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.7]{training_results_passthrough.png}
+\caption{\textbf{Training Results: passthrough recording and stimulation functions.} The significantly faster
+         training and higher recovery suggest the recording and stimulation functions make the task
+         more difficult for the co-processor.}
+\label{fig:results_passthrough}
+\end{figure}
+
+\subsection{Effect of observability on the recording function and training results.}
+\label{sup:observability}
+
+To study the effect of brain observability on the co-processor's training behavior
+we repeated the co-adaptive connection lesion experiment (\ref{sec:results_con})
+with varying recording functions. Here we varied the number of simulated electrodes
+in each mRNN module, and additionally performed an experiment with a passthrough recording
+function, as described in the prior section. Specifically, our original experiment used
+$20$ electrodes per module. Our additional experiments include $1$, $10$, and $60$ electrodes
+per module, and the passthrough recording function. This allows us to explore any
+gradient in the co-processor training results with respect to observability of the 
+brain. Additionally, we reduced the variance of the Gaussian governing
+the sensor model, such that an experiment with e.g. $1$ electrode per
+module measures effectively only 1-2 neurons per module. We held the number of
+neurons making up the CPN and EN fixed across these variations to ensure we
+are examining the effects of the recording function alone. We continued to stop
+learning at the point of $90\%$ recovery or a slow rate of recovery. In Fig.
+\ref{fig:results_observability} we present the results.
+
+We did not experiment with the placement of the electrodes, i.e. varying their
+location among the neurons. It wasn't necessary to do so for the purpose of this
+experiment. However, it's conceivable that results could improve considerably if we
+happened to observe a small handful of key neurons.
+
+Overall, learning is effectively the same regardless of the recording 
+function, up to the point that only a small number of neurons can be 
+observed. In all cases except the 
+case of a single electrode, the co-processor learned to improve task
+performance to the $90\%$ recovery point. Some variation in learning
+efficiency occurred, and perhaps notably the experiments with $1$, $10$
+electrodes observed the lowest two rates of recovery. The passthrough 
+function and $60$ electrode model observed the fastest recovery rates.
+Without additional computational power it remains difficult to hypothesis
+test the effect of electrode count against learning efficiency, but the
+results suggest that small variations in electrode count away from the
+count of $20$ used in our main experiments would not cause drastically
+different results.
+
+Class separation behavior remained largely the same across these variations. The one exception is the experiment
+involving only $1$ electrode per module. In that case, the co-processor
+slowly learned to treat the object classes similarly. That suggests that
+the classes could not be differentiated by the small amount of information
+attainable from a single electrode. In this case, the co-processor reduced
+towards a simpler closed-loop stimulator which modulates its behavior
+entirely on the user's volition to initiate movement, due to the hold
+signal being observable from every neuron.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.7]{training_results_observability.png}
+\caption{\textbf{Training Results: varying observability of the brain.}
+         (a) Increasing observability of the brain leads to faster learning. Here
+         we show results with a single simulated recording electrode in each module
+         ``1'', compared with ``20'' electrodes for the results reported in the main
+         portion of this paper, and compared with other electrode counts. ``pass''
+         refers to a passthrough recording function where every simulated neuron was
+         directly measured. Note that learning is generally the same as observability
+         increases, though perhaps somewhat faster. We did not hypothesis test that effect.
+         In all cases except ``1'', learning reached our $90\%$ recovery threshold where
+         we stop training. (b) As in our connection-based lesion experiment above
+         (\ref{sec:results_con}), class separation initially varied, then trended towards the
+         separation exhibited by the healthy mRNN network. However, in
+         the single electrode experiment ``1'', class separation steadily decreased, suggesting
+         that the information necessary to perform class separation cannot be determined from
+         the information available to the single electrode.}
+\label{fig:results_observability}
+\end{figure}
+
+\subsection{Table of losses and recovery.}
+\label{sup:results}
+
+\begin{table}[h]
 \begin{tabular}{|l|l|l|l|l|l|}
@@
 Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
-Pre-recovery*  & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
+Recovery*      & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
+Sensor drift   & 0.020719                                                          & 0.002966                                                        & 0.002825                                                              & 88.10\%            & 88.83\%                                                          \\ \hline
+Passthrough    & 0.020719                                                          & 0.001095                                                        & 0.001132                                                              & 97.37\%            & 97.19\%                                                          \\ \hline
 \end{tabular}
-\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
+\caption{\label{tab:results}Losses and recovery. *Lesioned loss and percent recoveries based on the post-recovery values.
+         ``Coadapt'' refers to co-adaptation. ``Con'' refers to F5-M1 connection lesions. ``Recovery'' refers to the
+         experiment involving recovery prior to co-processor training.
+         For further exploration of the results, see Results.}
 \end{table}
-
-With each experiment, we track two metrics to characterize the learning progress. First, we
-track the task loss, which is an MSE loss that measures the co-processor's ability to restore
-movement towards the target trajectory. We characterize that measure in terms of its percent
-difference from lesioned and healthy performances.
-
-Second, we measure the degree to which the co-processor's solution successfully
-differentiates object classes. We measure the ratio of within-class and total variation,
-and compare that to the healthy network. Our separation metric $S$ is defined as:
-
-\begin{equation}
-	S = \frac{\sigma_{a}}{\sigma_{w}} - \frac{\sigma_{a,h}}{\sigma_{w,h}}
-\end{equation}
-
-where $\sigma$ measures the mean variation, across time, and across all data points in
-a given sample. $\sigma_{a}$ is the total variation of a dataset, and $\sigma_{w}$
-is the average of the within-class variations. The $h$ indicates the healthy
-metric. An $S$ value of 0.0 indicates that the grasps for the various object classes
-vary to the same degree as the healthy network.
-
-This metric $S$ allows us to differentiate overall task performance improvement
-and the ability of the co-processor to condition the grasp based on visual information.
-The structure of the delayed reach-to-grasp task can be learned by the co-processor,
-i.e. to hold muscle velocities to 0.0 for the initial part of each trial, without it
-also learning to differentiate the various object shapes from the brain observations.
-$S$ is our metric for differentiating those two co-processor behaviors.
-
-\subsection{Experiment 1: AIP 75\% lesion}
-This experiment demonstrates the ability for the co-processor to succeed after the
-patient has lost some of the neural machinery responsible for representing the objects' shapes.
-In the non-coadaptive version of this experiment, the loss of AIP neurons involved
-in encoding the visual inputs results in the simulated brain being unable to
-condition its grasp for the object shape, though reaching largely still succeeds.
-Likewise, the co-processor doesn't have sufficient information to improve performance,
-for lack of that information.
-
-In the co-adaptive case, the brain and co-processor find a solution that largely recovers
-performance. This demonstrates that the co-processor can adapt to the non-stationarity
-of the mapping between the observed brain activity and the optimal stimulation. Note,
-though, that the simulation is not designed to indicate if the co-processor speeds up
-recovery, or results in greater results than natural recovery>
-
-Additionally, in Fig. \ref{fig:results} we see that initially the co-processor and
-brain did not strongly separate the object classes, but that they learned to as the
-training proceeded. This is again due to the co-processor and brain learning how
-to map the input visual information on to the task.
-
-\subsection{Experiment 2: M1 75\% lesion}
-In this experiment, 75\% of the output module of the simulated network has been
-lesioned. In the co-adaptive case, the co-processor needs to adapt to
-non-stationarity in the mapping between stimulation and task performance.
-In the non-coadaptive case, training quickly plateaus, and does not improve
-thereafter. That suggests we have lost some of our degrees of freedom to control
-the activations in the output layer, resulting in a reduced ability to recreate
-the target grasps.
-
-In the co-adapative case, performance likewise hits an inflexion point, but
-continues to improve slowly. The ability to improve further is tied to the
-simulated recovery restoring the degrees of freedom needed for success. The
-co-processor in this later phase of training is tasked with adapting to the
-non-stationarity that results from the recovery process, which it does
-successfully.
-
-\subsection{Experiment 3: F5 and M1 connectivity lesion}
-As in the other cases, the co-processor's loss in this experiment improves
-quickly, initially, but taskes much longer to refine. During that refinement,
-object differentiation occurs. Note that in this experiment, no information
-can propagate forward to the M1 module, or backwards from it, since we have
-completely severed the inter-module connections involving that module. It is
-the sole responsibility of the co-processor to perform that
-forward-propagation.
-
-In this case, co-adaptation only affects the M1 module, since we have nothing
-to use as a learning signal: there is no backpropagation to the AIP and
-F5 modules. As a result, this experiment demonstrates that the
-co-processor's learning algorithm is capable of adapting to non-stationarity
-in the mapping between stimulation parameters and task performance. Contrast
-that with the results of Experiment 1, where the non-stationarity affected
-the mapping between the visual inputs and the brain observations.
-
-\subsection{Experiment 4: Connectivity lesion with pre-recovery}
-In this case, we simulate some amount of recovery prior to the co-processor
-learning. Complete recovery is impossible, because the output module's
-connections to F5 (both forward and backward) are masked. As a result,
-the pre-recovery causes the simulated brain to learn an object-agnostic
-stereotyped grasp. The co-processor's job is to forward-propagate the
-information necessary to condition the grasp on the object shape.
-
-The co-processor in this scenario successfully improved task performance
-beyond the recovery point. Note that the object class separation metric
-$S$ returned to a healthy value towards the end of the training,
-indicating that the learned solution did not simply refine the
-stereotyped grasp.
-
-\section{Discussion and Future Work}
-\label{sec:discussion}
-
-We present here a first-of-its-kind demonstration of a novel design for neural co-processors,
-allowing for an AI agent to learn neural stimulation policies that improve a user's performance
-of an external task. The design revolves around the use of a stimulation model for training the
-agent, providing a proxy to the true function mapping stimulation and neural activity
-to task performance. We base our demonstration on a simulation of a neural circuit engaged in
-an external grasping task.
-
-Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
-75-90\% across our various experiments. To work this well, the co-processor needed to adapt to
-the long-running dynamics of the simulated network, as well as the long-running effects of
-stimulation. In some experiments we required it to additionally adapt to non-stationarity in
-the neural circuit, which was actively changing at the same time the co-processor was learning.
-It achieved that as well, co-adapting with the simulated brain to improve external task
-performance. Finally, it adapted to a simulated brain which had already undergone some amount of
-lesion recovery. In that case, the co-processor successfully identified the information upstream
-from the lesion which was necessary to stimulate the motor cortex downstream.
-
-Clearly, we need to discuss carefully what can and cannot be inferred from simulation studies
-such as these. As in Dura-Bernal \cite{bernal.sim}, we note that our simulation differs
-drastically from an organic brain, both in scale and structure. Today, predicting long-running
-neural responses to stimulation remains a difficult problem, implying targeted neural control
-is also difficult. Even moreso, identifying the neural correlates of complex task success
-remains largely out of reach \cite{khanna.openloop}. What, then, does our simulation provide us?
-
-Abstractly, the co-processor lends itself to any closed-loop neural stimulation problem
-where a relevant stimulation model can be identified. There are many lower dimensional problems
-today which we address with far simpler stimulation regimes than would be necessary for
-e.g. fine-grained finger control. In some applications, such as PD symptom relief,
-stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
-parameters, where the key challenge is learning when to apply the stimulation, and
-its shape and power. Our simulation demonstrates the ability of our training method to
-adapt to higher dimensional problems than these, where additionally the long-running
-effects of stimulation on the intended figure-of-merit must be modeled. These properties
-together suggest it may be useful in lower dimensional problems where a stimulation model
-can be estimated, and stimulation parameters can be tied to the target objective.
-
-Our future work will include a reinforcement learning (RL) approach, where the co-processor
-explicitly learns when to apply stimulation, in addition to parameterizing it. Its target
-objective involves not only symptom relief, but also minimizing energy use. In this approach,
-our CPN and EN may correspond to the Actor and Critic portions of an Actor-Critic model
-for example. Here, the Critic would learn to value symptom relief, while attaching a cost
-to stimulation, thus incentivizing an energy efficient approach.
-
-One challenge remaining with our current co-processor design is training effiency.
-On the whole, the co-processor quickly improved task performance, but required orders of
-magnitude more training examples to achieve its highest levels of performance. Even
-moderate amounts of recovery may be valuable to a user, but nevertheless we consider
-training efficiency to remain a problem. Due to limits on patient fatigue, time, implant
-battery life, and other concerns, it is not plausible to expect a learning algorithm
-to have access to unlimited amounts of training data. As a result, it is necessary to
-make efficient use of the data we can acquire. We believe there exist at least three
-mitigations for that:
-
-\begin{itemize}
-	\item Retraining an existing EN, rather than regularly creating a new one. As
-	      mentioned above, we encountered difficulty with this approach, but it remains
-	      to be seen if this is a fundamental problem with our training method, rather
-	      than a peculiarity of our simulation. Also, it remains unclear if a solution
-	      for our simulation exists as well, which we've simply yet to identify. This
-	      remains a future area of inquiry.
-	\item Making better use of what data we acquire. In this initial simulation we do
-	      not retain data beyond the present training epoch. In practice, we can likely
-	      retain training data for some time window. Non-stationarity requires us to
-	      regularly discard or discount data as it ages. However, in practice, training
-	      epochs operate on the order of seconds, suggesting that data can be retained
-	      and reused for multiple epochs. The `speed' of non-stationarity compared to
-	      the co-processor's learning therefore must be an area of future research.
-	\item Matching the dimensionality of the stimulation and observations
-	      to the amount of available data. If one chooses a stimulation regime with
-	      a small number of controllable parameters, where nevertheless task improvement
-	      from the stimulation is possible, that allows for a simpler co-processor which
-	      will likely require less training data.
-\end{itemize}
-
-Having now demonstrated our method on a higher dimensional problem, we will next demonstrate
-it on a lower dimensional problem, one which is likely to transfer more readily to
-an \textit{in vivo} trial. We hypothesize that this co-processor model will transition well
-to applications in PD, or essential tremor (ET). In Castaño-Candamil et al. \cite{castano.pd},
-we see an ML-based approach to closed-loop stimulation for ET relief. Here, a simple learning
-model extracts neural markers (NMs) from a human patient's M1 cortex. It uses those to
-condition DBS stimulation parameters. The stimulation controller adapts to non-stationarity
-of the mapping between NMs and optimal control, which is necessary due to the sensitivity
-of NMs to context (e.g. sitting or standing). The co-processor model is likely to
-transition well to such an application, where it may provide additional energy savings
-or reduction of stimulation side-effects. In that case, it will attempt to extract neural
-markers appropriate for each context, to detect transitions in context, and to apply
-optimal stimulation.
-
-\section{Conclusion}
-We have demonstrated a general framework for training a closed-loop neural stimulator,
-called a ``neural co-processor''. We outlined one specific co-processor design, and
-outlined an approach to training it using supervised learning. In simulation it
-successfully learned to alleviate some of the effects of a brain lesion. To do so, it
-needed to co-adapt with the simulated brain, which exhibited non-stationarity. Additionally,
-it needed to learn long-running effects of the stimulation it learned to apply. Because of
-these properites, the co-processor approach likely generalizes well to a wide range of
-clinical applications where closed-loop neural stimulation is appropriate.
-
-\section{Acknowledgements}
-TODO: funding sources
-
-The author would like to thank the following for discussions and insights related to
-the present topic:
-\begin{itemize}
-	\item Karunesh Ganguly, UC San Francisco
-	\item Priya Khanna, UC San Francisco
-	\item Anca Dragan, UC Berkeley
-	\item Justin Ong, University of Washington
-	\item Luciano De La Iglesia, University of Washington
-\end{itemize}
-
-\section{References}
-\bibliographystyle{iopart-num}
-\bibliography{refs}
 \end{document}

### 2023-03-18 cpn_epoch_cpn.py
Fix up tunables in cpn_epoch_cpn for recreating the results of older experiments.O

@@ -1,7 +1,7 @@
 import torch
 
 from cpn_utils import CPNENStats, calc_pred_loss, calc_train_loss, EpochType
-from experiment import utils
+from experiment import lesion, utils
 
 
 class CPNEpochCPN:
@@ -38,8 +38,16 @@ class CPNEpochCPN:
 
         if cfg.recover_after_lesion:
             self.set_opt_lr = self.lr_sched_aggressive_refine3
+            self.cpn_epoch_max_len = 200
+        elif (type(cfg.lesion_instance) is lesion.LesionType.outputs.value and
+                cfg.lesion_instance.start_idx == 0 and
+                cfg.lesion_instance.end_idx == 50):
+            # M1 is an odd beast...
+            self.set_opt_lr = self.lr_sched_aggressive_refine3
+            self.cpn_epoch_max_len = 100
         else:
             self.set_opt_lr = self.lr_sched_standard
+            self.cpn_epoch_max_len = 200
 
         self.reset()
 
@@ -210,7 +218,7 @@ class CPNEpochCPN:
         self.recent_task_losses.append(rtl)
         # if (loss_history.max_pct_recov > 0.99 and not self.cfg.dont_train and
         if (
-            loss_history.max_pct_recov > 0.9
+            loss_history.max_pct_recov > 0.99
             and not self.cfg.dont_train
             and loss_history.lesioned_loss > loss_history.healthy_loss
         ):
@@ -220,7 +228,7 @@ class CPNEpochCPN:
             we_are_done = True
             en_is_ready = False
         # For now: just run for awhile
-        elif self.cfg.dont_train and loss_history.eidx == 200000:
+        elif self.cfg.dont_train and loss_history.eidx == 250000:
             we_are_done = True
             en_is_ready = False
         else:
@@ -232,8 +240,7 @@ class CPNEpochCPN:
             elif self.recent_pred_loss > max(rtl / 10, 6e-4):
                 en_is_ready = False
                 self.reset_period()
-            # elif self.checkpoint_eidx >= 200:
-            elif self.checkpoint_eidx >= 100:
+            elif self.checkpoint_eidx >= self.cpn_epoch_max_len:
                 en_is_ready = False
                 self.reset_period()
             elif self.checkpoint_eidx > 30:

### 2023-03-18 experiment/config.py
Add plumbing for AIP/F5 connection lesion experiment, if we ever want to try.

@@ -37,6 +37,7 @@ class Config:
     drifting_obs: bool
     drop_m1: bool
     drop_f5: bool
+    drop_neuron_mask: torch.Tensor
     cuda: typing.Any
     holdout_pct: float
 
@@ -170,6 +171,7 @@ def get(
     obs_sigma=DEFAULT_OBS_SIGMA,
     out_dim=DEFAULT_OUT_DIM,
     holdout_pct=DEFAULT_HOLDOUT_PCT,
+    drop_neuron_mask=None,
     recover_after_lesion=False,
     coadapt=False,
     dont_train=False,
@@ -301,6 +303,9 @@ def get(
 
     dataset, loader_train, loader_test = get_dataset(holdout_pct=holdout_pct, cuda=cuda)
 
+    if cuda is not None:
+        drop_neuron_mask = drop_neuron_mask.cuda(cuda)
+
     cfg_out = Config(
         observer_instance,
         stimulus,
@@ -323,6 +328,7 @@ def get(
         drifting_obs,
         drop_m1,
         drop_f5,
+        drop_neuron_mask,
         cuda,
         holdout_pct,
     )

### 2023-03-18 experiment/experiment.py

@@ -87,14 +87,20 @@ def get_aip_f5_con_lesion_config(cuda=None, coadapt=True, **kwargs):
             (0, 1, 2, 3),
         ],
     )
+
+    drop_neuron_mask = torch.ones(300)
+    drop_neuron_mask[250:] = 0.0
+    drop_neuron_mask[100:150] = 0.0
+
     return get_config(
         cuda=cuda,
         coadapt=coadapt,
-        drop_f5=True,
         lesion_type=lesion_type,
         lesion_args=lesion_args,
-        stim_pad_left_neurons=100,
-        stim_pad_right_neurons=100,
+        num_stim_neurons=100,
+        stim_pad_left_neurons=150,
+        stim_pad_right_neurons=50,
+        drop_neuron_mask=drop_neuron_mask,
         **kwargs,
     )
 
@@ -273,10 +279,11 @@ class Experiment:
         if cfg.drop_m1:
             self.obs_drop_module_idx = 0
         elif cfg.drop_f5:
-            print("Dropping F5")
             self.obs_drop_module_idx = 1
         else:
             self.obs_drop_module_idx = None
+
+        self.drop_neuron_mask = cfg.drop_neuron_mask
 
     @property
     def cfg(self):
@@ -403,7 +410,8 @@ class Experiment:
 
             for tidx in range(steps - 1):
                 obs_raw = self.mike.observe(
-                    self.observer, drop_module_idx=self.obs_drop_module_idx
+                    self.observer, drop_module_idx=self.obs_drop_module_idx,
+                    drop_neuron_mask=self.drop_neuron_mask
                 )
                 obs = obs_raw + (trial_end[:, tidx, :],)
 

### 2023-03-18 experiment/mRNN.py

@@ -355,14 +355,19 @@ class MichaelsRNN(nn.Module):
         ret = self.fc(output[:, : self.num_neurons_per_module])
         return ret
 
-    def observe(self, obs_model, drop_module_idx=None):
+    def observe(self, obs_model, drop_module_idx=None, drop_neuron_mask=None):
         outputs = []
 
         for midx in range(self.num_modules):
             if midx == drop_module_idx:
                 continue
 
-            act = self.prev_output[
+            if drop_neuron_mask is not None:
+                po = self.prev_output * drop_neuron_mask.unsqueeze(0)
+            else:
+                po = self.prev_output
+
+            act = po[
                 :,
                 midx
                 * self.num_neurons_per_module : (midx + 1)

### 2023-03-18 experiment/observer.py

@@ -62,6 +62,37 @@ class ObserverGaussian1d(Observer):
         return f"gaussian{self.out_dim}.{self.sigma}"
 
 
+class ObserverGaussian1dMasked(Observer):
+    def __init__(self, in_dim, mask, out_dim=20, sigma=1.75, cuda=None):
+        super(ObserverGaussian1dMasked, self).__init__()
+        self._in_dim = in_dim
+        self._out_dim = out_dim
+        self._cuda = cuda
+        self.sigma = sigma
+
+        self.norm, self.weights = gaussian_array_weights(
+            in_dim, out_dim, sigma, normalize=True
+        )
+        self.weights = torch.tensor(
+            self.weights.reshape((1,) + self.weights.shape)
+        ).float()
+
+        self.mask = torch.tensor(mask).unsqueeze(0)
+
+        if cuda is not None:
+            self.weights = self.weights.cuda(cuda)
+            self.mask = self.mask.cuda(cuda)
+
+    def reduce(self, x):
+        # Note: weights may broadcast up if we have batches
+        reduced = self.weights @ x.reshape(x.shape + (1,))
+        masked = reduced * self._mask
+        return masked.squeeze(axis=-1).detach()
+
+    def __str__(self):
+        return f"gaussianMasked{self.out_dim}.{self.sigma}"
+
+
 class ObserverGaussian1dDrifting(ObserverGaussian1d):
     def __init__(self, in_dim, out_dim=20, sigma=1.75, drift_std=0.0005, cuda=None):
         super(ObserverGaussian1dDrifting, self).__init__(
@@ -110,5 +141,6 @@ class ObserverPassthrough(Observer):
 
 class ObserverType(enum.Enum):
     gaussian = ObserverGaussian1d
+    masked = ObserverGaussian1dMasked
     drifting_gaussian = ObserverGaussian1dDrifting
     passthrough = ObserverPassthrough

### 2023-03-18 paper/article.tex
v2 paper, after response to reviewers.

@@
 \usepackage{hyperref}
+\usepackage{microtype}
 \graphicspath{ {./figs/} }
@@
 % Authorship temporarily removed for double-anonymous review
-\author{Matthew J Bryan$^{1}$, Linxing Preston Jiang$^{1}$, Rajesh P N Rao$^{1}$}
+\author{Matthew J Bryan$^{1}$, Linxing Preston Jiang$^{1,2,3}$, Rajesh P N Rao$^{1,2,3}$}
 
 \address{$^{1}$ Neural Systems Laboratory, Paul G. Allen School of Computer
-Science \& Engineering, University of Washington, Box 352350,
-Seattle, WA 98105, USA}
+Science \& Engineering, University of Washington, Seattle, WA, USA}
+\address{$^{2}$ Center for Neurotechnology, University of Washington, Seattle, WA, USA}
+\address{$^{3}$ Computational Neuroscience Center, University of Washington, Seattle, WA, USA}
 
@@
 \begin{indented}
-\item[]November 2022
+\item[]March 2023
 \end{indented}
@@
 \begin{abstract}
-\textit{Objective} A major challenge in closed-loop brain-computer interfaces (BCIs)
+\textit{Objective} A major challenge in designing closed-loop brain-computer interfaces (BCIs)
 is finding optimal stimulation patterns as a function of ongoing neural activity for
-different subjects and for different objectives. Traditional approaches, such as
+different subjects and different objectives. Traditional approaches, such as
 those currently used for deep brain stimulation (DBS), have largely followed a manual
@@
 to shape neural activity and bridge injured neural circuits for targeted repair and
-rehabilitation.  Here we investigate a specific type of co-processor called a ``neural
+restoration of function.  Here we investigate a specific type of co-processor called a ``neural
 co-processor'' which uses artificial neural networks (ANNs) and deep learning to learn optimal
@@
 the biological circuit itself adapts to the stimulation, achieving a form of brain-device
-co-adaptation. We tested the neural co-processor's ability to restore function after
-stroke by simulating a variety of lesions in a previously published cortical model
-of grasping.  We additionally tested the ability of our co-processor
-to adapt its stimulation as the simulated cortical network and simulated sensors
-underwent changes.
-\textit{Main results} Our results show that a neural co-processor can restore
-reaching and grasping function after a simulated stroke in a cortical model, achieving
-recovery towards healthy function in the range 75-90\%. Our co-processor successfully
-co-adapted to accomplish the reach-and-grasp task after a variety of lesions to the
-simulated cortical circuit.
+co-adaptation. Here we use simulations to lay the groundwork for future \textit{in vivo}
+tests of neural co-processors. We leverage a previously published cortical model of grasping, to
+which we applied various forms of simulated lesions. We used our simulations to develop
+the critical learning algorithms and study adaptations to non-stationarity in preparation for
+future \textit{in vivo} tests.
+\textit{Main results} Our simulations show the ability of a neural co-processor to learn a
+stimulation policy using a supervised learning approach, and to adapt that policy as
+the underlying brain and sensors change. Our co-processor successfully co-adapted with
+the simulated brain to accomplish the reach-and-grasp task after a variety of lesions
+were applied, achieving recovery towards healthy function in the range 75-90\%.
 \textit{Significance} Our results provide the first proof-of-concept demonstration,
-using computer simulations, of a neural co-processor for activity-dependent closed-loop
-neurosimulation for optimizing a rehabilitation goal after injury. Our results provide
-insights on how such co-processors may eventually be developed for \textit{in vivo} use
-to learn complex adaptive stimulation policies for a variety of neural rehabilitation
-and neuroprosthetic applications. 
+using computer simulations, of a neural co-processor for adaptive activity-dependent closed-loop
+neurostimulation for optimizing a rehabilitation goal after injury. While a significant
+gap remains between simulations and \textit{in vivo} applications, our results provide
+insights on how such co-processors may eventually be developed for learning complex
+adaptive stimulation policies for a variety of neural rehabilitation and neuroprosthetic
+applications. 
 \end{abstract}
@@
 \vspace{2pc}
-\noindent{\it Keywords}: brain-computer interface, neural co-processor, AI, machine learning, deep learning, neural networks, stimulation, computational models
+\noindent{\it Keywords}: brain-computer interface, brain-machine interface, neurostimulation,
+neuromodulation, neural co-processor, AI, machine learning, deep learning, neural networks,
+computational models
 %
@@
 to the control of a wide variety of virtual and physical prostheses through neural signal decoding
-\cite{rao.bcibook, wolpaw.bcibook, moritz.neuro, lebedev.bmi}. Separately, advances in stimulation techniques and modeling have allowed
-us to probe neural circuit dynamics (e.g. \cite{walker.inception}) and learn to better drive neural
-circuits towards desired target dynamics by encoding and delivering information through
-stimulation \cite{niparko.cochlear, weiland.retinal, tomlinson.propr, tabot.tact, tyler.tact,
-dadarlat.tact, sharlene.tact, cronin.tact}. Bi-directional BCIs (BBCIs) allow stimulation to be
-conditioned on decoded brain activity and encoded sensor data for applications such as real-time,
+\cite{rao.bcibook, wolpaw.bcibook, moritz.neuro, lebedev.bmi}. Separately, advances in stimulation
+techniques and modeling have allowed us to probe neural circuit dynamics (e.g. \cite{walker.inception})
+and learn to better drive neural circuits towards desired target dynamics by encoding and delivering
+information through stimulation \cite{niparko.cochlear, weiland.retinal, tomlinson.propr, tabot.tact,
+tyler.tact, dadarlat.tact, sharlene.tact, cronin.tact}. Bi-directional BCIs (BBCIs) allow stimulation
+to be conditioned on decoded brain activity and encoded sensor data for applications such as real-time,
 fine-grained control of neural circuits and prosthetic devices (e.g., \cite{nicolelis.bmbi}).
@@
 Motivated by these advances, we investigate here a flexible framework for combining encoding
-and decoding using ``neural co-processors'' \cite{rao.coproc}, a type of brain co-processor \cite{rao.braincoproc}. Neural co-processors leverage artificial neural networks (ANNs) and deep
+and decoding using ``neural co-processors'' \cite{rao.coproc}, a type of brain co-processor \cite{rao.braincoproc}.
+Neural co-processors leverage artificial neural networks (ANNs) and deep
 learning to compute optimal closed-loop stimulation patterns. The approach can be used to not only drive neural activity
@@
 
-The co-processor framework also allows co-adaptation with biological circuits in the brain by updating its stimulation policy, while the brain updates its own response to the stimulation via adaptation and neural plasticity, or updates due to other reasons.
-This allows the co-processor to continually optimize its outputs for the desired optimization function in the presence of significant non-stationarity in the brain.
-
-Here we use computer simulations to demonstrate a neural co-processor that restores
-movement in a computational model of cortical networks controlling a limb, after a simulated stroke affects the ability to
-use that limb. Our demonstration combines:
+The co-processor framework also allows co-adaptation with biological circuits in the brain by updating its 
+stimulation policy, while the brain updates its own response to the stimulation via adaptation and neural
+plasticity, or modifies its response due to other reasons. The co-processor could potentially optimize its outputs 
+for a desired optimization function continually in the presence of significant non-stationarities in the brain.
+
+Here, to lay the groundwork for future \textit{in vivo} tests of the co-processor framework, we 
+use computer simulations to explore how co-processors can be trained to restore lost function and how they can adapt to 
+ non-stationarities. We demonstrate a neural co-processor that restores movement in a
+computational model of cortical networks involved in controlling a limb, after a simulated stroke affects the
+ability to use that limb. Our demonstration combines both components of a neural co-processor \cite{rao.coproc}:
 \begin{itemize}
-	\item An emulation model based on ANNs, which models the relationship between 
-	      stimulation, decoded brain activity, and task performance.
-	\item An AI agent based on ANNs which determines the best stimulation to apply in a closed-loop fashion in real time.
+	\item An emulation model based on ANNs, which learns a mapping from  
+	      stimulation and current neural activity to output variables such as task performance (or future neural activity).
+	\item An artificial intelligence (AI) ``agent'' based on ANNs which learns the best  closed-loop
+          activity-dependent stimulation to apply in real time to optimize a given task.
 \end{itemize}
@@
 \section{Background}
-Significant advances have been made in understanding and modeling the effects of electrical stimulation
-on the brain. Researchers have explored how information can be biomimetically or
+\label{sec:background}
+
+Significant advances have been made in understanding and modeling the effects of electrical and other forms of
+neural stimulation on the brain. Researchers have explored how information can be biomimetically or
 artificially encoded and delivered via stimulation to neuronal networks in the brain and
@@
 psychiatric disorders \cite{holtzheimer.psy, kisely.psy, fraint.psy}.
-In research more directly related to our work, Khanna et al. \cite{khanna.openloop}, investigated 
-the use of open loop stimulation in restoring dexterity after a lesion in a nonhuman primate's (NHP's) motor cortex.
-The authors demonstrate that the use of low-frequency alternating current, applied epidurally,
-can improve grasp performance.
-
-While open loop stimulation techniques have yielded clinically useful results, results in other domains have been mixed, such as in visual
-prostheses \cite{bosking.visual}, and in invoking somatosensory feedback
-\cite{cronin.tact}. We believe this is due to the stimulation not being conditioned
-on the ongoing dynamics of the neural circuit being stimulated. From moment to moment and
-throughout the day, a neuronal circuit in the brain can be expected to respond differently even when the same stimulation parameters are used,
-due to the multitude of different external and internal inputs influencing the circuit's ongoing activity. Stimulation
-therefore needs to be proactively adapted in response. This need is even
-greater over longer time scales as the effects of plasticity, changes in clinical conditions, and ageing change
-the dynamics and connectivity of the brain. Closed-loop stimulation may also provide means
-to better regulate the energy use of an implanted stimulator, allowing it to
-intelligently regulate when to apply stimulation, in order to preserve implant
-battery life. Another benefit is that closed-loop stimulation offers an opportunity to minimize the 
-side-effects of stimulation, through real time regulation
-of the stimulation parameters, such as in the use of deep brain stimulation (DBS) in
-PD patients \cite{little.park}. In recent years, closed-loop stimulation has been used to aid in learning
-new memories after some impairment \cite{berger.closedloop, kahana.biomarker},
-to replay visually-invoked activations \cite{tafazoli.acls}, and for optogenetic
-control of a thalamocortical circuit \cite{bolus.opto}, among others.
-
-A major open question is: how does one leverage closed-loop control for real-time
-co-adaptation with the brain to accomplish an external task? ``Co-adaptation'' here refers
-to the ability of a BCI to adapt its stimulation regime to the ongoing
-changes in the circuit it is stimulating, and to adapt with that circuit to accomplish
+In research more directly related to our work, Khanna et al. \cite{khanna.openloop}
+investigated the use of open loop stimulation in restoring dexterity after a lesion
+in nonhuman primate's (NHP) motor cortex. The authors demonstrate that the use
+of low-frequency alternating current, applied epidurally, can improve grasp performance.
+
+While open loop stimulation techniques have yielded clinically useful results, results
+in many domains have been mixed, such as in visual prostheses \cite{bosking.visual},
+and in invoking somatosensory feedback \cite{cronin.tact}. We believe this is due to the
+stimulation not being conditioned on the ongoing dynamics of the neural circuit being
+stimulated. From moment to moment and throughout the day, a neuronal circuit in the brain
+can be expected to respond differently even when the same stimulation parameters are used,
+due to the multitude of different external and internal inputs influencing the circuit's
+ongoing activity. Stimulation therefore needs to be closed-loop, i.e. proactively adapted
+in response. This need is even greater over longer time scales as the effects of plasticity,
+changes in clinical conditions, and ageing change the dynamics and connectivity of the brain.
+Closed-loop stimulation may also provide means to better regulate the energy use of an
+implanted stimulator, allowing it to intelligently regulate when to apply stimulation,
+in order to preserve implant battery life \cite{castano.pd}. Another benefit is that closed-loop
+stimulation offers an opportunity to minimize the side-effects of stimulation, through real
+time regulation of the stimulation parameters, such as in the use of deep brain stimulation
+(DBS) in PD patients \cite{little.park}. In recent years, closed-loop stimulation has been
+used to aid in learning new memories after some impairment \cite{berger.closedloop,
+kahana.biomarker}, to replay visually-invoked activations \cite{tafazoli.acls}, and for
+optogenetic control of a thalamocortical circuit \cite{bolus.opto}, among others.
+
+A major open question is: how does one leverage closed-loop stimulation for real-time
+co-adaptation with the brain to accomplish an external task such as restoration of a lost function?
+``Co-adaptation'' here refers to the ability of a BCI to adapt its stimulation regime to ongoing
+changes in neural circuits in the brain, and to adapt with the brain to accomplish
 the external task (e.g., grasping). The neural co-processor we present here provides one
-potential model for accomplishing that. Through the use of deep learning, a neural 
-co-processor model co-adapts its AI, which governs the
-stimulation, with the neural circuit being stimulated.
-
-For a neurologically complex task such as grasping, it is not possible to design
-\textit{a priori} a fixed real-time controller for the (potentially impaired) neural circuits involved in the task. That is
-due in large part to the variability of circuits from subject-to-subject,
-as well as variations in the placement of sensors and stimulators in the
-brain. The only plausible path to implementing a real-time controller is therefore to allow the device to adapt to the subject,
-long-term changes in their brain activities, and variability in the sensors, stimulators and hardware. Our proposed neural
-co-processors seek to accomplish such adaptation through ANNs and deep learning.
-
-To gain insights into neural co-processors before testing them in \textit{in vivo} experiments, we investigated a number of crucial design elements through the use of a neural network model of the cortical areas involved in
-grasping, presented previously by Michaels et al. \cite{michaels.mrnn}. We
-explored what properties of the co-processor allow
-successful adaptation to the short-term dynamics of the cortical model as it is being stimulated,
+potential approach to accomplishing this goal. Through the use of deep learning, a neural 
+co-processor co-adapts its AI, which controls stimulation, in synch with the biological circuits in the brain.
+
+For a neurologically complex task such as grasping, it is unlikely that there exists
+a fixed real-time controller which can be identified \textit{a priori} for stimulating the (potentially
+impaired) neural circuits involved in the task. This is due in large part to the variability
+in the placement and performance of sensors and stimulators in different brains, as well as variability
+in brain structure and function between subjects. The most plausible path to implementing a real-time
+controller is therefore to allow the device to adapt to the subject, and to the long-term changes in
+their brain activities, and variability in the sensors, stimulators and hardware. Our proposed neural
+co-processors seek to accomplish such adaptation through ANNs and deep learning, and a particular
+training paradigm described below.
+
+\subsection{Simulation as a Way to Gain Insights Prior to {\em in vivo} Experiments}
+To gain insights into neural co-processors before testing them \textit{in vivo}, we
+investigated a number of crucial design elements through the use of a previously published model by
+Michaels et al. \cite{michaels.mrnn} of the cortical areas involved in grasping; the model, based on
+multiple recurrent neural networks, is inspired by cortical anatomy and was fit to data from nonhuman
+primates performing grasping tasks. Using this cortical model as a ``simulated brain'' allowed us to
+rapidly iterate through different design and training methods to demonstrate key properties of the
+neural co-processor framework. These insights will help guide the co-processor training methodologies
+and experimental design for future \textit{in vivo} experiments. Additionally, a
+commonly-accepted maxim of animal experimentation is that animal use should be narrowly tailored
+to answering questions which cannot be answered in other ways. Consider, for example, ``the 3Rs
+alternatives'' approach to the use of animal experiments \cite{nrc.care}. In our case, we leverage
+computer simulations for initial investigations into co-processor design, gathering evidence in
+preparation for future \textit{in vivo} experiments. In the Discussion section, we explore
+potential paths for translation of this work to \textit{in vivo} experiments.
+
+A previous example of such a simulation approach is the work of Dura-Bernal et al.
+\cite{bernal.sim}. In this work, the authors used a simulated spiking neural network
+to train a stimulation agent. Their stimulation agent sought to restore the network's
+control of a simulated arm to reach a target, after a simulated lesion was
+applied. Similar to our approach, the authors simulated lesions by effectively removing parts
+of their simulated network, or by removing connections between parts of the network.
+As the authors point out, there exists only limited ability to probe a neural circuit
+\textit{in vivo} in order to perform learning. As a result, we first need to design
+our approach through the use of an admissible simulation. The key question then is:
+how can we design an admissible simulation, i.e. one which provides a test bed on
+which we can demonstrate key properties of our co-processor designs? We answer this question in our
+work by adopting a a previously published model of cortical grasping \cite{michaels.mrnn} that has
+been shown to replicate properties of the cortical circuit in the biological brain based on fits to
+nonhuman primate data.
+
+Through our simulation, we explored what properties of the co-processor allow successful
+adaptation to the short-term dynamics of the cortical model as it is being stimulated,
 as well as adaptation to longer-term connectivity changes in the cortical model.
-We present a training method for neural co-processors for learning optimal stimulation patterns that drive improvements in external task performance,
-while also adapting to the non-stationarity of the brain.
+We present a training method for neural co-processors for learning optimal stimulation
+patterns that drive improvements in external task performance, while also adapting to the
+non-stationarity of the stimulated neural circuits.
 
@@
     \centering
-    \begin{subfigure}[c]{0.75\textwidth}
+    \begin{subfigure}[c]{0.88\textwidth}
 		\centering
@@
 	\hfill
-    \begin{subfigure}[c]{0.75\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.png}
+    \begin{subfigure}[c]{0.88\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
 		\caption{}
@@
     The CPN outputs stimulation parameters, which are applied to the simulated
-    grasping circuit, according to a stimulation model. The EN models the relationship between the
-    stimulation, brain recordings, and external task.}
+    grasping circuit, according to a stimulation model. The EN models the relationship between
+    the     stimulation, brain recordings, and external task.}
     \label{fig:arch}
@@
 fundamental challenges in using neural stimulation to improve external task performance.
-First, to restore function for complex tasks, such as grasping, the mapping between
-neural activity, such as the activity encoding the intention to grasp, and the stimulation patterns
-to be applied to a downstream motor area to enable grasping cannot be pre-determined.
-As a result, the co-processor must learn what stimulation pattern is appropriate for achieving the
-external task. Unfortunately, in most cases, we do not know the optimal stimulation patterns to train
-the co-processor. This is because stimulation shapes the nonlinear dynamics of circuits in the brain
+First, to restore function for complex tasks (such as grasping), it is difficult to determine
+the mapping between observed neural activity and the stimulation patterns to be applied to a downstream
+motor area. As a result, the co-processor must learn what stimulation pattern is appropriate for achieving the
+external task given the current neural activity. Unfortunately, in most cases, we do not know what the correct
+stimulation pattern is for any given input activity pattern, a precondition for using supervised deep learning
+methods for training the co-processor. Stimulation shapes the nonlinear dynamics of circuits in the brain
 in complex ways, leading to complex effects in external behavior. It is therefore not obvious what
@@
 
-A neural co-processor attempts to solve these problems with a pair of artificial neural networks (ANNs)
+A neural co-processor attempts to solve these problems with a pair of artificial neural networks 
 (Fig.~\ref{fig:arch}):
 \begin{itemize}
-    \item a ``Co-Processor Network'' (CPN), which maps neural activity, and possibly
+    \item a ``Co-Processor Network'' (CPN), which is a recurrent neural network (RNN) which maps neural activity, and possibly
 	      data from external sensors, to appropriate stimulation parameters.
-    \item an ``Emulator Network'' (EN) which models the effect of stimulation on neural dynamics and
+    \item an ``Emulator Network'' (EN), also an RNN which models the effect of stimulation on neural dynamics and
           behavior for the external task.
@@
 between this desired behavior and the actual behavior caused by the CPN due to stimulation. How do we
-backpropagate this error, which is external to the CPN, to update the parameters of the CPN?
-
-We use a trained EN to backpropagate the external task error to the CPN. We train the EN to predict
-task-relevant parameters - a prediction of muscle velocities in our case - given the stimulation
-parameters output by the CPN and any measured neural activity. If the EN is trained to a sufficiently
-high level of precision, it can be used as a function approximator mapping stimulation to task output.
+backpropagate this external task error to update the parameters of the CPN so that it learns to produce
+stimulation patterns that are optimal for the task?
+
+We use a trained EN to backpropagate the external task error to the CPN, which then uses this backpropagated
+error to update its parameters. We train the EN to predict task-relevant parameters - a prediction of muscle
+velocities in our case - given randomly sampled stimulation patterns or stimulation
+patterns output by the CPN, and any measured neural activity. If the EN is trained to a sufficiently
+high level of precision, it can be used as a function approximator for the subject's true ``stimulation function''.
 When training the CPN, we treat the EN's output as the actual task output (e.g., actual grasp
-parameters).  We backpropagate through the EN (without changing its weights) the error between
-the EN's output and the desired task output and use this backpropagated error to train the
-CPN (see Fig. \ref{fig:arch}).
-
-In our experiments, the EN was a single layer fully connected long short-term memory (LSTM) recurrent
-neural network (RNN), with hyperbolic tangent ($tanh$) activations, and a linear readout. It has 87 LSTM neurons,
-which was chosen somewhat arbitrarily as a function of the input and output vector sizes. We found varying this neuron
-count did not drastically change results. Although other architectures could also be used, we found that this
-LSTM architecture allows the EN to continuously adapt to long-running dependencies in the simulated neural dynamics, 
-ar better than a vanilla RNN. The CPN had an almost identical architecture, but with 61 LSTM neurons. As with the CPN,
-this neuron count was chosen as a function of the input and output vector sizes, but varying it upward had little effect on results.
-There is no requirement for the EN and CPN to have similar network architectures, but we found that these choices worked
-well in our experiments.
+behavior or muscle velocities). We backpropagate through the EN (without changing its parameters) and then the
+CPN (changing its parameters) the error between the EN's output and the desired task output (see Fig. \ref{fig:arch}).
+
+In our experiments, the EN was a single-layer fully-connected long short-term memory (LSTM) recurrent
+neural network, with hyperbolic tangent ($tanh$) activations, and a linear readout. It had 87 LSTM neurons,
+the number being chosen somewhat arbitrarily as a function of the input and output vector sizes. We found that
+varying this neuron count did not drastically change results. Although other architectures could also be used,
+we found that this LSTM architecture allows the EN to continuously adapt to long-running dependencies in the
+simulated neural dynamics, far better than a vanilla RNN. The CPN had an almost identical architecture, but with
+61 LSTM neurons. As with the CPN, this neuron count was chosen as a function of the input and output vector sizes,
+and increasing it had little effect on results. There is no requirement for the EN and CPN to have similar network
+architectures, but we found that these choices worked well in our experiments.
 
 Note that the EN is more general than traditional models of neurostimulation which attempt to predict the effects of
-stimulation on \em{neural activity}. The EN in a neural co-processor predicts the effects of stimulation (taking into
-account ongoing neural dynamics) on \em{task performance}. This provides the key functionality needed to train the
+stimulation on { \em neural activity}. The EN in a neural co-processor predicts the effects of stimulation (taking into
+account ongoing neural dynamics) on {\em task performance}. This provides the key functionality needed to train the
 CPN. In the special case where the task involves driving neural circuits in the brain to desired neural activities,
@@
 For comparison, consider an EN architecture based on a traditional RNN with a nonlinearity. This is a nonlinear version
-of the common linear time-invariant state space model of stimulation (see, e.g.,  \cite{shanechi.stimmodel}). 
-We found, however, that compared to LSTM-based EN, the linear model and the traditional nonlinear RNN are both not
+of the common linear time-invariant state space stimulation model studied in previous research (see, e.g., \cite{shanechi.stimmodel}). 
+We found, however, that compared to an LSTM-based EN, the linear model and the vanilla nonlinear RNN are both not
 sufficiently powerful to capture the long-term dependencies in stimulation effects needed for the CPN to learn well.
-Our use of an LSTM-based EN builds on previous work modeling biological neural networks for predicting local field potentials
-\cite{kim.lstm}, and modeling stimulation \cite{guclu.lstm}.
+Our use of an LSTM for learning an EN builds on previous work using LSTMs for predicting the evolution of local field
+potentials \cite{kim.lstm} and blood-oxygen-level dependent (BOLD) hemodynamic responses \cite{guclu.lstm} over time.
 
 \subsection{Simulation Overview}
-To test the feasibility of the neural co-processor approach, we used a previously published model of a
-grasping circuit in a nonhuman primate (NHP) brain. Using such a model allows us to explore some of the critical
-architectural choices and training algorithms, enabling us to rapidly and cheaply iterate on our
-design, prior to any \textit{in vivo} experiments.
-
-In the present study, we use the grasping network model of Michaels et al. \cite{michaels.mrnn}
-(Fig.~\ref{fig:michaels}a). The network, representing multiple interconnected cortical areas, was trained
-to mimic the grasping circuits of NHP subjects engaged in a delayed reach-to-grasp task.
+To test the feasibility of the neural co-processor approach, we used a previously published cortical model for
+grasping in a nonhuman primate (NHP) brain. Using such a model allowed us to explore some of the critical
+architectural choices and training algorithms for neural co-processors, enabling us to rapidly and cheaply iterate on our
+design, laying the groundwork for future \textit{in vivo} experiments.
+
+Specifically, we use the cortical model of Michaels et al.\ \cite{michaels.mrnn}
+(Fig.~\ref{fig:michaels}a). The model, which uses multiple RNNs to represent multiple interconnected cortical areas,
+was trained to mimic the grasping circuits of NHP subjects engaged in a delayed reach-to-grasp task.
 The model's design draws on a body of literature focused on architectures and training methods for RNNs which
 seek to create artificial neural networks with activation dynamics similar to biological circuits,
-including circuits for delayed grasping tasks \cite{susillo.mrnn}. The model consists of three
-``modular'' vanilla RNNs (mRNNs), representing the cortical areas AIP, F5 and M1 respectively, and a linear
-readout layer (Fig.~\ref{fig:michaels}a). Each ``module'' consists of 100 vanilla
-RNN neurons, with a nonlinearity applied on the outputs. The modules are internally fully connected,
-and are connected to each other sparsely (10\% connectivity in our case). The inputs to the network are visual features
+including circuits for delayed grasping tasks \cite{sussillo.mrnn}. The model consists of three
+``modular'' vanilla RNNs (mRNNs), representing the cortical areas AIP (anterior intraparietal cortex), F5 (ventral
+premotor cortex) and M1 (primary motor cortex) respectively, and a linear readout layer from M1 producing muscle
+velocities (Fig.~\ref{fig:michaels}a). Each ``module'' consists of 100 vanilla RNN neurons, with a nonlinearity
+applied on the outputs. The modules are internally fully connected, and are connected to each other sparsely
+(10\% connectivity in our case). The inputs to the network are visual features
 representing the object to be grasped, as well as a hold signal. The visual features
@@
 VGGNet \cite{simonyan.vgg} from 3D renderings of the same objects which the subjects grasped. The hold signal is
-a boolean which encodes the point in the experiment when the subject began to reach. The outputs of the
-network are muscle length velocities for the shoulder, arm, and hand of the subject. The actual
-velocities were captured with a motion capture glove, and the Michaels et al. network model 
-was trained to recapitulate those grasping motions.\footnote{Data and trained models from this work
+a Boolean which encodes the point in the experiment when the subject began to reach. The outputs of the
+network are muscle velocities for the shoulder, arm, and hand of the subject. The actual
+velocities were captured with a motion capture glove. The Michaels et al.\ network model 
+was trained to recapitulate these grasping motions.\footnote{Data and trained models from this work
 were supplied to us by the lead author of \cite{michaels.mrnn}. We re-implemented their model in PyTorch,
@@
 		\centering
-		\includegraphics[width=\textwidth]{michaels.png}
+		\includegraphics[width=\textwidth]{michaels.eps}
 		\caption{}
@@
 	\hfill
-\caption{\textbf{Architecture of the Brain Model.} (a) Modular RNN (mRNN) used by Michaels et al.
+\caption{\textbf{Architecture of the Cortical Model for Grasping.} (a) Modular RNN (mRNN) used by Michaels et al.
         \cite{michaels.mrnn} to model cortical circuits involved in grasping objects.
@@
 		actual network weights are floating point values. White indicates non-connections.
-		Note the within-module connections (black squares along	the diagonal) are fully
-		connected while connections to adjacent modules are sparse
-		($10\%$ of possible connections).}
+		Note the each module is fully connected (black squares along the diagonal) while connections to adjacent
+        modules are sparse ($10\%$ of possible connections). All connections were trained using nonhuman primate
+        (NHP) data to recapitulate the experimentally recorded muscle velocity trajectories for grasping different
+        objects.}
 \label{fig:michaels}
@@
 
-The grasping model we use \cite{michaels.mrnn} implements a vision-to-grasp pipeline, from the visual
-processing needed to move the hand to the appropriate position for the grasp to shaping the hand for
-grasping an object of a particular shape. The emergent dynamics of the model's ``modules'', once trained,
-correspond roughly to the responses in the cortical areas AIP, F5, and M1 in a NHP subject's brain.
-Specifically, the activity of the first module (Fig.~\ref{fig:michaels}a), receiving the visual inputs in a trial resembles
-the activity in area AIP of the NHP subject from the same trial. Likewise, the second and
-third modules resemble the activity from the subjects's F5 and M1 cortical areas, respectively.
-The emergent network dynamics show a number of other correspondences to the subjects' 
-brain activity as well (see \cite{michaels.mrnn} for details). For convenience, we will refer to the three
-modules in the model using the cortical areas (AIP, F5, M1) they correspond to. For details on how this network
-was trained, refer to Supplementary Materials \ref{sup:michaelstraining}. For additional details on the task structure, see
+The Michaels et al.\ cortical model implements a vision-to-grasp pipeline, from the visual
+processing needed to move the hand to the appropriate position to shaping the hand for
+grasping an object of a particular shape. The emergent dynamics of the model's modules, once trained,
+correspond roughly to neural responses in the cortical areas AIP, F5, and M1 in a NHP subject's brain (Fig.~\ref{fig:michaels}a;
+see \cite{michaels.mrnn} for details). For convenience, we will refer to the three
+modules in the model using the cortical areas (AIP, F5, M1) they correspond to. For details on how this network model 
+was trained, please see Supplementary Materials \ref{sup:michaelstraining}. For additional details on the task structure, see
 Supplementary Materials \ref{sup:michaelstask}.
 
-An important attribute of the grasping network model is that  the simulated circuit's activity shows a relatively
-clear separation for different object shapes, i.e. the visual input to the network is leveraged by the model to
-successfully generate hand shape trajectories for grasping objects of particular shapes, and sizes. As we will note
-below, if the visual information is prevented from propagating through the modules, the model can at best learn a
-grasp stereotyped across all object sizes and shapes.
-
-\subsubsection{Lesioning the model causes real world failure modes.}
-Simulating a brain lesion in the grasping model described above results in error
-modes which resemble some natural lesions of the primate brain. For example, a ``lesion''
-involving zeroing some of the outputs of the first (AIP) module 
-leads to a reaching motion generally succeeding, but finger muscle velocities show a high
-degree of error - effectively implying that the subject can reach to grasp, but cannot
+An important attribute of this cortical model is that  the simulated circuit's activity shows a relatively
+clear separation for different object shapes, i.e., the  visual input is leveraged by the model to
+successfully generate hand shape trajectories for grasping objects of particular shapes and sizes. As noted
+below, if this visual information is prevented from propagating through the cortical modules (e.g., due to a
+simulated lesion), the model can at best learn a stereotyped grasp across all object sizes and shapes.
+
+\subsubsection{Lesioning the cortical model causes real world failure modes}
+As shown by Michaels et al. \cite{michaels.mrnn}, simulated lesions in the cortical model described
+above result in error modes which resemble the effects of some natural lesions in the primate brain.
+For example, a ``lesion'' involving zeroing some of the outputs of the first (AIP) module 
+leads to a reaching motion generally succeeding, but finger muscle velocities show a higher
+degree of error, effectively implying that the subject can reach to grasp, but cannot
 tailor the grasp to the current object. This is likely due to object shape information not
-being fully conveyed to the motor cortex (M1) module. Such a form of hand muscle spasticity
+being fully conveyed to the primary motor cortex (M1) module. Such a form of hand muscle spasticity
 is also a common symptom in certain strokes where the subject is able to position their
-hand, but is unable to properly form the appropriate grasp \cite{khanna.openloop,
-puthenveettil.hand}.
-
-On the other hand, if we lesion of a portion of the motor cortex (M1) module in the model,
+hand, but is unable to form the appropriate grasp \cite{khanna.openloop, puthenveettil.hand}.
+
+On the other hand, if we lesion of a portion of the M1 module in the model,
 we see a more complete loss of movement, affecting even the ability to reach for the grasp.
-Finally, if we ``disconnect'' communication between the F5 and M1 modules, we see a
-failure similar to an AIP lesion: the reaching movement is generally achieved, but we see
-a disproportionate impact on forming the appropriate hand grasp. See Fig. \ref{fig:lesion}
-for examples.
+Finally, if we ``disconnect'' communication between the F5 and M1 modules,
+we see a failure similar to an AIP lesion: the reaching movement is generally achieved, but
+we see a disproportionate impact on forming the appropriate hand grasp. See Fig.
+\ref{fig:lesion}a for examples of lesion impacts on velocities for individual muscles.
+In Fig. \ref{fig:lesion}b we see the ratio of mean squared error (MSE) loss for hand
+muscles to all muscles. Values greater than $1.0$ indicate that hand muscles are impacted more
+than other muscles. Note that with the AIP and connection lesions, hand muscles show significantly
+higher loss than overall loss. However, the M1 lesion shows the highest overall loss; see Table
+\ref{tab:results} in Supplement materials section \ref{sup:results} for detailed lesion loss data.
 
@@
              tend to cause a significant task loss on both.
-             (b) Ratio of hand muscle mean squared error (MSE) losses vs.
-             overall MSE losses, for different lesion types. The loss is measured
-             relative to the model's movement trajectories prior to the lesion.
-             Average of all trials, with +/- 1 stdev shown.}
+             (b) Ratio of hand muscle mean squared error (MSE) losses vs. overall MSE
+             losses, for different lesion types. The loss is measured relative to the
+             model's movement trajectories prior to the lesion. Average of all trials,
+             with $\pm1$ stdev shown. Higher ratios for AIP and connection lesions indicate those
+             lesions cause a larger loss for hand movement than overall movement, though
+             their overall loss is lower than for M1 (see Table \ref{tab:results}).}
 	\label{fig:lesion}
@@
 
-Given a particular type of lesion in the model, the co-processor's task is to
-generate the appropriate stimulation pattern given mRNN activities. The
+Given a particular type of lesion in the model, the goal of the co-processor is to learn
+to map ``neural recordings'' (derived from current mRNN  activity in the model)
+to the appropriate stimulation pattern for executing the required grasp. The
 co-processor thus seeks to effectively bridge across the lesion and deliver
@@
 \begin{itemize}
-	\item \textbf{Loss of AIP Neurons:} We force the output of some proportion of ``AIP''
+	\item \textbf{Loss of AIP Neurons:} We force the output of some proportion of the AIP module's
 	      neurons to zero, effectively removing them from the network. This results in some amount of 
@@
 	\item \textbf{Loss of F5-M1 Connections:} We prevent the propagation of
-	      information between the ``F5'' and ``M1'' modules, effectively representing a
+	      information between the F5 and M1 modules, effectively representing a
 	      severing of the connections between the two modules. Note that the connections are sparse
 	      and run in both directions, and we lesion the connections in both directions.
-	\item \textbf{Loss of M1 Neurons:} We force the output of some proportion of ``M1''
+	\item \textbf{Loss of M1 Neurons:} We force the output of some proportion of the M1 module's
 	      neurons to zero. Here, the lesion may make it impossible for the co-processor
-	      to find a perfect solution, since the loss of M1 neurons may make it impossible to
-	      activate muscles in the same ways. However, in Results we will see that some recovery
-	      is possible.
+	      to find a perfect solution since the loss of M1 neurons may make it impossible to
+	      activate muscles in the same ways as before the lesion. However, in the Results section,
+        we show that some recovery is still possible.
 \end{itemize}
 
-\subsubsection{Simulated network exhibits long running and stable dynamics.}
+\subsubsection{Simulated network exhibits long running and stable dynamics}
 To compute stimulation patterns that optimize for a task, the co-processor must
@@
 networks as well as our simulated grasping network exhibit long range changes in dynamics
-due to stimulation. A perturbation of the network (i.e. due to stimulation) will cause
+due to stimulation. A perturbation of the network (i.e., due to stimulation) will cause
 changes in neuronal activations long after the stimulation has been applied, sometimes far
-from the site of stimulation. Our model of the cortical grasping network exhibits the same
+from the site of stimulation. Our cortical grasping network model exhibits the same
 behavior.
@@
 	         respectively. Probability distribution shown to +/- 2 stdevs. Network exhibits similar
-	         behavior when perturbance is performed at other times.}
+	         behavior when perturbance is applied at other time steps.}
 	\label{fig:dynamics}
@@
 
-\subsubsection{Stimulation model.}
+\subsubsection{Stimulation model}
 In our experiments, we stimulate neurons only in the output area (here M1) since the co-processor's
@@
 biophysical model of stimulation (our model does not arrange neurons in a volume to allow for such detailed
-simulation). Instead, the stimulation model approximates the effects of in-vivo biological stimulation (e.g.
-extracellular electrical stimulation) by diffusing the effects of stimulation across neural tissue space and time.
-Including this model implies that our EN must approximate this stimulation function, in addition to approximating
-the cortical dynamics of the brain and the mapping of those dynamics onto the grasping task. In Supplementary Materials
-\ref{sup:passthrough}, we compare results to the case with no stimulation or recording function.
-
-Specifically, we use a stimulation function $S$ which receives as input the stimulation parameter vector $\theta$
-from the CPN. $S$ performs temporal smoothing using a simple exponential decay model by adding its current input to
-an exponentially decaying sum of inputs that decays towards $0.0$ at some rate (see Equation ~\ref{eq:temporalstim} below).
-The effect of stimulation is thus not instantaneous, but rather decays with time. Likewise, to approximate how charge
-dissipates into a surrounding volume, we applied Gaussian smoothing to map $\theta$ (16-D in our experiments) to changes
-in the activations of a large number of the simulated M1 neurons (100 neurons in our experiments). If we assume each element
-of the $\theta$ represents the stimulation parameter for a single electrode, our Gaussian smoothing operation emulates how
-stimulation may affect the neurons in its vicinity more than it affects neurons further away. To accomplish that, we assume
-our model neurons are aligned along a spatial dimension arbitrarily and fix $\sigma=1.75$.
-
-The resulting equations defining our stimulation model are:
+simulation). Instead, the stimulation model approximates the effects of \textit{in vivo} biological
+stimulation (e.g., extracellular electrical stimulation) by diffusing the effects of stimulation across
+many neurons and across time. Our EN must thus learn to approximate this stimulation model,
+in addition to approximating the cortical dynamics of the brain and the mapping of those dynamics onto
+the grasping outputs.\footnote{In Supplementary Materials section \ref{sup:passthrough}, we explore  
+other stimulation models.}
+
+Our stimulation model uses a function which receives as input the stimulation parameter
+vector $\theta$ from the CPN. The function performs temporal smoothing using a simple exponential decay model by
+adding its current input to an exponentially decaying sum of inputs that decays towards $0.0$ at some
+rate (see Equation ~\ref{eq:temporalstim} below). The effect of stimulation is thus not instantaneous,
+but rather decays with time. Likewise, to simulate the difficulty of stimulating only single neurons,
+we applied Gaussian smoothing to map $\theta$ (16-D in our experiments) to changes in the activations of
+a large number of the simulated M1 neurons (100 neurons in our experiments). If we assume each element
+of the $\theta$ represents the stimulation parameter for a single electrode, our Gaussian smoothing
+operation emulates how stimulation may affect the neurons in its vicinity more than it affects neurons
+further away. To accomplish this, we assume our model neurons are aligned along a spatial dimension
+arbitrarily and fix $\sigma=1.75$.\\
+
+\noindent The resulting equations which define our stimulation model are:
 
@@
 
-The governing equations of an mRNN with stimulation then become:
+\noindent The governing equations of an mRNN with stimulation then become:
 
@@
 
-	\caption{\textbf{Simulating the effects of stimulation on a recurrent network across space and time.} The stimulation model performs spatial and temporal smoothing of a 16-dimensional
+	\caption{\textbf{Simulating the effects of stimulation on a recurrent network across space and time.}
+          The stimulation model performs spatial and temporal smoothing of a 16-dimensional
 	         vector of stimulation parameters $\theta$ onto 100 neurons in a recurrent network.
@@
 	         arbitrarily. We show here the effects of two randomized $\theta$ with two respective smoothing
-	         parameters $\sigma$. After $t=0$,
-		     $\theta$ is the zero vector. Color values indicate the magnitude of the value summed into
-		     each neuron's hidden state at that time step. }
+	         parameters $\sigma$ (a. $\sigma=1.875$, b. $\sigma=2.875$). After $t=0$, $\theta$ is the zero
+		   vector. Color values indicate the magnitude of the value summed into each neuron's hidden
+		   state at that time step.}
 	\label{fig:stim_single}
@@
 
-Fig. \ref{fig:stim_single} depicts an example where
+\noindent Fig. \ref{fig:stim_single} depicts an example where
 $\theta$ is non-zero at $t=0$ and zero for all other times, illustrating the effects of 
@@
 	has been lesioned 50\%, indicated by hidden states of M1 neurons with a value of zero (bottom
-	left panel, lowest part of the plot) . We record neural activities (recording model) only
-	from the AIP and F5 modules, as explained in Section \ref{sec:experiments}. Stimulation is applied
-	to M1 to drive the network's output. Stimulation is mostly constant until approximately t=275,
-	when the hold signal was lifted and reach began.}
+	left panel, lowest part of the plot). We record neural activities (using the recording model) from only
+	 the AIP and F5 modules, as explained in Section \ref{sec:experiments}. Stimulation is applied
+	to M1 to drive the network's output. Stimulation is mostly constant until approximately t=250,
+	when the hold signal was lifted and the reach began.}
 	\label{fig:stim_and_obs}
@@
 
-\subsubsection{Recording model.}
-To simulate recordings of our model neurons, we used a recording model that assumed the model neurons
-are laid out along a single spatial dimension with the given number of electrodes spread evenly apart.
+\subsubsection{Recording model}
+To simulate recordings of our model neurons, we used a recording model that assumed that the model neurons
+are laid out along a single spatial dimension with a given number of electrodes spread evenly apart.
 We then applied Gaussian convolution, where the Gaussian kernel is centered at each electrode position.
-Thus, the recording from each `electrode' is a Gaussian-weighted average of the activities of all neurons
-in the given module. Fig.~\ref{fig:stim_and_obs} (bottom panels) show an example of recordings obtained
-using this recording model; note that the Gaussian-averaging of activities provides only a low-pass
-filtered, potentially ambiguous view of the activities of the underlying neurons, making it more challenging
-for the co-processor to interpret the neural activity and produce an appropriate stimulation pattern. Our
-simulated recordings are similar in spirit, though not actually modeling, local field potentials recorded
-extracellularly in biological neural tissue. In Supplemental section
-\ref{sup:observability} we show that as we vary the observability of the
-brain via this recording model the results are largely the same, up to a
-point where objects can no longer be distinguished by the recording data.
-
-\subsubsection{Simulating co-adaptation by the brain.}
-To demonstrate the co-processor's ability to adapt with the brain as it adapts to the co-processor's
-stimulation, we modify the cortical grasping network's parameters (synaptic weights and biases) throughout
-the co-processor's training. We use the standard error backpropagation algorithm to adapt the grasping network
-(using PyTorch's implementation of the $AdamW$ optimizer). With each trial, task loss is calculated and
-backpropagated through the mRNNs as they receive stimulation. The learning rate was set arbitrarily
-to a relatively low rate of $1\mathrm{e}{-7}$ so that the network adapts more slowly than the co-processor.
+Thus, the recording from each ``electrode'' is a Gaussian-weighted average of the activities of all model neurons
+in the given module. Fig.~\ref{fig:stim_and_obs} (bottom right panel) shows an example of recordings obtained
+using this recording model; note that the Gaussian averaging provides a potentially ambiguous view of the
+activities of the underlying neurons due to the inherently destructive nature of averaging, making it
+more challenging for the co-processor to interpret the neural activity and produce an appropriate
+stimulation pattern. Our simulated recordings are similar in spirit, though not actually modeling,
+local field potentials (LFPs) recorded extracellularly in biological neural tissue. In Supplementary Materials section
+\ref{sup:observability}, we show that as we vary the observability of our ``simulated brain'' via this recording
+model, the results remain largely the same, up to the point where the objects to be grasped can no longer be
+distinguished from the neural recordings due to excessive averaging.
+
+\subsubsection{Simulating co-adaptation by the brain}
+To demonstrate the co-processor's ability to co-adapt with the brain as it adapts to the co-processor's
+stimulation, we modify the cortical grasping network model's parameters (synaptic weights and biases) throughout
+the co-processor's training. We use the standard error backpropagation algorithm to adapt the grasping network to
+minimize task loss for the object grasping task (using $PyTorch$'s implementation of the $AdamW$ optimizer). With each
+trial, task loss is calculated based on the NHP training data and backpropagated through the mRNNs as they receive
+stimulation. The learning rate was set arbitrarily to a relatively low rate of $1\mathrm{e}{-7}$ so that the network
+adapts more slowly than the co-processor.
 
@@
 affected by the stroke. We simulated this ability in our grasping network model by re-training the 
-network for the grasping task after lesioning it. For our simulated lesions which zero
-the outputs of neurons, the mRNN often cannot supply a partial recovery model.
-This is because the mRNN model has sufficient redundancy built into it that lesioning it by
+network for the grasping task after lesioning it. For our simulated lesions which zero-out 
+the outputs of neurons, we found that the cortical model has sufficient redundancy built into it that lesioning it by
 inactivating large numbers of neurons often leaves enough remaining degrees of freedom that a
-nearly full recovery can occur, unless so many neurons are lesioned that no stimulation could be
-effective.
+nearly full recovery can occur. We therefore explored lesions of  the model (specifically lesions of AIP and M1 modules)
+under two conditions: (a) No co-adaptation: the cortical model was not re-trained after the lesion, allowing us to study
+how much of lost function the co-processor can learn to restore on its own, and (b) Co-adaptation: the cortical model
+was re-trained after the lesion, allowing us to study how the co-processor copes with non-stationarity as our simulated
+brain recovers from its lesion.
 
@@
 the input object's shape cannot propagate forward in the network to allow shaping of the hand for
-grasping. However, the mRNN can learn a stereotyped grasp during the recovery period. After this recovery
-period, a co-processor can further boost grasp accuracy by forward-propagating object shape information
+grasping. However, the cortical model can learn a stereotyped grasp during the recovery period. After this recovery
+period, a co-processor can help further boost grasp accuracy by forward-propagating object shape information
 from AIP and F5 to M1, acting as an artificial neural bridge. We explore this application of the
@@
 
-\subsubsection{Simulating a non-stationary recording function, e.g. sensor drift}
-We also demonstrate the co-processor's ability to adapt to non-stationarity in the recording
-function. Over time, implanted sensors may drift from true readings. To test the co-processor's
-ability to adapt to that drift, we perform an experiment where the observation function
-has a bias term which changes over time, according to a random process. Each epoch, we
-add a random value to each element of the bias, drawn from a mean-0 Gaussian distribution. That
-causes the ``sensors'' to have a drifting zero point over time.
-See Supplementary Materials \ref{sup:drift} for further details.
+\subsubsection{Simulating a non-stationary recording function, e.g., sensor drift}
+Over time, implanted sensors may drift from true readings. We explored the co-processor's ability to adapt to
+non-stationarity in the recording function. We performed an experiment where the recording function
+includes an added bias term which changes over time, according to a random process. Each epoch, we
+add a random value to each element of the bias term, drawn from a zero-mean Gaussian distribution, causing the
+``recordings'' to drift over time. See Supplementary Materials section \ref{sup:drift} for further details.
 
@@
 To train the CPN to generate appropriate stimulation patterns for grasping a given input object of a
-particular shape, we can use the backpropagation algorithm to minimize the errors between the current
-generated grasp and the desired target grasp. However, this error is in terms of grasp error instead of
-error in stimulation patterns (which is what the CPN needs), and we have no way of backpropagating grasp
-error through biological networks in the brain. This motivates our use of the EN as a model for the
-transformation from stimulation to muscle velocities: we backpropagate grasp error through the EN,
-while keeping its parameters fixed, and use this backpropagated error to change the parameters of the
-CPN towards minimizing grasp error.
+particular shape, we would like to use a learning algorithm such as the backpropagation algorithm and
+minimize the errors between the current generated grasp (in terms of muscle velocities) and the desired target
+grasp. However, backpropagation requires the error to be in terms of CPN output stimulation error. We cannot  
+backpropagate grasp error through biological networks in the brain to get the error needed to train the CPN. 
+This motivates our use of the EN as a model for the transformation from stimulation to muscle velocities:
+we backpropagate grasp error through the EN, while keeping its parameters fixed, and use this
+backpropagated error to change the parameters of the CPN in order to minimize grasp error.
 
 Using the EN to train the CPN requires careful interleaving of EN and CPN training epochs. EN training
-epochs concentrate on updating the EN, based on observations of the effect of stimulation on the grasping
+epochs concentrate on updating the EN, based on observations of the effect of stimulation on the cortical
 network model's output. The loss function when training the EN is the mean squared error (MSE) loss between
-the EN's prediction and the  actual output of the grasping network (i.e. muscle velocities). During the CPN
+the EN's prediction and the  actual output of the cortical model (i.e., muscle velocities). During the CPN
 training epochs, we keep the EN's parameters fixed and use the EN to train the CPN. Specifically, we
@@
 		\centering
-		\includegraphics[width=\textwidth]{backprop_en.png}
+		\includegraphics[width=\textwidth]{backprop_en.eps}
 		\caption{}
@@
 		\centering
-		\includegraphics[width=\textwidth]{backprop_cpn.png}
+		\includegraphics[width=\textwidth]{backprop_cpn.eps}
 		\caption{}
@@
 \caption{\textbf{Training the neural co-processor.} (a) EN training phase: 
-         backpropagate actual vs predicted muscle velocity MSE loss.
-         (b) CPN training phase: backpropagate actual vs target
-         MSE loss, via EN, to CPN. This effectively treats the EN
-         output as the actual, and thus the EN must be trained first.}
+         backpropagate the error between actual vs predicted muscle velocities through the EN to update the EN's parameters.
+         (b) CPN training phase: backpropagate the error between EN's predicted output vs target muscle velocities, via the
+         EN (without changing its parameters) and through the CPN, to update the CPN's parameters. Note that we are effectively
+         treating the EN's predicted output as the actual muscle velocity output, and therefore it is important to have a good
+         pre-trained EN.}
 \label{fig:training}
@@
 
-\subsubsection{Training and testing data sets.}
+\subsubsection{Training and testing data sets}
 For training and evaluation of our model, we use the same data as in Michaels et al.
-\cite{michaels.mrnn}, which was provided to us by the lead author. The data consist of:
+\cite{michaels.mrnn}. The data consist of:
 \begin{itemize}
     \item The input visual features and hold signal. We input these to the simulated
-          cortical circuit;
+          cortical model;
     \item The object identity. We use this to calculate how well the co-processor
           can differentiate objects; see Results section below;
-    \item The muscle velocity data for the trial, as extracted from the data glove
-          during the NHP trial, and as processed. This is our target output.
+    \item The muscle velocity data for the trial, as extracted from the data glove of a nohhuman primate 
+          during the NHP trial, and as processed by Michaels et al.
+\cite{michaels.mrnn}. This is our target output.
 \end{itemize}
@@
 
-\subsubsection{EN training.}
+\subsubsection{EN training}
+\label{sec:en_training}
+
 For the EN to be useful for training the CPN, it must accurately predict the behavioral effects of stimulation
 produced by the CPN. In addition, backpropagating through the EN must yield gradients which are useful for
-changing the weights of the CPN in order to produce better stimulation patterns or minimizing the error.
+changing the weights of the CPN in order to produce better stimulation patterns for minimizing the error.
 We discovered that this latter property does not occur simply by virtue of the former. Specifically, an EN
 can be trained to high levels of predictive power, even for random stimulation, and to orders of magnitude
-lower loss than the task loss, but backpropagation through the EN can still yield gradients which lead to
-unstable training of the CPN.
+lower loss than the task loss, while at the same time,  backpropagation through the EN still yields gradients
+which lead to unstable training of the CPN.
 
@@
 the high dimensionality of our stimulation parameters $\theta$, and the dynamics of the
-network being stimulated, the function learned by the EN may be biased and provide gradients not suitable
+network being stimulated, the function learned by the EN may not provide gradients suitable
 for the stimulation inputs generated by the CPN during its training.
@@
 To address this problem, we first structure the training dataset for each EN training epoch
-in a specific way. Each epoch, we include:
+in a specific way. In each epoch, we include:
 \begin{itemize}
@@
 here refers to the region of CPN parameter space near the current CPN. This may cause the EN to be
-overfit, i.e. fit to the local area of CPN parameter space but exhibiting poor predictive power in
+overfit, i.e., fit to the local area of CPN parameter space but exhibiting poor predictive power in
 other areas of the space. However, we solve this problem by retraining the EN, interleaved with CPN
-training. We show the effect of this dataset composition in one results section below
-(\ref{sec:results_con}).
+training. We show the effect of this paradigm in our results below
+(section~\ref{sec:results_con}).
 
@@
 data set. We found that using rates much lower or higher than these rates in any training phase
-may cause the EN learning to not converge. EN training proceeds until the network's output error
-exceed a threshold, defined as a fraction of the current CPN's task loss. See Section
-\ref{sec:interleaving} for further details on when we switch between training the EN and CPN.
+may cause the EN learning to not converge. EN training proceeds until the network's prediction error
+is lower than a threshold, defined as a fraction of the current CPN's task loss. Section
+\ref{sec:interleaving} provides further details on when we switch between training the EN and CPN.
 
 \subsubsection{CPN training}
-Once an EN is properly trained, we can use the EN as a surrogate for the part of the grasping model
-that transforms stimulation patterns and neural activity to grasping behavior. We use a large number
-of randomly sampled trials from the original Michaels et al. task \cite{michaels.mrnn} and generate
+Once an EN is properly trained, we can use the EN as a surrogate for the cortical model's transformation
+of stimulation patterns and neural activity to grasping behavior (i.e., muscle velocities). We use a large number
+of randomly sampled trials from the original Michaels et al.\ task \cite{michaels.mrnn} and generate
 stimulation patterns using the current CPN alone. These stimulation patterns are passed through the EN
@@
 We found that the CPN appears to train in two phases. In the first phase, it is largely learning the structure
-of the reach-to-grasp task, e.g. that the muscles need to remain at a position during the hold period until the
+of the reach-to-grasp task, e.g., that the muscles need to remain at a position during the hold period until the
 reach begins. During this early training phase, the learning rate can be quite high ($1\mathrm{e}{-3}$). Later, the CPN
@@
 the target grasp for that object. This later phase of training requires a learning rate 2-3 orders of magnitude
-lower than in the first phase ($1\mathrm{e}{-6} - 5\mathrm{e}{-5}$). See Section~\ref{sec:results} for more details.
+lower than in the first phase ($1\mathrm{e}{-6}$ to $5\mathrm{e}{-5}$). See Section~\ref{sec:results} for more details.
 
@@
 Having defined the training procedures for the EN and CPN, we can define a training
-algorithm combining those two. We train the two in alternation, creating
+protocol combining those two. We train the two in alternation, creating
 a new EN each time we enter an EN training phase. We explored the possibility
 of reusing an existing EN but found that retraining an EN was no more fast than training a new one.
-Training a new EN also allows us to adapt to the brain's non-stationarity. That requires us to
+Training a new EN also allows us to adapt to the brain's non-stationarity, which requires us to
 determine when the current EN is no longer suitable for training the current CPN. 
@@
 
-We performed 8 experiments to investigate the co-processor's ability to learn under a variety of
-conditions: we studied three types of lesions to the grasping network (simulated stroke) and for
-each lesion type, we either turned on or off brain co-adaptation, natural recovery before co-processor
-use, and non-stationarity due to sensor drift. The following table summarizes the experiments
-(a blank entry under a condition (e.g., Simulate co-adaptation?) means ``No'' while an X means ``Yes''): \\
+We performed eight experiments to investigate the co-processor's ability to learn under a variety of
+conditions. We studied three types of lesions to the cortical network model (simulated stroke) and for
+each lesion type, we turned on or off (a) brain/co-processor co-adaptation (brain recovery during
+co-processor use), (b) brain recovery {\em before} co-processor use, and (c) non-stationarity due to
+sensor drift during co-processor use. The eight experiments are summarized in the following table 
+(a blank entry under a condition (e.g., Co-adaptation?) means ``No'' while an X means ``Yes''): \\
 
@@
 \hline
- & \textbf{Lesion}                                           & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ co-adaptation?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ recovery?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Simulate\\ sensor\\ drift?\end{tabular}}}\\ \hline
+ & \textbf{Lesion}                                           & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Co-adaptation?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Prior recovery?\end{tabular}}} & \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Sensor\\ drift?\end{tabular}}}\\ \hline
 1              & 50\% AIP loss                                             &                                                                                                 &                                                                                                &                                                                                                 \\ \hline
@@
 
-For each experiment, we used the mRNN grasping model from Michaels et al.\cite{michaels.mrnn}, 
-and the same set of input and output data used in their work. The dataset contains a total of
-502 trials, sampled uniformly across 42 object classes. In each experiment, we held out 
-a random sample of 20\% of the dataset for validation.
-
-\subsection{Stopping condition}
-In these experiments, we train the co-processor until one of two
-conditions is satisfied:
+Each of the above experiments used the same set of input and output data as was used by
+Michaels et al. \cite{michaels.mrnn} to train the cortical model that we use. The dataset
+contains a total of 502 trials, sampled uniformly across 42 object classes. In each experiment,
+we held out a random sample of 20\% of the dataset for validation.
+
+\subsection{Stopping criteria}
+In these experiments, we train the co-processor until one of the following two stopping criteria is
+satisfied (in Supplementary Materials section \ref{sup:stopping_criteria}, we probe how these criteria
+affect the results). The two criteria are:
 
@@
 	\item The percent change in average task loss between two consecutive ranges of 500 epochs
-	      is below a threshold of 0.01\%, indicating minimal benefits of further training. 
-	\item We have exceeded 250k epochs, which acts as a hard stopping criterion. 
+	      is below a threshold, indicating minimal benefits of further training. The choice of
+        threshold varies by experiment, due to the significant differences in setup. See Supplementary
+        Material section \ref{sup:stopping_criteria} for further details.
+	\item The training run has exceeded some number of epochs. This also varied by experiment.
 \end{itemize}
@@
 track the task loss, which is an MSE loss that measures the co-processor's ability to restore
-movement towards the target trajectory. We define it over the muscle length trajectory vector
-output from the lesioned grasping network model, compared to the ground truth muscle length data captured
-during the Michaels et al. experiment \cite{michaels.mrnn}. In graphs, we characterize this loss in terms of 
-percent recovery - i.e. difference between lesioned and healthy performance. Figs.
-\ref{fig:results_aip}-\ref{fig:results_drift} and Table \ref{tab:results} provide the detailed results.
-
-Second, we measure the degree to which lesioned grasping network, when coupled with the
-co-processor, successfully differentiates object classes. We compared the ratio of within-class
-to total variation between the lesioned and the healthy network. This metric $S$ is defined as:
+movement towards the target trajectory. We compute the loss based on the error between the muscle
+velocity trajectory output by the lesioned cortical network model and the ground truth muscle velocity
+data from the Michaels et al. experiment \cite{michaels.mrnn}. In our results, we characterize this loss
+in terms of percent loss recovery, i.e., difference between lesioned and healthy performance.
+Specifically the percent loss recovery metric is defined as:
+
+\begin{equation}
+  \mbox{percent loss recovery} = (1 - \frac{L - L_{healthy}}{L_{lesioned} - L_{healthy}}) \times 100 
+\end{equation}
+
+\noindent where $L$ is the current task loss, and $L_{healthy}$ and $L_{lesioned}$ are the task losses before the
+lesion was applied and after the lesion was applied (but before any recovery) respectively. Thus 0\% represents
+percent loss equal to the lesioned brain (prior to any recovery) while 100\% represents percent loss equal to the
+healthy brain. Note that it is possible for this metric to  become negative if current task loss $L$ exceeds
+$L_{lesioned}$ (e.g., if the stimulation is not beneficial and makes task performance worse than the
+performance with the lesioned brain).
+
+Second, we measure the degree to which the lesioned cortical model, when coupled with the
+co-processor, successfully differentiates the grasps for different object classes. We compare the ratio of
+total variation to within-class variation of muscle velocities, for grasps executed by the lesioned and
+the healthy cortical network. We define the grasp separability metric $S$ as:
 
@@
 
-where $\sigma$ measures the mean variation, across time, and across all data points in
-a given sample, $\sigma_{a}$ is the total variation of a dataset, and $\sigma_{w}$
-is the average of the within-class variations ($h$ indicates healthy). An $S$ value of 0.0 indicates that the
-grasps for the different object classes vary to the same degree as the healthy network.
-
-The metric $S$ allows us to differentiate overall task performance improvement
+\noindent where $\sigma$ measures the mean standard deviation, across time, and across all data points in
+a given sample, $\sigma_{a}$ is the standard deviation of a dataset, and $\sigma_{w}$
+is the average of the within-class standard deviations ($h$ indicates healthy). An $S$ value of 0.0
+indicates that the grasps for the different object classes vary to the same degree as the healthy network.
+A negative value indicates the grasping motions are more similar between the object classes; likewise a
+positive value indicates the grasps are more dissimilar.
+
+The grasp separability metric $S$ allows us to differentiate between overall task performance improvement
 and the ability of the co-processor to condition the grasp on the input object based on visual information.
@@
 the shape of the object being grasped, and must close around it appropriately.
-Note that the structure of the delayed reach-to-grasp task can potentially be learned by the co-processor,
-(e.g. to hold muscle velocities to 0.0 for the initial part of each trial) without the co-processor
-also learning to differentiate the various object shapes from the brain observations.
-Thus, $S$ is useful for differentiating improvement in object differentiation as opposed only overall
-decrease in task loss.
-
-\subsection{Experiment 1: AIP 75\% lesion}
+Note that it is possible for the structure of the delayed reach-to-grasp task to be potentially learned
+by the co-processor, (e.g., to hold muscle velocities to 0.0 for the initial part of each trial) without
+learning to also differentiate the various object shapes. $S$ provides a useful metric in this regard for
+tracking changes in grasp separability as opposed to only overall changes in task loss.
+
+\subsection{Experiment 1: AIP 50\% lesion}
 \begin{figure}[h]
 \centering
-\includegraphics[scale=1.3]{training_results_aip.png}
-\caption{Training Results: AIP 50\% Lesion}
+\includegraphics[scale=1]{training_results_aip.png}
+\caption{Co-Processor Performance Results: AIP 50\% Lesion. a. Percent loss recovery for the non-coadaptive
+         experiment for the task and validation datasets. 0\% represents task loss equal to the loss for the
+         lesioned cortical model, prior to any recovery. 100\% represents loss equal to that for the healthy
+         cortical model. Negative initial values are due to the newly-initialized CPN causing
+         worse performance than the lesioned model prior to recovery. b. Grasp separability metric ($S$) for the non-coadaptive
+         experiment. Negative $S$ values indicate grasps are more similar between object types while 
+         positive values indicate they are more distinct. c., d. show the performance of the neural co-processor for the
+         co-adaptive experiment.}
 \label{fig:results_aip}
@@
 Our first experiment tested the ability of the co-processor to compensate for the loss of a significant
-fraction of neurons in the area AIP of grasping model, which is responsible for encoding object shapres
-from visual inputs. Loss of AIP neurons results in the network model, without further adaptation, being unable to
-condition its grasp on the object shape, after largely succeeding in reaching toward the object.\footnote{Note
-in Supplementary Materials \ref{sup:results} that lesioned loss is lower in this experiment than
-in others: the user continues to reach successfully, but not to form the hand properly, as
-explore above. Likewise, the co-processor doesn't have sufficient information to improve
-performance, for lack of that information.} See Fig. \ref{fig:results_aip}a. Likewise, the co-processor cannot
-observe the object shape encoding in the brain, because that encoding is no longer happening. As a result,
-further recovery is not possible. Note in Fig. \ref{fig:results_aip}b that the co-processor never learns to
-separate object classes, which would have been indicated by an $S$ value near $0.0$.
-
-In the co-adaptive case, the brain and co-processor find a solution that reaches 76\% recovery towards non-lesioned
+fraction (50\%) of the model neurons in the AIP module of the cortical model; this module is responsible for encoding
+object shapes from visual inputs. Loss of AIP model neurons results in the cortical model, without further adaptation,
+being unable to condition its grasp on the object shape, after largely succeeding in reaching toward the object.\footnote{
+In Supplementary Materials section \ref{sup:results}, lesioned loss is lower in this experiment than in others: the subject
+continues to reach successfully, but does not form the hand properly.} Likewise, the co-processor cannot observe sufficient
+object shape encoding information due to the AIP lesion. As a result, as shown in Fig. \ref{fig:results_aip}a, significant
+recovery is not possible. Fig. \ref{fig:results_aip}b shows that the co-processor never learns to separate object classes
+($S$ value never approaches or exceeds $0.0$).
+
+In the co-adaptive case, the brain and co-processor together find a solution that reaches $76\%$ recovery towards healthy
 performance. This demonstrates that the co-processor can adapt to the non-stationarity
-of the mapping between the observed brain activity and the CPN-delivered stimulation.
+of the mapping between the muscle outputs and the CPN-delivered stimulation.
 As the brain adapts, it learns to encode the input visual information, which in turn allows the co-processor
-to observe that information through brain recordings. The co-processor can then leverage that information to condition
-stimulation. See Fig. \ref{fig:results_aip}c. Note, though, that the simulation is not designed to indicate if
-the co-processor speeds up recovery, or provides better results than natural recovery, since we aren't modeling the
-timeline of natural recovery. With this lesion design, simulated recovery would in-fact allow near-complete recovery
-of task performance. As a result, this experiment intends only to demonstrate co-adaptation.
-
-In Fig.~\ref{fig:results_aip}d, which plots the object separability metric $S$ discussed above, we see that initially
-the co-processor and brain do not strongly separate the object classes, resulting in negative $S$ values, but as the
-training proceeds, the metric exceeds 0.0, indicating separability. This can be attributed to the co-processor and the
-simulated ``brain'' learning how to map the input visual information about object shape to the appropriate grasp for the object.
-
-In Figs \ref{fig:results_aip}a and c, we see the co-processor initially causes worse performance, as
-indicated by the negative recovery rate. At this point the CPN is newly initialized, and has therefore
-not yet learned an interpretation of the recordings, or the problem structure.
+to observe that information through its ``recordings.'' The co-processor can then leverage that information to condition
+stimulation, improving task performance (Fig. \ref{fig:results_aip}c). Note, though, that this simulation is not designed
+to indicate if the co-processor speeds up recovery, or provides better results than natural recovery, since we aren't
+modeling the timeline of natural recovery. With this lesion design, simulated recovery would in-fact allow near-complete
+recovery of task performance. As a result, this experiment is designed  only to demonstrate co-adaptation.
+
+In Fig.~\ref{fig:results_aip}d, which plots the grasp  separability metric $S$, we see that initially
+the co-processor and the simulated brain do not strongly separate the object classes, resulting in negative $S$ values,
+but as the co-processor  training proceeds, the metric exceeds 0.0, indicating separability. This can be attributed to
+the co-processor and the simulated brain learning how to map the input visual information about object shape to the
+appropriate grasp for
+the object.
+
+In Figs.\ \ref{fig:results_aip}a and c, the initial percent loss recovery values are negative: this is because the
+newly-initialized CPN in the co-processor delivers stimulation that initially causes task performance to be worse
+than the performance of the lesioned simulated brain; this is corrected with further training of the CPN.
 
@@
 \centering
-\includegraphics[scale=1.3]{training_results_m1.png}
-\caption{Training Results: M1 50\% Lesion}
+\includegraphics[scale=1]{training_results_m1.png}
+\caption{Co-Processor Performance Results: M1 50\% Lesion. a. Percent loss recovery
+         for the non-coadaptive experiment for the task and validation datasets. See caption for Fig.~\ref{fig:results_aip}
+         for interpretation of positive and negative values. b. Grasp separability metric (S) for the non-coadaptive experiment.
+         See caption for Fig.~\ref{fig:results_aip} for interpretation of positive and negative values. c., d. show the
+         performance of the neural co-processor for the co-adaptive experiment. 
+         }
 \label{fig:results_m1}
@@
 
-In this experiment, we lesioned 50\% of the M1 module of the grasping network model. 
-In the non-coadaptive case, training quickly plateaus at 75\%, and does not improve
+In this experiment, we lesioned 50\% of the M1 module of the cortical model.
+As shown Fig.\ \ref{fig:results_m1}a, in the non-coadaptive case, training quickly plateaus at 75\%, and does not improve
 thereafter. This suggests the lesion inactivated some of the degrees of freedom needed to
 control the output layer, resulting in a reduced ability to recreate the different target grasps
-for different objects. See Fig. \ref{fig:results_m1}a.
+for different objects (Fig.\ \ref{fig:results_m1}b).
 
@@
 
-In the co-adaptive case, performance likewise hits an inflexion point at 75\%, but
+In the co-adaptive case (Fig. \ref{fig:results_m1}c), performance likewise hits an inflexion point at 75\%, but
 continues to improve slowly. The ability to improve further is tied to the
-``natural'' recovery occurring in the co-adapting network model, which restores some of the
-degrees of freedom needed for improving grasping performance. The co-processor in this case adapts
-successfully to this non-stationarity due to the recovery process. See Fig. \ref{fig:results_m1}c.
+``natural'' recovery occurring in the co-adapting cortical model, which restores some of the
+degrees of freedom needed for improving grasping performance (Fig.\ \ref{fig:results_m1}d). The co-processor 
+n this case adapts successfully to this non-stationarity being caused by the recovery process.
 
@@
 \includegraphics[scale=1]{training_results_con.png}
-\caption{Training Results: F5-M1 Connection Lesion}
+\caption{Co-Processor Performance Results: F5-M1 Connection Lesion. 
+          a. Percent loss recovery for the non-coadaptive experiment for the task and validation datasets.
+          See caption for Fig.~\ref{fig:results_aip} for interpretation of positive and negative values.
+          b. Grasp separability metric (S) for the non-coadaptive experiment. See caption for Fig.~\ref{fig:results_aip} 
+          or interpretation of positive and negative values. c., d. show the performance of the neural
+          co-processor for the co-adaptive experiment.
+         }
 \label{fig:results_con}
@@
 \includegraphics[scale=1]{training_results_no_explore.png}
-\caption{Training Results: F5-M1 Connection Lesion. Use of exploratory
+\caption{Co-Processor Performance Results: F5-M1 Connection Lesion: Use of exploratory
          and randomized examples in the training data set lead to stabilized
-         and faster training. (a) Without exploratory examples in the
+         and faster training. a., b. Without exploratory examples in the
          neighborhood of the CPN, we have unstable and slower training. In
          this case, the data set is composed of $10\%$ examples drawn from the
-         CPN, and $90\%$ randomized examples (no exploration).
-         (b) Using only examples from the CPN and/or exploratory examples also
-         leads to unstable training. Here we additionally show experiments with
-         $90\%$ exploratory examples (no noise), and $100\%$ examples drawn from
-         the CPN (neither).}
+         CPN, and $90\%$ randomized examples (``No exploration'').
+         c., d. Using only examples from the CPN (``Neither''; no exploratory or randomized examples) or $10\%$ CPN
+         and $90\%$ exploratory examples (``No noise''; no randomized examples) also
+         leads to unstable training. In c.\ and d., we truncate the experiments and use different
+         axes to allow for clearer plots, but show the data from a.\ and b.\ for
+         comparison.
+         }
 \label{fig:results_con_no_explore}
@@
 act as a bridge between the two areas to appropriately convey information required for the grasping
-task from one area to the other. As in the other experiments, the co-processor's loss in this experiment
-improved quickly, but took much longer to refine the stimulation patterns to enable 
-object differentiation. See Fig. \ref{fig:results_con}a and c.
+task from one area to the other. As in the other experiments, the co-processor's task performance in this
+experiment improved quickly (Fig.~\ref{fig:results_con}a and c), but it took much longer to refine the
+stimulation patterns to enable object differentiation (Fig.~\ref{fig:results_con}b and d).
 
 Since feedback from the output cannot ``backpropagate'' to the F5 and AIP modules, co-adaptation and
-learning in this experiment only affects the M1 module. As a result, this experiment demonstrates
-that the co-processor's learning algorithm is capable of adapting to non-stationarity
-in the network model's mapping between stimulation parameters and the behavioral output (muscle velocities
+learning in this experiment only affects the M1 module. As a result, the co-processor's performance in
+this experiment demonstrates that its learning algorithm is capable of adapting to non-stationarity
+in the cortical model's mapping between stimulation parameters and the behavioral output (muscle velocities
 for grasping). These results contrast  with the results of Experiment 1, where the non-stationarity affected
@@
 
-As seen in Fig.~\ref{fig:results_con}b and d, object separation is initially low and unstable, as the
-co-processor begins to learn the task. Afterwards, it stabilizes as the co-processor begins to
-leverage observed brain activity to differentiate object shapes, gradually converging towards the
+As seen in Fig.~\ref{fig:results_con}b and d, grasp separation for objects is initially low and unstable,
+as the co-processor begins to learn the task, but later stabilizes as the co-processor begins to
+leverage observed brain activity to differentiate object shapes, gradually meeting or exceeding the
 pre-lesion grasping performance.
 
-To illustrate the effect of EN training data set composition as described in \ref{sec:training}, we
-performed additional experiments where we vary the data set composition. We chose to base those
-experiments on the F5-M1 connection lesion arbitrarily. As outlined in Training (\ref{sec:training}),
-the training data set for the EN is composed of stimulation examples drawn from the current CPN,
-examples drawn from other CPNs with parameters in the neighborhood of the current CPN in parameter
-space (referred to as ``exploratory examples''), and randomized stimulation. In Fig.
-\ref{fig:results_con_no_explore} we see the effect of removing the latter two of those.
-Exploratory examples proved necessary to stabilize training, and to achieve higher levels of recovery,
-as in Fig. \ref{fig:results_con_no_explore}a. Removing the examples of randomized stimulation results in
-even more unstable training, where no improvement in task performance occurs.
+To illustrate the effect of the EN training dataset composition as described in section~\ref{sec:training}, we
+performed additional experiments on the F5-M1 connection lesion where we varied the dataset composition. As
+outlined in section~\ref{sec:training}), the training data set for the EN is composed of stimulation examples
+drawn from the current CPN, examples drawn from other CPNs with parameters in the neighborhood of the current
+CPN in parameter space (referred to as ``exploratory examples''), and randomized stimulation. In
+Fig.~\ref{fig:results_con_no_explore}, we see the effect of removing the latter two types of examples.
+Exploratory examples proved necessary to stabilize training, and to achieve higher levels of recovery. Removing
+the examples of randomized stimulation resulted in even more unstable training, where no improvement in task
+performance occurs.
 
 \subsection{Experiment 4: Connection lesion with recovery}
+It is often the case that after a stroke, the brain recovers some of the function lost immediately after the
+stroke. We simulated this natural recovery in our cortical  network model by lesioning the F5-M1 connections
+and then training the lesioned network on the grasping task. Complete recovery is impossible because the M1 module's
+connections to F5 (both forward and backward) are disconnected. The cortical network model in this case learns
+an object-agnostic stereotypical grasp. We tested whether a co-processor can subsequently improve grasping performance
+beyond natural recovery by learning to convey to M1 via stimulation the object-related information from AIP and F5
+necessary to tailor the grasp to the current input object's shape.
 
@@
 \includegraphics[scale=1]{training_results_recov.png}
-\caption{Training Results: Connection Lesion with Recovery}
+\caption{Co-Processor Performance Results: Connection Lesion and Natural Recovery before Co-Processor Use.
+         a. Percent loss recovery for the task and validation datasets. The horizontal dashed line is the task
+         performance after both lesioning and natural recovery but before co-processor use. Initially the
+         co-processor's stimulation causes task performance to be worse than the performance after natural
+         recovery, but eventually the co-processor's performance exceeds this level. b. Grasp separability
+         metric (S). Negative S values indicate grasps are more similar between object types while positive values
+        indicate they are more distinct.
+        }
 \label{fig:results_recov}
@@
 
-It is often the case that after a stroke, the brain recovers some of the function lost immediately after the
-stroke. We simulated this recovery in the grasping network model after lesioning the F5-M1 connections by
-training the lesioned network on the grasping task. Complete recovery is impossible because the M1 module's
-connections to F5 (both forward and backward) are disconnected. The grasping network model in this case learns
-an object-agnostic stereotypical grasp. We tested whether a co-processor can improve grasping performance by
-propagating to M1 the object-related information from F5 necessary to tailor the grasp to the current
-input object's shape.
-
-We found that the co-processor can successfully improve task performance beyond the ``natural''
-recovery after stroke (Fig.~\ref{fig:results_recov}a). The dashed line
-indicates the task loss after initial recovery. As in some experiments above, loss is
-initially higher at first, as the co-processor trains. Eventually however it learns
-to drive the task loss lower, as it learns to forward-propagate information from the
-``earlier'' parts of the simulated brain. Also, note the object class separation metric $S$
-returned to a healthy value towards the end of co-processor training
-(Fig.~\ref{fig:results_recov})b.
+We found that the co-processor can successfully improve task performance beyond the natural
+recovery after the simulated ``stroke'' (Fig.~\ref{fig:results_recov}a). The dashed line
+indicates task loss after initial recovery. As in some experiments above, task performance is
+initially worse than after recovery, due to stimulation by an  initially untrained co-processor. However
+the co-processor gradually learns to drive the task loss lower, as it learns to forward-propagate information
+from the ``earlier'' parts (AIP and F5) of the simulated brain. Also, note that the grasp separation metric $S$
+returns to a healthy value towards the end of co-processor training
+(Fig.~\ref{fig:results_recov}b).
 
 \subsection{Experiment 5: Connection lesion with sensor drift}
+Sensor-related non-stationarities are
+often seen in real-world neural recording systems due to a variety of factors, from impedance
+changes due to sensor movement to scar tissue formation. In our final experiment, we again lesioned the F5-M1
+connections and allowed co-adaptation by the simulated cortical model but additionally, we allowed the sensor
+readings to drift over time. This drift was modeled in the recording function, where the readings had a non-stationary
+zero point across epochs. We added a bias term which we updated between epochs according to a random process
+(see Supplementary Materials section~\ref{sup:drift} for details).
+
 \begin{figure}[h]
@@
 \includegraphics[scale=1]{training_results_drift.png}
-\caption{Training Results: Connection Lesion with Sensor Drift}
+\caption{Co-Processor Performance Results: Connection Lesion with Sensor Drift. a. 
+         Percent loss recovery for the task and validation datasets. 0\% represents task loss equal to the loss for
+         the lesioned cortical model, prior to any recovery. 100\% represents
+         loss equal to that for the healthy cortical model. 
+         b. Grasp separability metric (S) which reflects the co-processor's ability
+         to tailor grasps to specific objects. A negative value indicates the grasps are less
+         differentiated between object shapes. Due to sensor drift, the co-processor
+         is much slower to differentiate objects and the trend exhibits 
+         higher noise compared to other experiments.}
 \label{fig:results_drift}
@@
 
-In our final experiment, we again lesioned the F5-M1 connections and allowed co-adaptation by the
-simulated model network but additionally, allowed the sensor readings to drift over time. This
-drift was modeled in the recording function, where the readings had a non-stationary zero point
-across epochs. We added a bias term which we updated between epochs according to a random process.
-See Supplementary Materials \ref{sup:drift} for details. Sensor-related non-stationarities are
-often seen in real-world neural recording systems due to a variety of factors, 
-from impedance changes due to sensor movement to biological scar tissue formation.
-
 As seen in Fig.~\ref{fig:results_drift}, the co-processor was able to quickly
-recover the reach-to-grasp part of the task, and gradually learned to
-condition the grasp on the object information. Because the recording model is non-stationary
+recover the reach-to-grasp part of the task, and then gradually learned to
+condition the grasp on the object information.\footnote{Note that we relaxed our 250k epoch
+         stopping criterion for these results because sensor drift causes a longer
+         convergence time.} Because the recording model is non-stationary
 in this experiment, the object class separation exhibits far greater epoch-to-epoch
-variance in the later epochs. The co-processor presumably becomes more
-reliant on upstream visual information over time, as it learns how to leverage this
-in order to differentiate object shapes.\footnote{For this experiment, we opted to stop training
-after running the experiment long enough to get results near to the version
-of this experiment without sensor drift.}
-
-As expected, training efficiency was decreased relative to
-the experiment without sensor drift: with the sensor drift,
-282K training epochs was needed to reach $85\%$ recovery while
-without sensor drift,  only 93K was required.
+variability in the later epochs. The co-processor presumably becomes more
+reliant on upstream visual information over time, as it learns to leverage this information 
+in order to differentiate object shapes.
+
+As expected, training efficiency was decreased relative to the experiment without sensor
+drift: with the sensor drift, 282K training epochs were needed to reach $85\%$ recovery
+(Fig.~\ref{fig:results_drift}a) while without sensor drift, only 93K were required
+(Fig.~\ref{fig:results_con}a).
 
@@
 We present here a first-of-its-kind demonstration of a novel design for neural co-processors,
-allowing for an AI agent to learn neural stimulation policies that improve a user's performance
-of an external task. The design revolves around the use of a stimulation model for training the
-agent, providing a proxy to the true function mapping stimulation and neural activity
-to task performance. We base our demonstration on a simulation of a neural circuit engaged in
-an external grasping task.
-
-Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
-75-90\% across our various experiments. To work this well, the co-processor needed to adapt to
-the long-running dynamics of the simulated network, as well as the long-running effects of
-stimulation. In some experiments we required it to additionally adapt to non-stationarity in
-the neural circuit, which was actively changing at the same time the co-processor was learning, which it achieved as well.
-In one experiment, it also adapted to a simulated brain which had already undergone some amount of
-lesion recovery. In that case, the co-processor successfully identified the information upstream
-from the lesion which was necessary to stimulate the motor cortex (M1) module
-downstream. In the final experiment, it also successfully adapted to a non-stationary recording model.
-
-A previous example of such a simulation approach is the work of Dura-Bernal et al. \cite{bernal.sim}. In this work, the
-authors used a simulated spiking neural network to train a stimulation agent.
-Their stimulation agent sought to restore the network's
-control of a simulated arm to reach a target, after a simulated lesion was
-applied. As the authors point out, we have a limited ability to
-probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
-we first need to design our approach through the use of an admissible simulation.
-As we did, the authors in this case simulated lesions by effectively removing parts
-of their simulated network, or by cutting connections between parts of the network.
-
-Clearly, we need to discuss what can and cannot be inferred from simulation studies
-such as these. We note that our simulation differs drastically from an organic brain, both in scale and structure. Today, predicting long-running neural responses
-to stimulation remains a difficult problem, implying targeted neural control is also difficult. Even
-more so, identifying the neural correlates of complex task success remains largely out of reach
-\cite{khanna.openloop}. What, then, does our simulation provide us?
-
-Our simulation does not constitute evidence that our method will, for example, immediately translate
-to true restoration of fine-grained control for individual fingers of a stroke patient. One clear
-challenge with doing so would be the sheer dimensionality of natural brain dynamics, and how
-that compares to the resolution of stimulation the co-processor can learn to apply.
-There exists a  mismatch between the dimensionality of the underlying problem, sensor and
-stimulator technologies, and the amount of training data which can reasonably be collected to
-train a closed-loop neural controller. We argue that while this fact clearly creates a challenge
-for learning-based closed-loop stimulation, that nevertheless the insights we generate through this
-simulation are likely to be applicable.
-
-First, this simulation requires the co-processor to contend with issues of long-running
-dynamics, high dimensionality, and non-stationarity, which are likely to be key issues
-facing real world deployments. We designed our training algorithm to contend with dimensionality
-by training the CPN with a local stimulation model. The algorithm adapts to non-stationarity by
-regularly updating that stimulation model, allowing it to adapt to changes in the brain, in
-addition to ``following'' the CPN through stimulation parameter space.
-
-Second, the co-processor abstractly lends itself to any closed-loop neural stimulation problem
-where a relevant stimulation model can be identified. There are many lower dimensional problems
-today which we address with far simpler stimulation regimes than would be necessary for
-e.g. fine-grained finger control. In some applications, such as Parkinson's disease symptom relief,
-stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
-parameters, where the key challenge is learning when to apply the stimulation, and
-its shape and power. Our simulation demonstrates the ability of our training method to
-adapt to higher dimensional problems than these, where additionally the long-running
-effects of stimulation on the intended figure-of-merit must be modeled. These properties
-together suggest it may be useful in lower dimensional problems where a stimulation model
-can be estimated, and stimulation parameters can be tied to the target objective.
-
-Our future work will include a reinforcement learning (RL) approach, where the co-processor
-explicitly learns when to apply stimulation, in addition to parameterizing it. Its target
-objective involves not only symptom relief, but also minimizing energy use. In this approach,
-our CPN and EN may correspond to the Actor and Critic portions of an Actor-Critic model
-for example. Here, the Critic would learn to value symptom relief, while attaching a cost
-to stimulation, thus incentivizing an energy efficient approach.
-
-One challenge remaining with our current co-processor design is training efficiency.
+showing in simulation crucial training and design methodologies that lay the groundwork for
+co-processors to be demonstrated \textit{in vivo} in the future. Our results show that a deep
+learning-based co-processor network (CPN) can learn neural stimulation policies that improve
+performance of an external task after simulated lesions in different parts of a cortical model.
+The training of the CPN relies on the use of a second neural network, called an emulator network (EN), 
+which learns to approximate the function mapping stimulation and neural activity to task performance.
+Our experiments revealed the effects of different parameter choices and training paradigms for CPNs and
+ENs, providing new insights for future {\em in vivo} studies of neural co-processors.
+
+Our co-processor design adapted well to a variety of simulated lesion types, reducing task loss
+$75-90\%$ across our various experiments. The co-processor successfully adapted to
+the long-running dynamics of the simulated cortical model, as well as the long-range effects of
+stimulation. In some experiments, we required the co-processor to additionally adapt to non-stationarity in
+the neural circuit, which was actively changing at the same time the co-processor was learning. In one
+experiment, it also adapted to a simulated brain which had already undergone some amount of natural recovery
+after a lesion. In this case, the co-processor successfully identified the information upstream from the 
+lesion which was necessary to stimulate the motor cortex (M1) module downstream. In the final experiment,
+the co-processor also successfully adapted to sensor drift simulated by a non-stationary recording model.
+
+An important question that needs to be addressed is what can and cannot be inferred from simulation studies
+such as ours. Today, predicting neural responses to stimulation over long time scales and complex stimulation
+patterns remains a difficult problem, implying targeted control of neural activity is also difficult. Our
+simulation results  do not constitute evidence that our method will, for example, immediately translate to
+restoration of fine-grained control for grasping in a stroke patient. One clear challenge is the sheer
+dimensionality of the problem. There exists a mismatch between the dimensionality of the brain and current
+sensor and stimulator technologies. We are also severely limited in the amount of training data which can
+reasonably be collected to train a closed-loop neural controller. We argue that while this fact clearly
+creates a challenge for learning-based closed-loop stimulation, the insights we have gained through our
+simulation studies are likely to be useful for {\em in vivo} testing of neural co-processors.
+
+As mentioned in the Background section~\ref{sec:background}, it is important to test 
+the feasibility of applying a complex AI method such as a neural co-processor prior to testing it
+\textit{in vivo}. Our simulations allowed us to design and test training methods around which
+future \textit{in vivo} experiments can be developed. Additionally, iterating over different design
+choices in simulation minimizes the use of animal experimentation, supporting a commonly-accepted
+maxim regarding animal welfare in experimentation \cite{nrc.care}.
+
+Our simulation studies required the co-processor to contend with issues of long-running dynamics and
+non-stationarity, which are likely to be key issues facing real-world {\em in vivo} deployments. We designed
+our training algorithm to contend with dimensionality by training the CPN with an EN that learns the effects
+of stimulation. The training algorithm adapts to non-stationarity by regularly updating the EN, allowing it
+to adapt to changes in the brain, in addition to ``following'' the CPN through stimulation parameter space.
+In the next section, we briefly explore the possibility of further addressing the dimensionality problem by
+using neural co-processors for optimizing low-dimensional neural correlates of task performance.
+
+\subsection{Co-processors for optimizing low-dimensional neural correlates of task performance}
+Though our simulation involves a neural co-processor directly optimizing stimulation for
+external task performance, the co-processor concept also generalizes to optimization for
+\textit{neural correlates} (NCs) of task performance. We present this concept here to illustrate
+the generality of our approach, and to provide a possible path towards tackling the dimensionality issue.
+NCs are defined over measured brain activity and can be correlated with the desired patient outcome measures
+such as improved dexterity. If such a correlate additionally possesses a causal relationship to the patient
+outcome, i.e., if driving the NC through stimulation towards particular regimes also improves patient outcome,
+then it may serve as a co-processor's optimization target. For example, Khanna et al. \cite{khanna.openloop}
+suggest that ``changes in task-related ensemble firing are linked to improvements in dexterity.''
+The authors then provide a measurement of ensemble co-firing which constitutes a candidate NC of dexterity
+and which may potentially be driven through stimulation.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.88\textwidth]{cofiring_model.png}
+	\caption{\textbf{A co-processor leveraging a neural correlate of dexterity rather
+    than external (e.g., muscle) measurements.} Here, a measure $M$, serving as a neural correlate of
+    dexterity, is calculated based on observed brain activity $O$. The emulator network (EN) learns
+    to predict the future values of $M$ as a function of $O$ and stimulation parameters
+     $\theta$. The co-processor network (CPN) maps the observed brain activity $O$ onto stimulation
+    parameters. The CPN is trained using the co-processor approach presented in this article with the
+    objective of maximizing the dexterity measure $M$ rather than minimizing a task loss defined in terms
+    of grasping motions or muscle velocities.}
+	\label{fig:correlate}
+\end{figure}
+
+A co-processor optimizing for an NCoD may provide a plausible path towards \textit{in vivo}
+application. Driving a wide variety of grasping behaviors by altering the firing rates of a small number of
+individual neurons or multiunits may be infeasible in practice, and may not generalize well beyond
+the set of training examples. On the other hand, optimizing a neural correlate may generalize across a wide
+variety of behaviors. Additionally, such an optimization problem may be lower dimensional and require less
+training data in practice. We illustrate a design for such a co-processor in Fig.~\ref{fig:correlate}.
+
+% Many of the neural stimulation techniques cited throughout this work
+% involve stimulation parameterized with few variables. Castaño-Candamil et al.
+% \cite{castano.pd} used an ML-based approach to closed-loop stimulation for essential
+% tremor (ET) where a simple learning model extracted neural markers (NMs) from a human
+% patient's M1 cortex and used the NMs to determine deep brain stimulation
+% (DBS) stimulation parameters. In this case, the model used as input a 20 dimensional
+% vector sourced from eCog and tremor incidence data, and outputted a single scalar
+% stimulation parameter. The authors used a simple linear model which required few
+% parameters, yet it was effective. Our co-processor model is likely to transition
+% well to such an application by being able to extract neural markers appropriate for
+% various contexts (e.g. sitting or standing), and doing so where hand-designed features are
+% difficult to identify. Note that such a co-processor would not be modeling the firing
+% rates of the underlying brain's neurons as it does in our simulation, but would
+% rather model a higher-level quantity such as beta band power.
+
+\subsection{Future work}
+Identification of neural correlates such as NCoDs for different behaviors remains an open problem whose
+solution may enable efficient types of future co-processors. Similar to Khanna et al.\ \cite{khanna.openloop},
+Heimbuch et al.\ recently presented work suggesting that there exist low-dimensional neural correlates of
+dexterity which can be measured during and after stroke recovery \cite{heimbuch.sfn}. They further hypothesize
+that neural stimulation could drive brain activity in a stroke patient towards maximization of those metrics,
+and that doing so would cause an improvement in dexterity. A co-processor could allow us to test
+these hypotheses by attempting to learn a CPN stimulation policy which optimizes the candidate
+NCoDs. Future work will involve collaborations with experimental neuroscientists to identify candidate NCs of
+desirable behavior, and to perform {\em in vivo} experiments involving co-processors that seek to optimize
+these NCs.
+
+Additional future work will explore reinforcement learning (RL) approaches to training co-processors: the
+neural co-processor we explored in this article already includes a ``policy network'' (the CPN) and a
+``world model'' (the EN). The proposed framework is therefore well-suited to model-based reinforcement
+learning algorithms such as model-based Actor-Critic learning. We will explore optimizing reward functions
+that involve not only symptom relief, but also minimizing energy use by attaching a cost to stimulation.
+
+One additional challenge remaining with our current co-processor design is data efficiency.
 On the whole, the co-processor quickly improved task performance, but required orders of
@@
 moderate amounts of recovery may be valuable to a user, but nevertheless we consider
-training efficiency to remain a problem. Due to limits on patient fatigue, time, implant
+data efficiency to remain a problem. Due to limits on patient fatigue, time, implant
 battery life, and other concerns, it is not plausible to expect a learning algorithm
 to have access to unlimited amounts of training data. As a result, it is necessary to
-make efficient use of the data we can acquire. We believe there exist at least three
-mitigations for that:
+make efficient use of the data we can acquire. We believe there exist at least four
+mitigation strategies for this:
 
 \begin{itemize}
-	\item Retraining an existing EN, rather than regularly creating a new one. As
-	      mentioned above, we encountered difficulty with this approach, but it remains
-	      to be seen if this is a fundamental problem with our training method, rather
-	      than a peculiarity of our simulation. Also, it remains unclear if a solution
-	      for our simulation exists as well, which we've simply yet to identify. This
+	\item Retraining an existing EN, rather than repeatedly creating a new one. As
+	      mentioned above, the former approach did not yield good results, but it remains
+	      to be seen if this is a fundamental problem with our training method or
+	    a peculiarity of our simulation which we have yet to identify. This
 	      remains a future area of inquiry.
-	\item Making better use of what data we acquire. In this initial simulation we do
+	\item Making better use of what data we acquire. In this initial simulation, we do
 	      not retain data beyond the present training epoch because non-stationarity requires us to
-	      regularly discard or discount data as it ages. We found that in the presence
+	      regularly discard data as it ages. We found that in the presence
 	      of a rapidly learning CPN, and especially in our co-adaptive experiments,
-	      data retention in-fact caused training instability such that training
-	      eventually required more data collection than otherwise. However, in
+	      data retention in fact caused training instability. However, in
 	      practice, training epochs operate on the order of seconds, suggesting
-	      that data can be retained and reused for multiple epochs. The `speed' of 
-	      non-stationarity compared to the co-processor's learning therefore is an
-	      area for future research.
-	\item Matching the dimensionality of the stimulation and recording parameters
-	      to the amount of available data. A stimulation regime with
+	      that data could be retained and reused for multiple epochs. The ``speed'' of 
+	      non-stationarity compared to the co-processor's learning is an
+	      area worthy of future investigation.
+	\item Matching the dimensionality of the stimulation parameters and the optimization
+        target to the amount of available data. A stimulation paradigm with
 	      a small number of controllable parameters that nevertheless allows
-	      improvements in task performance  will likely require less training data.
+	      improvements in task performance will likely require less training data. Likewise,
+        optimizing low-dimensional targets (e.g., NCoDs) may require less training
+        data.  
+    \item Sharing data across subjects and transfer learning. Recent work has shown that
+        it is possible to train a deep neural network using data from multiple subjects and
+        transfer that knowledge to decode neural data from new subjects \cite{peterson.transfer}.
+        A similar technique could potentially be used to train ENs, reducing the amount of
+        data needed from a new subject and improving data efficiency.
 \end{itemize}
 
-Future work will test the neural co-processor framework on  lower dimensional
-problems that are likely to transfer more readily to \textit{in vivo} experiments
-and human trials. We hypothesize that the co-processor framework will transition well
-to treatments for conditions such as Parkinson's disease (PD), and essential tremor
-(ET). For example, Castaño-Candamil et al. \cite{castano.pd} used an ML-based approach
-to closed-loop stimulation for ET where a simple learning model extracted neural
-markers (NMs) from a human patient's M1 cortex and used the NMs to determine deep
-brain stimulation (DBS) stimulation parameters. Our co-processor model is
-likely to transition well to such an application by being able to extract neural
-markers appropriate for various contexts (e.g. sitting or standing) and applying 
-stimulation appropriate to the current context.
-
 \section{Conclusion}
-Using a simulated model of circuits in the primate brain involved in grasping, we
-demonstrated that a closed-loop neural stimulator, called a ``neural co-processor''
-can be trained to restore grasping function after different types of lesions to the
-model. We showed a neural co-processor can be  trained using supervised learning
-through backpropagation with the help of two networks: a co-processor network (CPN)
-that learns to generate stimulation patterns to optimize task performance, and an
-emulator network (EN) that learns to predict the effects of stimulation. Our results
-show that a neural co-processor can improve task performance after a lesion by
-co-adapting with the simulated brain model in the presence of nonstationarities in the
-brain and sensors. Given the generality of the framework, we expect neural
-co-processors  to be applicable to a wide range of clinical applications that
-required adaptive closed-loop neural stimulation.
-
-\section{Data availability statement}
+Using a simulated model of cortical circuits in the primate brain that are involved in grasping
+behaviors, we demonstrated a training method based on deep learning for a closed-loop neural
+stimulator called a ``neural co-processor'' \cite{rao.coproc,rao.braincoproc}. We created a
+learning paradigm which allowed the co-processor to adapt to the distributed neural activity
+in the simulated brain, as well as to the non-stationarity of the neural activity and the 
+sensors, key properties which are likely necessary for future \textit{in vivo} applications.
+We showed that the co-processor could be trained to restore grasping function after we applied
+varying types of simulated lesions. Specifically, a neural co-processor can be trained using
+deep learning through backpropagation with the help of two networks: a co-processor
+network (CPN) that learns to generate stimulation patterns to optimize task performance, and an
+emulator network (EN) that learns to predict the effects of stimulation. Though the results we presented 
+involved optimization of an external task, we believe that neural co-processors will successfully
+generalize to the optimization of neural correlates of health, successfully driving neural activity
+towards target regimes which are known to correlate with positive clinical outcomes. Given the generality
+of the framework, we expect neural co-processors to be applicable to a wide range of clinical applications
+that require adaptive closed-loop neural stimulation for treating sensorimotor disorders and
+neuropsychiatric conditions.
+
+\section*{Code and data availability}
 Code including analysis code used to
@@
 
-\section{Conflict of interest statement}
+\section*{Conflict of interest}
 The authors of this work are not aware of any conflicts of interest
@@
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 This work was supported by a Weill Neurohub Investigator grant, a CJ and Elizabeth Hwang
-endowed professorship (RPNR), and National Science Foundation grant no. EEC-1028725.
-The authors would like to thank Karunesh Ganguly (UC San Francisco), Priya Khanna
-(UC Berkeley), Anca Dragan (UC Berkeley), Justin Ong, and Luciano De La Iglesia
-for discussions and insights related to this work.
-
-\section{References}
+endowed professorship (RPNR), National Science Foundation (NSF) grant no.\ EEC-1028725 and NSF EFRI
+grant no.\ 2223495. The authors would like to thank Karunesh Ganguly (UC San Francisco), Priya Khanna
+(UC Berkeley), Anca Dragan (UC Berkeley), Justin Ong, Ian Heimbuch, and Luciano De La Iglesia for
+discussions and insights related to this work.
+
+\section*{References}
 \bibliographystyle{iopart-num}
@@
 \section{Supplementary Materials}
-\subsection{Details of the simulated task.}
+\subsection{Details of the simulated task}
 \label{sup:michaelstask}
 
-The reach-to-grasp task on which the mRNN was trained involves multiple phases, which can be identified by the
+The reach-to-grasp task on which the mRNN cortical model was trained involves multiple phases, which can be identified by the
 mRNN's inputs. Initially, the NHP subject sat in a dark room, with only a red cue visible. After a short period,
-an object was presented to the subject visually, after which the overhead light was turned off and once again only the red
-cue was visible. After another brief period, the red cue was turned off, which cued the subject to perform a grasp and hold of
-the object (still in the dark). This task is presented in more detail in the work of Susillo \cite{susillo.mrnn} and Michaels et al.
-\cite{michaels.mrnn}.
-
-The inputs to the mRNN network encode the task phase information. It consists of a twenty dimensional vector of visual
-features, which are presented only to the input (AIP) end of the network, and a 1D hold signal, which is presented to all
-neurons across the network. The authors of Michaels et al. \cite{michaels.mrnn} derived the hold signal
-from the grasp data, and it represents a point in time shortly before grasping began. As explained above, the visual features are
-drawn from VGGNet \cite{simonyan.vgg}, based on rendered images. The images consist of some combination of the red cue
-light and the object image. Note that the visual input vector and hold signal are not observed directly by the co-processor,
-but rather indirectly via brain recordings.
-
-The phases of the task are depicted in Table \ref{tab:phases}, and the timeline of a single trial is depicted in Fig. \ref{fig:trial_timeline}.
+an object was presented to the subject visually, after which the overhead light was turned off and once again
+only the red cue was visible. After another brief period, the red cue was turned off, which cued the subject
+to perform a grasp and hold of the object (still in the dark). This task is presented in more detail in the
+work of Sussillo \cite{sussillo.mrnn} and Michaels et al. \cite{michaels.mrnn}.
+
+The inputs to the mRNN network encode the task phase information. It consists of a twenty dimensional vector
+of visual features, which are presented only to the input (AIP) end of the network, and a 1D hold signal,
+which is presented to all neurons across the network. The authors of Michaels et al. \cite{michaels.mrnn}
+derived the hold signal from the grasp data, and it represents a point in time shortly before grasping began.
+As explained above, the visual features are drawn from VGGNet \cite{simonyan.vgg}, based on rendered images.
+The images consist of some combination of the red cue light and the object image. Note that the visual input
+vector and hold signal are not observed directly by the co-processor, but rather indirectly via brain
+recordings.
+
+The phases of the task are depicted in Table \ref{tab:phases}, and the timeline of a single trial is
+depicted in Fig. \ref{fig:trial_timeline}.
 
@@
 \end{tabular}
-\caption{\label{tab:phases} Phases of the reach to grasp task. During each phase, a visual stimulus as well as a binary hold signal are presented.
-The visual stimulus encodes an image, consisting of a red cue light and/or an image of an object.}
+\caption{\label{tab:phases} \textbf{Phases of the reach to grasp task.} During each phase, a visual stimulus as well
+         as a binary hold signal are presented. The visual stimulus encodes an image, consisting of a red cue light
+         and/or an image of an object.}
 \end{table}
@@
 
-\subsection{Training the grasping model network.}
+\subsection{Training the cortical model for grasping}
 \label{sup:michaelstraining}
 
-For our simulation, we leverage a pre-trained mRNN model, provided by the
-lead author of Michaels et al. \cite{michaels.mrnn}, and used with their permission.
+For our simulation, we leverage a pre-trained mRNN cortical model, provided by the
+lead author of Michaels et al. \cite{michaels.mrnn} and used with their permission.
 The training method for that model is based on prior work by the same authors, and
-others \cite{susillo.mrnn}.
+others \cite{sussillo.mrnn}.
 
@@
 
-\subsection{EN/CPN Interleaving.}
+\subsection{EN/CPN Interleaving}
 \label{sup:encpninter}
@@
 \begin{itemize}
-	\item The EN prediction loss is greater than $min(6\mathrm{e}{-4}, L/10)$, where $L$ is the most recent
+	\item The EN prediction loss is greater than $\min(6\mathrm{e}{-4}, L/10)$, where $L$ is the most recent
 	      task loss. EN prediction loss is an MSE loss between the EN's prediction of muscle
-	      trajectories, and the actual muscle trajectories output by the grasping model network.
+	      velocities, and the actual muscle velocities output by the grasping model network.
 	\item Task loss increased across 15 of the previous 30 training epochs.
@@
 Likewise, the EN training period ends when its prediction loss $L$ on the validation data set drops
-below a threshold of $max(3\mathrm{e}{-4}, L/50)$.
-
-\subsection{Sensor drift.}
+below a threshold of $\max(3\mathrm{e}{-4}, L/50)$.
+
+\subsection{Sensor drift}
 \label{sup:drift}
@@
 
-\subsection{Passthrough reference model.}
+\subsection{Passthrough reference model}
 \label{sup:passthrough}
@@
 To understand the effect of the recording and stimulation functions on the co-processor's performance, 
-we repeated our experiment with the F5-M1 connection lesion and coadaptation, with passthrough recording
+we repeated our experiment with the F5-M1 connection lesion and coadaptation using passthrough recording
 and stimulation functions. ``Passthrough'' here means that the recording and stimulation functions have
@@
 we increased the CPN from 61 to 200 neurons, and the EN from 87 to 351. We measured whether the increase in
-neurons alone caused in the different results we see here, and found it did not; we omit those results for
+neurons alone caused the different results we see here, and found it did not; we omit those results for
 brevity.
@@
 In Fig. \ref{fig:results_passthrough} we see that the co-processor performed significantly better with
-the passthrough functions. It learned faster, and to a higher recovery level: $90\%$ compared with $97\%$.
+the passthrough functions. It learned faster, and to a higher recovery level: $97\%$ compared with $90\%$.
 Note: we truncate the results to the point we stopped the passthrough experiment. The reader can find the
@@
 
-\subsection{Effect of observability on the recording function and training results.}
+\subsection{Effect of observability on the recording function and training results}
 \label{sup:observability}
@@
 function, up to the point that only a small number of neurons can be 
-observed. In all cases except the 
-case of a single electrode, the co-processor learned to improve task
-performance to the $90\%$ recovery point. Some variation in learning
-efficiency occurred, and perhaps notably the experiments with $1$, $10$
-electrodes observed the lowest two rates of recovery. The passthrough 
+observed. In all cases except the case of a single electrode, the co-processor
+learned to improve task performance to the $90\%$ recovery point. Some variation
+in learning efficiency occurred, and perhaps notably the experiments with $1$ and
+$10$ electrodes observed the lowest two rates of recovery. The passthrough 
 function and $60$ electrode model observed the fastest recovery rates.
@@
 
-Class separation behavior remained largely the same across these variations. The one exception is the experiment
-involving only $1$ electrode per module. In that case, the co-processor
-slowly learned to treat the object classes similarly. That suggests that
-the classes could not be differentiated by the small amount of information
-attainable from a single electrode. In this case, the co-processor reduced
-towards a simpler closed-loop stimulator which modulates its behavior
-entirely on the user's volition to initiate movement, due to the hold
-signal being observable from every neuron.
+Class separation behavior remained largely the same across these variations.
+The one exception is the experiment involving only $1$ electrode per module.
+In that case, the co-processor learned to treat the object classes
+similarly. That suggests that the classes could not be differentiated by
+the small amount of information attainable from a single electrode. In this
+case, the co-processor reduced towards a simpler closed-loop stimulator
+which modulates its behavior entirely on the user's volition to initiate
+movement, due to the hold signal being observable from every neuron.
 
@@
          increases, though perhaps somewhat faster. We did not hypothesis test that effect.
-         In all cases except ``1'', learning reached our $90\%$ recovery threshold where
+         In all cases except ``1'', learning reached a $90\%$ recovery threshold where
          we stop training. (b) As in our connection-based lesion experiment above
@@
 
-\subsection{Table of losses and recovery.}
+\subsection{Stopping criteria}
+\label{sup:stopping_criteria}
+
+Both stopping criteria affected the results presented above. We cannot economically perform
+statistical analysis to show e.g. the percentage of runs of each experiment type which
+are stopped by one criteria versus the other. However, in Table \ref{tab:stopping_criteria}
+we present the stopping criterion which caused the end of each experiment whose results
+we presented herein.
+
+We used an alternate stopping condition for the sensor drift experiment, since recovery improved
+more slowly in that experiment due to the sensor drift. We allowed the experiment to run until
+roughly 500k epochs - essentially double the 250k epoch criterion.
+
+We also used an alternate stopping condition for the passthrough functions experiment. This
+was due to the computational expense: running the models at full resolution required far
+greater run time. As a result, we stopped this experiment once $97\%$ recovery on the
+validation data was achieved.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\multicolumn{1}{|c|}{\textbf{Experiment}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Loss\\ change \%\\ (default: 0.1\%)\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Run length\\ (default: 250k\\ epochs)\end{tabular}}} & \multicolumn{1}{c|}{\textbf{Notes}}                                                                       \\ \hline
+AIP No-coadapt                            & X                                                                                                      &                                                                                                                &                                                                                                           \\ \hline
+AIP Coadapt                               &                                                                                                        & X                                                                                                              &                                                                                                           \\ \hline
+M1 No-coadapt                             &                                                                                                        & X                                                                                                              & \begin{tabular}[c]{@{}l@{}}Pct change criteria relaxed to\\show longer-run S trend\end{tabular}            \\ \hline
+M1 Coadapt                                &                                                                                                        & X                                                                                                              &                                                                                                           \\ \hline
+Con No-coadapt                            & X                                                                                                      &                                                                                                                & Loss chg threshold: 0.05\%                                                                                \\ \hline
+Con Coadapt                               &                                                                                                        & X                                                                                                              & Loss chg threshold: 0.05\%                                                                                \\ \hline
+Recovery                                  &                                                                                                        & X                                                                                                              &                                                                                                           \\ \hline
+Sensor drift                              & X                                                                                                      &                                                                                                                & \begin{tabular}[c]{@{}l@{}}Allowed to run to $\sim$500k epochs,\\due to slower convergence.\end{tabular} \\ \hline
+Passthrough                               &                                                                                                        &                                                                                                                & \begin{tabular}[c]{@{}l@{}}Stopped at 97\% recovery,\\due to computational expense.\end{tabular}         \\ \hline
+\end{tabular}
+\caption{\label{tab:stopping_criteria} \textbf{Stopping criterion which bound each experiment.}
+         ``Coadapt'' refers to co-adaptation. ``Con'' refers to F5-M1 connection lesions.
+         ``Recovery'' refers to the experiment involving recovery prior to co-processor training.}
+\end{table}
+
+\subsection{Prediction Accuracy of Stimulation Model (EN)}
+\label{sup:pred_losses}
+
+As mentioned in Section~\ref{sec:en_training} above, we train our stimulation model, i.e. EN, to a high
+level of predictive power before using it to train the CPN. We expire it and create a new one when training
+performance begins to suffer, or prediction error significantly disimproves. To depict the EN's predictions
+visually, Fig. \ref{fig:pred_loss_examples} shows example muscle velocities for a highly trained CPN and EN,
+for one example output channel.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.8]{pred_loss_min_max.png}
+\caption{\textbf{Prediction errors for four trials, on TMAJ muscle (retracts the shoulder).}
+         The EN's predicted trajectory for the TMAJ muscle (blue), closely tracks the actual
+         trajectory (orange). Sourced from the co-adaptive connection lesion experiment.
+         (a) depicts the trial with minimum prediction MSE. (d) maximum prediction
+         MSE. (b, c) two other arbitrary-chosen trials. Standard deviation across all trials:
+         $1.91\mathrm{e}{-5}$.}
+\label{fig:pred_loss_examples}
+\end{figure}
+
+In Fig. \ref{fig:pred_loss_cv} we see the variability of prediction losses across muscle
+channels. Prediction loss typically varies $\pm34.0\%$ for a given channel across all
+trials. Variability is roughly consistent across channels, suggesting that the statistic
+is not driven by extrema. However, there exist a handful of channels with high
+coefficients of variation (CVs). Those appear to be the channels with the lowest overall
+prediction losses, suggesting their CVs are driven by the low mean, rather than an unusually
+high variability.
+
+\begin{figure}
+	\centering
+	\begin{subfigure}[c]{0.48\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{pred_loss_cv_hist.png}
+		\caption{\textbf{Histogram of coefficients of variability (CVs) across prediction MSE
+                 losses for each muscle channel.} Mean: $0.340$; i.e. per-channel loss
+                 typically varies $\pm34.0\%$.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[c]{0.48\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{pred_loss_cv_v_mse.png}
+		\caption{\textbf{Prediction MSE loss vs CV for each channel.} High CV extrema are from channels
+                 with the lowest losses. i.e. the highest CV values in (a) are due to extreme
+                 low losses, suggesting their CVs are driven by the low mean, rather than an
+                 unusually high variability.}
+	\end{subfigure}
+	\hfill
+\caption{\textbf{Prediction loss variability across muscle channels}}
+\label{fig:pred_loss_cv}
+\end{figure}
+
+\subsection{Table of losses and recovery}
 \label{sup:results}
@@
 \hline
-\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
-AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
-AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
+\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
+AIP No-coadapt & 0.004507                                                          & 0.003916                                                         & 0.003960                                                             & 15.01\%            & 13.89\%                                                          \\ \hline
+AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.23\%            & 76.26\%                                                          \\ \hline
 M1 No-coadapt  & 0.021136                                                          & 0.005765                                                         & 0.005604                                                             & 74.73\%            & 75.51\%                                                          \\ \hline
-M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
-Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
-Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
-Recovery*      & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
-Sensor drift   & 0.020719                                                          & 0.002966                                                        & 0.002825                                                              & 88.10\%            & 88.83\%                                                          \\ \hline
-Passthrough    & 0.020719                                                          & 0.001095                                                        & 0.001132                                                              & 97.37\%            & 97.19\%                                                          \\ \hline
+M1 Coadapt     & 0.021136                                                          & 0.004383                                                         & 0.004385                                                             & 81.39\%            & 81.41\%                                                          \\ \hline
+Con No-coadapt & 0.020719                                                          & 0.002540                                                         & 0.002490                                                             & 90.21\%            & 90.45\%                                                          \\ \hline
+Con Coadapt    & 0.020719                                                          & 0.002350                                                         & 0.002521                                                             & 91.15\%            & 90.30\%                                                          \\ \hline
+Recovery*      & 0.003834                                                          & 0.003136                                                         & 0.003099                                                             & 18.20\%            & 19.17\%                                                          \\ \hline
+Sensor drift   & 0.020719                                                          & 0.002966                                                        & 0.002825                                                              & 88.13\%            & 88.83\%                                                          \\ \hline
+Passthrough    & 0.020719                                                          & 0.001095                                                        & 0.001132                                                              & 97.00\%            & 96.83\%                                                          \\ \hline
 \end{tabular}
@@
          ``Coadapt'' refers to co-adaptation. ``Con'' refers to F5-M1 connection lesions. ``Recovery'' refers to the
-         experiment involving recovery prior to co-processor training.
-         For further exploration of the results, see Results.}
+         experiment involving recovery prior to co-processor training. For further exploration of the results, see Results.}
 \end{table}

### 2023-03-18 paper/v1.tex
v2 paper, after response to reviewers.

--- 2022-10-22 paper/article.tex
@@
 % Authorship temporarily removed for double-anonymous review
-\author{Matthew J Bryan$^{1}$, Linxing Preston Jiang$^{1}$, Rajesh P N Rao$^{1}$}
-
-\address{$^{1}$ Neural Systems Laboratory, Paul G. Allen School of Computer
-Science \& Engineering, University of Washington, Box 352350,
-Seattle, WA 98105, USA}
-
-\ead{\{mmattb,prestonj,rao\}@cs.washington.edu}
+%\author{Matthew J Bryan$^{1}$, Linxing Preston Jiang$^{1}$, Rajesh P N Rao$^{1}$}
+
+%\address{$^{1}$ Neural Systems Laboratory, Paul G. Allen School of Computer
+%Science \& Engineering, University of Washington, Box 352350,
+%Seattle, WA 98105, USA}
+
+%\ead{\{mmattb,prestonj,rao\}@cs.washington.edu}
 \vspace{10pt}
@@
 		\centering
-		\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.png}
+		\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
 		\caption{}
@@
 \section{Data availability statement}
-Code including analysis code used to
-generate figures can be found at
-\url{https://github.com/mmattb/coproc-poc}.
+%Code including analysis code used to
+%generate figures can be found at
+%\url{https://github.com/mmattb/coproc-poc}.
 Data is available upon reasonable request to the authors.
@@
 
-\section{Acknowledgements}
-This work was supported by a Weill Neurohub Investigator grant, a CJ and Elizabeth Hwang
-endowed professorship (RPNR), and National Science Foundation grant no. EEC-1028725.
-The authors would like to thank Karunesh Ganguly (UC San Francisco), Priya Khanna
-(UC Berkeley), Anca Dragan (UC Berkeley), Justin Ong, and Luciano De La Iglesia
-for discussions and insights related to this work.
+%\section{Acknowledgements}
+%This work was supported by a Weill Neurohub Investigator grant, a CJ and Elizabeth Hwang
+%endowed professorship (RPNR), and National Science Foundation grant no. EEC-1028725.
+%The authors would like to thank Karunesh Ganguly (UC San Francisco), Priya Khanna
+%(UC Berkeley), Anca Dragan (UC Berkeley), Justin Ong, and Luciano De La Iglesia
+%for discussions and insights related to this work.
 

### 2023-03-18 cpn_epoch_cpn.py
Paint it black

@@ -39,9 +39,11 @@ class CPNEpochCPN:
         if cfg.recover_after_lesion:
             self.set_opt_lr = self.lr_sched_aggressive_refine3
             self.cpn_epoch_max_len = 200
-        elif (type(cfg.lesion_instance) is lesion.LesionType.outputs.value and
-                cfg.lesion_instance.start_idx == 0 and
-                cfg.lesion_instance.end_idx == 50):
+        elif (
+            type(cfg.lesion_instance) is lesion.LesionType.outputs.value
+            and cfg.lesion_instance.start_idx == 0
+            and cfg.lesion_instance.end_idx == 50
+        ):
             # M1 is an odd beast...
             self.set_opt_lr = self.lr_sched_aggressive_refine3
             self.cpn_epoch_max_len = 100

### 2023-03-18 cpn_epoch_en.py

@@ -160,7 +160,9 @@ class CPNEpochEN:
             en_is_ready = False
 
         # Every 10 epochs let's validate/test
-        next_is_validation = not en_is_ready and not is_validation and (self.checkpoint_eidx % 10) == 0
+        next_is_validation = (
+            not en_is_ready and not is_validation and (self.checkpoint_eidx % 10) == 0
+        )
 
         user_data = CPNENStats(
             "en_offline" if reused_data else "en",

### 2023-03-18 experiment/experiment.py

@@ -410,8 +410,9 @@ class Experiment:
 
             for tidx in range(steps - 1):
                 obs_raw = self.mike.observe(
-                    self.observer, drop_module_idx=self.obs_drop_module_idx,
-                    drop_neuron_mask=self.drop_neuron_mask
+                    self.observer,
+                    drop_module_idx=self.obs_drop_module_idx,
+                    drop_neuron_mask=self.drop_neuron_mask,
                 )
                 obs = obs_raw + (trial_end[:, tidx, :],)
 

### 2023-03-18 experiment/mRNN.py

@@ -99,16 +99,16 @@ class MichaelsRNN(nn.Module):
                 for j in range(npm):
                     possible_connections.append((i, j))
 
-            for c12 in random.sample(possible_connections, int(sparsity * npm ** 2)):
+            for c12 in random.sample(possible_connections, int(sparsity * npm**2)):
                 sparse_mask[c12[0] + npm, c12[1]] = 1
 
-            for c23 in random.sample(possible_connections, int(sparsity * npm ** 2)):
+            for c23 in random.sample(possible_connections, int(sparsity * npm**2)):
                 sparse_mask[c23[0] + (npm * 2), c23[1] + npm] = 1
 
-            for c21 in random.sample(possible_connections, int(sparsity * npm ** 2)):
+            for c21 in random.sample(possible_connections, int(sparsity * npm**2)):
                 sparse_mask[c21[0], c21[1] + npm] = 1
 
-            for c32 in random.sample(possible_connections, int(sparsity * npm ** 2)):
+            for c32 in random.sample(possible_connections, int(sparsity * npm**2)):
                 sparse_mask[c32[0] + npm, c32[1] + (npm * 2)] = 1
 
             # Zero grad mask: used to 0 out gradients during training,

### 2023-03-18 experiment/observer.py

@@ -4,6 +4,7 @@ import torch
 
 from .utils import gaussian_array_weights
 
+
 class Observer(object):
     def __init__(self):
         self._out_dim = None

### 2023-03-18 run_nocoadap_phase1.py

@@ -16,10 +16,10 @@ import experiment
 import cpn
 
 # Connection
-#cfg = experiment.experiment.get_config(coadapt=False, cuda="1")
+# cfg = experiment.experiment.get_config(coadapt=False, cuda="1")
 
 # AIP
-#cfg = experiment.experiment.get_aip_lesion_config(cuda="3", coadapt=True)
+# cfg = experiment.experiment.get_aip_lesion_config(cuda="3", coadapt=True)
 
 # M1
 cfg = experiment.experiment.get_m1_lesion_config(coadapt=False, cuda="3")
@@ -31,4 +31,3 @@ my_coproc = cpn.CPN_EN_CoProc(cfg, log_dir=LOG_DIR)
 my_experiment = experiment.experiment.stage(my_coproc, cfg)
 
 loss_history = my_experiment.run()
-

### 2023-03-18 run_nocoadap_phase2.py

@@ -16,7 +16,9 @@ import experiment
 import cpn
 
 # Connection
-cfg = experiment.experiment.get_m1_lesion_config(coadapt=True, dont_train=True, cuda="1")
+cfg = experiment.experiment.get_m1_lesion_config(
+    coadapt=True, dont_train=True, cuda="1"
+)
 
 
 LOG_DIR = "/home/mbryan/coproc-poc/models/gaussian20.1.75_outputsIdxs0.50_gaussianExp16.sig2.175.decay0.3_enActTanh_cpnActTanh/d57bd818-f5f7-43d9-aef3-c075075f415d"
@@ -26,4 +28,3 @@ my_coproc = cpn.CPN_EN_CoProc(cfg, log_dir=LOG_DIR)
 my_experiment = experiment.experiment.stage(my_coproc, cfg)
 
 loss_history = my_experiment.run()
-

### 2023-05-04 paper/article.tex
Final revision updates.
.tex may not match final text online, due to back-and-forth with editors.

@@
 		\centering
-		\includegraphics[width=\textwidth]{weill_arch.png}
+		\includegraphics[width=\textwidth]{weill_arch.eps}
 		\caption{}
@@
 		\centering
-		\includegraphics[width=\textwidth]{mRNN_J.png}
+		\includegraphics[width=\textwidth]{mRNN_J.eps}
 		\caption{}
@@
 modules in the model using the cortical areas (AIP, F5, M1) they correspond to. For details on how this network model 
-was trained, please see Supplementary Materials \ref{sup:michaelstraining}. For additional details on the task structure, see
-Supplementary Materials \ref{sup:michaelstask}.
+was trained, please see Appendix \ref{sup:michaelstraining}. For additional details on the task structure, see
+Appendix \ref{sup:michaelstask}.
 
@@
 higher loss than overall loss. However, the M1 lesion shows the highest overall loss; see Table
-\ref{tab:results} in Supplement materials section \ref{sup:results} for detailed lesion loss data.
+\ref{tab:results} in Appendix \ref{sup:results} for detailed lesion loss data.
 
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
+	    \includegraphics[width=\textwidth]{lesion_trajs.eps}
 	    \caption{}
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
+	    \includegraphics[width=\textwidth]{lesion_hand.eps}
 	    \caption{}
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{perturbe_single_tmaj.pdf}
+	    \includegraphics[width=\textwidth]{perturbe_single_tmaj.eps}
 	    \caption{}
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{perturbe_dist_tmaj.pdf}
+	    \includegraphics[width=\textwidth]{perturbe_dist_tmaj.eps}
 	    \caption{}
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{perturbe_single_eip.pdf}
+	    \includegraphics[width=\textwidth]{perturbe_single_eip.eps}
 	    \caption{}
@@
 	    \centering
-	    \includegraphics[width=\textwidth]{perturbe_dist_eip.pdf}
+	    \includegraphics[width=\textwidth]{perturbe_dist_eip.eps}
 	    \caption{}
@@
 in addition to approximating the cortical dynamics of the brain and the mapping of those dynamics onto
-the grasping outputs.\footnote{In Supplementary Materials section \ref{sup:passthrough}, we explore  
+the grasping outputs.\footnote{In Appendix \ref{sup:passthrough}, we explore
 other stimulation models.}
@@
 		\centering
-		\includegraphics[width=\textwidth]{stim_single.png}
+		\includegraphics[width=\textwidth]{stim_single1.eps}
 		\caption{$\sigma=1.875$}
@@
 		\centering
-		\includegraphics[width=\textwidth]{stim_single2.png}
+		\includegraphics[width=\textwidth]{stim_single2.eps}
 		\caption{$\sigma=2.875$}
@@
 	\centering
-	\includegraphics[width=\textwidth]{stim_and_obs.png}
+	\includegraphics[width=\textwidth]{stim_and_obs.eps}
 	\caption{\textbf{Example of stimulation and recording for a single trial.} Here, M1
@@
 stimulation pattern. Our simulated recordings are similar in spirit, though not actually modeling,
-local field potentials (LFPs) recorded extracellularly in biological neural tissue. In Supplementary Materials section
+local field potentials (LFPs) recorded extracellularly in biological neural tissue. In Appendix
 \ref{sup:observability}, we show that as we vary the observability of our ``simulated brain'' via this recording
@@
 add a random value to each element of the bias term, drawn from a zero-mean Gaussian distribution, causing the
-``recordings'' to drift over time. See Supplementary Materials section \ref{sup:drift} for further details.
+``recordings'' to drift over time. See Appendix \ref{sup:drift} for further details.
 
@@
 (this is a common stopping criterion in machine learning). Additional details can be found in the
-Supplementary Materials section ~\ref{sup:encpninter}.
+Appendix ~\ref{sup:encpninter}.
 
@@
 In these experiments, we train the co-processor until one of the following two stopping criteria is
-satisfied (in Supplementary Materials section \ref{sup:stopping_criteria}, we probe how these criteria
+satisfied (in Appendix \ref{sup:stopping_criteria}, we probe how these criteria
 affect the results). The two criteria are:
@@
 	      is below a threshold, indicating minimal benefits of further training. The choice of
-        threshold varies by experiment, due to the significant differences in setup. See Supplementary
-        Material section \ref{sup:stopping_criteria} for further details.
+              threshold varies by experiment, due to the significant differences in setup. See Appendix
+              \ref{sup:stopping_criteria} for further details.
 	\item The training run has exceeded some number of epochs. This also varied by experiment.
@@
 \centering
-\includegraphics[scale=1]{training_results_aip.png}
+\includegraphics[scale=1]{training_results_aip.eps}
 \caption{Co-Processor Performance Results: AIP 50\% Lesion. a. Percent loss recovery for the non-coadaptive
@@
 being unable to condition its grasp on the object shape, after largely succeeding in reaching toward the object.\footnote{
-In Supplementary Materials section \ref{sup:results}, lesioned loss is lower in this experiment than in others: the subject
+In Appendix \ref{sup:results}, lesioned loss is lower in this experiment than in others: the subject
 continues to reach successfully, but does not form the hand properly.} Likewise, the co-processor cannot observe sufficient
@@
 \centering
-\includegraphics[scale=1]{training_results_m1.png}
+\includegraphics[scale=1]{training_results_m1.eps}
 \caption{Co-Processor Performance Results: M1 50\% Lesion. a. Percent loss recovery
@@
 \centering
-\includegraphics[scale=1]{training_results_con.png}
+\includegraphics[scale=1]{training_results_con.eps}
 \caption{Co-Processor Performance Results: F5-M1 Connection Lesion. 
@@
 \centering
-\includegraphics[scale=1]{training_results_no_explore.png}
+\includegraphics[scale=1]{training_results_no_explore.eps}
 \caption{Co-Processor Performance Results: F5-M1 Connection Lesion: Use of exploratory
@@
 \centering
-\includegraphics[scale=1]{training_results_recov.png}
+\includegraphics[scale=1]{training_results_recov.eps}
 \caption{Co-Processor Performance Results: Connection Lesion and Natural Recovery before Co-Processor Use.
@@
 zero point across epochs. We added a bias term which we updated between epochs according to a random process
-(see Supplementary Materials section~\ref{sup:drift} for details).
+(see Appendix \ref{sup:drift} for details).
 
@@
 \centering
-\includegraphics[scale=1]{training_results_drift.png}
+\includegraphics[scale=1]{training_results_drift.eps}
 \caption{Co-Processor Performance Results: Connection Lesion with Sensor Drift. a.
@@
 	\centering
-	\includegraphics[width=0.88\textwidth]{cofiring_model.png}
+	\includegraphics[width=0.88\textwidth]{cofiring_model.eps}
 	\caption{\textbf{A co-processor leveraging a neural correlate of dexterity rather
@@
 
-\section{Supplementary Materials}
+\section{Appendix}
 \subsection{Details of the simulated task}
@@
 \centering
-\includegraphics[scale=0.7]{trial_timeline.png}
+\includegraphics[scale=0.7]{trial_timeline.eps}
 \caption{Timeline of a Single Trial}
@@
 Note: we truncate the results to the point we stopped the passthrough experiment. The reader can find the
-final results in Supplementary Materials \ref{sup:results}.
+final results in Appendix \ref{sup:results}.
 
@@
 \centering
-\includegraphics[scale=0.7]{training_results_passthrough.png}
+\includegraphics[scale=0.7]{training_results_passthrough.eps}
 \caption{\textbf{Training Results: passthrough recording and stimulation functions.} The significantly faster
@@
 \centering
-\includegraphics[scale=0.7]{training_results_observability.png}
+\includegraphics[scale=0.7]{training_results_observability.eps}
 \caption{\textbf{Training Results: varying observability of the brain.}
@@
 \centering
-\includegraphics[scale=0.8]{pred_loss_min_max.png}
+\includegraphics[scale=0.8]{pred_loss_min_max.eps}
 \caption{\textbf{Prediction errors for four trials, on TMAJ muscle (retracts the shoulder).}
@@
 		\centering
-		\includegraphics[width=\textwidth]{pred_loss_cv_hist.png}
+		\includegraphics[width=\textwidth]{pred_loss_cv_hist.eps}
 		\caption{\textbf{Histogram of coefficients of variability (CVs) across prediction MSE
@@
 		\centering
-		\includegraphics[width=\textwidth]{pred_loss_cv_v_mse.png}
+		\includegraphics[width=\textwidth]{pred_loss_cv_v_mse.eps}
 		\caption{\textbf{Prediction MSE loss vs CV for each channel.} High CV extrema are from channels
